### 2023-12-04 Code/MetaDataImport.Rmd
created rmd to read and filter culture catalog and save as .Rds

---
title: "MetaDataImport"
author:
- Mireille Savoie
- Douglas A. Campbell
date: "`r format(Sys.Date())`"
output:
bookdown::html_document2:
    code_folding: show
    keep_md: yes
    toc: TRUE
    toc_float: TRUE
    toc_depth: 6
    fig_caption: yes
bibliography: Prochlorococcus_O2_NPQ.bib
csl: plos-one.csl
editor_options: 
  markdown: 
    wrap: 72
---

# Set Options

## Set figure caption font size

```{css, echo=FALSE}
p.caption {
  font-size: 18px;
}
```

## Set Chunk Options

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
knitr::opts_chunk$set(fig.path='Figs/')
```

# Introduction

This RMD imports culture meta data from a google sheet.

# Load Libraries and set Project Variables

```{r load libraries}
# libraries; Note check actual dependencies
library(tidyverse)
library(lubridate)
library(knitr)
library(data.table)
library(googledrive)
library(googlesheets4)

```

```{r set project variables}
#"..", takes up a level in the directory path
Project <- "PICO"
DataOut <- file.path("..","Data")

```

# Import MetaData 
```{r load Catalog direct from googlesheet, results = "hide"}
# #implement read with googlesheet name instead of full url

gs4_deauth()

# this is the URL or ID of a Sheet readable by anyone (with a link)

CultureCatalog <- read_sheet("https://docs.google.com/spreadsheets/d/1ZXpwR7Gfto-uRzVdXzMpQF4frbrvMLH_IyLqonFZRSw/edit#gid=0") %>%

  # read_sheet has an annoying "feature" to set the type of columns it can't parse to a list.
  # ggplot/dplyr doesn't like working with a dataframe of lists.
  # In this case WL is set to a list since some values are numbers, some are strings, some are blank.
  # To fix this, first drop all rows missing WL, then unlist.
  # Must first drop NA rows since unlist will collapse NULL lists, then the unlisted WL is a shorter length than original WL column, which mutate doesn't like.
  drop_na(WL) %>%
  mutate(WL = unlist(WL)) %>%
   mutate(ExpDate = ymd(ExpDate),
          ExpEndDate = ymd_hms(`ExpEndDate`)) %>% 
  filter(str_detect(ID,  "MiSa|JuNa|SaAr")) %>% 
  select(c("PrimaryOperator", "SampleID", "ID", "Strain", "ExpCul", "ExpDate", "ExpStartTime", "Par_ue", "Photoperiod", "Temp_c", "MC", "Tube", "EndDate","O2", "O2_Category", "WL","LightShape", "Optode","OptodeCh","OptodeMeasure", "ExpEndDate"))

saveRDS(object = CultureCatalog, file = file.path("..", "Data", "CultureCatalog.Rds"))

```