### 2023-12-04 Code/JazDataImport.Rmd
Fixed information extracted from filename column due to new data file path. Generated Emission spectra .Rds

---
title: "Import Jaz data"
output: html_notebook
---

```{r libraries}
#library(dplyr) 
library(tidyverse)
library(lubridate)
library(googledrive)
library(googlesheets4)

library(gghighlight)
library(ggrepel)

#library(ggplot2)
library(scales)
library(photobiology)
library(photobiologyWavebands)
library(ggspectra)
library(ggrepel)
```

```{r import the data}
Project <- "PICO"

DataIn <- file.path("..", "Data","RawData", "MCEmissionSpectra", fsep = .Platform$file.sep)

DataOut<- file.path("..", "Data", "ProcessedData", fsep = .Platform$file.sep)

#set variables for file import & processing of fluorescence files Jaz
#DataPathCS <- file.path("..", "Jaz", fsep = .Platform$file.sep)
#FileIDSingle <- "202205201054_PICO_SySl1246_445_LL_Subt6_10-59-24-712"

Jazfile <- list.files(path = DataIn, pattern = ".txt", full.names = TRUE)

Jazfile[1:10]

unique(duplicated(Jazfile))
SkipCS <- 14
FileEncodeCS <- "UTF-8"
DelimCS <- "\t"
RollWindow <- 10

```

```{r set colours}
Wavelengths_nm = c(405, 450, 470, 530, 620, 660, 720, "WW")
Colours_nm = c("yellow", "dodgerblue", "darkblue", "yellowgreen", "darkgreen", "Red",  "darkorange", "black")

names(Colours_nm) <- Wavelengths_nm
Colours_nm
```

```{r read multiculti PAR data}

MultiCultiPAR <- read.csv(file.path(DataIn, "MultiCulti_PAR.csv"))

```

```{r read Jaz data using map_df}

fread_plus <- function(Flnm, Skip, FileEncode, Delim){data.table::fread(file = Flnm, skip = Skip, encoding = FileEncode, sep = Delim, header = FALSE) %>%
    mutate(Filename = Flnm, CDateTime = ymd_hms(file.info(Flnm)$ctime))
}

JazSpec <- Jazfile %>%
  map_df(~fread_plus(Flnm = ., Skip = SkipCS, FileEncode = FileEncodeCS, Delim = DelimCS))

# JazSpec <- JazSpec %>%
#   filter(!str_detect(V2, pattern = "Content"))

colnames(JazSpec)[1] <- "Wavelength"
colnames(JazSpec)[2] <- "Counts"

JazSpec[1:10]

# JazSpec |>
#   ggplot() +
#   geom_point(aes(x = Wavelength, y = Counts)) +
#   theme_bw()

```

```{r filename separation}

JazSpec <- JazSpec %>% 
  mutate(Filename = str_remove(Filename, pattern = "../Data/RawData/MCEmissionSpectra/")) |>
  separate(col = Filename, into = c("Date", "MultiCulti", "Par", "WL"), sep = "([\\/\\/\\_\\_\\_\\_\\_\\_\\.])", remove = FALSE)

# JazSpecTrim <- JazSpecTrim %>% 
# select(-fp1, -fp2, -fp3, -txt, -date_experiment, -device2, -Project, -CDateTime, -Subt, -subseconde)

JazSpec <- JazSpec %>% 
mutate(Date = ymd(Date),
       Par = as.numeric(Par),
       WL = as.factor(WL))

```

```{r remove wl < 350 and > 750}

JazSpecTrim <- JazSpec %>% 
  subset(JazSpec$Wavelength >= 350 & JazSpec$Wavelength <= 750)

```

Remove all NA columns & empty rows, fix column types
```{r remove missing columns}

#https://community.rstudio.com/t/drop-all-na-columns-from-a-dataframe/5844/2
not_all_na <- function(x) {!all(is.na(x))}

JazSpecTrim <- JazSpecTrim %>%
  select_if(not_all_na)

# JazSpecTrim <- JazSpecTrim %>%
#   mutate(JazTimeHMS = lubridate::hms(stringr::str_sub(CDateTime, start = 1L, end = 8L)), .after = CDateTime)

# JazData2 <- JazData %>%
#   filter(!is.na(Wavelength))
# 
# JazSpecTrim <- JazSpecTrim %>%
#   mutate(Duration_s = as.numeric(Duration_s)) %>% 
#   mutate(AquisitionTime_ms = as.numeric(AquisitionTime_ms))

```

Find nm at max count and compare with rolling mean of counts
```{r, fig.height = 6, fig.width = 8}

JazSpecTrim <- JazSpecTrim %>% 
  group_by(WL, MultiCulti) %>%
   mutate(rollmeanCounts = zoo::rollmean(Counts, k = RollWindow, align = "left", fill = NA),
          is_nm = round(Wavelength[which.max(Counts)]),
          is_nm_rollmean= round(Wavelength[which.max(rollmeanCounts)])) %>% 
  ungroup()

```

```{r nest the spectral data}

JazSpecNest <- JazSpecTrim %>%
  nest(Spectra = c(Wavelength, Counts, rollmeanCounts))

```

```{r remove redundant DF}

remove(JazSpecTrim)
remove(JazSpec)

```

```{r annotate measured nm, fig.height = 6, fig.width = 8}

JazSpecNest %>%
  unnest(Spectra) %>%
    ggplot() +
  geom_point(aes(x = Wavelength, y = Counts, colour = as.factor(WL))) +
  # geom_point(aes(x = Wavelength, y = is_nm, colour = as.factor(WL))) +
  geom_text(aes(x = is_nm, y= max(Counts), label= is_nm), alpha = 0.5 ,hjust = 0.5, vjust=-0.5) +
  scale_color_manual(values = Colours_nm) +
  guides(colour = "none") +
  geom_vline(aes(xintercept = is_nm), linetype = "dashed") +
  scale_y_continuous(limits = c(0,100000)) +
  ggh4x::facet_nested(rows = vars(MultiCulti, WL), labeller = labeller(WL = label_both, Par = label_both, MultiCulti = label_both)) +
  theme_bw()

```

```{r estimate peak ratios}
# JazSpecWideNest <- JazSpecWide %>%
#   nest(Spectra = c(One_Wavelength, One_Counts, Two_Counts, Three_Counts, Four_Counts, CountsAv, CountsMax, CountsMin)) %>%
#   mutate(CountsBack = map_dbl(.x = Spectra, ~mean(.$CountsAv[.$One_Wavelength > 740])),
#     Counts680690 = map_dbl(.x = Spectra, ~mean(.$CountsAv[.$One_Wavelength > 680 & .$One_Wavelength < 690])),
#          Counts715725 = map_dbl(.x = Spectra, ~mean(.$CountsAv[.$One_Wavelength > 715 & .$One_Wavelength < 725])),
#          Counts685_720 = Counts680690/Counts715725,
#     Counts685_720back = (Counts680690 - CountsBack)/(Counts715725 - CountsBack)
#     )
# 
# JazSpecWideNest %>%
#   filter(Par_ue == 60) %>%
#   #filter(Ex_WL != 445) %>%
#   ggplot() +
#   geom_point(aes(x = Ex_WL, y = Counts685_720)) +
#   geom_point(aes(x = Ex_WL, y = Counts685_720back), colour = "red") +
#   scale_color_manual(values = Colours_nm) +
#   guides(colour = "none") +
#   #coord_cartesian(xlim = c(660, 750), ylim =  c(0,5000)) +
#   ggh4x::facet_nested(cols = vars(Strain, Par_ue, WL), labeller = labeller(strain = label_value, Par_ue = label_both, WL = label_both))  +
#   theme_bw()
```

XXXX6 June 2022XXX

Adding the maximum (peak) value in ggplot for geom_smoth
https://stackoverflow.com/questions/71162252/adding-the-maximum-peak-value-in-ggplot-for-geom-smoth
```{r}

# sm_max <- JazSpecTrim %>% 
#   group_by(WL) %>%
#   mutate(smooth =predict(loess(Counts~as.numeric(Wavelength), span=.05))) %>%
#   slice_max(order_by = smooth)
# 
# JazSpecTrim |>
#   group_by(WL) %>%
#   mutate(is_nm = Wavelength[which.max(Counts)]) %>%
#   ggplot(df, mapping = aes(Wavelength, Counts, color = MultiCulti)) +
#   geom_point(aes(x = Wavelength, y = Counts, colour = as.factor(MultiCulti))) +
#   scale_color_manual(values = Colours_nm) +
#   guides(colour = FALSE) +
#   scale_y_continuous(limits = c(0,60000)) +
#   ggh4x::facet_nested(rows = vars(WL), labeller = labeller(WL = label_both, strain = label_value, Par = label_both, WL = label_both))  +
#   geom_smooth(span=.05, se=F) +
#   geom_point(size = 3) +
#   gghighlight(is_max,
#               n = 1,
#               unhighlighted_params = list(alpha = 0),
#               label_key = Counts)

```

ggspectra:
https://cran.r-project.org/web/packages/ggspectra/vignettes/userguide1-grammar.html
https://cran.r-project.org/web/packages/ggspectra/vignettes/userguide2-autoplot-methods.html
```{r}

# JazSpecTrim |> 
#   # group_by(Filename) %>%
#     filter(
#          Ex_WL != "470",
#          Ex_WL != "505") %>%
#   #mutate(is_max = Counts == max(Counts)) %>%
#   ggplot(df, mapping = aes(Wavelength, Counts, color = Filename)) +
# 
#   geom_point(aes(x = Wavelength, y = Counts, colour = as.factor(Ex_WL))) +
#   scale_color_manual(values = Colours_nm) +
#   guides(colour = FALSE) +
# 
#   scale_y_continuous(limits = c(0,60000)) +
#   ggh4x::facet_nested(rows = vars(Ex_WL), cols = vars(Strain, Par_ue, WL), labeller = labeller(Ex_WL = label_both, strain = label_value, Par_ue = label_both, WL = label_both))  +
#   #geom_point(size = 3) +
#   geom_line() +
#   stat_peaks(geom = "point",
#              span = 51,
#              color = "red",
#              vjust = "bottom",
#              position = position_nudge(y = 0.01)) +
#   theme_bw() +
#   theme(axis.text.x = element_text(angle = 45))
# # 
# 
# 
# #Showed y-value on spectra
# ggplot(sun.spct) + geom_line() + 
#   stat_peaks(shape = 21, span = 35, size = 2) + 
#   stat_label_peaks(aes(label = stat(y.label)),
#                    span = 35, geom = "label", size = 3,
#                    position = position_nudge(y = 0.04),
#                    label.fmt = "%1.2f") +
#   expand_limits(y = 1) +
#   scale_fill_identity() + scale_color_identity()

```

```{r save rds}

 saveRDS(JazSpecNest, file.path(DataOut, paste(Project, "MultiCultiSpectra.Rds", sep = "_"), fsep = .Platform$file.sep))

```