### 2023-12-07 Code/MultiDataProcessPico.Rmd
Removed some extra plots

---
title: "MultiDataProcessPico"
author:
- Laurel Genge
- Carlie Barnhill
- Max Berthold
- Douglas A. Campbell
- Mireille Savoie
date: "`r format(Sys.Date())`"
output:
bookdown::html_document2:
    code_folding: show
    keep_md: yes
    toc: TRUE
    toc_float: TRUE
    toc_depth: 6
    fig_caption: yes
bibliography: Prochlorococcus_O2_NPQ.bib
csl: plos-one.csl
editor_options: 
  markdown: 
    wrap: 72
---

## Set Chunk Options
```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
knitr::opts_chunk$set(fig.path='Figs/')
```

# Introduction

Implement logistic and other growth curve fits to MultiCulti growth trajectories.

# Load Libraries and set project variables
Run only first time in session

```{r load libraries}
# libraries; Note check actual dependencies
library(tidyverse)
library(lubridate)
library(stringr)
library(broom)
library(knitr)
library(zoo)
library(minpack.lm)
library(photobiologyWavebands)

```

Run only first time in session

```{r set project variables}
#"..", takes up a level in the directory path
Project <- "PICO"
DataIn <- file.path("..", "Data", "ImportedData", "ImportedMCData")
DataOut <- file.path("..", "Data","ProcessedData", "ProcessedMCData")

InitialTime <- 24 #hours to filter off start of data set to eliminate initial fluctuations common in SS120

ReplaceOD <- 0.0001 #replace OD720 and/or deltaOD when log of original OD results in NAN or infinite #

TruncFactor = 0.5 # fraction of max OD for truncating full growth trajectory to compare mu.

GrowthFlagError <- 0.20 # percentage of growth parameter is flagged as growth when std error is less than xx of mu.

RollWindow = 20 #measurements every 5 min; 20 points is 1 h to generate rolling averages

OD720GrowthFlagValue = 0.010 # OD720 absolute change for setting GrowthFlag = 1 and DeclineFlag = 1; did culture show  growth;

deltaODGrowthFlagValue = 0.007 #deltaOD absolute change for setting GrowthFlag = 1 and DeclineFlag = 1; did culture show  growth;

GrowthAmpWindow = 14 * 24 # maximum hours to determine if cultures doubled.

MaxAmpHour <- 48  # used to determine if max OD was achieved within xx hours which then is used to calculate growthAmp on decline growth.

DeclineWindow <- 12 #duration in hours overwhich we test for decline from max by amplitude GrowthFlagValue in OD720 or deltaOD

```

```{r set colours}
MyWavelengths = c(405, 445, 470, 530, 620, 660, 730, "WW")
MCMIXColours = c(c(w_length2rgb(405), w_length2rgb(445), w_length2rgb(470), w_length2rgb(530),            w_length2rgb(620), w_length2rgb(660), w_length2rgb(730), "black"))

names(MCMIXColours) <- MyWavelengths
MCMIXColours
```

# List previously processed files

Run only first time in session

```{r previously processed files}
list.files(path = DataOut, pattern = Project, full.names = TRUE)

```

# List available tidied, organized .Rds files of MultiCulti data.

Exported from runs of MultiDataImport.Rmd Run only first time in session

```{r list tidied data files}
list.files(path = DataIn, pattern = Project, full.names = TRUE)

```

# Select and Read ProcessFile

Run for each file in a session; all chunks below run for each file.

Moving this template to a working project folder will need editing of this file path to a relative path within the project folder.
```{r read ProcessFile}
ProcessFile <- "../Data/ImportedData/ImportedMCData/20230210_PICO_MCMIX006_RUN108_TargetDataMetaFilter.Rds"

ProcessFileName <- str_split(string = ProcessFile, "/")[[1]][5] %>%
  str_remove(pattern = ".Rds") %>%
  str_remove(pattern = "_TargetDataMetaFilter")

ProcessFileName

ProcessData <- readRDS(ProcessFile)  %>%
  ungroup()

colnames(ProcessData)
```

Remove extraneous columns from ProcessData; could have been done at end of MultiDataImport.Rmd

SS120 show anomalous jump in OD720 within first 24 h after inoculation.
Therefore we add a filter to remove the first InitialTime  from all SS120 runs.
```{r filter initial data  ProcessData}

# ProcessData <- ProcessData %>%
#   filter(Strain == "SS120" & time >= InitialTime |
#          Strain == "MIT9313" |
#          Strain =="MED4"  )

```

```{r remove problem tube}

# ProcessData <- ProcessData %>%
#   filter(Tube !=4)

```

```{r filter time tube specific}

# ProcessData <- ProcessData %>%
#   filter(Tube == 1 |
#          Tube == 2 |
#          Tube == 3 & time >= 24 & time <= 125  |
#          Tube == 4 |
#          Tube == 5 |
#          Tube == 6 |
#          Tube == 7 |
#          Tube == 8 )

```

Add deltaOD columns; filter for failed ln; add lognorm columns; add lagseed for each tube
```{r add columns for fitting}

ProcessData <- ProcessData %>%
  mutate(deltaOD = OD680 - OD720) %>%
  filter(!is.infinite(log(OD720))) %>%
  filter(!is.infinite(log(OD680))) %>%
  filter(!is.nan(log(OD720))) %>%
  filter(!is.nan(log(OD680))) %>%
  mutate(#log of zero results in -infinite #'s. Replace with 0.0001. Replacing with 0 creates very low (10e-20) rollmeans which results in very high log normalized values further on in this chunk.
       deltaOD = if_else(!is.infinite(log(deltaOD)), deltaOD, ReplaceOD),
       deltaOD = if_else(!is.nan(log(deltaOD)), deltaOD, ReplaceOD)
        )

#lognormOD720 and deltaOD for Gompertz
#using RcppRoll functions to find the minimum of a k point window to lower noise in normalizations

ProcessData <- ProcessData %>%
  group_by(Tube) %>%
  mutate(rollmeanOD720 = zoo::rollmean(OD720, k = RollWindow, align = "left", fill = NA),
         rollmeandeltaOD = zoo::rollmean(deltaOD, k = RollWindow, align = "left", fill = NA),
         lognormOD720 = if_else(OD720 == (min(rollmeanOD720, na.rm = TRUE)), ReplaceOD, log(OD720/min(rollmeanOD720, na.rm = TRUE))),
         lognormdeltaOD = if_else(deltaOD == (min(rollmeandeltaOD, na.rm = TRUE)), ReplaceOD, log(deltaOD/min(rollmeandeltaOD, na.rm = TRUE)))) %>%
  mutate(LagSeedOD720 = time[which.min(rollmeanOD720)],
         LagSeeddeltaOD = time[which.min(rollmeandeltaOD)]
         ) %>%
  ungroup()

```

# Generate a plot of the data in the selected tidied MultiCulti file.
# Remember data has been filtered to remove first InitialTime h after start of Multiculti program, which is actually InitialTime + ~18 h after inoculation.

```{r prelim ProcessPlot}

ProcessData %>%
  ggplot() +
  geom_point(aes(x = time, y = OD720), size = 0.1) +
  #geom_point(aes(x = time, y = Actinic_par/1000, colour = as.factor(WL)), size = 0.05) +
   geom_vline(aes(xintercept = 1 * 24), linetype = "dashed") + 
  scale_colour_manual(values = MCMIXColours) +
  facet_grid(rows = vars(O2), cols = vars(as.factor(Tube), Strain, Par_ue)) +
  theme_bw()

ProcessData %>%
 # filter(Tube == 1) %>%
  ggplot() +
  geom_point(aes(x = time, y = deltaOD), size = 0.1) +
   geom_vline(aes(xintercept = 1 * 24), linetype = "dashed") + 
  #geom_point(aes(x = time, y = Actinic_par/100, colour = as.factor(WL)), size = 0.05) +
  scale_colour_manual(values = MCMIXColours) +
  facet_grid(rows = vars(O2), cols = vars(as.factor(Tube), Strain, Par_ue)) +
  theme_bw()

```

# For split files only; select the interval that covers before the change in actinic light
```{r Interval Data for split files}
#even multiples of 24 + StartHour to start at photoperiod
# LowerTime <- 24 * 0
# UpperTime <- 24 * 17
# #
# ProcessData <- ProcessData %>%
#   filter(time <= UpperTime)
```

```{r,  remove  decline growth data. Filter x hr after max OD720 and deltaOD }

ProcessData <- ProcessData %>%
  group_by(Tube) %>%
   mutate(TrimRollMeanOD720 = zoo::rollmean(OD720, k = RollWindow, align = "right", fill = NA),
          TrimRollMeandeltaOD = zoo::rollmean(deltaOD, k = RollWindow, align = "right", fill = NA)
          ) %>%
  mutate(DeclineFlagdeltaOD = if_else(time > (time[which.max(TrimRollMeandeltaOD)] + DeclineWindow) & (max(TrimRollMeandeltaOD, na.rm = TRUE) - min(TrimRollMeandeltaOD, na.rm = TRUE)) > deltaODGrowthFlagValue, 1, 0),
         DeclineFlagOD720 = if_else(time > (time[which.max(TrimRollMeanOD720)] + DeclineWindow) & (max(TrimRollMeanOD720, na.rm = TRUE) - min(TrimRollMeanOD720, na.rm = TRUE)) > OD720GrowthFlagValue, 1, 0)) %>%
    filter(DeclineFlagdeltaOD == 0,
           DeclineFlagOD720 == 0) %>%
  select(-c("TrimRollMeanOD720", "TrimRollMeandeltaOD")) %>%
  ungroup()

glimpse(ProcessData)

```

```{r add min meanOD}

ProcessData <- ProcessData %>%
  group_by(Tube) %>%
   mutate(minRollMeanOD720 = min(rollmeanOD720, na.rm = TRUE),
          minRollMeandeltaOD = min(rollmeandeltaOD, na.rm = TRUE)) %>% 
  ungroup()

glimpse(ProcessData)

```

```{r plot filtered data}

ProcessData %>%
  ggplot() +
  geom_point(aes(x = time, y = OD720), size = 0.1) +
 # geom_point(aes(x = time, y = Actinic_par/1000, colour = as.factor(WL)), size = 0.05) +
   geom_vline(aes(xintercept = 1 * 24), linetype = "dashed") + 
  geom_vline(aes(xintercept = 7 * 24), linetype = "dashed") + 
  geom_vline(aes(xintercept = 14 * 24), linetype = "dashed") + 
  scale_colour_manual(values = MCMIXColours) +
  facet_grid(rows = vars(O2), cols = vars(as.factor(Tube), Strain, Par_ue)) +
  theme_bw()

ProcessData %>%
 # filter(Tube == 1) %>%
  ggplot() +
  geom_point(aes(x = time, y = deltaOD), size = 0.1) +
  #geom_point(aes(x = time, y = Actinic_par/100, colour = as.factor(WL)), size = 0.05) +
   geom_vline(aes(xintercept = 1 * 24), linetype = "dashed") + 
  geom_vline(aes(xintercept = 7 * 24), linetype = "dashed") + 
  geom_vline(aes(xintercept = 14 * 24), linetype = "dashed") + 
  scale_colour_manual(values = MCMIXColours) +
  facet_grid(rows = vars(O2), cols = vars(as.factor(Tube), Strain, Par_ue)) +
  theme_bw()

```

# Fits of Data
Create R function for logistic equation.
```{r fitting functions, message = FALSE}
#define a logistic equation as a function.
#x will be taken from 'time' when we run the fit.
logistic_eqn <- function(x, Lmax, Lmu, Lintercept){(Lmax*Lintercept*exp(Lmu*x))/(Lmax + (Lintercept*(exp(Lmu*x)-1)))}

#Modified Gompertz equation
ModGompertzEqn <- function(x,GLag,Gmax,Gmu){(Gmax*(exp(-exp((Gmu*exp(1))/Gmax*(GLag-x)+1))))}

# Exponential equation 
exp_eqn <- function(x, Emu, Eintercept){(Eintercept*exp(x*Emu))}

#Estimate absolute amplitude of growth
growth_amp <- function(y){(max(as.numeric(y), na.rm = TRUE) - min(as.numeric(y), na.rm = TRUE))}

#returns a negative absolute amplitude of growth if OD decline
growth_amp_decline <- function(y){(min(as.numeric(y), na.rm = TRUE)) - max(as.numeric(y), na.rm = TRUE)}

```

```{r nested regressions }

#nest must include all variables that have more than one value for a given tube nest.
#filtering to remove negative lognorm data only upon transition to ProcessDataNestGrowth and fitting; could do earlier?
ProcessDataNestGrowth  <- ProcessData %>%
  filter(lognormdeltaOD >= 0) %>%
  filter(lognormOD720 >= 0) %>%
nest(tubedata = c(time, ToD, Day, Actinic_par, OD680, OD720, deltaOD, ActinicMid_parOD720, ActinicMid_dayOD720,rollmeanOD720,rollmeandeltaOD,lognormOD720, lognormdeltaOD, LagSeedOD720, LagSeeddeltaOD))

#test of nesting
nrow(ProcessDataNestGrowth)

ProcessDataNestGrowth <- ProcessDataNestGrowth %>%
  mutate(OD720_logistic = map(tubedata, possibly(~nlsLM(OD720 ~ logistic_eqn(x = time, Lmax, Lmu, Lintercept),
                            data = .x,
                            start = list(Lmax = max(.$rollmeanOD720, na.rm = TRUE),  Lmu = (log(max(.$rollmeanOD720, na.rm = TRUE)) - log(min(.$rollmeanOD720, na.rm = TRUE)))/max(.$time), Lintercept = min(.$rollmeanOD720, na.rm = TRUE)),
                            control = list(maxiter = 500)), otherwise = NULL)),
         OD720_logistic_tidied =  map(OD720_logistic, possibly(tidy, otherwise = NULL)),
         OD720_logistic_param = map(OD720_logistic,possibly(glance, otherwise = NULL)),
         OD720_logistic_predict = map(OD720_logistic, possibly(augment, otherwise = NULL))
         ) %>%
  mutate(OD720_TimetoMax = map(.x = OD720_logistic_predict, possibly(~.$time[which.max(.$`.fitted`)], otherwise = NULL))  
         ) %>% 
  mutate(OD720_TimetoMax = as.numeric(ifelse(OD720_TimetoMax == "NULL" , NA, OD720_TimetoMax))
         ) %>%
  mutate(OD720trunc_logistic = map(tubedata, possibly(~nlsLM(OD720 ~ logistic_eqn(x = time, Lmax, Lmu, Lintercept),
                            data = .x %>%
                            filter(OD720 <= (max(OD720) * TruncFactor)),
                            start = list(Lmax = max(.$rollmeanOD720, na.rm = TRUE),  Lmu = (log(max(.$rollmeanOD720, na.rm = TRUE)) - log(min(.$rollmeanOD720, na.rm = TRUE)))/max(.$time), Lintercept = min(.$rollmeanOD720, na.rm = TRUE)),
                            control = list(maxiter = 500)), otherwise = NULL)),
         OD720trunc_logistic_tidied =  map(OD720trunc_logistic, possibly(tidy, otherwise = NULL)),
         OD720trunc_logistic_param = map(OD720trunc_logistic,possibly(glance, otherwise = NULL)),
         OD720trunc_logistic_predict = map(OD720trunc_logistic, possibly(augment, otherwise = NULL))
         ) %>%
  mutate(OD720trunc_TimetoMax = map(.x = OD720trunc_logistic_predict, possibly(~.$time[which.max(.$`.fitted`)], otherwise = NULL))  
         ) %>%  
  mutate(OD720trunc_TimetoMax = as.numeric(ifelse(OD720trunc_TimetoMax == "NULL" , NA, OD720trunc_TimetoMax))
         ) %>%
   mutate(deltaOD_logistic = map(tubedata, possibly(~nlsLM(deltaOD ~ logistic_eqn(x = time, Lmax, Lmu, Lintercept),
                            data = .x,
                            start = list(Lmax = max(.$rollmeandeltaOD, na.rm = TRUE),  Lmu = (log(max(.$rollmeandeltaOD, na.rm = TRUE)) - log(min(.$rollmeandeltaOD, na.rm =  TRUE)))/max(.$time), Lintercept = min(.$rollmeandeltaOD, na.rm = TRUE)),
                            control = list(maxiter = 500)), otherwise = NULL)),
         deltaOD_logistic_tidied =  map(deltaOD_logistic, possibly(tidy, otherwise = NULL)),
         deltaOD_logistic_param = map(deltaOD_logistic,possibly(glance, otherwise = NULL)),
         deltaOD_logistic_predict = map(deltaOD_logistic, possibly(augment, otherwise = NULL))
         ) %>%
  mutate(deltaOD_TimetoMax = map(.x = deltaOD_logistic_predict, possibly(~.$time[which.max(.$`.fitted`)], otherwise = NA))  
         ) %>% 
  mutate(deltaOD_TimetoMax = as.numeric(ifelse(deltaOD_TimetoMax == "NULL" , NA, deltaOD_TimetoMax))
         ) %>%
   mutate(deltaODtrunc_logistic = map(tubedata, possibly(~nlsLM(deltaOD ~ logistic_eqn(x = time, Lmax, Lmu, Lintercept),
                            data = .x %>%
                            filter(deltaOD <= (max(deltaOD) * TruncFactor)),
                            start = list(Lmax = max(.$rollmeandeltaOD, na.rm = TRUE),  Lmu = (log(max(.$rollmeandeltaOD, na.rm = TRUE)) - log(min(.$rollmeandeltaOD, na.rm = TRUE)))/max(.$time), Lintercept = min(.$rollmeandeltaOD, na.rm = TRUE)),
                            control = list(maxiter = 500)), otherwise = NULL)),
         deltaODtrunc_logistic_tidied =  map(deltaODtrunc_logistic, possibly(tidy, otherwise = NULL)),
         deltaODtrunc_logistic_param = map(deltaODtrunc_logistic, possibly(glance, otherwise = NULL)),
         deltaODtrunc_logistic_predict = map(deltaODtrunc_logistic, possibly(augment, otherwise = NULL))
         ) %>%
   mutate(deltaODtrunc_TimetoMax = map(.x = deltaODtrunc_logistic_predict, possibly(~.$time[which.max(.$`.fitted`)], otherwise = NULL))  
         ) %>%  
  mutate(deltaODtrunc_TimetoMax = as.numeric(ifelse(deltaODtrunc_TimetoMax == "NULL" , NA, deltaOD_TimetoMax))
         ) %>%
   mutate(GrowthAmpdeltaOD = if_else(deltaOD_TimetoMax <= MaxAmpHour | is.na(deltaOD_TimetoMax), map(tubedata, ~growth_amp_decline(y = .$rollmeandeltaOD)), map(tubedata, ~growth_amp(y = .$rollmeandeltaOD))),
          GrowthAmpOD720 = if_else(OD720_TimetoMax <= MaxAmpHour | is.na(OD720_TimetoMax), map(tubedata, ~growth_amp_decline(y = .$rollmeanOD720)), map(tubedata, ~growth_amp(y = .$rollmeanOD720)))
         ) %>%
   mutate(GrowthAmpdeltaODFlag = if_else(GrowthAmpdeltaOD >= deltaODGrowthFlagValue , 1, 0),
           GrowthAmpOD720Flag = if_else(GrowthAmpOD720 >= OD720GrowthFlagValue, 1, 0)
           )

#Growthamp flags for growth up to 14 days 
ProcessDataNestGrowth2 <- ProcessDataNestGrowth %>%
  unnest(c(tubedata)) %>%
   filter(time <= GrowthAmpWindow) %>%
  nest(tubedata = c(time, ToD, Day, Actinic_par, OD680, OD720, deltaOD, ActinicMid_parOD720,
                  ActinicMid_dayOD720,rollmeanOD720,rollmeandeltaOD,lognormOD720, lognormdeltaOD, LagSeedOD720, LagSeeddeltaOD)) %>%
   mutate(GrowthAmpdeltaOD_14days = if_else(deltaOD_TimetoMax <= MaxAmpHour | is.na(deltaOD_TimetoMax), map(tubedata, ~growth_amp_decline(y = .$rollmeandeltaOD)), map(tubedata, ~growth_amp(y = .$rollmeandeltaOD))),
          GrowthAmpOD720_14days = if_else(OD720_TimetoMax <= MaxAmpHour| is.na(OD720_TimetoMax) , map(tubedata, ~growth_amp_decline(y = .$rollmeanOD720)), map(tubedata, ~growth_amp(y = .$rollmeanOD720))),
          minRollMeanOD720_14days = map(tubedata, ~min(.$rollmeanOD720, na.rm = TRUE)),
          maxRollMeanOD720_14days = map(tubedata, ~max(.$rollmeanOD720, na.rm = TRUE)),
          minRollMeandeltaOD_14days = map(tubedata, ~min(.$rollmeandeltaOD, na.rm = TRUE)),
          maxRollMeandeltaOD_14days = map(tubedata, ~max(.$rollmeandeltaOD, na.rm = TRUE)),
          minRollMeanOD720_14days = as.numeric(minRollMeanOD720_14days),
          maxRollMeanOD720_14days = as.numeric(maxRollMeanOD720_14days),
          minRollMeandeltaOD_14days = as.numeric(minRollMeandeltaOD_14days),
          maxRollMeandeltaOD_14days = as.numeric(maxRollMeandeltaOD_14days)
           ) %>%
      mutate(GrowthAmpdeltaODFlag_14days = if_else(GrowthAmpdeltaOD_14days >= deltaODGrowthFlagValue, 1, 0),
             GrowthAmpOD720Flag_14days = if_else(GrowthAmpOD720_14days >= OD720GrowthFlagValue, 1, 0)
      ) %>% 
       mutate(DoubleOD720Flag_14days = if_else(maxRollMeanOD720_14days >= minRollMeanOD720_14days * 2 & !is.na(OD720_TimetoMax) >= MaxAmpHour & !is.na(OD720_TimetoMax) & GrowthAmpOD720_14days >= 0, 1, 0),
             DoubledeltaODFlag_14days = if_else(maxRollMeandeltaOD_14days >= minRollMeandeltaOD_14days * 2 & !is.na(deltaOD_TimetoMax) >= MaxAmpHour & !is.na(deltaOD_TimetoMax) & GrowthAmpdeltaOD_14days >= 0, 1, 0)
           )

ProcessDataNestGrowth <- ProcessDataNestGrowth %>%
  left_join(., ProcessDataNestGrowth2 %>% select(Tube, GrowthAmpdeltaOD_14days, GrowthAmpOD720_14days,  minRollMeanOD720_14days, maxRollMeanOD720_14days, minRollMeandeltaOD_14days, maxRollMeandeltaOD_14days, GrowthAmpdeltaODFlag_14days, GrowthAmpOD720Flag_14days, DoubledeltaODFlag_14days, DoubleOD720Flag_14days), by = "Tube")

remove(ProcessDataNestGrowth2)

#failed attempt with map2
# test = map2(.x = lognormOD720_Gompertz_tidied, .y = lognormOD720_Gompertz_predict, ~if_else(as.numeric(~pluck(.x, 2, 1)) <= as.numeric(~.y$time[which.max(.$`.fitted`)]), 1, 0))
```
```{r}
glimpse(ProcessDataNestGrowth)
```

# Save .Rds of ProcessDataNestGrowth containing logistic fits of growth data from time resolved and whole interval fits.
This .Rds is the substrate for the 'MultiDataGrowthAssessPico.Rmd to compare information across multiple runs.
```{r save rds}

 saveRDS(object = ProcessDataNestGrowth, file = file.path(DataOut, paste(ProcessFileName,  "ProcessDataNestGrowth.Rds",  sep = "_")), ascii = FALSE, version = NULL, compress = "xz", refhook = NULL)

```

```{r plot logistics predictions}
#improve with automated annotate placement based upon data traces; difficulty is that data is nested until expansion for ggplot b/c we need to conserve memory so data traces are not easily accessible
# segmentYstart <- as.numeric(IntervalDataTube1[IntervalDataTube1$time == 52, "lnOD680"])
# segmentYend <- as.numeric(IntervalDataTube1[IntervalDataTube1$time == 54, "lnOD680"])

OD_x = 7
OD_y = 0.15

Par_x = 15
Par_y = 0.04

Resid_x = 3
Resid_y = -0.03

Predict_x = 4
Predict_y = 0.06

# #predictions of condition specific fits
ProcessDataNestGrowth %>%
#drop_na(OD720_logistic) %>%
  mutate(OD720_logistic_predict = map(OD720_logistic, possibly(augment, otherwise = NULL))) %>%
  unnest(cols = c(tubedata, OD720_logistic_predict),names_sep = "_", keep_empty = TRUE) %>%
  ggplot() +
  geom_point(aes(x = tubedata_time, y = tubedata_OD720), colour = "darkgreen", size = 0.1) +
  scale_colour_manual(values = MCMIXColours) +
  geom_point(aes(x = tubedata_time, y = OD720_logistic_predict_.fitted), colour = "black", size = 0.1) +
  geom_point(aes(x = tubedata_time, y = OD720_logistic_predict_.resid), colour = "gold", size = 0.05) +
  geom_point(aes(x = tubedata_time, y = tubedata_Actinic_par/10000, colour = as.factor(WL)), size = 0.05, alpha = 0.1) +
  #coord_cartesian(xlim = c(-5, 675)) +
#scale_x_continuous(breaks=seq(0, 25, by = 5)) +
  facet_grid(rows = vars(O2), cols = vars(Tube, WL, Par_ue, Strain, Photoperiod)) +
  labs(subtitle = "OD720 Logistic:  Tube; Growth Light (nm; µE); Strain; Photoperiod", caption = ProcessFileName, y = "Optical Density (OD720)", x = "Elapsed Time (h)") +
  theme_bw() +
  labs(colour = "Actinic PAR (nm)")+
  guides(colour = guide_legend(override.aes = list(size = 3,  alpha = 1)))

ProcessDataNestGrowth %>%
  mutate(deltaOD_logistic_predict = map(deltaOD_logistic, possibly(augment, otherwise = NULL))) %>%
  unnest(cols = c(tubedata, deltaOD_logistic_predict),names_sep = "_", keep_empty = TRUE) %>%
  ggplot() +
  geom_point(aes(x = tubedata_time, y = tubedata_deltaOD), colour = "darkgreen", size = 0.1) +
  scale_colour_manual(values = MCMIXColours) +
  geom_point(aes(x = tubedata_time, y = deltaOD_logistic_predict_.fitted), colour = "black", size = 0.1) +
  geom_point(aes(x = tubedata_time, y = deltaOD_logistic_predict_.resid), colour = "gold", size = 0.05) +
  geom_point(aes(x = tubedata_time, y = tubedata_Actinic_par/10000, colour = as.factor(WL)), size = 0.05, alpha = 0.1) +
  #coord_cartesian(xlim = c(-5, 675)) +
  #scale_x_continuous(breaks=seq(0, 675, by = 125)) +
  facet_grid(rows = vars(O2), cols = vars(Tube, WL, Par_ue, Strain, Photoperiod)) +
  labs(subtitle = "DeltaOD Logistic:   Tube; Growth Light (nm; µE); Strain; Photoperiod", caption = ProcessFileName, y = "Delta Optical Density (deltaOD)", x = "Elapsed Time (h)") +
  theme_bw() +
  labs(colour = "Actinic PAR (nm)")+
  guides(colour = guide_legend(override.aes = list(size = 3,  alpha = 1)))

# ProcessDataNestGrowth %>%
#   filter(Tube == 5) %>%
#   mutate(OD720_logistic_predict = map(OD720_logistic, possibly(augment, otherwise = NULL))) %>%
#   unnest(cols = c(tubedata, OD720_logistic_predict),names_sep = "_") %>%
#   ggplot() +
#   geom_point(aes(x = tubedata_time, y = tubedata_OD720), colour = "darkgreen", size = 0.1) +
#   scale_colour_manual(values = MCMIXColours) +
#   geom_point(aes(x = tubedata_time, y = OD720_logistic_predict_.fitted), colour = "black", size = 0.1) +
#   geom_point(aes(x = tubedata_time, y = OD720_logistic_predict_.resid), colour = "gold", size = 0.05, alpha = 0.5) +
#   geom_point(aes(x = tubedata_time, y = tubedata_Actinic_par/10000, colour = as.factor(WL)), size = 0.01, alpha = 0.1) +
#   #coord_cartesian(xlim = c(0, 675)) +
#   #scale_x_continuous(breaks=seq(0, 675, by = 125)) +
#   facet_grid(rows = vars(O2), cols = vars(Tube, WL, Par_ue, Strain, Photoperiod)) +
#   annotate(geom = "text", x = OD_x, y = OD_y, label = "OD720", size = 5, colour = "darkgreen") +
#   annotate(geom = "text", x = Par_x, y = Par_y, label = "Light level", size = 5, colour = "darkblue") +
#   annotate(geom = "text", x = Resid_x, y = Resid_y, label = "Model residuals", size = 5, colour = "red") +
#   annotate(geom = "text", x = Predict_x, y = Predict_y, label = "Logistic Regression", size = 5, colour = "black") +
#   labs(subtitle = "Tube; Growth Light (nm; µE); Strain; Photoperiod", caption = ProcessFileName , y = "Optical Density (OD720)", x = "Elapsed Time (h)") +
#   theme_bw() +
#   labs(colour = "Actinic PAR (nm)") +
#   guides(colour = guide_legend(override.aes = list(size = 3,  alpha = 1)))

```

```{r exclude poor fits}
# TestProcessDataNestGrowth %>%
#   filter(Tube != 4)
```

Truncated Logistic fits vs. Logistic fits
```{r compare trunc to full plots}

ProcessDataNestGrowth %>%
unnest(cols = c(OD720_logistic_tidied), names_sep = "_", keep_empty = TRUE) %>%
  pivot_wider(names_from =  OD720_logistic_tidied_term, values_from = c(OD720_logistic_tidied_estimate, OD720_logistic_tidied_std.error, OD720_logistic_tidied_statistic, OD720_logistic_tidied_p.value), names_sep = "_")  %>%
  rename(OD720_Lmax = OD720_logistic_tidied_estimate_Lmax,
         OD720_Lmax_se = OD720_logistic_tidied_std.error_Lmax,
         OD720_Lmu = OD720_logistic_tidied_estimate_Lmu,
         OD720_Lmu_se = OD720_logistic_tidied_std.error_Lmu,
         OD720_Lintercept = OD720_logistic_tidied_estimate_Lintercept,
         OD720_Lintercept_se = OD720_logistic_tidied_std.error_Lintercept)  %>%
  unnest(cols = c(OD720trunc_logistic_tidied), names_sep = "_", keep_empty = TRUE) %>%
  pivot_wider(names_from =  OD720trunc_logistic_tidied_term, values_from = c(OD720trunc_logistic_tidied_estimate, OD720trunc_logistic_tidied_std.error, OD720trunc_logistic_tidied_statistic, OD720trunc_logistic_tidied_p.value), names_sep = "_") %>%
  rename(OD720_Trunc_Lmax = OD720trunc_logistic_tidied_estimate_Lmax,
         OD720_Trunc_Lmax_se = OD720trunc_logistic_tidied_std.error_Lmax,
         OD720_Trunc_Lmu = OD720trunc_logistic_tidied_estimate_Lmu,
         OD720_Trunc_Lmu_se = OD720trunc_logistic_tidied_std.error_Lmu)  %>%
   ggplot() +
  geom_point(aes(x = OD720_Lmu, y = OD720_Trunc_Lmu, colour = as.factor(WL)), size = 4)  +
  ggrepel::geom_text_repel(aes(x = OD720_Lmu, y = OD720_Trunc_Lmu, label= Tube),hjust=0, vjust=0) +
scale_colour_manual(values = MCMIXColours) +
  scale_shape_manual(values = c(13, 16)) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_vline(xintercept = 0, linetype = "dashed") +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed") +
  coord_fixed(ratio = 1) +
  coord_cartesian(xlim = c (0, 0.3), ylim = c(0, 0.3)) +
  scale_x_continuous(breaks=seq(0, 0.3, by = 0.1)) +
  labs(subtitle = "Strain, O2, Actinic PAR (nm)", y = "Truncated Logistic Growth Rate (OD720_Lmu)", x = "Logistic Growth Rate (OD720_Lmu)") +
  facet_grid(rows = vars(Strain), cols = vars(O2, WL)) + 
  theme_bw() + 
  labs(title = "Logistic vs. Truncated Logistic Cell Specific Growth Rates (µ" ~ h^-1~")- All Growth",
       colour = "Actinic PAR (nm)")

ProcessDataNestGrowth %>%
unnest(cols = c(deltaOD_logistic_tidied), names_sep = "_", keep_empty = TRUE) %>%
  pivot_wider(names_from =  deltaOD_logistic_tidied_term, values_from = c(deltaOD_logistic_tidied_estimate, deltaOD_logistic_tidied_std.error, deltaOD_logistic_tidied_statistic, deltaOD_logistic_tidied_p.value), names_sep = "_")  %>%
  rename(deltaOD_Lmax = deltaOD_logistic_tidied_estimate_Lmax,
         deltaOD_Lmax_se = deltaOD_logistic_tidied_std.error_Lmax,
         deltaOD_Lmu = deltaOD_logistic_tidied_estimate_Lmu,
         deltaOD_Lmu_se = deltaOD_logistic_tidied_std.error_Lmu,
         deltaOD_Lintercept = deltaOD_logistic_tidied_estimate_Lintercept,
         deltaOD_Lintercept_se = deltaOD_logistic_tidied_std.error_Lintercept)  %>%
  unnest(cols = c(deltaODtrunc_logistic_tidied), names_sep = "_", keep_empty = TRUE) %>%
  pivot_wider(names_from =  deltaODtrunc_logistic_tidied_term, values_from = c(deltaODtrunc_logistic_tidied_estimate, deltaODtrunc_logistic_tidied_std.error, deltaODtrunc_logistic_tidied_statistic, deltaODtrunc_logistic_tidied_p.value), names_sep = "_") %>%
  rename(deltaOD_Trunc_Lmax = deltaODtrunc_logistic_tidied_estimate_Lmax,
         deltaOD_Trunc_Lmax_se = deltaODtrunc_logistic_tidied_std.error_Lmax,
         deltaOD_Trunc_Lmu = deltaODtrunc_logistic_tidied_estimate_Lmu,
         deltaOD_Trunc_Lmu_se = deltaODtrunc_logistic_tidied_std.error_Lmu)  %>%
   ggplot() +
  geom_point(aes(x = deltaOD_Lmu, y = deltaOD_Trunc_Lmu, colour = as.factor(WL)), size = 4)  +
  ggrepel::geom_text_repel(aes(x = deltaOD_Lmu, y = deltaOD_Trunc_Lmu, label= Tube), hjust=0, vjust=0) +
scale_colour_manual(values = MCMIXColours) +
  scale_shape_manual(values = c(13, 16)) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_vline(xintercept = 0, linetype = "dashed") +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed") +
  coord_fixed(ratio = 1) +
  coord_cartesian(xlim = c (0, 0.1), ylim = c(0, 0.1)) +
  scale_x_continuous(breaks=seq(0, 0.1, by = 0.02)) +
  labs(subtitle = "Strain, O2, Actinic PAR (nm)", y = "Truncated Logistic Growth Rate (deltaOD_Lmu)", x = "Logistic Growth Rate (deltaOD_Lmu)") +
  facet_grid(rows = vars(Strain), cols = vars(O2, WL)) + 
  theme_bw() + 
  labs(title = "Logistic vs. Truncated Logistic Chlorophyll Specific Growth Rates (µ" ~ h^-1~")- All Growth",
       colour = "Actinic PAR (nm)")

```

Plot all Fit Terms vs. conditions- no filters
```{r plot all}
ProcessDataNestGrowth %>%
 unnest(cols = c(OD720_logistic_tidied),names_sep = "_") %>%
  select(c(Filename, Tube, Strain, Par_ue, Photoperiod, Temp_c, O2, WL, LightShape, OD720_logistic_tidied_term, OD720_logistic_tidied_estimate, OD720_logistic_tidied_std.error)) %>%
  pivot_wider(names_from =  OD720_logistic_tidied_term, values_from = c(OD720_logistic_tidied_estimate, OD720_logistic_tidied_std.error), names_sep = "_") %>%
  rename(OD720_Lmax = OD720_logistic_tidied_estimate_Lmax,
         OD720_Lmax_se = OD720_logistic_tidied_std.error_Lmax,
         OD720_Lmu = OD720_logistic_tidied_estimate_Lmu,
         OD720_Lmu_se = OD720_logistic_tidied_std.error_Lmu,
         OD720_Lintercept = OD720_logistic_tidied_estimate_Lintercept,
         OD720_Lintercept_se = OD720_logistic_tidied_std.error_Lintercept) %>%
  ggplot() +
  geom_point(aes(x = Par_ue, y = OD720_Lmu, colour = as.factor(WL))) +
  geom_errorbar(aes(x = Par_ue, ymin = OD720_Lmu - OD720_Lmu_se, ymax = OD720_Lmu + OD720_Lmu_se, colour = as.factor(WL))) +
  scale_colour_manual(values = MCMIXColours) +
  facet_grid (cols = vars(Strain), rows = vars(O2)) +
  theme_bw() +
  labs(title = "Logistic cell Specific Growth Rate - All data")

ProcessDataNestGrowth %>%
 unnest(cols = c(deltaOD_logistic_tidied),names_sep = "_") %>%
  select(c(Filename, Tube, Strain, Par_ue, Photoperiod, Temp_c, O2, WL, LightShape, deltaOD_logistic_tidied_term, deltaOD_logistic_tidied_estimate, deltaOD_logistic_tidied_std.error)) %>%
  pivot_wider(names_from =  deltaOD_logistic_tidied_term, values_from = c(deltaOD_logistic_tidied_estimate, deltaOD_logistic_tidied_std.error), names_sep = "_") %>%
  rename(deltaOD_Lmax = deltaOD_logistic_tidied_estimate_Lmax,
         deltaOD_Lmax_se = deltaOD_logistic_tidied_std.error_Lmax,
         deltaOD_Lmu = deltaOD_logistic_tidied_estimate_Lmu,
         deltaOD_Lmu_se = deltaOD_logistic_tidied_std.error_Lmu,
         deltaOD_Lintercept = deltaOD_logistic_tidied_estimate_Lintercept,
         deltaOD_Lintercept_se = deltaOD_logistic_tidied_std.error_Lintercept) %>%
  ggplot() +
  geom_point(aes(x = Par_ue, y = deltaOD_Lmu, colour = as.factor(WL))) +
  geom_errorbar(aes(x = Par_ue, ymin = deltaOD_Lmu - deltaOD_Lmu_se, ymax = deltaOD_Lmu + deltaOD_Lmu_se, colour = as.factor(WL))) +
  scale_colour_manual(values = MCMIXColours) +
  facet_grid (cols = vars(Strain), rows = vars(O2)) +
  theme_bw() +
  labs(title = "Logistic chlorophyll specific - All data")

```

Plot Fit Terms vs. conditions - High Std error filtered out by flag argument
```{r logistic fit terms plot}
ProcessDataNestGrowth %>%
 unnest(cols = c(OD720_logistic_tidied),names_sep = "_") %>%
  select(c(Filename, Tube, Strain, Par_ue, Photoperiod, Temp_c, O2, WL, LightShape, OD720_logistic_tidied_term, OD720_logistic_tidied_estimate, OD720_logistic_tidied_std.error)) %>%
  pivot_wider(names_from =  OD720_logistic_tidied_term, values_from = c(OD720_logistic_tidied_estimate, OD720_logistic_tidied_std.error), names_sep = "_") %>%
  rename(OD720_Lmax = OD720_logistic_tidied_estimate_Lmax,
         OD720_Lmax_se = OD720_logistic_tidied_std.error_Lmax,
         OD720_Lmu = OD720_logistic_tidied_estimate_Lmu,
         OD720_Lmu_se = OD720_logistic_tidied_std.error_Lmu,
         OD720_Lintercept = OD720_logistic_tidied_estimate_Lintercept,
         OD720_Lintercept_se = OD720_logistic_tidied_std.error_Lintercept) %>%
  mutate(LmuOD720Flag = if_else(OD720_Lmu_se < OD720_Lmu * GrowthFlagError, 1, 0)) %>%
# mutate(GrowthFlag = if_else((OD720_Lintercept + OD720_Lintercept_se) < ((OD720_Lmax - OD720_Lmax_se) & (OD720_Lmu_se < OD720_Lmu * GrowthFlagError) & (OD720_Lmax_se < OD720_Lmax * GrowthFlagError) ), 1, 0)) %>%
  filter(LmuOD720Flag == 1) %>%
  ggplot() +
  geom_point(aes(x = Par_ue, y = OD720_Lmu, colour = as.factor(WL))) +
  geom_errorbar(aes(x = Par_ue, ymin = OD720_Lmu - OD720_Lmu_se, ymax = OD720_Lmu + OD720_Lmu_se, colour = as.factor(WL))) +
  scale_colour_manual(values = MCMIXColours) +
  facet_grid (cols = vars(Strain), rows = vars(O2)) +
  theme_bw() +
  labs(title = "Logistic cell scatter growth rates - High std error fits omitted")

ProcessDataNestGrowth %>%
 unnest(cols = c(deltaOD_logistic_tidied),names_sep = "_") %>%
  select(c(Filename, Tube, Strain, Par_ue, Photoperiod, Temp_c, O2, WL, LightShape, deltaOD_logistic_tidied_term, deltaOD_logistic_tidied_estimate, deltaOD_logistic_tidied_std.error)) %>%
  pivot_wider(names_from =  deltaOD_logistic_tidied_term, values_from = c(deltaOD_logistic_tidied_estimate, deltaOD_logistic_tidied_std.error), names_sep = "_") %>%
  rename(deltaOD_Lmax = deltaOD_logistic_tidied_estimate_Lmax,
         deltaOD_Lmax_se = deltaOD_logistic_tidied_std.error_Lmax,
         deltaOD_Lmu = deltaOD_logistic_tidied_estimate_Lmu,
         deltaOD_Lmu_se = deltaOD_logistic_tidied_std.error_Lmu,
         deltaOD_Lintercept = deltaOD_logistic_tidied_estimate_Lintercept,
         deltaOD_Lintercept_se = deltaOD_logistic_tidied_std.error_Lintercept) %>%
  mutate(LmudeltaODFlag = if_else(deltaOD_Lmu_se < deltaOD_Lmu * GrowthFlagError, 1, 0)) %>%
 # mutate(GrowthFlag = if_else((deltaOD_Lintercept + deltaOD_Lintercept_se) < ((deltaOD_Lmax - deltaOD_Lmax_se) & (deltaOD_Lmu_se < deltaOD_Lmu * GrowthFlagError) & (deltaOD_Lmax_se < deltaOD_Lmax * GrowthFlagError) ), 1, 0)) %>%
  filter(LmudeltaODFlag == 1) %>%
  ggplot() +
  geom_point(aes(x = Par_ue, y = deltaOD_Lmu, colour = as.factor(WL))) +
  geom_errorbar(aes(x = Par_ue, ymin = deltaOD_Lmu - deltaOD_Lmu_se, ymax = deltaOD_Lmu + deltaOD_Lmu_se, colour = as.factor(WL))) +
  scale_colour_manual(values = MCMIXColours) +
  facet_grid (cols = vars(Strain), rows = vars(O2)) +
  theme_bw() +
  labs(title = "Logistic chlorophyll specific growth rates - High Std error fits omitted")

```