### 2023-12-07 Code/MultiDataGrowthAssessPico.Rmd
Added MultiCultiGrowthAssessPico to Github

---
title: "MultiDataGrowthAssess"
author:
- Laurel Genge
- Carlie Barnhill
- Max Berthold
- Douglas A. Campbell
- Mireille Savoie
date: "`r format(Sys.Date())`"
output:
  bookdown::html_document2:
    code_folding: show
    keep_md: yes
    toc: TRUE
    toc_float: TRUE
    toc_depth: 6
    fig_caption: yes
bibliography: Prochlorococcus_O2_NPQ.bib
csl: plos-one.csl
editor_options: 
  markdown: 
    wrap: 72
---

## Set figure caption font size
```{css, echo=FALSE}
p.caption {
  font-size: 25px;
}
```

## Set Chunk Options
```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
knitr::opts_chunk$set(fig.path='Figs/')
```

# Introduction

# Materials and Methods

## Load Libraries and set project variables

```{r load libraries, message= TRUE}
# libraries; Note check actual dependencies

library(lubridate)
library(stringr)
library(broom)
library(knitr)
library(zoo)
library(RColorBrewer)
library(grid)
library(tidyverse)
#library(tidyquant)
#library(data.table)
#library(googledrive)
#library(googlesheets4)
#library(dplyr)
```

```{r set project variables}
#"..", takes up a level in the directory path
Project <- "PICO"
DataIn1 <- file.path("..","ProcessedData", "LogGrowthFits1")
DataIn2 <- file.path("..","ProcessedData", "LogGrowthFits2")
PlotsPath <- file.path("..","Plots")
ShinyPath <- file.path("..", "ShinyPico", "data")
DataOut <- file.path("..","GrowthAssessData")
MetaCatalog <- file.path("..","PicoCatalog.csv")

GrowthFlagError <- 0.2       # percentage of mu that std error must be less than

OD720GrowthAmpFlagValue = 0.010 # OD720 absolute change for setting GrowthFlag = 1 and DeclineFlag = 1; did culture show  growth;

deltaODGrowthAmpFlagValue = 0.007 # deltaOD absolute change for setting GrowthFlag = 1 and DeclineFlag = 1; did culture show  growth;

GrowthTerm <- "Gompertz Growth Rate"

```

```{r set colours}
MyWavelengths = c(405, 450, 475, 530, 615, 660, 730, "WW")
MCMIXColours = c("violet", "darkblue", "dodgerblue", "darkgreen","orange","red","purple", "black")

names(MCMIXColours) <- MyWavelengths
MCMIXColours

```

## List previously processed data

Note: Split files will not be correct for Par_ue, but Actinic_par should be OK Fix: replace MetaData Par_ue with max(Actinic_par) for each file or split file

```{r previous ProcessedData}
ProcessedFiles1 <- list.files(path = DataIn1, pattern = Project, full.names = TRUE)

ProcessedFiles1

ProcessedFiles2 <- list.files(path = DataIn2, pattern = Project, full.names = TRUE)

ProcessedFiles2
```

# Read .Rds, Summarize, Combine to single data frame

## Create function using to read in .Rds, adding source filename as a column

```{r readRDS_plus}
readRDS_plus <- function(Flnm){readRDS(file = Flnm) %>%
    mutate(Filename = Flnm)
}
```

## Read in ProcessedFiles .Rds 
## April 2023, Had to split the read in into 2 as there was not enough memory to read all files at once.

```{r read selected ProcessedData}

# TestData <- readRDS_plus(Flnm = file.path("../ProcessedData/LogGrowthFits/20200731_PICO_MCMIX004_RUN17_ProcessDataNestGrowth.Rds"))

LogFitsData1 <-  ProcessedFiles1 %>%
    map_df(~readRDS_plus(Flnm = .)) %>%
    select(-c("Motivation", "doi", "SourceSalinity", "EndDate", "Salinity",  "InocpH","lognormdeltaOD_Gompertz", "lognormOD720_Gompertz", "OD720_logistic", "deltaOD_logistic","OD720trunc_logistic", "deltaODtrunc_logistic", "OD720trunc_exp", "deltaODtrunc_exp",
              "lognormOD720_Gompertz_param",
              "lognormdeltaOD_Gompertz_param",
              "OD720_logistic_param",
              "OD720trunc_exp_param",
              "deltaODtrunc_logistic_param",
              "OD720trunc_logistic_param",
              "deltaOD_logistic_param",
              "deltaODtrunc_exp_param"
              ))

LogFitsData2 <-  ProcessedFiles2 %>%
    map_df(~readRDS_plus(Flnm = .)) %>%
    select(-c("Motivation", "doi", "SourceSalinity", "EndDate", "Salinity",  "InocpH","lognormdeltaOD_Gompertz", "lognormOD720_Gompertz", "OD720_logistic", "deltaOD_logistic","OD720trunc_logistic", "deltaODtrunc_logistic", "OD720trunc_exp", "deltaODtrunc_exp",
              "lognormOD720_Gompertz_param",
              "lognormdeltaOD_Gompertz_param",
              "OD720_logistic_param",
              "OD720trunc_exp_param",
              "deltaODtrunc_logistic_param",
              "OD720trunc_logistic_param",
              "deltaOD_logistic_param",
              "deltaODtrunc_exp_param"
              ))

glimpse(LogFitsData1)
glimpse(LogFitsData2)

LogFitsData <- rbind(LogFitsData1, LogFitsData2)

remove(LogFitsData1)
remove(LogFitsData2)

```

#Unnest data
# assign growth rate to zero if model returns mu < 0 or if growth amplitude flag = 0

```{r unnest fits}

# OD720 gompertz fits
# assign 0 to growth fits that are NA's and for tubes with absolute amplitude change of less than 0.007.Defined at line 75 
lognormOD720_unnest <- LogFitsData %>%
   select(c(Filename, CDateTime, Tube, ExpDate, MC, PrimaryOperator, Run, Description, ProjectID, ID, Strain, Inoc_mL, Media_mL, ExpCul, InnocDate, ExpStartTime, Par_ue, Photoperiod, Temp_c, O2, WL, LightShape, Media, Source, Optode, OptodeCh, FinalpH, Par_ueAdjusted, DateOfAdjustment, ElaspedHoursAtAdjustment, ExpEndDate, ExpEndHour, SetActinic_day, tubedata, IsOD720_GLagNeg, lognormOD720_TimetoMax, lognormdeltaOD_TimetoMax, OD720trunc_TimetoMax, deltaODtrunc_TimetoMax, IsdeltaOD_GLagNeg, lognormOD720_GompLag, lognormdeltaOD_GompLag, GrowthAmpOD720, GrowthAmpOD720_14days, GrowthAmpdeltaOD, GrowthAmpdeltaOD_14days, lognormOD720_Gompertz_tidied, lognormOD720_Gompertz_predict, DoubledeltaODFlag_14days, DoubleOD720Flag_14days, O2_Category)
          ) %>%
  unnest(cols = c(lognormOD720_Gompertz_tidied), names_sep = "_", keep_empty = TRUE) %>%
  pivot_wider(names_from =  lognormOD720_Gompertz_tidied_term, values_from = c(lognormOD720_Gompertz_tidied_estimate, lognormOD720_Gompertz_tidied_std.error, lognormOD720_Gompertz_tidied_statistic, lognormOD720_Gompertz_tidied_p.value), names_sep = "_") %>%
  rename(OD720_Gmax = lognormOD720_Gompertz_tidied_estimate_Gmax,
         OD720_Gmax_se = lognormOD720_Gompertz_tidied_std.error_Gmax,
         OD720_Gmu_raw = lognormOD720_Gompertz_tidied_estimate_Gmu,
         OD720_Gmu_se = lognormOD720_Gompertz_tidied_std.error_Gmu,
         OD720_GLag = lognormOD720_Gompertz_tidied_estimate_GLag,             
         OD720_GLag_se = lognormOD720_Gompertz_tidied_std.error_GLag)  %>%
   mutate(OD720_Gmu_corr = if_else(!is.na(OD720_Gmu_raw), OD720_Gmu_raw, 0),
          OD720_Gmu_se = if_else(!is.na(OD720_Gmu_se), OD720_Gmu_se, 0),
          OD720_GLag = if_else(!is.na(OD720_GLag), OD720_GLag, 0),
          OD720_GLag_se = if_else(!is.na(OD720_GLag_se), OD720_GLag_se, 0),
          OD720_Gmax = if_else(!is.na(OD720_Gmax), OD720_Gmax, 0),
          OD720_Gmax_se = if_else(!is.na(OD720_Gmax_se), OD720_Gmax_se, 0),
          GrowthAmpOD720Flag = if_else(GrowthAmpOD720 >= OD720GrowthAmpFlagValue, 1, 0),
          GrowthAmpOD720Flag_14days = if_else(GrowthAmpOD720_14days >= OD720GrowthAmpFlagValue, 1, 0),
          OD720_Gmu_corr = if_else(GrowthAmpOD720Flag_14days == 1 & OD720_Gmu_corr >= 0, OD720_Gmu_corr, 0),
          # OD720_GseGrowthFlag = case_when(OD720_Gmu_corr > 0 & OD720_Gmu_corr * GrowthFlagError > OD720_Gmu_se ~ 1,
          #                                   OD720_Gmu_corr > 0 & OD720_Gmu_corr * GrowthFlagError < OD720_Gmu_se ~ 0,
          #                                   OD720_Gmu_corr == 0 ~ 1),

           OD720_GseGrowthFlag = if_else(OD720_Gmu_se < OD720_Gmu_corr * GrowthFlagError, 1, 0),
   )

# deltaOD gompertz fits
# assign 0 to growth fits that are NA's and for tubes with absolute amplitude change of less than 0.007.Defined at line 75 
lognormdeltaOD_unnest <- LogFitsData %>% 
select(c(Filename, CDateTime, Tube, ExpDate, MC, PrimaryOperator, Run, Description, ProjectID, ID, Strain, Inoc_mL, Media_mL, ExpCul, InnocDate, ExpStartTime, Par_ue, Photoperiod, Temp_c, O2, WL, LightShape, Media, Source, Optode, OptodeCh, FinalpH, Par_ueAdjusted, DateOfAdjustment, ElaspedHoursAtAdjustment, ExpEndDate, ExpEndHour, SetActinic_day, tubedata, IsOD720_GLagNeg, lognormOD720_TimetoMax, lognormdeltaOD_TimetoMax, OD720trunc_TimetoMax, deltaODtrunc_TimetoMax, IsdeltaOD_GLagNeg, lognormOD720_GompLag, lognormdeltaOD_GompLag, GrowthAmpOD720, GrowthAmpOD720_14days, GrowthAmpdeltaOD, GrowthAmpdeltaOD_14days, lognormdeltaOD_Gompertz_tidied, lognormdeltaOD_Gompertz_predict, DoubledeltaODFlag_14days, DoubleOD720Flag_14days, O2_Category)
       ) %>%
unnest(cols = c(lognormdeltaOD_Gompertz_tidied), names_sep = "_", keep_empty = TRUE) %>%
  pivot_wider(names_from =  lognormdeltaOD_Gompertz_tidied_term, values_from = c(lognormdeltaOD_Gompertz_tidied_estimate, lognormdeltaOD_Gompertz_tidied_std.error, lognormdeltaOD_Gompertz_tidied_statistic, lognormdeltaOD_Gompertz_tidied_p.value), names_sep = "_")  %>%
  rename(deltaOD_Gmax = lognormdeltaOD_Gompertz_tidied_estimate_Gmax,
         deltaOD_Gmax_se = lognormdeltaOD_Gompertz_tidied_std.error_Gmax,
         deltaOD_Gmu_raw = lognormdeltaOD_Gompertz_tidied_estimate_Gmu,
         deltaOD_Gmu_se = lognormdeltaOD_Gompertz_tidied_std.error_Gmu,
         deltaOD_GLag = lognormdeltaOD_Gompertz_tidied_estimate_GLag,
         deltaOD_GLag_se = lognormdeltaOD_Gompertz_tidied_std.error_GLag) %>%
   mutate(deltaOD_Gmu_corr = if_else(!is.na(deltaOD_Gmu_raw), deltaOD_Gmu_raw, 0),
          deltaOD_Gmu_se = if_else(!is.na(deltaOD_Gmu_se), deltaOD_Gmu_se, 0),
          deltaOD_GLag = if_else(!is.na(deltaOD_GLag), deltaOD_GLag, 0),
          deltaOD_GLag_se = if_else(!is.na(deltaOD_GLag_se), deltaOD_GLag_se, 0),
          deltaOD_Gmax = if_else(!is.na(deltaOD_Gmax), deltaOD_Gmax, 0),
          deltaOD_Gmax_se = if_else(!is.na(deltaOD_Gmax_se), deltaOD_Gmax_se, 0),
          GrowthAmpdeltaODFlag = if_else(GrowthAmpdeltaOD >= deltaODGrowthAmpFlagValue , 1, 0),
          GrowthAmpdeltaODFlag_14days = if_else(GrowthAmpdeltaOD_14days >= deltaODGrowthAmpFlagValue, 1, 0),
          deltaOD_Gmu_corr = if_else(GrowthAmpdeltaODFlag_14days == 1 & deltaOD_Gmu_corr >= 0, deltaOD_Gmu_corr, 0),
          # deltaOD_GseGrowthFlag = case_when(deltaOD_Gmu_corr > 0 & deltaOD_Gmu_corr * GrowthFlagError > deltaOD_Gmu_se ~ 1,
          #                                   deltaOD_Gmu_corr > 0 & deltaOD_Gmu_corr * GrowthFlagError < deltaOD_Gmu_se ~ 0,
          #                                   deltaOD_Gmu_corr == 0 ~ 1),

         deltaOD_GseGrowthFlag = if_else(deltaOD_Gmu_se < deltaOD_Gmu_corr * GrowthFlagError, 1, 0),
          )

# OD720 logistic fits
# assign 0 to growth fits that are NA's and for tubes with absolute amplitude change of less than 0.007.Defined at line 75 
OD720_unnest <- LogFitsData %>%
  select(c(Filename, CDateTime, Tube, ExpDate, MC, PrimaryOperator, Run, Description, ProjectID, ID, Strain, Inoc_mL, Media_mL, ExpCul, InnocDate, ExpStartTime, Par_ue, Photoperiod, Temp_c, O2, WL, LightShape, Media, Source, Optode, OptodeCh, FinalpH, Par_ueAdjusted, DateOfAdjustment, ElaspedHoursAtAdjustment, ExpEndDate, ExpEndHour, SetActinic_day, tubedata, IsOD720_GLagNeg, lognormOD720_TimetoMax, lognormdeltaOD_TimetoMax, OD720trunc_TimetoMax, deltaODtrunc_TimetoMax, IsdeltaOD_GLagNeg, lognormOD720_GompLag, lognormdeltaOD_GompLag, GrowthAmpOD720, GrowthAmpOD720_14days, GrowthAmpdeltaOD, GrowthAmpdeltaOD_14days, OD720_logistic_tidied, OD720_logistic_predict, DoubledeltaODFlag_14days, DoubleOD720Flag_14days, O2_Category)
         ) %>%
  unnest(cols = c(OD720_logistic_tidied), names_sep = "_", keep_empty = TRUE) %>%
  pivot_wider(names_from =  OD720_logistic_tidied_term, values_from = c(OD720_logistic_tidied_estimate, OD720_logistic_tidied_std.error, OD720_logistic_tidied_statistic, OD720_logistic_tidied_p.value), names_sep = "_") %>%
  rename(OD720_Lmax = OD720_logistic_tidied_estimate_Lmax,
         OD720_Lmax_se = OD720_logistic_tidied_std.error_Lmax,
         OD720_Lmu_raw = OD720_logistic_tidied_estimate_Lmu,
         OD720_Lmu_se = OD720_logistic_tidied_std.error_Lmu,
         OD720_Lintercept = OD720_logistic_tidied_estimate_Lintercept,             
         OD720_Lintercept_se = OD720_logistic_tidied_std.error_Lintercept)  %>%
   mutate(OD720_Lmu_corr = if_else(!is.na(OD720_Lmu_raw), OD720_Lmu_raw, 0),
          OD720_Lmu_se = if_else(!is.na(OD720_Lmu_se), OD720_Lmu_se, 0),
          OD720_Lintercept = if_else(!is.na(OD720_Lintercept), OD720_Lintercept, 0),
          OD720_Lintercept_se = if_else(!is.na(OD720_Lintercept_se), OD720_Lintercept_se, 0),
          OD720_Lmax = if_else(!is.na(OD720_Lmax), OD720_Lmax, 0),
          OD720_Lmax_se = if_else(!is.na(OD720_Lmax_se), OD720_Lmax_se, 0),
          GrowthAmpOD720Flag = if_else(GrowthAmpOD720 >= OD720GrowthAmpFlagValue, 1, 0),
          GrowthAmpOD720Flag_14days = if_else(GrowthAmpOD720_14days >= OD720GrowthAmpFlagValue, 1, 0),
          OD720_Lmu_corr = if_else(GrowthAmpOD720Flag_14days == 1 & OD720_Lmu_corr >= 0 , OD720_Lmu_corr, 0),
          # OD720_LseGrowthFlag = case_when(OD720_Lmu_corr > 0 & OD720_Lmu_corr * GrowthFlagError > OD720_Lmu_se ~ 1,
          #                                   OD720_Lmu_corr > 0 & OD720_Lmu_corr * GrowthFlagError < OD720_Lmu_se ~ 0,
          #                                   OD720_Lmu_corr == 0 ~ 1),
          OD720_LseGrowthFlag = if_else(OD720_Lmu_se < OD720_Lmu_corr * GrowthFlagError, 1, 0),
          OD720_TimeHalfPmax_h = (log(2-(OD720_Lintercept/(OD720_Lmax - OD720_Lintercept))))/OD720_Lmu_corr,
          OD720_TimeHalfPmax_h = if_else(!is.infinite(OD720_TimeHalfPmax_h), OD720_TimeHalfPmax_h, 0),
          OD720_TimeHalfPmax_h = if_else(!is.nan(OD720_TimeHalfPmax_h), OD720_TimeHalfPmax_h, 0)
          )

# deltaOD logistic fits
# assign 0 to growth fits that are NA's and for tubes with absolute amplitude change of less than 0.007 and less than 14 days.Defined at line 75 
deltaOD_unnest <- LogFitsData %>%
   select(c(Filename, CDateTime, Tube, ExpDate, MC, PrimaryOperator, Run, Description, ProjectID, ID, Strain, Inoc_mL, Media_mL, ExpCul, InnocDate, ExpStartTime, Par_ue, Photoperiod, Temp_c, O2, WL, LightShape, Media, Source, Optode, OptodeCh, FinalpH, Par_ueAdjusted, DateOfAdjustment, ElaspedHoursAtAdjustment, ExpEndDate, ExpEndHour, SetActinic_day, tubedata, IsOD720_GLagNeg, lognormOD720_TimetoMax, lognormdeltaOD_TimetoMax, OD720trunc_TimetoMax, deltaODtrunc_TimetoMax, IsdeltaOD_GLagNeg, lognormOD720_GompLag, lognormdeltaOD_GompLag, GrowthAmpOD720,  GrowthAmpOD720_14days, GrowthAmpdeltaOD, GrowthAmpdeltaOD_14days, deltaOD_logistic_tidied, deltaOD_logistic_predict, DoubledeltaODFlag_14days, DoubleOD720Flag_14days, O2_Category)
          ) %>%
unnest(cols = c(deltaOD_logistic_tidied), names_sep = "_", keep_empty = TRUE) %>%
  pivot_wider(names_from =  deltaOD_logistic_tidied_term, values_from = c(deltaOD_logistic_tidied_estimate, deltaOD_logistic_tidied_std.error, deltaOD_logistic_tidied_statistic, deltaOD_logistic_tidied_p.value), names_sep = "_")  %>%
  rename(deltaOD_Lmax = deltaOD_logistic_tidied_estimate_Lmax,
         deltaOD_Lmax_se = deltaOD_logistic_tidied_std.error_Lmax,
         deltaOD_Lmu_raw = deltaOD_logistic_tidied_estimate_Lmu,
         deltaOD_Lmu_se = deltaOD_logistic_tidied_std.error_Lmu,
         deltaOD_Lintercept = deltaOD_logistic_tidied_estimate_Lintercept,
         deltaOD_Lintercept_se = deltaOD_logistic_tidied_std.error_Lintercept) %>%
   mutate(deltaOD_Lmu_corr = if_else(!is.na(deltaOD_Lmu_raw), deltaOD_Lmu_raw, 0),
          deltaOD_Lmu_se = if_else(!is.na(deltaOD_Lmu_se), deltaOD_Lmu_se, 0),
          deltaOD_Lintercept = if_else(!is.na(deltaOD_Lintercept), deltaOD_Lintercept, 0),
          deltaOD_Lintercept_se = if_else(!is.na(deltaOD_Lintercept_se), deltaOD_Lintercept_se, 0),
          deltaOD_Lmax = if_else(!is.na(deltaOD_Lmax), deltaOD_Lmax, 0),
          deltaOD_Lmax_se = if_else(!is.na(deltaOD_Lmax_se), deltaOD_Lmax_se, 0),
          GrowthAmpdeltaODFlag = if_else(GrowthAmpdeltaOD >= deltaODGrowthAmpFlagValue , 1, 0),
          GrowthAmpdeltaODFlag_14days = if_else(GrowthAmpdeltaOD_14days >= deltaODGrowthAmpFlagValue, 1, 0),
          deltaOD_Lmu_corr = if_else(GrowthAmpdeltaODFlag_14days == 1 & deltaOD_Lmu_corr >= 0, deltaOD_Lmu_corr, 0),
          # deltaOD_LseGrowthFlag = case_when(deltaOD_Lmu_corr > 0 & deltaOD_Lmu_corr * GrowthFlagError > deltaOD_Lmu_se ~ 1,
                                            # deltaOD_Lmu_corr > 0 & deltaOD_Lmu_corr * GrowthFlagError < deltaOD_Lmu_se ~ 0,
                                            # deltaOD_Lmu_corr == 0 ~ 1),
         deltaOD_LseGrowthFlag = if_else(deltaOD_Lmu_se < deltaOD_Lmu_corr * GrowthFlagError, 1, 0),
          deltaOD_TimeHalfPmax_h = (log(2-(deltaOD_Lintercept/(deltaOD_Lmax - deltaOD_Lintercept))))/deltaOD_Lmu_corr,
          deltaOD_TimeHalfPmax_h = if_else(!is.infinite(deltaOD_TimeHalfPmax_h), deltaOD_TimeHalfPmax_h, 0),
          deltaOD_TimeHalfPmax_h = if_else(!is.nan(deltaOD_TimeHalfPmax_h), deltaOD_TimeHalfPmax_h, 0)
          )

# OD720 truncated logistic fits
# assign 0 to growth fits that are NA's and for tubes with absolute amplitude change of less than 0.007.Defined at line 75 
OD720truncLog_unnest <- LogFitsData %>%
   select(c(Filename, CDateTime, Tube, ExpDate, MC, PrimaryOperator, Run, Description, ProjectID, ID, Strain, Inoc_mL, Media_mL, ExpCul, InnocDate, ExpStartTime, Par_ue, Photoperiod, Temp_c, O2, WL, LightShape, Media, Source, Optode, OptodeCh, FinalpH, Par_ueAdjusted, DateOfAdjustment, ElaspedHoursAtAdjustment, ExpEndDate, ExpEndHour, SetActinic_day, tubedata, IsOD720_GLagNeg, lognormOD720_TimetoMax, lognormdeltaOD_TimetoMax, OD720trunc_TimetoMax, deltaODtrunc_TimetoMax, IsdeltaOD_GLagNeg, lognormOD720_GompLag, lognormdeltaOD_GompLag, GrowthAmpOD720, GrowthAmpOD720_14days, GrowthAmpdeltaOD, GrowthAmpdeltaOD_14days, OD720trunc_logistic_tidied, OD720trunc_logistic_predict, DoubledeltaODFlag_14days, DoubleOD720Flag_14days, O2_Category)
          ) %>%
  unnest(cols = c(OD720trunc_logistic_tidied), names_sep = "_", keep_empty = TRUE) %>%
  pivot_wider(names_from =  OD720trunc_logistic_tidied_term, values_from = c(OD720trunc_logistic_tidied_estimate, OD720trunc_logistic_tidied_std.error, OD720trunc_logistic_tidied_statistic, OD720trunc_logistic_tidied_p.value), names_sep = "_") %>%
  rename(OD720trunc_Lmax = OD720trunc_logistic_tidied_estimate_Lmax,
         OD720trunc_Lmax_se = OD720trunc_logistic_tidied_std.error_Lmax,
         OD720trunc_Lmu_raw = OD720trunc_logistic_tidied_estimate_Lmu,
         OD720trunc_Lmu_se = OD720trunc_logistic_tidied_std.error_Lmu,
         OD720trunc_Lintercept = OD720trunc_logistic_tidied_estimate_Lintercept,             
         OD720trunc_Lintercept_se = OD720trunc_logistic_tidied_std.error_Lintercept)  %>%
   mutate(OD720trunc_Lmu_corr = if_else(!is.na(OD720trunc_Lmu_raw), OD720trunc_Lmu_raw, 0),
          OD720trunc_Lmu_se = if_else(!is.na(OD720trunc_Lmu_se), OD720trunc_Lmu_se, 0),
          OD720trunc_Lintercept = if_else(!is.na(OD720trunc_Lintercept), OD720trunc_Lintercept, 0),
          OD720trunc_Lintercept_se = if_else(!is.na(OD720trunc_Lintercept_se), OD720trunc_Lintercept_se, 0),
          OD720trunc_Lmax = if_else(!is.na(OD720trunc_Lmax), OD720trunc_Lmax, 0),
          OD720trunc_Lmax_se = if_else(!is.na(OD720trunc_Lmax_se), OD720trunc_Lmax_se, 0),
          GrowthAmpOD720Flag = if_else(GrowthAmpOD720 >= OD720GrowthAmpFlagValue, 1, 0),
          GrowthAmpOD720Flag_14days = if_else(GrowthAmpOD720_14days >= OD720GrowthAmpFlagValue, 1, 0),
          OD720trunc_Lmu_corr = if_else(GrowthAmpOD720Flag_14days == 1 & OD720trunc_Lmu_corr >= 0 , OD720trunc_Lmu_corr, 0),
          # OD720trunc_LseGrowthFlag = case_when(OD720trunc_Lmu_corr > 0 & OD720trunc_Lmu_corr * GrowthFlagError > OD720trunc_Lmu_se ~ 1,
          #                                   OD720trunc_Lmu_corr > 0 & OD720trunc_Lmu_corr * GrowthFlagError < OD720trunc_Lmu_se ~ 0,
          #                                   OD720trunc_Lmu_corr == 0 ~ 1),

           OD720trunc_LseGrowthFlag = if_else(OD720trunc_Lmu_se < OD720trunc_Lmu_corr * GrowthFlagError, 1, 0),
          OD720trunc_TimeHalfPmax_h = (log(2-(OD720trunc_Lintercept/(OD720trunc_Lmax - OD720trunc_Lintercept))))/OD720trunc_Lmu_corr,
          OD720trunc_TimeHalfPmax_h = if_else(!is.infinite(OD720trunc_TimeHalfPmax_h), OD720trunc_TimeHalfPmax_h, 0),
          OD720trunc_TimeHalfPmax_h = if_else(!is.nan(OD720trunc_TimeHalfPmax_h), OD720trunc_TimeHalfPmax_h, 0)
          )

# deltaOD truncated logistic fits
# assign 0 to growth fits that are NA's and for tubes with absolute amplitude change of less than 0.007.Defined at line 75 
deltaODtruncLog_unnest <- LogFitsData %>%
   select(c(Filename, CDateTime, Tube, ExpDate, MC, PrimaryOperator, Run, Description, ProjectID, ID, Strain, Inoc_mL, Media_mL, ExpCul, InnocDate, ExpStartTime, Par_ue, Photoperiod, Temp_c, O2, WL, LightShape, Media, Source, Optode, OptodeCh, FinalpH, Par_ueAdjusted, DateOfAdjustment, ElaspedHoursAtAdjustment, ExpEndDate, ExpEndHour, SetActinic_day, tubedata, IsOD720_GLagNeg, lognormOD720_TimetoMax, lognormdeltaOD_TimetoMax, OD720trunc_TimetoMax, deltaODtrunc_TimetoMax, IsdeltaOD_GLagNeg, lognormOD720_GompLag, lognormdeltaOD_GompLag, GrowthAmpOD720, GrowthAmpOD720_14days, GrowthAmpdeltaOD, GrowthAmpdeltaOD_14days, deltaODtrunc_logistic_tidied, deltaODtrunc_logistic_predict, DoubledeltaODFlag_14days, DoubleOD720Flag_14days, O2_Category)
          ) %>%
unnest(cols = c(deltaODtrunc_logistic_tidied), names_sep = "_", keep_empty = TRUE) %>%
  pivot_wider(names_from =  deltaODtrunc_logistic_tidied_term, values_from = c(deltaODtrunc_logistic_tidied_estimate, deltaODtrunc_logistic_tidied_std.error, deltaODtrunc_logistic_tidied_statistic, deltaODtrunc_logistic_tidied_p.value), names_sep = "_")  %>%
  rename(deltaODtrunc_Lmax = deltaODtrunc_logistic_tidied_estimate_Lmax,
         deltaODtrunc_Lmax_se = deltaODtrunc_logistic_tidied_std.error_Lmax,
         deltaODtrunc_Lmu_raw = deltaODtrunc_logistic_tidied_estimate_Lmu,
         deltaODtrunc_Lmu_se = deltaODtrunc_logistic_tidied_std.error_Lmu,
         deltaODtrunc_Lintercept = deltaODtrunc_logistic_tidied_estimate_Lintercept,
         deltaODtrunc_Lintercept_se = deltaODtrunc_logistic_tidied_std.error_Lintercept) %>%
   mutate(deltaODtrunc_Lmu_corr = if_else(!is.na(deltaODtrunc_Lmu_raw), deltaODtrunc_Lmu_raw, 0),
          deltaODtrunc_Lmu_se = if_else(!is.na(deltaODtrunc_Lmu_se), deltaODtrunc_Lmu_se, 0),
          deltaODtrunc_Lintercept = if_else(!is.na(deltaODtrunc_Lintercept), deltaODtrunc_Lintercept, 0),
          deltaODtrunc_Lintercept_se = if_else(!is.na(deltaODtrunc_Lintercept_se), deltaODtrunc_Lintercept_se, 0),
          deltaODtrunc_Lmax = if_else(!is.na(deltaODtrunc_Lmax), deltaODtrunc_Lmax, 0),
          deltaODtrunc_Lmax_se = if_else(!is.na(deltaODtrunc_Lmax_se), deltaODtrunc_Lmax_se, 0),
          GrowthAmpdeltaODFlag = if_else(GrowthAmpdeltaOD >= deltaODGrowthAmpFlagValue , 1, 0),
          GrowthAmpdeltaODFlag_14days = if_else(GrowthAmpdeltaOD_14days >= deltaODGrowthAmpFlagValue, 1, 0),
          deltaODtrunc_Lmu_corr = if_else(GrowthAmpdeltaODFlag_14days == 1 & deltaODtrunc_Lmu_corr >= 0, deltaODtrunc_Lmu_corr, 0),
          # deltaODtrunc_LseGrowthFlag = case_when(deltaODtrunc_Lmu_corr > 0 & deltaODtrunc_Lmu_corr * GrowthFlagError > deltaODtrunc_Lmu_se ~ 1,
          #                                   deltaODtrunc_Lmu_corr > 0 & deltaODtrunc_Lmu_corr * GrowthFlagError < deltaODtrunc_Lmu_se ~ 0,
          #                                   deltaODtrunc_Lmu_corr == 0 ~ 1),

          deltaODtrunc_LseGrowthFlag = if_else(deltaODtrunc_Lmu_se < deltaODtrunc_Lmu_corr * GrowthFlagError, 1, 0),
          deltaODtrunc_TimeHalfPmax_h = (log(2-(deltaODtrunc_Lintercept/(deltaODtrunc_Lmax - deltaODtrunc_Lintercept))))/deltaODtrunc_Lmu_corr,
          deltaODtrunc_TimeHalfPmax_h = if_else(!is.infinite(deltaODtrunc_TimeHalfPmax_h), deltaODtrunc_TimeHalfPmax_h, 0),
          deltaODtrunc_TimeHalfPmax_h = if_else(!is.nan(deltaODtrunc_TimeHalfPmax_h), deltaODtrunc_TimeHalfPmax_h, 0)
          )

# OD720 truncated exponential fits
# assign 0 to growth fits that are NA's and for tubes with absolute amplitude change of less than 0.007.Defined at line 75 
OD720truncExp_unnest <- LogFitsData %>%
   select(c(Filename, CDateTime, Tube, ExpDate, MC, PrimaryOperator, Run, Description, ProjectID, ID, Strain, Inoc_mL, Media_mL, ExpCul, InnocDate, ExpStartTime, Par_ue, Photoperiod, Temp_c, O2, WL, LightShape, Media, Source, Optode, OptodeCh, FinalpH, Par_ueAdjusted, DateOfAdjustment, ElaspedHoursAtAdjustment, ExpEndDate, ExpEndHour, SetActinic_day, tubedata, IsOD720_GLagNeg, lognormOD720_TimetoMax, lognormdeltaOD_TimetoMax, OD720trunc_TimetoMax, deltaODtrunc_TimetoMax, IsdeltaOD_GLagNeg, lognormOD720_GompLag, lognormdeltaOD_GompLag, GrowthAmpOD720, GrowthAmpOD720_14days, GrowthAmpdeltaOD, GrowthAmpdeltaOD_14days, OD720trunc_exp_tidied, OD720trunc_exp_predict, DoubledeltaODFlag_14days, DoubleOD720Flag_14days, O2_Category)
          ) %>%
  unnest(cols = c(OD720trunc_exp_tidied), names_sep = "_", keep_empty = TRUE) %>%
  pivot_wider(names_from =  OD720trunc_exp_tidied_term, values_from = c(OD720trunc_exp_tidied_estimate, OD720trunc_exp_tidied_std.error, OD720trunc_exp_tidied_statistic, OD720trunc_exp_tidied_p.value), names_sep = "_") %>%
  rename(OD720trunc_Emu_raw = OD720trunc_exp_tidied_estimate_Emu,
         OD720trunc_Emu_se = OD720trunc_exp_tidied_std.error_Emu,
         OD720trunc_Eintercept = OD720trunc_exp_tidied_estimate_Eintercept,             
         OD720trunc_Eintercept_se = OD720trunc_exp_tidied_std.error_Eintercept)  %>%
   mutate(OD720trunc_Emu_corr = if_else(!is.na(OD720trunc_Emu_raw), OD720trunc_Emu_raw, 0),
          OD720trunc_Emu_se = if_else(!is.na(OD720trunc_Emu_se), OD720trunc_Emu_se, 0),
          OD720trunc_Eintercept = if_else(!is.na(OD720trunc_Eintercept), OD720trunc_Eintercept, 0),
          OD720trunc_Eintercept_se = if_else(!is.na(OD720trunc_Eintercept_se), OD720trunc_Eintercept_se, 0),
          GrowthAmpOD720Flag = if_else(GrowthAmpOD720 >= OD720GrowthAmpFlagValue, 1, 0),
          GrowthAmpOD720Flag_14days = if_else(GrowthAmpOD720_14days >= OD720GrowthAmpFlagValue, 1, 0),
          OD720trunc_Emu_corr = if_else(GrowthAmpOD720Flag_14days == 1 & OD720trunc_Emu_corr >= 0 , OD720trunc_Emu_corr, 0),
          # OD720trunc_EseGrowthFlag = case_when(OD720trunc_Emu_corr > 0 & OD720trunc_Emu_corr * GrowthFlagError > OD720trunc_Emu_se ~ 1,
          #                                   OD720trunc_Emu_corr > 0 & OD720trunc_Emu_corr * GrowthFlagError < OD720trunc_Emu_se ~ 0,
          #                                   OD720trunc_Emu_corr == 0 ~ 1)
         OD720trunc_EseGrowthFlag = if_else(OD720trunc_Emu_se < OD720trunc_Emu_corr * GrowthFlagError, 1, 0),
          )

# deltaOD truncated exponential fits
# assign 0 to growth fits that are NA's and for tubes with absolute amplitude change of less than 0.007.Defined at line 75 
deltaODtruncExp_unnest <- LogFitsData %>%
   select(c(Filename, CDateTime, Tube, ExpDate, MC, PrimaryOperator, Run, Description, ProjectID, ID, Strain, Inoc_mL, Media_mL, ExpCul, InnocDate, ExpStartTime, Par_ue, Photoperiod, Temp_c, O2, WL, LightShape, Media, Source, Optode, OptodeCh, FinalpH, Par_ueAdjusted, DateOfAdjustment, ElaspedHoursAtAdjustment, ExpEndDate, ExpEndHour, SetActinic_day, tubedata, IsOD720_GLagNeg, lognormOD720_TimetoMax, lognormdeltaOD_TimetoMax, OD720trunc_TimetoMax, deltaODtrunc_TimetoMax, IsdeltaOD_GLagNeg, lognormOD720_GompLag, lognormdeltaOD_GompLag, GrowthAmpOD720, GrowthAmpOD720_14days, GrowthAmpdeltaOD, GrowthAmpdeltaOD_14days, deltaODtrunc_exp_tidied, deltaODtrunc_exp_predict, DoubledeltaODFlag_14days, DoubleOD720Flag_14days, O2_Category)
          ) %>%
unnest(cols = c(deltaODtrunc_exp_tidied), names_sep = "_", keep_empty = TRUE) %>%
  pivot_wider(names_from =  deltaODtrunc_exp_tidied_term, values_from = c(deltaODtrunc_exp_tidied_estimate, deltaODtrunc_exp_tidied_std.error, deltaODtrunc_exp_tidied_statistic, deltaODtrunc_exp_tidied_p.value), names_sep = "_")  %>%
  rename(deltaODtrunc_Emu_raw = deltaODtrunc_exp_tidied_estimate_Emu,
         deltaODtrunc_Emu_se = deltaODtrunc_exp_tidied_std.error_Emu,
         deltaODtrunc_Eintercept = deltaODtrunc_exp_tidied_estimate_Eintercept,
         deltaODtrunc_Eintercept_se = deltaODtrunc_exp_tidied_std.error_Eintercept) %>%
   mutate(deltaODtrunc_Emu_corr = if_else(!is.na(deltaODtrunc_Emu_raw), deltaODtrunc_Emu_raw, 0),
          deltaODtrunc_Emu_se = if_else(!is.na(deltaODtrunc_Emu_se), deltaODtrunc_Emu_se, 0),
          deltaODtrunc_Eintercept = if_else(!is.na(deltaODtrunc_Eintercept), deltaODtrunc_Eintercept, 0),
          deltaODtrunc_Eintercept_se = if_else(!is.na(deltaODtrunc_Eintercept_se), deltaODtrunc_Eintercept_se, 0),
          GrowthAmpdeltaODFlag = if_else(GrowthAmpdeltaOD >= deltaODGrowthAmpFlagValue , 1, 0),
          GrowthAmpdeltaODFlag_14days = if_else(GrowthAmpdeltaOD_14days >= deltaODGrowthAmpFlagValue, 1, 0),
          deltaODtrunc_Emu_corr = if_else(GrowthAmpdeltaODFlag_14days == 1 & deltaODtrunc_Emu_corr >= 0, deltaODtrunc_Emu_corr, 0),
          # deltaODtrunc_EseGrowthFlag = case_when(deltaODtrunc_Emu_corr > 0 & deltaODtrunc_Emu_corr * GrowthFlagError > deltaODtrunc_Emu_se ~ 1,
          #                                   deltaODtrunc_Emu_corr > 0 & deltaODtrunc_Emu_corr * GrowthFlagError < deltaODtrunc_Emu_se ~ 0,
          #                                   deltaODtrunc_Emu_corr == 0 ~ 1),
           deltaODtrunc_EseGrowthFlag = if_else(deltaODtrunc_Emu_se < deltaODtrunc_Emu_corr * GrowthFlagError, 1, 0),
          )

glimpse(lognormdeltaOD_unnest)

```

```{r combine unnested fits for PICO Shiny App}

# OD720 gompertz fits
lognormOD720_unnest <- lognormOD720_unnest %>%
select(-c(lognormOD720_Gompertz_tidied_estimate_NA,
            lognormOD720_Gompertz_tidied_p.value_NA,
            lognormOD720_Gompertz_tidied_statistic_NA,
            lognormOD720_Gompertz_tidied_std.error_NA,
            lognormOD720_Gompertz_tidied_statistic_GLag,
            lognormOD720_Gompertz_tidied_statistic_Gmax,
            lognormOD720_Gompertz_tidied_statistic_Gmu,
            lognormOD720_Gompertz_tidied_p.value_GLag,
            lognormOD720_Gompertz_tidied_p.value_Gmax,
            lognormOD720_Gompertz_tidied_p.value_Gmu)
       )

# deltaOD gompertz fits
lognormdeltaOD_unnest <- lognormdeltaOD_unnest %>%
  select(-c(lognormdeltaOD_Gompertz_tidied_estimate_NA,
            lognormdeltaOD_Gompertz_tidied_p.value_NA,
            lognormdeltaOD_Gompertz_tidied_statistic_NA,
            lognormdeltaOD_Gompertz_tidied_std.error_NA,
            lognormdeltaOD_Gompertz_tidied_statistic_GLag,
            lognormdeltaOD_Gompertz_tidied_statistic_Gmax,
            lognormdeltaOD_Gompertz_tidied_statistic_Gmu,
            lognormdeltaOD_Gompertz_tidied_p.value_GLag,
            lognormdeltaOD_Gompertz_tidied_p.value_Gmax,
            lognormdeltaOD_Gompertz_tidied_p.value_Gmu)
         )

# OD720 logistic fits
OD720_unnest <- OD720_unnest %>%
select(-c(OD720_logistic_tidied_estimate_NA,
          OD720_logistic_tidied_p.value_NA,
          OD720_logistic_tidied_statistic_NA,
          OD720_logistic_tidied_std.error_NA,
          OD720_logistic_tidied_statistic_Lintercept,
          OD720_logistic_tidied_statistic_Lmax,
          OD720_logistic_tidied_statistic_Lmu,
          OD720_logistic_tidied_p.value_Lintercept,
          OD720_logistic_tidied_p.value_Lmax,
          OD720_logistic_tidied_p.value_Lmu)
       )

# deltaOD logistic fits
deltaOD_unnest <- deltaOD_unnest %>%
  select(-c(deltaOD_logistic_tidied_estimate_NA,
            deltaOD_logistic_tidied_p.value_NA,
            deltaOD_logistic_tidied_statistic_NA,
            deltaOD_logistic_tidied_std.error_NA,
            deltaOD_logistic_tidied_statistic_Lintercept,
            deltaOD_logistic_tidied_statistic_Lmax,
            deltaOD_logistic_tidied_statistic_Lmu,
            deltaOD_logistic_tidied_p.value_Lintercept,
            deltaOD_logistic_tidied_p.value_Lmax,
            deltaOD_logistic_tidied_p.value_Lmu)
         )

# OD720 truncated logistic fits
OD720truncLog_unnest <- OD720truncLog_unnest %>%
select(-c(OD720trunc_logistic_tidied_estimate_NA,
           OD720trunc_logistic_tidied_p.value_NA,
           OD720trunc_logistic_tidied_statistic_NA,
           OD720trunc_logistic_tidied_std.error_NA,
            OD720trunc_logistic_tidied_statistic_Lintercept,
            OD720trunc_logistic_tidied_statistic_Lmax,
            OD720trunc_logistic_tidied_statistic_Lmu,
            OD720trunc_logistic_tidied_p.value_Lintercept,
            OD720trunc_logistic_tidied_p.value_Lmax,
            OD720trunc_logistic_tidied_p.value_Lmu)
       )

# deltaOD truncated logistic fits
deltaODtruncLog_unnest <- deltaODtruncLog_unnest %>%
  select(-c(deltaODtrunc_logistic_tidied_estimate_NA,
            deltaODtrunc_logistic_tidied_p.value_NA,
            deltaODtrunc_logistic_tidied_statistic_NA,
            deltaODtrunc_logistic_tidied_std.error_NA,
            deltaODtrunc_logistic_tidied_statistic_Lintercept,
            deltaODtrunc_logistic_tidied_statistic_Lmax,
            deltaODtrunc_logistic_tidied_statistic_Lmu,
            deltaODtrunc_logistic_tidied_p.value_Lintercept,
            deltaODtrunc_logistic_tidied_p.value_Lmax,
            deltaODtrunc_logistic_tidied_p.value_Lmu)
         )

# OD720 truncated exponential fits
OD720truncExp_unnest <- OD720truncExp_unnest %>%
select(-c(OD720trunc_exp_tidied_estimate_NA,
           OD720trunc_exp_tidied_p.value_NA,
           OD720trunc_exp_tidied_statistic_NA,
           OD720trunc_exp_tidied_std.error_NA,
            OD720trunc_exp_tidied_statistic_Eintercept,
            OD720trunc_exp_tidied_statistic_Emu,
            OD720trunc_exp_tidied_p.value_Eintercept,
            OD720trunc_exp_tidied_p.value_Emu)
       )

# deltaOD truncated exponential fits
deltaODtruncExp_unnest <- deltaODtruncExp_unnest %>%
  select(-c(deltaODtrunc_exp_tidied_estimate_NA,
            deltaODtrunc_exp_tidied_p.value_NA,
            deltaODtrunc_exp_tidied_statistic_NA,
            deltaODtrunc_exp_tidied_std.error_NA,
            deltaODtrunc_exp_tidied_statistic_Eintercept,
            deltaODtrunc_exp_tidied_statistic_Emu,
            deltaODtrunc_exp_tidied_p.value_Eintercept,
            deltaODtrunc_exp_tidied_p.value_Emu)
         )

glimpse(deltaOD_unnest)

# Join Full Growth Rate Data
UnnestedGompertzData <- left_join(x = lognormOD720_unnest, y = lognormdeltaOD_unnest)

UnnestedLogData <- left_join(x = OD720_unnest, y = deltaOD_unnest)

UnnestedFullData <- left_join(x = UnnestedLogData, y = UnnestedGompertzData)

glimpse(UnnestedFullData)

# Join Truncated Growth Rate Data
UnnestedTruncLogData <- left_join(x = OD720truncLog_unnest, y =  deltaODtruncLog_unnest)

UnnestedTruncExpData <- left_join(x = OD720truncExp_unnest, y =  deltaODtruncExp_unnest)

UnnestedTruncData <- left_join(x = UnnestedTruncLogData, y = UnnestedTruncExpData)

#combine Full and Truncated unnested data
NestedFitsData <- left_join(x = UnnestedFullData, y = UnnestedTruncData)

glimpse(NestedFitsData)

#remove redundant objects
remove(lognormOD720_unnest,
       lognormdeltaOD_unnest,
       OD720_unnest,
       deltaOD_unnest,
       UnnestedGompertzData,
       UnnestedLogData,
       OD720truncLog_unnest,
       deltaODtruncLog_unnest,
       OD720truncExp_unnest,
       deltaODtruncExp_unnest,
       UnnestedTruncLogData,
       UnnestedTruncExpData,
       UnnestedTruncData,
       UnnestedFullData
       )

```

<!-- Chunk to convert categorize oxygen levels based on input in metacatalog.  Have to do this here as this column was added after many runs were already processed and the merge with metacatalog already occured.  Run numbers >= 36 and <= 91 and 0% oxygen are labelled 'Low', runs <= 35  and 0% oxygen are labelled 'Intermediate'.  All runs >= 92 are from the metacatelog. All runs with 21% oxygen are labelled 'High'.  Assign [O2] to O2_uM based on category -->

```{r}
NestedFitsData <- NestedFitsData %>% 
  mutate(O2_Category = case_when(Run <= 35 & O2 == 0 ~ "Intermediate",
                                 Run >= 36 & Run <= 91 & O2 == 0 ~ "Low",
                                 O2 == 21 ~ "High",
                                 Run >= 92 ~ O2_Category),
         O2_uM = case_when(O2_Category == "Low" ~ 2.5,
                            O2_Category == "Intermediate" ~ 25,
                            O2_Category == "High" ~ 250))

```

<!-- Chunk to calculate the rate per hour of actinic light rise to peak. -->
```{r}
AllNestedFitsData <- NestedFitsData %>% 
  mutate(RateActinic_h = case_when(LightShape == "Sine"  ~ Par_ue/(Photoperiod/2)))

glimpse(AllNestedFitsData)

```

```{r remove trunc and gompertz from nested DF}

# remove gompertz and truncated growth data
NestedFitsData <- NestedFitsData %>% 
  select(-contains(c("trunc", "Gompertz", "Gomp", "OD720_G", "deltaOD_G")))

# need the deltaOD trunc growth rate for future comparisons
onlydeltaODtrunc<- AllNestedFitsData %>% 
  select((c("ID", "deltaODtrunc_Lmu_corr")))

# add deltaOD truncated growth rates to DF
NestedFitsData <- full_join(NestedFitsData, onlydeltaODtrunc)

glimpse(NestedFitsData)

remove(onlydeltaODtrunc)

```

```{r, pivot Nested data longer for GAM}

# OD720 pivot longer Gmax, Gmu, Glag, Lmu, Lmax and Lintercept estimates
Pivot_Longer_OD720_est <- AllNestedFitsData %>%
  select(c(Filename, Tube, ID, Run, PrimaryOperator, Strain, Par_ue,RateActinic_h, Photoperiod, SetActinic_day, Temp_c, O2, O2_Category,O2_uM, WL, LightShape, IsOD720_GLagNeg, lognormOD720_TimetoMax, lognormOD720_GompLag ,GrowthAmpOD720, GrowthAmpOD720Flag, GrowthAmpOD720_14days, GrowthAmpOD720Flag_14days, DoubleOD720Flag_14days, OD720_GseGrowthFlag, OD720_LseGrowthFlag, OD720trunc_LseGrowthFlag, OD720_Gmu_corr, OD720_Gmax, OD720_GLag, OD720_Lmu_corr, OD720_Lmax, OD720_Lintercept, OD720trunc_Lmu_corr, OD720trunc_Lmax, OD720trunc_Lintercept
            )) %>%
   pivot_longer(c(OD720_Gmu_corr, OD720_Gmax, OD720_GLag, OD720_Lmu_corr, OD720_Lmax, OD720_Lintercept, OD720trunc_Lmu_corr, OD720trunc_Lmax, OD720trunc_Lintercept), names_to = c("Wavelength", "term"), values_to = c("estimate"), names_sep = '\\_')

# OD720 pivot longer Gmax, Gmu, Glag , Lmu, Lmax and Lintercept standard error
Pivot_Longer_OD720_se <- AllNestedFitsData %>%
  select(c(Filename, Tube, ID, Run, PrimaryOperator, Strain, Par_ue, RateActinic_h, Photoperiod, SetActinic_day, Temp_c, O2, O2_Category,O2_uM, WL, LightShape, IsOD720_GLagNeg, lognormOD720_TimetoMax, lognormOD720_GompLag ,GrowthAmpOD720, GrowthAmpOD720Flag, GrowthAmpOD720_14days, GrowthAmpOD720Flag_14days, DoubleOD720Flag_14days, OD720_GseGrowthFlag, OD720_LseGrowthFlag, OD720trunc_LseGrowthFlag,  OD720_Gmu_se, OD720_Gmax_se, OD720_GLag_se, OD720_Lmu_se, OD720_Lmax_se, OD720_Lintercept_se,OD720trunc_Lmu_se, OD720trunc_Lmax_se, OD720trunc_Lintercept_se
            )) %>%
   pivot_longer(c(OD720_Gmu_se, OD720_Gmax_se, OD720_GLag_se, OD720_Lmu_se, OD720_Lmax_se, OD720_Lintercept_se, OD720trunc_Lmu_se, OD720trunc_Lmax_se, OD720trunc_Lintercept_se), names_to = c("Wavelength", "term"), values_to = c("se"), names_sep = '\\_')

# Join OD720 estimate and se DF 
Longer_OD720_Joined <- full_join(Pivot_Longer_OD720_est, Pivot_Longer_OD720_se)

remove(Pivot_Longer_OD720_est, Pivot_Longer_OD720_se)

# rename common columns for full join
Longer_OD720_Joined <- Longer_OD720_Joined %>%
  rename(Is_GLagNeg = 'IsOD720_GLagNeg',
         lognorm_TimetoMax = 'lognormOD720_TimetoMax',
         lognorm_GompLag = 'lognormOD720_GompLag',
         GrowthAmp = 'GrowthAmpOD720',
         GrowthAmpFlag = 'GrowthAmpOD720Flag',
         GrowthAmp_14days = 'GrowthAmpOD720_14days',
         GrowthAmp_14daysFlag = 'GrowthAmpOD720Flag_14days',
         Doubled_14daysFlag = 'DoubleOD720Flag_14days',
         Gmu_seFlag = 'OD720_GseGrowthFlag',
         Lmu_seFlag = 'OD720_LseGrowthFlag',
         Lmutrunc_seFlag = 'OD720trunc_LseGrowthFlag'
         )

# deltaOD pivot longer Gmax, Gmu, Glag, Lmu, Lmax and Lintercept estimates
Pivot_Longer_deltaOD_est <- AllNestedFitsData %>%
  select(c(Filename, Tube, ID, Run, PrimaryOperator, Strain, Par_ue, RateActinic_h, Photoperiod, SetActinic_day, Temp_c, O2, O2_Category,O2_uM, WL, LightShape, IsdeltaOD_GLagNeg, lognormdeltaOD_TimetoMax, lognormdeltaOD_GompLag ,GrowthAmpdeltaOD, GrowthAmpdeltaODFlag, GrowthAmpdeltaOD_14days, GrowthAmpdeltaODFlag_14days, DoubledeltaODFlag_14days, deltaOD_GseGrowthFlag, deltaOD_LseGrowthFlag, deltaODtrunc_LseGrowthFlag,  deltaOD_Gmu_corr, deltaOD_Gmax, deltaOD_GLag, deltaOD_Lmu_corr, deltaOD_Lmax, deltaOD_Lintercept, deltaODtrunc_Lmu_corr, deltaODtrunc_Lmax, deltaODtrunc_Lintercept
            )) %>%
   pivot_longer(c(deltaOD_Gmu_corr, deltaOD_Gmax, deltaOD_GLag, deltaOD_Lmu_corr, deltaOD_Lmax, deltaOD_Lintercept, deltaODtrunc_Lmu_corr, deltaODtrunc_Lmax, deltaODtrunc_Lintercept), names_to = c("Wavelength", "term"), values_to = c("estimate"), names_sep = '\\_')

# deltaOD pivot longer Gmax, Gmu, Glag , Lmu, Lmax and Lintercept standard error
Pivot_Longer_deltaOD_se <- AllNestedFitsData %>%
  select(c(Filename, Tube, ID, Run, PrimaryOperator, Strain, Par_ue, RateActinic_h, Photoperiod, SetActinic_day, Temp_c, O2, O2_Category,O2_uM, WL, LightShape, IsdeltaOD_GLagNeg, lognormdeltaOD_TimetoMax, lognormdeltaOD_GompLag ,GrowthAmpdeltaOD, GrowthAmpdeltaODFlag, GrowthAmpdeltaOD_14days, GrowthAmpdeltaODFlag_14days, DoubledeltaODFlag_14days, deltaOD_GseGrowthFlag, deltaOD_LseGrowthFlag, deltaODtrunc_LseGrowthFlag, deltaOD_Gmu_se, deltaOD_Gmax_se, deltaOD_GLag_se, deltaOD_Lmu_se, deltaOD_Lmax_se, deltaOD_Lintercept_se, deltaODtrunc_Lmu_se, deltaODtrunc_Lmax_se, deltaODtrunc_Lintercept_se
            )) %>%
   pivot_longer(c(deltaOD_Gmu_se, deltaOD_Gmax_se, deltaOD_GLag_se, deltaOD_Lmu_se, deltaOD_Lmax_se, deltaOD_Lintercept_se, deltaODtrunc_Lmu_se, deltaODtrunc_Lmax_se, deltaODtrunc_Lintercept_se), names_to = c("Wavelength", "term"), values_to = c("se"), names_sep = '\\_')

# Join deltaOD estimate and se DF 
Longer_deltaOD_Joined <- full_join(Pivot_Longer_deltaOD_est, Pivot_Longer_deltaOD_se)

remove(Pivot_Longer_deltaOD_est, Pivot_Longer_deltaOD_se)

# rename common columns for full join
Longer_deltaOD_Joined <- Longer_deltaOD_Joined %>%
  rename(Is_GLagNeg = 'IsdeltaOD_GLagNeg',
         lognorm_TimetoMax = 'lognormdeltaOD_TimetoMax',
         lognorm_GompLag = 'lognormdeltaOD_GompLag',
         GrowthAmp = 'GrowthAmpdeltaOD',
         GrowthAmpFlag = 'GrowthAmpdeltaODFlag',
         GrowthAmp_14days = 'GrowthAmpdeltaOD_14days',
         GrowthAmp_14daysFlag = 'GrowthAmpdeltaODFlag_14days',
         Doubled_14daysFlag = 'DoubledeltaODFlag_14days',
         Gmu_seFlag = 'deltaOD_GseGrowthFlag',
         Lmu_seFlag = 'deltaOD_LseGrowthFlag',
         Lmutrunc_seFlag = 'deltaODtrunc_LseGrowthFlag'
         )

#Join OD720 and deltaOD fits data
LongerFitsData <- rbind(Longer_OD720_Joined, Longer_deltaOD_Joined) %>%
  mutate_if(is.numeric, round, digits = 3)

remove(Longer_OD720_Joined, Longer_deltaOD_Joined)

glimpse(LongerFitsData)

```

Save Long DF for GAM and Nested for further analysis 
```{r save files for GAM and Shiny App}

saveRDS(object = LongerFitsData, file = file.path(DataOut, paste(Project, "MiSa_LongerFitsData.Rds",  sep = "_")), ascii = FALSE, version = NULL, compress = TRUE, refhook = NULL)

# write_csv(LongerFitsData, file = file.path(DataOut, paste(Project, "MiSa_LongerFitsData",".csv",sep = "_"),fsep = .Platform$file.sep))

saveRDS(object = NestedFitsData, file = file.path(DataOut, paste(Project, "MiSa_NestedFitsData.Rds",  sep = "_")), ascii = FALSE, version = NULL, compress = TRUE, refhook = NULL)

saveRDS(NestedFitsData, file = file.path(ShinyPath, paste(Project, "MiSa_GrowthLogs", ".Rds",sep = "_"),fsep = .Platform$file.sep))

```

```{r LogisticdeltaODPlots, fig.height= 10, fig.width= 10}

SS120deltaODLmuMatrixPlot <- NestedFitsData %>%
   filter(Strain == "SS120") %>%
  ggplot() +
  geom_point(aes(x = Par_ue, y = deltaOD_Lmu_corr, colour = as.factor(WL)), size = 3, alpha = 0.5) +
 geom_line(aes(x = Par_ue, y = deltaOD_Lmu_corr, colour = as.factor(WL)), linewidth = 0.1, alpha = 0.4) +
   labs(subtitle = "Strain, [Oxygen], Photoperiod, Actinic PAR (nm)", y = "deltaOD [~Chl] " ~ "Logistic Growth Rate" ~ "("~h^-1~")" , x = "Growth Light (µmol photons" ~m^-2 ~s^-1~")") +
 geom_text(aes(x = Par_ue, y = deltaOD_Lmu_corr, label= Run), alpha = 0.5 ,hjust = 0.5, vjust=-0.5) +
  geom_errorbar(aes(x = Par_ue, ymin = (deltaOD_Lmu_corr - deltaOD_Lmu_se), ymax = (deltaOD_Lmu_corr + deltaOD_Lmu_se)), size = 0.3, color = 'black', data = . %>% filter(GrowthAmpdeltaODFlag_14days  == 1)) +
  scale_x_continuous(breaks=seq(0, 200, by = 50)) +
  coord_cartesian(xlim = c (0, 200)) +
  #   scale_y_continuous(breaks=seq(-0.03, 0.07, by = 0.01)) +
  # coord_cartesian(ylim = c (-0.03, 0.07)) +
  scale_colour_manual(values = MCMIXColours) +
  ggh4x::facet_nested(cols = vars(Strain, O2_uM, Photoperiod), rows = vars(WL), 
                    labeller = labeller(WL = label_both, O2_uM = label_both, Strain = label_both, Photoperiod = label_value)) +
  theme_bw() +
  labs(colour = "Actinic PAR (nm)")

SS120deltaODLmuMatrixPlot

 NestedFitsData %>%
   filter(Strain == "SS120",
          deltaOD_LseGrowthFlag == 1) %>%
  ggplot() +
  geom_point(aes(x = Par_ue, y = deltaOD_Lmu_corr, colour = as.factor(WL)), size = 3, alpha = 0.5) +
 geom_line(aes(x = Par_ue, y = deltaOD_Lmu_corr, colour = as.factor(WL)), linewidth = 0.1, alpha = 0.4) +
   labs(subtitle = "Strain, [Oxygen],Photoperiod, Actinic PAR (nm) - Outliers filtered by standard error > 20% of growth rate", y = "deltaOD [~Chl] " ~ "Logistic Growth Rate" ~ "("~h^-1~")" , x = "Growth Light (µmol photons" ~m^-2 ~s^-1~")") +
 geom_text(aes(x = Par_ue, y = deltaOD_Lmu_corr, label= Run), alpha = 0.5 ,hjust = 0.5, vjust=-0.5) +
  geom_errorbar(aes(x = Par_ue, ymin = (deltaOD_Lmu_corr - deltaOD_Lmu_se), ymax = (deltaOD_Lmu_corr + deltaOD_Lmu_se)), size = 0.3, color = 'black', data = . %>% filter(GrowthAmpdeltaODFlag_14days  == 1)) +
  scale_x_continuous(breaks=seq(0, 200, by = 50)) +
  coord_cartesian(xlim = c (0, 200)) +
  #   scale_y_continuous(breaks=seq(-0.03, 0.07, by = 0.01)) +
  # coord_cartesian(ylim = c (-0.03, 0.07)) +
  scale_colour_manual(values = MCMIXColours) +
  ggh4x::facet_nested(cols = vars(Strain, O2_uM, Photoperiod), rows = vars(WL), 
                    labeller = labeller(WL = label_both, O2_uM = label_both, Strain = label_both, Photoperiod = label_value)) +
  theme_bw() +
  labs(colour = "Actinic PAR (nm)")

 NestedFitsData %>%
   filter(Strain == "SS120",
          deltaOD_Lmu_corr == 0 |
          deltaOD_Lmu_corr != 0 & deltaOD_LseGrowthFlag == 1) %>%
  ggplot() +
  geom_point(aes(x = Par_ue, y = deltaOD_Lmu_corr, colour = as.factor(WL)), size = 3, alpha = 0.5) +
 geom_line(aes(x = Par_ue, y = deltaOD_Lmu_corr, colour = as.factor(WL)), linewidth = 0.1, alpha = 0.4) +
   labs(subtitle = "Strain, [Oxygen],Photoperiod, Actinic PAR (nm) - Of cultures that grew, outliers filtered by standard error > 20% of growth rate.", y = "deltaOD [~Chl] " ~ "Logistic Growth Rate" ~ "("~h^-1~")" , x = "Growth Light (µmol photons" ~m^-2 ~s^-1~")") +
 geom_text(aes(x = Par_ue, y = deltaOD_Lmu_corr, label= Run), alpha = 0.5 ,hjust = 0.5, vjust=-0.5) +
  geom_errorbar(aes(x = Par_ue, ymin = (deltaOD_Lmu_corr - deltaOD_Lmu_se), ymax = (deltaOD_Lmu_corr + deltaOD_Lmu_se)), size = 0.3, color = 'black', data = . %>% filter(GrowthAmpdeltaODFlag_14days  == 1)) +
  scale_x_continuous(breaks=seq(0, 200, by = 50)) +
  coord_cartesian(xlim = c (0, 200)) +
  #   scale_y_continuous(breaks=seq(-0.03, 0.07, by = 0.01)) +
  # coord_cartesian(ylim = c (-0.03, 0.07)) +
  scale_colour_manual(values = MCMIXColours) +
  ggh4x::facet_nested(cols = vars(Strain, O2_uM, Photoperiod), rows = vars(WL), 
                    labeller = labeller(WL = label_both, O2_uM = label_both, Strain = label_both, Photoperiod = label_value)) +
  theme_bw() +
  labs(colour = "Actinic PAR (nm)")

MIT9313deltaODLmuMatrixPlot <- NestedFitsData %>%
   filter(Strain == "MIT9313",
          deltaOD_Lmu_corr == 0 |
          deltaOD_Lmu_corr != 0 & deltaOD_LseGrowthFlag == 1) %>%
  ggplot() +
  geom_point(aes(x = Par_ue, y = deltaOD_Lmu_corr, colour = as.factor(WL)), size = 3, alpha = 0.4) +
 geom_line(aes(x = Par_ue, y = deltaOD_Lmu_corr, colour = as.factor(WL)), linewidth = 0.1, alpha = 0.4) +
 labs(subtitle = "Strain, [Oxygen],Photoperiod, Actinic PAR (nm) - Of cultures that grew, outliers filtered by standard error > 20% of growth rate.", y = "deltaOD [~Chl] " ~ "Logistic Growth Rate" ~ "("~h^-1~")" , x = "Growth Light (µmol photons" ~m^-2 ~s^-1~")") +
   geom_text(aes(x = Par_ue, y = deltaOD_Lmu_corr, label= Run), alpha = 0.5,hjust = 0.5, vjust=-0.5) +
  # geom_jitter(aes(x = Par_ue, y = deltaOD_Lmu_corr, colour = as.factor(WL)), size = 6) +
  geom_errorbar(aes(x = Par_ue, ymin = deltaOD_Lmu_corr - deltaOD_Lmu_se, ymax = deltaOD_Lmu_corr + deltaOD_Lmu_se), size = 0.3, color = 'black', data = . %>% filter(GrowthAmpdeltaODFlag_14days  == 1)) +
  scale_x_continuous(breaks=seq(0, 200, by = 50)) +
  coord_cartesian(xlim = c (0, 200)) +
  #   scale_y_continuous(breaks=seq(0, 0.07, by = 0.01)) +
  # coord_cartesian(ylim = c (0, 0.07)) +
  scale_colour_manual(values = MCMIXColours) +
  ggh4x::facet_nested(cols = vars(Strain, O2_uM, Photoperiod), rows = vars(WL), 
                    labeller = labeller(WL = label_both, O2_uM = label_both, Strain = label_both, Photoperiod = label_value)) +
  theme_bw() +
  labs(colour = "Actinic PAR (nm)")

MIT9313deltaODLmuMatrixPlot

NestedFitsData %>%
   filter(Strain == "MIT9313",
          deltaOD_LseGrowthFlag == 1) %>%
  ggplot() +
  geom_point(aes(x = Par_ue, y = deltaOD_Lmu_corr, colour = as.factor(WL)), size = 3, alpha = 0.4) +
 geom_line(aes(x = Par_ue, y = deltaOD_Lmu_corr, colour = as.factor(WL)), linewidth = 0.1, alpha = 0.4) +
 labs(subtitle = "Strain, [Oxygen], Photoperiod, Actinic PAR (nm) - Outliers filtered by standard error > 20% of growth rate", y = "deltaOD [~Chl] " ~ "Logistic Growth Rate" ~ "("~h^-1~")" , x = "Growth Light (µmol photons" ~m^-2 ~s^-1~")") +
 geom_text(aes(x = Par_ue, y = deltaOD_Lmu_corr, label= Run), alpha = 0.5,hjust = 0.5, vjust=-0.5) +
  geom_errorbar(aes(x = Par_ue, ymin = deltaOD_Lmu_corr - deltaOD_Lmu_se, ymax = deltaOD_Lmu_corr + deltaOD_Lmu_se), size = 0.3, color = 'black', data = . %>% filter(GrowthAmpdeltaODFlag_14days  == 1)) +
  scale_x_continuous(breaks=seq(0, 200, by = 50)) +
  coord_cartesian(xlim = c (0, 200)) +
  #   scale_y_continuous(breaks=seq(0, 0.07, by = 0.01)) +
  # coord_cartesian(ylim = c (0, 0.07)) +
  scale_colour_manual(values = MCMIXColours) +
  ggh4x::facet_nested(cols = vars(Strain, O2_uM, Photoperiod), rows = vars(WL), 
                    labeller = labeller(WL = label_both, O2_uM = label_both, Strain = label_both, Photoperiod = label_value)) +
  theme_bw() +
  labs(colour = "Actinic PAR (nm)")

 MED4deltaODLmuMatrixPlot <- NestedFitsData %>%
   filter(Strain == "MED4",
          deltaOD_Lmu_corr == 0 |
          deltaOD_Lmu_corr != 0 & deltaOD_LseGrowthFlag == 1) %>%
  ggplot() +
  geom_point(aes(x = Par_ue, y = deltaOD_Lmu_corr, colour = as.factor(WL)), size = 3, alpha = 0.4) +
 geom_line(aes(x = Par_ue, y = deltaOD_Lmu_corr, colour = as.factor(WL)), linewidth = 0.1, alpha = 0.4) +
   labs(subtitle = "Strain, [Oxygen],Photoperiod, Actinic PAR (nm) - Of cultures that grew, outliers filtered by standard error > 20% of growth rate.", y = "deltaOD [~Chl] " ~ "Logistic Growth Rate" ~ "("~h^-1~")" , x = "Growth Light (µmol photons" ~m^-2 ~s^-1~")") +
 geom_text(aes(x = Par_ue, y = deltaOD_Lmu_corr, label= Run), alpha = 0.5,hjust = 0.5, vjust=-0.5) +
  geom_errorbar(aes(x = Par_ue, ymin = deltaOD_Lmu_corr - deltaOD_Lmu_se, ymax = deltaOD_Lmu_corr + deltaOD_Lmu_se), size = 0.3, color = 'black', data = . %>% filter(GrowthAmpdeltaODFlag_14days  == 1)) +
  scale_x_continuous(breaks=seq(0, 200, by = 50)) +
  coord_cartesian(xlim = c (0, 200)) +
  #   scale_y_continuous(breaks=seq(0, 0.115, by = 0.01)) +
  # coord_cartesian(ylim = c (0, 0.115)) +
  scale_colour_manual(values = MCMIXColours) +
 ggh4x::facet_nested(cols = vars(Strain, O2_uM, Photoperiod), rows = vars(WL), 
                    labeller = labeller(WL = label_both, O2_uM = label_both, Strain = label_both, Photoperiod = label_value)) +
  theme_bw() +
  labs(colour = "Actinic PAR (nm)")

MED4deltaODLmuMatrixPlot

 NestedFitsData %>%
   filter(Strain == "MED4",
          deltaOD_LseGrowthFlag == 1) %>%
  ggplot() +
  geom_point(aes(x = Par_ue, y = deltaOD_Lmu_corr, colour = as.factor(WL)), size = 3, alpha = 0.4) +
 geom_line(aes(x = Par_ue, y = deltaOD_Lmu_corr, colour = as.factor(WL)), linewidth = 0.1, alpha = 0.4) +
   labs(subtitle = "Strain, [Oxygen], Photoperiod, Actinic PAR (nm) - Outliers filtered by standard error > 20% of growth rate", y = "deltaOD [~Chl] " ~ "Logistic Growth Rate" ~ "("~h^-1~")" , x = "Growth Light (µmol photons" ~m^-2 ~s^-1~")") +
 geom_text(aes(x = Par_ue, y = deltaOD_Lmu_corr, label= Run), alpha = 0.5,hjust = 0.5, vjust=-0.5) +
  geom_errorbar(aes(x = Par_ue, ymin = deltaOD_Lmu_corr - deltaOD_Lmu_se, ymax = deltaOD_Lmu_corr + deltaOD_Lmu_se), size = 0.3, color = 'black', data = . %>% filter(GrowthAmpdeltaODFlag_14days  == 1)) +
  scale_x_continuous(breaks=seq(0, 200, by = 50)) +
  coord_cartesian(xlim = c (0, 200)) +
  #   scale_y_continuous(breaks=seq(0, 0.115, by = 0.01)) +
  # coord_cartesian(ylim = c (0, 0.115)) +
  scale_colour_manual(values = MCMIXColours) +
 ggh4x::facet_nested(cols = vars(Strain, O2_uM, Photoperiod), rows = vars(WL), 
                    labeller = labeller(WL = label_both, O2_uM = label_both, Strain = label_both, Photoperiod = label_value)) +
  theme_bw() +
  labs(colour = "Actinic PAR (nm)")

deltaOD_LmuSummaryGrowthPlot <- NestedFitsData %>%
   filter(Strain == "MIT9313" |
         Strain == "SS120" |
         Strain == "MED4",
         deltaOD_LseGrowthFlag == 1) %>%
  ggplot() +
  geom_point(aes(x = Par_ue, y = deltaOD_Lmu_corr, colour = as.factor(WL)), size = 3, alpha = 0.5) +
  geom_line(aes(x = Par_ue, y = deltaOD_Lmu_corr, colour = as.factor(WL)), linewidth = 0.1, alpha = 0.4) +
labs(subtitle = "Strain, [Oxygen], Photoperiod, Actinic PAR (nm)", y = "deltaOD [~Chl] " ~ "Logistic Growth Rate" ~ "("~h^-1~")" , x = "Growth Light (µmol photons" ~m^-2 ~s^-1~")") +
 geom_text(aes(x = Par_ue, y = deltaOD_Lmu_corr, label= Run),alpha = 0.5 ,hjust = 0.5, vjust=-0.5) +
  geom_errorbar(aes(x = Par_ue, ymin = deltaOD_Lmu_corr - deltaOD_Lmu_se, ymax = deltaOD_Lmu_corr + deltaOD_Lmu_se), size = 0.3, color = 'black', data = . %>% filter(GrowthAmpdeltaODFlag_14days  == 1)) +
  scale_x_continuous(breaks=seq(0, 200, by = 50)) +
  coord_cartesian(xlim = c (0, 200)) +
 #   scale_y_continuous(breaks=seq(-0.03, 0.07, by = 0.01)) +
  # coord_cartesian(ylim = c (-0.03, 0.07)) +
  scale_colour_manual(values = MCMIXColours) +
  ggh4x::facet_nested(cols = vars(Strain, O2_uM, Photoperiod), rows = vars(WL), 
                    labeller = labeller(WL = label_both, O2_uM = label_both, Strain = label_both, Photoperiod = label_value)) +
  theme_bw() +
  labs(colour = "Actinic PAR (nm)")

deltaOD_LmuSummaryGrowthPlot

#plotly::ggplotly(deltaOD_LmuSummaryGrowthPlot)

 NestedFitsData %>%
   filter(WL == "WW") %>%
  ggplot() +
  geom_point(aes(x = Par_ue, y = deltaOD_Lmu_corr, colour = as.factor(O2_uM)), size = 4, alpha = 0.5) +
 geom_line(aes(x = Par_ue, y = deltaOD_Lmu_corr, colour = as.factor(O2_uM)), linewidth = 0.1, alpha = 0.4) +
   labs(subtitle = "Actinic PAR (nm), Strain, Photoperiod", y = "deltaOD [~Chl] " ~ "Logistic Growth Rate" ~ "("~h^-1~")" , x = "Growth Light (µmol photons" ~m^-2 ~s^-1~")") +
 geom_text(aes(x = Par_ue, y = deltaOD_Lmu_corr, label= Run), alpha = 0.5 ,hjust = 0.5, vjust=-0.5) +
  geom_errorbar(aes(x = Par_ue, ymin = (deltaOD_Lmu_corr - deltaOD_Lmu_se), ymax = (deltaOD_Lmu_corr + deltaOD_Lmu_se)), size = 0.3, color = 'black', data = . %>% filter(GrowthAmpdeltaODFlag_14days  == 1)) +
  scale_x_continuous(breaks=seq(0, 200, by = 50)) +
  coord_cartesian(xlim = c (0, 200)) +
  #   scale_y_continuous(breaks=seq(-0.03, 0.07, by = 0.01)) +
  # coord_cartesian(ylim = c (-0.03, 0.07)) +
  #scale_colour_manual(values = MCMIXColours) +
  ggh4x::facet_nested(cols = vars(WL, Strain), rows = vars(Photoperiod), 
                    labeller = labeller(WL = label_both, O2_uM = label_both, Strain = label_both, Photoperiod = label_value)) +
  theme_bw() +
  labs(colour = "Oxygen Concentration")

NestedFitsData %>%
   filter(WL == "450") %>%
  ggplot() +
  geom_point(aes(x = Par_ue, y = deltaOD_Lmu_corr, colour = as.factor(O2_uM)), size = 4, alpha = 0.5) +
 geom_line(aes(x = Par_ue, y = deltaOD_Lmu_corr, colour = as.factor(O2_uM)), linewidth = 0.1, alpha = 0.4) +
   labs(subtitle = "Actinic PAR (nm), Strain, Photoperiod", y = "deltaOD [~Chl] " ~ "Logistic Growth Rate" ~ "("~h^-1~")" , x = "Growth Light (µmol photons" ~m^-2 ~s^-1~")") +
 geom_text(aes(x = Par_ue, y = deltaOD_Lmu_corr, label= Run), alpha = 0.5 ,hjust = 0.5, vjust=-0.5) +
  geom_errorbar(aes(x = Par_ue, ymin = (deltaOD_Lmu_corr - deltaOD_Lmu_se), ymax = (deltaOD_Lmu_corr + deltaOD_Lmu_se)), size = 0.3, color = 'black', data = . %>% filter(GrowthAmpdeltaODFlag_14days  == 1)) +
  scale_x_continuous(breaks=seq(0, 200, by = 50)) +
  coord_cartesian(xlim = c (0, 200)) +
  #   scale_y_continuous(breaks=seq(-0.03, 0.07, by = 0.01)) +
  # coord_cartesian(ylim = c (-0.03, 0.07)) +
  #scale_colour_manual(values = MCMIXColours) +
  ggh4x::facet_nested(cols = vars(WL, Strain), rows = vars(Photoperiod), 
                    labeller = labeller(WL = label_both, O2_uM = label_both, Strain = label_both, Photoperiod = label_value)) +
  theme_bw() +
  labs(colour = "Oxygen Concentration")

NestedFitsData %>%
   filter(WL == "660") %>%
  ggplot() +
  geom_point(aes(x = Par_ue, y = deltaOD_Lmu_corr, colour = as.factor(O2_uM)), size = 4, alpha = 0.5) +
 geom_line(aes(x = Par_ue, y = deltaOD_Lmu_corr, colour = as.factor(O2_uM)), linewidth = 0.1, alpha = 0.4) +
   labs(subtitle = "Actinic PAR (nm), Strain, Photoperiod", y = "deltaOD [~Chl] " ~ "Logistic Growth Rate" ~ "("~h^-1~")" , x = "Growth Light (µmol photons" ~m^-2 ~s^-1~")") +
 geom_text(aes(x = Par_ue, y = deltaOD_Lmu_corr, label= Run), alpha = 0.5 ,hjust = 0.5, vjust=-0.5) +
  geom_errorbar(aes(x = Par_ue, ymin = (deltaOD_Lmu_corr - deltaOD_Lmu_se), ymax = (deltaOD_Lmu_corr + deltaOD_Lmu_se)), size = 0.3, color = 'black', data = . %>% filter(GrowthAmpdeltaODFlag_14days  == 1)) +
  scale_x_continuous(breaks=seq(0, 200, by = 50)) +
  coord_cartesian(xlim = c (0, 200)) +
  #   scale_y_continuous(breaks=seq(-0.03, 0.07, by = 0.01)) +
  # coord_cartesian(ylim = c (-0.03, 0.07)) +
  #scale_colour_manual(values = MCMIXColours) +
  ggh4x::facet_nested(cols = vars(WL, Strain), rows = vars(Photoperiod), 
                    labeller = labeller(WL = label_both, O2_uM = label_both, Strain = label_both, Photoperiod = label_value)) +
  theme_bw() +
  labs(colour = "Oxygen Concentration")

```

```{r save summary growth plots}

# ggsave(file = file.path(PlotsPath, paste("MIT9313deltaODLmuMatrixPlot",".png",sep = "")), plot = MIT9313deltaODLmuMatrixPlot, device = NULL, scale = 1, height=15, width= 25, units = c("cm"),dpi = 300, limitsize = TRUE)
# 
# ggsave(file = file.path(PlotsPath, paste("SS120deltaODLmuMatrixPlot",".png",sep = "")), plot = SS120deltaODLmuMatrixPlot, device = NULL, scale = 1, height=15, width= 25, units = c("cm"),dpi = 300, limitsize = TRUE)
# 
# ggsave(file = file.path(PlotsPath, paste("MED4deltaODLmuMatrixPlot",".png",sep = "")), plot = MED4deltaODLmuMatrixPlot, device = NULL, scale = 1, height=15, width= 25, units = c("cm"),dpi = 300, limitsize = TRUE)

```

Compare truncated runs
```{r CompareTruncated, fig.height= 15, fig.width= 15}

 NestedFitsData %>%
  filter(Strain == "MIT9313" |
         Strain == "SS120" |
         Strain == "MED4",
           WL == 450 |
           WL == 530 |
           WL == 660 |
           WL == "WW",
          Run >= 25 & 
           Run < 80, ) %>%
  ggplot(aes(x = OD720_Lmu_corr, y = OD720trunc_Lmu_corr, colour = WL)) +
  theme(panel.spacing.x = unit(10, "cm"), panel.spacing.y = unit(2, "cm")) +
  geom_point(aes(labels = c(Run))) +
  geom_errorbar(aes(xmin = OD720_Lmu_corr - OD720_Lmu_se, xmax = OD720_Lmu_corr + OD720_Lmu_se, ymin = OD720trunc_Lmu_corr - OD720trunc_Lmu_se, ymax = OD720trunc_Lmu_corr + OD720trunc_Lmu_se), size = 0.05, color = 'black', data = . %>% filter(GrowthAmpOD720Flag_14days == 1)) +
  scale_colour_manual(values = MCMIXColours) +
  geom_text(aes(x = OD720_Lmu_corr, y = OD720trunc_Lmu_corr, label= Run),hjust=0, vjust=0) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_vline(xintercept = 0, linetype = "dashed") +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed") +
  coord_fixed(ratio = 1, xlim = c (0, 0.05), ylim = c(0, 0.05) ) +
 # coord_cartesian() +
  scale_x_continuous(breaks=seq(0, 0.05, by = 0.01)) +
  labs(subtitle = "Strain, O2, Actinic PAR (nm)  All Growth Data From Runs >= #25", y = "Corrected Truncated Logistic Growth Rate (OD720trunc_Lmu)", x = "Corrected Logistic Growth Rate (OD720_Lmu)") +
   ggh4x::facet_nested(cols = vars(O2, WL), rows = vars(Strain), 
                    labeller = labeller(WL = label_both, Strain = label_value, O2 = label_both)) +
  theme_bw() +
  labs(colour = "Actinic PAR (nm)")

NestedFitsData %>%
  filter(Strain == "MIT9313" |
         Strain == "SS120" |
         Strain == "MED4",
           WL == 450 |
           WL == 530 |
           WL == 660 |
           WL == "WW",
          Run >= 25 & 
           Run < 80, ) %>%
  ggplot(aes(x = deltaOD_Lmu_corr, y = deltaODtrunc_Lmu_corr, colour = WL)) +
  theme(panel.spacing.x = unit(10, "cm"), panel.spacing.y = unit(2, "cm")) +
  geom_point(aes(labels = c(Run))) +
   geom_errorbar(aes(xmin = deltaOD_Lmu_corr - deltaOD_Lmu_se, xmax = deltaOD_Lmu_corr + deltaOD_Lmu_se, ymin = deltaODtrunc_Lmu_corr - deltaODtrunc_Lmu_se, ymax = deltaODtrunc_Lmu_corr + deltaODtrunc_Lmu_se), size = 0.3, color = 'black', data = . %>% filter(GrowthAmpdeltaODFlag_14days == 1)) +
  scale_colour_manual(values = MCMIXColours) +
  geom_text(aes(x = deltaOD_Lmu_corr, y = deltaODtrunc_Lmu_corr, label= Run),hjust=0, vjust=0) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_vline(xintercept = 0, linetype = "dashed") +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed") +
  coord_fixed(ratio = 1, xlim = c (0, 0.05), ylim = c(0, 0.05) ) +
 # coord_cartesian() +
  scale_x_continuous(breaks=seq(0, 0.05, by = 0.01)) +
  labs(subtitle = "Strain, O2, Actinic PAR (nm) All Growth Data From Runs >= #25", y = "Corrected Truncated Logistic Growth Rate (deltaODtrunc_Lmu)", x = "Corrected Logistic Growth Rate (deltaOD_Lmu)") +
   ggh4x::facet_nested(cols = vars(O2, WL), rows = vars(Strain), 
                    labeller = labeller(WL = label_both, Strain = label_value, O2 = label_both)) +
  theme_bw() +
  labs(colour = "Actinic PAR (nm)")

NestedFitsData %>%
  filter(Strain == "MIT9313" |
         Strain == "SS120" |
         Strain == "MED4",
           WL == 450 |
           WL == 530 |
           WL == 660 |
           WL == "WW",
          Run >= 25 & 
           Run < 80, 
          OD720trunc_Lmu_corr == 0 |
          OD720trunc_Lmu_corr != 0 & OD720trunc_LseGrowthFlag == 1,
          OD720_Lmu_corr == 0 |
          OD720_Lmu_corr != 0 & OD720_LseGrowthFlag == 1 ) %>%
  ggplot(aes(x = OD720_Lmu_corr, y = OD720trunc_Lmu_corr, colour = WL)) +
  theme(panel.spacing.x = unit(10, "cm"), panel.spacing.y = unit(2, "cm")) +
  geom_point(aes(labels = c(Run))) +
  geom_errorbar(aes(xmin = OD720_Lmu_corr - OD720_Lmu_se, xmax = OD720_Lmu_corr + OD720_Lmu_se, ymin = OD720trunc_Lmu_corr - OD720trunc_Lmu_se, ymax = OD720trunc_Lmu_corr + OD720trunc_Lmu_se), size = 0.3, color = 'black', data = . %>% filter(GrowthAmpOD720Flag_14days == 1)) +
  scale_colour_manual(values = MCMIXColours) +
  geom_text(aes(x = OD720_Lmu_corr, y = OD720trunc_Lmu_corr, label= Run),hjust=0, vjust=0) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_vline(xintercept = 0, linetype = "dashed") +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed") +
  coord_fixed(ratio = 1, xlim = c (0, 0.05), ylim = c(0, 0.05) ) +
 # coord_cartesian() +
  scale_x_continuous(breaks=seq(0, 0.05, by = 0.01)) +
  labs(subtitle = "Strain, O2, Actinic PAR (nm) Standard Error Filtered Showing Runs >= #25", y = "Corrected Truncated Logistic Growth Rate  (OD720trunc_Lmu)", x = "Corrected Logistic Growth Rate (OD720_Lmu)") +
   ggh4x::facet_nested(cols = vars(O2, WL), rows = vars(Strain), 
                    labeller = labeller(WL = label_both, Strain = label_value, O2 = label_both)) +
  theme_bw() +
  labs(colour = "Actinic PAR (nm)")

NestedFitsData %>%
  filter(Strain == "MIT9313" |
         Strain == "SS120" |
         Strain == "MED4",
           WL == 450 |
           WL == 530 |
           WL == 660 |
           WL == "WW",
          Run >= 25 & 
           Run < 80, 
          deltaODtrunc_Lmu_corr == 0 |
          deltaODtrunc_Lmu_corr != 0 & deltaODtrunc_LseGrowthFlag == 1,
          deltaOD_Lmu_corr == 0 |
          deltaOD_Lmu_corr != 0 & deltaOD_LseGrowthFlag == 1 ) %>%
  ggplot(aes(x = deltaOD_Lmu_corr, y = deltaODtrunc_Lmu_corr, colour = WL)) +
  theme(panel.spacing.x = unit(10, "cm"), panel.spacing.y = unit(2, "cm")) +
  geom_point(aes(labels = c(Run))) +
  geom_errorbar(aes(xmin = deltaOD_Lmu_corr - deltaOD_Lmu_se, xmax = deltaOD_Lmu_corr + deltaOD_Lmu_se, ymin = deltaODtrunc_Lmu_corr - deltaODtrunc_Lmu_se, ymax = deltaODtrunc_Lmu_corr + deltaODtrunc_Lmu_se), size = 0.3, color = 'black', data = . %>% filter(GrowthAmpdeltaODFlag_14days == 1)) +
  scale_colour_manual(values = MCMIXColours) +
  geom_text(aes(x = deltaOD_Lmu_corr, y = deltaODtrunc_Lmu_corr, label= Run),hjust=0, vjust=0) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_vline(xintercept = 0, linetype = "dashed") +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed") +
  coord_fixed(ratio = 1, xlim = c (0, 0.05), ylim = c(0, 0.05) ) +
 # coord_cartesian() +
  scale_x_continuous(breaks=seq(0, 0.05, by = 0.01)) +
  labs(subtitle = "Strain, O2, Actinic PAR (nm) Standard Error Filtered Showing Runs >= #25", y = "Corrected Truncated Logistic Growth Rate  (deltaODtrunc_Lmu)", x = "Corrected Logistic Growth Rate (deltaOD_Lmu)") +
   ggh4x::facet_nested(cols = vars(O2, WL), rows = vars(Strain), 
                    labeller = labeller(WL = label_both, Strain = label_value, O2 = label_both)) +
  theme_bw() +
  labs(colour = "Actinic PAR (nm)")

NestedFitsData %>%
  filter(Strain == "MIT9313" |
         Strain == "SS120" |
         Strain == "MED4",
           WL == 450 |
           WL == 530 |
           WL == 660 |
           WL == "WW",
          Run >= 25 & 
           Run < 80, ) %>%
  ggplot(aes(x = deltaOD_Lmu_corr, y = deltaODtrunc_Lmu_corr, colour = WL)) +
  theme(panel.spacing.x = unit(10, "cm"), panel.spacing.y = unit(2, "cm")) +
  geom_point(aes(labels = c(Run))) +
  geom_errorbar(aes(xmin = deltaOD_Lmu_corr - deltaOD_Lmu_se, xmax = deltaOD_Lmu_corr + deltaOD_Lmu_se, ymin = deltaODtrunc_Lmu_corr - deltaODtrunc_Lmu_se, ymax = deltaODtrunc_Lmu_corr + deltaODtrunc_Lmu_se), size = 0.3, color = 'black', data = . %>% filter(GrowthAmpdeltaODFlag_14days == 1)) +
  scale_colour_manual(values = MCMIXColours) +
  geom_text(aes(x = deltaOD_Lmu_corr, y = deltaODtrunc_Lmu_corr, label= Run),hjust=0, vjust=0) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_vline(xintercept = 0, linetype = "dashed") +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed") +
  coord_fixed(ratio = 1, xlim = c (0, 0.05), ylim = c(0, 0.05) ) +
 # coord_cartesian() +
  scale_x_continuous(breaks=seq(0, 0.05, by = 0.01)) +
  labs(subtitle = "Strain, O2, Actinic PAR (nm)  All Growth Data From Runs >= #25", y = "Corrected Truncated Logistic Growth Rate  (deltaODtrunc_Lmu)", x = "Corrected Logistic Growth Rate (deltaOD_Lmu)") +
  #facet_grid(rows = vars(Strain), cols = vars(O2, WL)) +
  theme_bw() +
  labs(colour = "Actinic PAR (nm)")

NestedFitsData %>%
  filter(Strain == "MIT9313" |
         Strain == "SS120" |
         Strain == "MED4",
           WL == 450 |
           WL == 530 |
           WL == 660 |
           WL == "WW",
          Run >= 25 & 
           Run < 80, 
          deltaODtrunc_Lmu_corr == 0 |
          deltaODtrunc_Lmu_corr != 0 & deltaODtrunc_LseGrowthFlag == 1,
          deltaOD_Lmu_corr == 0 |
          deltaOD_Lmu_corr != 0 & deltaOD_LseGrowthFlag == 1) %>%
  ggplot(aes(x = deltaOD_Lmu_corr, y = deltaODtrunc_Lmu_corr, colour = WL)) +
  theme(panel.spacing.x = unit(10, "cm"), panel.spacing.y = unit(2, "cm")) +
  geom_point(aes(labels = c(Run))) +
  geom_errorbar(aes(xmin = deltaOD_Lmu_corr - deltaOD_Lmu_se, xmax = deltaOD_Lmu_corr + deltaOD_Lmu_se, ymin = deltaODtrunc_Lmu_corr - deltaODtrunc_Lmu_se, ymax = deltaODtrunc_Lmu_corr + deltaODtrunc_Lmu_se), size = 0.3, color = 'black', data = . %>% filter(GrowthAmpdeltaODFlag_14days == 1)) +
  scale_colour_manual(values = MCMIXColours) +
  geom_text(aes(x = deltaOD_Lmu_corr, y = deltaODtrunc_Lmu_corr, label= Run),hjust=0, vjust=0) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_vline(xintercept = 0, linetype = "dashed") +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed") +
  coord_fixed(ratio = 1, xlim = c (0, 0.05), ylim = c(0, 0.05) ) +
 # coord_cartesian() +
  scale_x_continuous(breaks=seq(0, 0.05, by = 0.01)) +
  labs(subtitle = "Strain, O2, Actinic PAR (nm) Standard Error Filtered From Runs >= #25 ", y = "Corrected Truncated Logistic Growth Rate  (deltaODtrunc_Lmu)", x = "Corrected Logistic Growth Rate (deltaOD_Lmu)") +
  #facet_grid(rows = vars(Strain), cols = vars(O2, WL)) +
  theme_bw() +
  labs(colour = "Actinic PAR (nm)")

NestedFitsData %>%
  filter(Strain == "MIT9313" |
         Strain == "SS120" |
         Strain == "MED4",
           WL == 450 |
           WL == 530 |
           WL == 660 |
           WL == "WW",
          Run >= 25 & 
           Run < 80, 
          deltaODtrunc_Lmu_corr == 0 |
          deltaODtrunc_Lmu_corr != 0 & deltaODtrunc_LseGrowthFlag == 1,
          deltaOD_Lmu_corr == 0 |
          deltaOD_Lmu_corr != 0 & deltaOD_LseGrowthFlag == 1) %>%
  ggplot() +
  theme(panel.spacing.x = unit(10, "cm"), panel.spacing.y = unit(2, "cm")) +
  geom_point(aes(x = deltaOD_TimeHalfPmax_h, y = deltaODtrunc_TimetoMax, colour = WL), size = 3) +
  # geom_errorbar(aes(xmin = deltaOD_Lmu_corr - deltaOD_Lmu_se, xmax = deltaOD_Lmu_corr + deltaOD_Lmu_se, ymin = deltaODtrunc_Lmu_corr - deltaODtrunc_Lmu_se, ymax = deltaODtrunc_Lmu_corr + deltaODtrunc_Lmu_se), size = 0.3, color = 'black', data = . %>% filter(GrowthAmpdeltaODFlag_14days == 1)) +
  scale_colour_manual(values = MCMIXColours) +
  # geom_text(aes(x = deltaOD_Lmu_corr, y = deltaODtrunc_Lmu_corr, label= Run),hjust=0, vjust=0) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_vline(xintercept = 0, linetype = "dashed") +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed") +
 coord_fixed(ratio = 1, xlim = c (0, 600), ylim = c(0, 600) ) +
 # coord_cartesian() +
  # scale_x_continuous(breaks=seq(0, 0.05, by = 0.01)) +
  # labs(subtitle = "Strain, O2, Actinic PAR (nm) Standard Error Filtered From Runs >= #25 ", y = "deltaOD time to half pmax (Calculate from full growth trajectory)", x = "deltaOD truncated time to pmax") +
  #facet_grid(rows = vars(Strain), cols = vars(O2, WL)) +
  theme_bw() +
  labs(colour = "Actinic PAR (nm)")

```