### 2023-12-08 Code/MultiDataGrowthAssessPico.Rmd
Removed all redundant varibles

---
title: "MultiDataGrowthAssessPico"
author:
- Laurel Genge
- Carlie Barnhill
- Max Berthold
- Douglas A. Campbell
- Mireille Savoie
date: "`r format(Sys.Date())`"
output:
  bookdown::html_document2:
    code_folding: show
    keep_md: yes
    toc: TRUE
    toc_float: TRUE
    toc_depth: 6
    fig_caption: yes
bibliography: Prochlorococcus_O2_NPQ.bib
csl: plos-one.csl
editor_options: 
  markdown: 
    wrap: 72
---

## Set figure caption font size
```{css, echo=FALSE}
p.caption {
  font-size: 25px;
}
```

## Set Chunk Options
```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
knitr::opts_chunk$set(fig.path='Figs/')
```

# Introduction

# Materials and Methods

## Load Libraries and set project variables

```{r load libraries, message= TRUE}
# libraries; Note check actual dependencies

library(lubridate)
library(broom)
library(knitr)
library(zoo)
library(RColorBrewer)
library(grid)
library(photobiologyWavebands)
library(tidyverse)

```

```{r set project variables}
#"..", takes up a level in the directory path

Project <- "PICO"

DataOut <- file.path("..", "Data", "CleanData", "CleanedMCData")
DataIn <- file.path("..", "Data","ProcessedData", "ProcessedMCData")

GrowthFlagError <- 0.2     # percentage of mu that std error must be less than

OD720GrowthAmpFlagValue = 0.010 # OD720 absolute change for setting GrowthFlag = 1 and DeclineFlag = 1; did culture show  growth;

deltaODGrowthAmpFlagValue = 0.007 # deltaOD absolute change for setting GrowthFlag = 1 and DeclineFlag = 1; did culture show  growth;

```

```{r set colours}

MyWavelengths = c(405, 450, 470, 530, 620, 660, 730, "WW")
MCMIXColours = c(c(w_length2rgb(405), w_length2rgb(450), w_length2rgb(470), w_length2rgb(530),
                   w_length2rgb(620), w_length2rgb(660), w_length2rgb(730), "black"))

names(MCMIXColours) <- MyWavelengths
MCMIXColours

```

## List processed data

```{r previous ProcessedData}

ProcessedFiles <- list.files(path = DataIn, pattern = Project, full.names = TRUE)

ProcessedFiles

```

# Read .Rds, Summarize, Combine to single data frame

## Create function using to read in .Rds, adding source filename as a column

```{r readRDS_plus}
readRDS_plus <- function(Flnm){readRDS(file = Flnm) %>%
    mutate(Filename = Flnm)
}
```

## Read in ProcessedFiles .Rds

```{r read selected ProcessedData}

LogFitsData <-  ProcessedFiles %>%
    map_df(~readRDS_plus(Flnm = .))

glimpse(LogFitsData)

```

#Unnest data
# assign growth rate to zero if model returns mu < 0 or if growth amplitude flag = 0

```{r unnest fits}

# OD720 logistic fits
# assign 0 to growth fits that are NA's and for tubes with absolute amplitude change defined in project variables chunk.

OD720_unnest <- LogFitsData %>%
  select(c(Filename, Tube, ExpDate, MC, PrimaryOperator, Run,  SampleID, Strain, ExpCul, ExpStartTime, Par_ue, Photoperiod, Temp_c, O2, WL, LightShape, Optode, OptodeCh, OptodeMeasure, ExpEndDate, ExpEndHour, SetActinic_day, tubedata, minRollMeanOD720, maxRollMeanOD720, minRollMeandeltaOD, maxRollMeandeltaOD,  GrowthAmpOD720_14days,  GrowthAmpdeltaOD_14days, OD720_logistic_tidied, OD720_logistic_predict, O2_Category)
         ) %>%
  unnest(cols = c(OD720_logistic_tidied), names_sep = "_", keep_empty = TRUE) %>%
  pivot_wider(names_from =  OD720_logistic_tidied_term, values_from = c(OD720_logistic_tidied_estimate, OD720_logistic_tidied_std.error, OD720_logistic_tidied_statistic, OD720_logistic_tidied_p.value), names_sep = "_") %>%
  rename(OD720_Lmax = OD720_logistic_tidied_estimate_Lmax,
         OD720_Lmax_se = OD720_logistic_tidied_std.error_Lmax,
         OD720_Lmu_raw = OD720_logistic_tidied_estimate_Lmu,
         OD720_Lmu_se = OD720_logistic_tidied_std.error_Lmu,
         OD720_Lintercept = OD720_logistic_tidied_estimate_Lintercept,             
         OD720_Lintercept_se = OD720_logistic_tidied_std.error_Lintercept)  %>%
   mutate(OD720_Lmu_corr = if_else(!is.na(OD720_Lmu_raw), OD720_Lmu_raw, 0),
          OD720_Lmu_se = if_else(!is.na(OD720_Lmu_se), OD720_Lmu_se, 0),
          OD720_Lintercept = if_else(!is.na(OD720_Lintercept), OD720_Lintercept, 0),
          OD720_Lintercept_se = if_else(!is.na(OD720_Lintercept_se), OD720_Lintercept_se, 0),
          OD720_Lmax = if_else(!is.na(OD720_Lmax), OD720_Lmax, 0),
          OD720_Lmax_se = if_else(!is.na(OD720_Lmax_se), OD720_Lmax_se, 0),
          GrowthAmpOD720Flag_14days = if_else(GrowthAmpOD720_14days >= OD720GrowthAmpFlagValue, 1, 0),
          OD720_Lmu_corr = if_else(GrowthAmpOD720Flag_14days == 1 & OD720_Lmu_corr >= 0 , OD720_Lmu_corr, 0),
          OD720_LseGrowthFlag = if_else(OD720_Lmu_se < OD720_Lmu_corr * GrowthFlagError, 1, 0)
          )

# deltaOD logistic fits
# assign 0 to growth fits that are NA's and for tubes with absolute amplitude change defined in project variables chunk.

deltaOD_unnest <- LogFitsData %>%
   select(c(Filename,  Tube, ExpDate, MC, PrimaryOperator, Run,  SampleID, Strain, ExpCul, ExpStartTime, Par_ue, Photoperiod, Temp_c, O2, WL, LightShape,  Optode, OptodeCh, OptodeMeasure,  ExpEndDate, ExpEndHour, SetActinic_day, tubedata, minRollMeanOD720, maxRollMeanOD720, minRollMeandeltaOD, maxRollMeandeltaOD, GrowthAmpOD720_14days,  GrowthAmpdeltaOD_14days, deltaOD_logistic_tidied, deltaOD_logistic_predict, O2_Category)
          ) %>%
unnest(cols = c(deltaOD_logistic_tidied), names_sep = "_", keep_empty = TRUE) %>%
  pivot_wider(names_from =  deltaOD_logistic_tidied_term, values_from = c(deltaOD_logistic_tidied_estimate, deltaOD_logistic_tidied_std.error, deltaOD_logistic_tidied_statistic, deltaOD_logistic_tidied_p.value), names_sep = "_")  %>%
  rename(deltaOD_Lmax = deltaOD_logistic_tidied_estimate_Lmax,
         deltaOD_Lmax_se = deltaOD_logistic_tidied_std.error_Lmax,
         deltaOD_Lmu_raw = deltaOD_logistic_tidied_estimate_Lmu,
         deltaOD_Lmu_se = deltaOD_logistic_tidied_std.error_Lmu,
         deltaOD_Lintercept = deltaOD_logistic_tidied_estimate_Lintercept,
         deltaOD_Lintercept_se = deltaOD_logistic_tidied_std.error_Lintercept) %>%
   mutate(deltaOD_Lmu_corr = if_else(!is.na(deltaOD_Lmu_raw), deltaOD_Lmu_raw, 0),
          deltaOD_Lmu_se = if_else(!is.na(deltaOD_Lmu_se), deltaOD_Lmu_se, 0),
          deltaOD_Lintercept = if_else(!is.na(deltaOD_Lintercept), deltaOD_Lintercept, 0),
          deltaOD_Lintercept_se = if_else(!is.na(deltaOD_Lintercept_se), deltaOD_Lintercept_se, 0),
          deltaOD_Lmax = if_else(!is.na(deltaOD_Lmax), deltaOD_Lmax, 0),
          deltaOD_Lmax_se = if_else(!is.na(deltaOD_Lmax_se), deltaOD_Lmax_se, 0),
          GrowthAmpdeltaODFlag_14days = if_else(GrowthAmpdeltaOD_14days >= deltaODGrowthAmpFlagValue, 1, 0),
          deltaOD_Lmu_corr = if_else(GrowthAmpdeltaODFlag_14days == 1 & deltaOD_Lmu_corr >= 0, deltaOD_Lmu_corr, 0),
         deltaOD_LseGrowthFlag = if_else(deltaOD_Lmu_se < deltaOD_Lmu_corr * GrowthFlagError, 1, 0)
          )

# OD720 truncated logistic fits
# assign 0 to growth fits that are NA's and for tubes with absolute amplitude change defined in project variables chunk.

OD720truncLog_unnest <- LogFitsData %>%
   select(c(Filename, Tube, ExpDate, MC, PrimaryOperator, Run,  SampleID, Strain, ExpCul, ExpStartTime, Par_ue, Photoperiod, Temp_c, O2, WL, LightShape,  Optode, OptodeCh, OptodeMeasure,  ExpEndDate, ExpEndHour, SetActinic_day, tubedata, minRollMeanOD720, maxRollMeanOD720, minRollMeandeltaOD, maxRollMeandeltaOD, GrowthAmpOD720_14days,  GrowthAmpdeltaOD_14days,OD720trunc_logistic_tidied, OD720trunc_logistic_predict, O2_Category)
          ) %>%
  unnest(cols = c(OD720trunc_logistic_tidied), names_sep = "_", keep_empty = TRUE) %>%
  pivot_wider(names_from =  OD720trunc_logistic_tidied_term, values_from = c(OD720trunc_logistic_tidied_estimate, OD720trunc_logistic_tidied_std.error, OD720trunc_logistic_tidied_statistic, OD720trunc_logistic_tidied_p.value), names_sep = "_") %>%
  rename(OD720trunc_Lmax = OD720trunc_logistic_tidied_estimate_Lmax,
         OD720trunc_Lmax_se = OD720trunc_logistic_tidied_std.error_Lmax,
         OD720trunc_Lmu_raw = OD720trunc_logistic_tidied_estimate_Lmu,
         OD720trunc_Lmu_se = OD720trunc_logistic_tidied_std.error_Lmu,
         OD720trunc_Lintercept = OD720trunc_logistic_tidied_estimate_Lintercept,             
         OD720trunc_Lintercept_se = OD720trunc_logistic_tidied_std.error_Lintercept)  %>%
   mutate(OD720trunc_Lmu_corr = if_else(!is.na(OD720trunc_Lmu_raw), OD720trunc_Lmu_raw, 0),
          OD720trunc_Lmu_se = if_else(!is.na(OD720trunc_Lmu_se), OD720trunc_Lmu_se, 0),
          OD720trunc_Lintercept = if_else(!is.na(OD720trunc_Lintercept), OD720trunc_Lintercept, 0),
          OD720trunc_Lintercept_se = if_else(!is.na(OD720trunc_Lintercept_se), OD720trunc_Lintercept_se, 0),
          OD720trunc_Lmax = if_else(!is.na(OD720trunc_Lmax), OD720trunc_Lmax, 0),
          OD720trunc_Lmax_se = if_else(!is.na(OD720trunc_Lmax_se), OD720trunc_Lmax_se, 0),
          GrowthAmpOD720Flag_14days = if_else(GrowthAmpOD720_14days >= OD720GrowthAmpFlagValue, 1, 0),
          OD720trunc_Lmu_corr = if_else(GrowthAmpOD720Flag_14days == 1 & OD720trunc_Lmu_corr >= 0 , OD720trunc_Lmu_corr, 0),
          OD720trunc_LseGrowthFlag = if_else(OD720trunc_Lmu_se < OD720trunc_Lmu_corr * GrowthFlagError, 1, 0)
          )

# deltaOD truncated logistic fits
# assign 0 to growth fits that are NA's and for tubes with absolute amplitude change defined in project variables chunk.

deltaODtruncLog_unnest <- LogFitsData %>%
   select(c(Filename, Tube, ExpDate, MC, PrimaryOperator, Run,  SampleID, Strain, ExpCul, ExpStartTime, Par_ue, Photoperiod, Temp_c, O2, WL, LightShape,  Optode, OptodeCh, OptodeMeasure,  ExpEndDate, ExpEndHour, SetActinic_day, tubedata, minRollMeanOD720, maxRollMeanOD720, minRollMeandeltaOD, maxRollMeandeltaOD, GrowthAmpOD720_14days,  GrowthAmpdeltaOD_14days, deltaODtrunc_logistic_tidied, deltaODtrunc_logistic_predict, O2_Category)
          ) %>%
unnest(cols = c(deltaODtrunc_logistic_tidied), names_sep = "_", keep_empty = TRUE) %>%
  pivot_wider(names_from =  deltaODtrunc_logistic_tidied_term, values_from = c(deltaODtrunc_logistic_tidied_estimate, deltaODtrunc_logistic_tidied_std.error, deltaODtrunc_logistic_tidied_statistic, deltaODtrunc_logistic_tidied_p.value), names_sep = "_")  %>%
  rename(deltaODtrunc_Lmax = deltaODtrunc_logistic_tidied_estimate_Lmax,
         deltaODtrunc_Lmax_se = deltaODtrunc_logistic_tidied_std.error_Lmax,
         deltaODtrunc_Lmu_raw = deltaODtrunc_logistic_tidied_estimate_Lmu,
         deltaODtrunc_Lmu_se = deltaODtrunc_logistic_tidied_std.error_Lmu,
         deltaODtrunc_Lintercept = deltaODtrunc_logistic_tidied_estimate_Lintercept,
         deltaODtrunc_Lintercept_se = deltaODtrunc_logistic_tidied_std.error_Lintercept) %>%
   mutate(deltaODtrunc_Lmu_corr = if_else(!is.na(deltaODtrunc_Lmu_raw), deltaODtrunc_Lmu_raw, 0),
          deltaODtrunc_Lmu_se = if_else(!is.na(deltaODtrunc_Lmu_se), deltaODtrunc_Lmu_se, 0),
          deltaODtrunc_Lintercept = if_else(!is.na(deltaODtrunc_Lintercept), deltaODtrunc_Lintercept, 0),
          deltaODtrunc_Lintercept_se = if_else(!is.na(deltaODtrunc_Lintercept_se), deltaODtrunc_Lintercept_se, 0),
          deltaODtrunc_Lmax = if_else(!is.na(deltaODtrunc_Lmax), deltaODtrunc_Lmax, 0),
          deltaODtrunc_Lmax_se = if_else(!is.na(deltaODtrunc_Lmax_se), deltaODtrunc_Lmax_se, 0),
          GrowthAmpdeltaODFlag_14days = if_else(GrowthAmpdeltaOD_14days >= deltaODGrowthAmpFlagValue, 1, 0),
          deltaODtrunc_Lmu_corr = if_else(GrowthAmpdeltaODFlag_14days == 1 & deltaODtrunc_Lmu_corr >= 0, deltaODtrunc_Lmu_corr, 0),
          deltaODtrunc_LseGrowthFlag = if_else(deltaODtrunc_Lmu_se < deltaODtrunc_Lmu_corr * GrowthFlagError, 1, 0)
          )

```

```{r combine unnested fits for PICO Shiny App}

# OD720 logistic fits
OD720_unnest <- OD720_unnest %>%
select(-c(OD720_logistic_tidied_estimate_NA,
          OD720_logistic_tidied_p.value_NA,
          OD720_logistic_tidied_statistic_NA,
          OD720_logistic_tidied_std.error_NA,
          OD720_logistic_tidied_statistic_Lintercept,
          OD720_logistic_tidied_statistic_Lmax,
          OD720_logistic_tidied_statistic_Lmu,
          OD720_logistic_tidied_p.value_Lintercept,
          OD720_logistic_tidied_p.value_Lmax,
          OD720_logistic_tidied_p.value_Lmu)
       )

# deltaOD logistic fits
deltaOD_unnest <- deltaOD_unnest %>%
  select(-c(deltaOD_logistic_tidied_estimate_NA,
            deltaOD_logistic_tidied_p.value_NA,
            deltaOD_logistic_tidied_statistic_NA,
            deltaOD_logistic_tidied_std.error_NA,
            deltaOD_logistic_tidied_statistic_Lintercept,
            deltaOD_logistic_tidied_statistic_Lmax,
            deltaOD_logistic_tidied_statistic_Lmu,
            deltaOD_logistic_tidied_p.value_Lintercept,
            deltaOD_logistic_tidied_p.value_Lmax,
            deltaOD_logistic_tidied_p.value_Lmu)
         )

# OD720 truncated logistic fits
OD720truncLog_unnest <- OD720truncLog_unnest %>%
select(-c(OD720trunc_logistic_tidied_estimate_NA,
           OD720trunc_logistic_tidied_p.value_NA,
           OD720trunc_logistic_tidied_statistic_NA,
           OD720trunc_logistic_tidied_std.error_NA,
            OD720trunc_logistic_tidied_statistic_Lintercept,
            OD720trunc_logistic_tidied_statistic_Lmax,
            OD720trunc_logistic_tidied_statistic_Lmu,
            OD720trunc_logistic_tidied_p.value_Lintercept,
            OD720trunc_logistic_tidied_p.value_Lmax,
            OD720trunc_logistic_tidied_p.value_Lmu)
       )

# deltaOD truncated logistic fits
deltaODtruncLog_unnest <- deltaODtruncLog_unnest %>%
  select(-c(deltaODtrunc_logistic_tidied_estimate_NA,
            deltaODtrunc_logistic_tidied_p.value_NA,
            deltaODtrunc_logistic_tidied_statistic_NA,
            deltaODtrunc_logistic_tidied_std.error_NA,
            deltaODtrunc_logistic_tidied_statistic_Lintercept,
            deltaODtrunc_logistic_tidied_statistic_Lmax,
            deltaODtrunc_logistic_tidied_statistic_Lmu,
            deltaODtrunc_logistic_tidied_p.value_Lintercept,
            deltaODtrunc_logistic_tidied_p.value_Lmax,
            deltaODtrunc_logistic_tidied_p.value_Lmu)
         )

# Join Full Growth Rate Data

UnnestedLogData <- left_join(x = OD720_unnest, y = deltaOD_unnest)

# Join Truncated Growth Rate Data
UnnestedTruncLogData <- left_join(x = OD720truncLog_unnest, y =  deltaODtruncLog_unnest)

#combine Full and Truncated unnested data
NestedFitsData <- left_join(x = UnnestedLogData, y = UnnestedTruncLogData)

#remove redundant objects
remove(OD720_unnest,
       deltaOD_unnest,
       UnnestedLogData,
       OD720truncLog_unnest,
       deltaODtruncLog_unnest,
       UnnestedTruncLogData,
       UnnestedFullData
       )

```

<!-- Chunk to convert categorize oxygen levels based on input in metacatalog. Have to do this here as this column was added after many runs were already processed and the merge with metacatalog already occured. Run numbers >= 36 and <= 91 and 0% oxygen are labelled 'Low', runs <= 35  and 0% oxygen are labelled 'Intermediate'.  All runs >= 92 are from the metacatelog. All runs with 21% oxygen are labelled 'High'.  Assign [O2] to O2_uM based on category -->

```{r}

NestedFitsData <- NestedFitsData %>% 
  mutate(O2_uM = case_when(O2_Category == "Low" ~ 2.5,
                            O2_Category == "Intermediate" ~ 25,
                            O2_Category == "High" ~ 250))

```

<!-- Chunk to calculate the rate per hour of actinic light rise to peak. -->
```{r}
NestedFitsData <- NestedFitsData %>% 
  mutate(RateActinic_h = case_when(LightShape == "Sine"  ~ Par_ue/(Photoperiod/2)))

glimpse(NestedFitsData)

```

```{r, pivot Nested data longer for GAM}

# OD720 pivot longer Gmax, Gmu, Glag, Lmu, Lmax and Lintercept estimates
Pivot_Longer_OD720_est <- NestedFitsData %>%
  select(c(Filename, Tube, SampleID, Run, PrimaryOperator, Strain, Par_ue,RateActinic_h, Photoperiod, SetActinic_day, Temp_c, O2, O2_Category,O2_uM, WL, LightShape,  GrowthAmpOD720_14days, GrowthAmpOD720Flag_14days,  OD720_LseGrowthFlag, OD720trunc_LseGrowthFlag,  OD720_Lmu_corr, OD720_Lmax, OD720_Lintercept, OD720trunc_Lmu_corr, OD720trunc_Lmax, OD720trunc_Lintercept
            )) %>%
   pivot_longer(c(OD720_Lmu_corr, OD720_Lmax, OD720_Lintercept, OD720trunc_Lmu_corr, OD720trunc_Lmax, OD720trunc_Lintercept), names_to = c("Wavelength", "term"), values_to = c("estimate"), names_sep = '\\_')

# OD720 pivot longer Gmax, Gmu, Glag , Lmu, Lmax and Lintercept standard error
Pivot_Longer_OD720_se <- NestedFitsData %>%
  select(c(Filename, Tube, SampleID, Run, PrimaryOperator, Strain, Par_ue, RateActinic_h, Photoperiod, SetActinic_day, Temp_c, O2, O2_Category,O2_uM, WL, LightShape, GrowthAmpOD720_14days, GrowthAmpOD720Flag_14days,  OD720_LseGrowthFlag, OD720trunc_LseGrowthFlag,   OD720_Lmu_se, OD720_Lmax_se, OD720_Lintercept_se,OD720trunc_Lmu_se, OD720trunc_Lmax_se, OD720trunc_Lintercept_se
            )) %>%
   pivot_longer(c(OD720_Lmu_se, OD720_Lmax_se, OD720_Lintercept_se, OD720trunc_Lmu_se, OD720trunc_Lmax_se, OD720trunc_Lintercept_se), names_to = c("Wavelength", "term"), values_to = c("se"), names_sep = '\\_')

# Join OD720 estimate and se DF 
Longer_OD720_Joined <- full_join(Pivot_Longer_OD720_est, Pivot_Longer_OD720_se)

remove(Pivot_Longer_OD720_est, Pivot_Longer_OD720_se)

# rename common columns for full join
Longer_OD720_Joined <- Longer_OD720_Joined %>%
  rename(GrowthAmp_14days = 'GrowthAmpOD720_14days',
         GrowthAmp_14daysFlag = 'GrowthAmpOD720Flag_14days',
         Lmu_seFlag = 'OD720_LseGrowthFlag',
         Lmutrunc_seFlag = 'OD720trunc_LseGrowthFlag'
         )

# deltaOD pivot longer Gmax, Gmu, Glag, Lmu, Lmax and Lintercept estimates
Pivot_Longer_deltaOD_est <- NestedFitsData %>%
  select(c(Filename, Tube, SampleID, Run, PrimaryOperator, Strain, Par_ue, RateActinic_h, Photoperiod, SetActinic_day, Temp_c, O2, O2_Category,O2_uM, WL, LightShape,  GrowthAmpdeltaOD_14days, GrowthAmpdeltaODFlag_14days,  deltaOD_LseGrowthFlag, deltaODtrunc_LseGrowthFlag,  deltaOD_Lmu_corr, deltaOD_Lmax, deltaOD_Lintercept, deltaODtrunc_Lmu_corr, deltaODtrunc_Lmax, deltaODtrunc_Lintercept
            )) %>%
   pivot_longer(c( deltaOD_Lmu_corr, deltaOD_Lmax, deltaOD_Lintercept, deltaODtrunc_Lmu_corr, deltaODtrunc_Lmax, deltaODtrunc_Lintercept), names_to = c("Wavelength", "term"), values_to = c("estimate"), names_sep = '\\_')

# deltaOD pivot longer Gmax, Gmu, Glag , Lmu, Lmax and Lintercept standard error
Pivot_Longer_deltaOD_se <- NestedFitsData %>%
  select(c(Filename, Tube, SampleID, Run, PrimaryOperator, Strain, Par_ue, RateActinic_h, Photoperiod, SetActinic_day, Temp_c, O2, O2_Category,O2_uM, WL, LightShape,  GrowthAmpdeltaOD_14days, GrowthAmpdeltaODFlag_14days,  deltaOD_LseGrowthFlag, deltaODtrunc_LseGrowthFlag,  deltaOD_Lmu_se, deltaOD_Lmax_se, deltaOD_Lintercept_se, deltaODtrunc_Lmu_se, deltaODtrunc_Lmax_se, deltaODtrunc_Lintercept_se
            )) %>%
   pivot_longer(c(deltaOD_Lmu_se, deltaOD_Lmax_se, deltaOD_Lintercept_se, deltaODtrunc_Lmu_se, deltaODtrunc_Lmax_se, deltaODtrunc_Lintercept_se), names_to = c("Wavelength", "term"), values_to = c("se"), names_sep = '\\_')

# Join deltaOD estimate and se DF 
Longer_deltaOD_Joined <- full_join(Pivot_Longer_deltaOD_est, Pivot_Longer_deltaOD_se)

remove(Pivot_Longer_deltaOD_est, Pivot_Longer_deltaOD_se)

# rename common columns for full join
Longer_deltaOD_Joined <- Longer_deltaOD_Joined %>%
  rename(GrowthAmp_14days = 'GrowthAmpdeltaOD_14days',
         GrowthAmp_14daysFlag = 'GrowthAmpdeltaODFlag_14days',
         Lmu_seFlag = 'deltaOD_LseGrowthFlag',
         Lmutrunc_seFlag = 'deltaODtrunc_LseGrowthFlag'
         )

#Join OD720 and deltaOD fits data
LongerFitsData <- rbind(Longer_OD720_Joined, Longer_deltaOD_Joined) %>%
  mutate_if(is.numeric, round, digits = 3)

remove(Longer_OD720_Joined, Longer_deltaOD_Joined)

glimpse(LongerFitsData)

```

Save Long DF for GAM and Nested for further analysis 
```{r save files}

saveRDS(object = LongerFitsData, file = file.path(DataOut, paste(Project, "LongerFitsData.Rds",  sep = "_")), ascii = FALSE, version = NULL, compress = "xz", refhook = NULL)

saveRDS(object = NestedFitsData, file = file.path(DataOut, paste(Project, "NestedFitsData.Rds",  sep = "_")), ascii = FALSE, version = NULL, compress = TRUE, refhook = NULL)

```

Plot All
```{r LogisticdeltaODPlots, fig.height= 10, fig.width= 10}

NestedFitsData %>%
   filter(Strain == "SS120") %>%
  ggplot() +
  geom_point(aes(x = Par_ue, y = deltaOD_Lmu_corr, colour = as.factor(WL)), size = 3, alpha = 0.5) +
 geom_line(aes(x = Par_ue, y = deltaOD_Lmu_corr, colour = as.factor(WL)), linewidth = 0.1, alpha = 0.4) +
   labs(subtitle = "Strain, [Oxygen], Photoperiod, Actinic PAR (nm)", y = "deltaOD [~Chl] " ~ "Logistic Growth Rate" ~ "("~h^-1~")" , x = "Growth Light (µmol photons" ~m^-2 ~s^-1~")") +
 geom_text(aes(x = Par_ue, y = deltaOD_Lmu_corr, label= Run), alpha = 0.5 ,hjust = 0.5, vjust=-0.5) +
  geom_errorbar(aes(x = Par_ue, ymin = (deltaOD_Lmu_corr - deltaOD_Lmu_se), ymax = (deltaOD_Lmu_corr + deltaOD_Lmu_se)), size = 0.3, color = 'black', data = . %>% filter(GrowthAmpdeltaODFlag_14days  == 1)) +
  scale_x_continuous(breaks=seq(0, 200, by = 50)) +
  coord_cartesian(xlim = c (0, 200)) +
  #   scale_y_continuous(breaks=seq(-0.03, 0.07, by = 0.01)) +
  # coord_cartesian(ylim = c (-0.03, 0.07)) +
  scale_colour_manual(values = MCMIXColours) +
  ggh4x::facet_nested(cols = vars(Strain, O2_uM, Photoperiod), rows = vars(WL), 
                    labeller = labeller(WL = label_both, O2_uM = label_both, Strain = label_both, Photoperiod = label_value)) +
  theme_bw() +
  labs(colour = "Actinic PAR (nm)")

NestedFitsData %>%
   filter(Strain == "MIT9313",
          deltaOD_Lmu_corr == 0 |
          deltaOD_Lmu_corr != 0 & deltaOD_LseGrowthFlag == 1) %>%
  ggplot() +
  geom_point(aes(x = Par_ue, y = deltaOD_Lmu_corr, colour = as.factor(WL)), size = 3, alpha = 0.4) +
 geom_line(aes(x = Par_ue, y = deltaOD_Lmu_corr, colour = as.factor(WL)), linewidth = 0.1, alpha = 0.4) +
 labs(subtitle = "Strain, [Oxygen],Photoperiod, Actinic PAR (nm) - Of cultures that grew, outliers filtered by standard error > 20% of growth rate.", y = "deltaOD [~Chl] " ~ "Logistic Growth Rate" ~ "("~h^-1~")" , x = "Growth Light (µmol photons" ~m^-2 ~s^-1~")") +
   geom_text(aes(x = Par_ue, y = deltaOD_Lmu_corr, label= Run), alpha = 0.5,hjust = 0.5, vjust=-0.5) +
  # geom_jitter(aes(x = Par_ue, y = deltaOD_Lmu_corr, colour = as.factor(WL)), size = 6) +
  geom_errorbar(aes(x = Par_ue, ymin = deltaOD_Lmu_corr - deltaOD_Lmu_se, ymax = deltaOD_Lmu_corr + deltaOD_Lmu_se), size = 0.3, color = 'black', data = . %>% filter(GrowthAmpdeltaODFlag_14days  == 1)) +
  scale_x_continuous(breaks=seq(0, 200, by = 50)) +
  coord_cartesian(xlim = c (0, 200)) +
  #   scale_y_continuous(breaks=seq(0, 0.07, by = 0.01)) +
  # coord_cartesian(ylim = c (0, 0.07)) +
  scale_colour_manual(values = MCMIXColours) +
  ggh4x::facet_nested(cols = vars(Strain, O2_uM, Photoperiod), rows = vars(WL), 
                    labeller = labeller(WL = label_both, O2_uM = label_both, Strain = label_both, Photoperiod = label_value)) +
  theme_bw() +
  labs(colour = "Actinic PAR (nm)")

 NestedFitsData %>%
   filter(Strain == "MED4",
          deltaOD_Lmu_corr == 0 |
          deltaOD_Lmu_corr != 0 & deltaOD_LseGrowthFlag == 1) %>%
  ggplot() +
  geom_point(aes(x = Par_ue, y = deltaOD_Lmu_corr, colour = as.factor(WL)), size = 3, alpha = 0.4) +
 geom_line(aes(x = Par_ue, y = deltaOD_Lmu_corr, colour = as.factor(WL)), linewidth = 0.1, alpha = 0.4) +
   labs(subtitle = "Strain, [Oxygen],Photoperiod, Actinic PAR (nm) - Of cultures that grew, outliers filtered by standard error > 20% of growth rate.", y = "deltaOD [~Chl] " ~ "Logistic Growth Rate" ~ "("~h^-1~")" , x = "Growth Light (µmol photons" ~m^-2 ~s^-1~")") +
 geom_text(aes(x = Par_ue, y = deltaOD_Lmu_corr, label= Run), alpha = 0.5,hjust = 0.5, vjust=-0.5) +
  geom_errorbar(aes(x = Par_ue, ymin = deltaOD_Lmu_corr - deltaOD_Lmu_se, ymax = deltaOD_Lmu_corr + deltaOD_Lmu_se), size = 0.3, color = 'black', data = . %>% filter(GrowthAmpdeltaODFlag_14days  == 1)) +
  scale_x_continuous(breaks=seq(0, 200, by = 50)) +
  coord_cartesian(xlim = c (0, 200)) +
  #   scale_y_continuous(breaks=seq(0, 0.115, by = 0.01)) +
  # coord_cartesian(ylim = c (0, 0.115)) +
  scale_colour_manual(values = MCMIXColours) +
 ggh4x::facet_nested(cols = vars(Strain, O2_uM, Photoperiod), rows = vars(WL), 
                    labeller = labeller(WL = label_both, O2_uM = label_both, Strain = label_both, Photoperiod = label_value)) +
  theme_bw() +
  labs(colour = "Actinic PAR (nm)")

 NestedFitsData %>%
   filter(Strain == "MIT9313" |
         Strain == "SS120" |
         Strain == "MED4",
          deltaOD_Lmu_corr == 0 |
          deltaOD_Lmu_corr != 0 & deltaOD_LseGrowthFlag == 1) %>%
  ggplot() +
  geom_point(aes(x = Par_ue, y = deltaOD_Lmu_corr, colour = as.factor(WL)), size = 3, alpha = 0.5) +
  geom_line(aes(x = Par_ue, y = deltaOD_Lmu_corr, colour = as.factor(WL)), linewidth = 0.1, alpha = 0.4) +
labs(subtitle = "Strain, [Oxygen], Photoperiod, Actinic PAR (nm)", y = "deltaOD [~Chl] " ~ "Logistic Growth Rate" ~ "("~h^-1~")" , x = "Growth Light (µmol photons" ~m^-2 ~s^-1~")") +
 geom_text(aes(x = Par_ue, y = deltaOD_Lmu_corr, label= Run),alpha = 0.5 ,hjust = 0.5, vjust=-0.5) +
  geom_errorbar(aes(x = Par_ue, ymin = deltaOD_Lmu_corr - deltaOD_Lmu_se, ymax = deltaOD_Lmu_corr + deltaOD_Lmu_se), size = 0.3, color = 'black', data = . %>% filter(GrowthAmpdeltaODFlag_14days  == 1)) +
  scale_x_continuous(breaks=seq(0, 200, by = 50)) +
  coord_cartesian(xlim = c (0, 200)) +
 #   scale_y_continuous(breaks=seq(-0.03, 0.07, by = 0.01)) +
  # coord_cartesian(ylim = c (-0.03, 0.07)) +
  scale_colour_manual(values = MCMIXColours) +
  ggh4x::facet_nested(cols = vars(Strain, O2_uM, Photoperiod), rows = vars(WL), 
                    labeller = labeller(WL = label_both, O2_uM = label_both, Strain = label_both, Photoperiod = label_value)) +
  theme_bw() +
  labs(colour = "Actinic PAR (nm)")

```