### 2023-12-21 Code/Process_OlisJazEmData.Rmd

---
title: "Process_OlisSpectraJazEmData"
author:
- Sylwia Sliwinska-Wilczewska
- Douglas A. Campbell
date: "`r format(Sys.Date())`"
output:
bookdown::html_document2:
    code_folding: show
    keep_md: yes
    toc: TRUE
    toc_float: TRUE
    toc_depth: 6
    fig_caption: yes
bibliography: BalticPhotoperiod.bib
csl: plos-one.csl
editor_options: 
  markdown: 
    wrap: 72
---

# Set Chunk Options

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
knitr::opts_chunk$set(fig.path='Figs/')
```

# Introduction

Process_OlisSpectraJazEmData.Rmd processes and combines Baltic_Photoperiod_Imported_JazEmData.Rds from Data/ImportedData/ImportedJazEmData folder and Baltic_Photoperiod_Imported_OlisData.Rds from Data/ImportedData/ImportedOlisData folder. This .Rmd generates Baltic_Photoperiod_Processed_OlisSpectraAll.Rds and Baltic_Photoperiod_Processed_OlisSpectraTidy.Rds (both stored in Data/ProcessedData/ProcessedOlisJazData folder) and OlisSpectra_Plot.png (stored in Output/Plots folder).

# Load Libraries and set Project Variables

```{r load libraries, warning = FALSE, echo=FALSE} 
library(tidyverse)
library(lubridate)
library(stringr)
library(broom)
library(knitr)
library(OneR)
library(zoo)
library(strucchange)
library(ggspectra)
library(ggpubr)
library(caret)
library(photobiologyWavebands)
library(reshape2)
library(photobiology)
library(gcookbook)
```

```{r set project variables}
Project <- "Baltic_Photoperiod"
DataOut <- file.path("..", "Data", "ProcessedData", "ProcessedOlisJazData")
DataInOlis <- file.path("..", "Data", "ImportedData", "ImportedOlisData", fsep = .Platform$file.sep)
DataInJaz <- file.path("..", "Data", "ImportedData", "ImportedJazEmData", fsep = .Platform$file.sep)

PlotsPath <- file.path("..", "Output", "Plots")
RDSPlotPath <- file.path("..", "Output", "PlotsRDS")

FileEncode <- "UTF-8" 
Delimiter <- ""
HeaderRows <- 0
```

# List and read imported Olis and Jaz files

```{r exported Rmd only first time in session}
list.files(path = DataInOlis, pattern = Project, full.names = TRUE)
list.files(path = DataInJaz, pattern = Project, full.names = TRUE)
```

```{r read imported Olis file}
OLISSpectraFile <- "../Data/ImportedData/ImportedOlisData/Baltic_Photoperiod_Imported_OlisData.Rds"
OLISSpectraFileName <- str_split(string = OLISSpectraFile, "/")[[1]][3] %>%
  str_remove(pattern = ".Rds")
OLISSpectra <- readRDS(OLISSpectraFile)  %>%
  ungroup()

JazFile <- "../Data/ImportedData/ImportedJazEmData/Baltic_Photoperiod_Imported_JazEmData.Rds"
JazFileName <- str_split(string = JazFile, "/")[[1]][3] %>%
  str_remove(pattern = ".Rds")
JazData <- readRDS(JazFile)  %>%
  ungroup()
```

# Preparing df for further analysis

```{r round nm}
JazData <- JazData %>%
   mutate(WavelengthRound = round(Wavelength))
colnames(JazData)

JazDataRound <- JazData %>%
group_by(FilenameJaz, WavelengthRound) %>%
  summarize(Counts, FilenameJaz, ObsDate, Strain, Par_ue, WL, WavelengthRound, CountsMean = mean(Counts)) %>%
ungroup()

JazDataRound <- JazDataRound %>%
  summarize(FilenameJaz, Strain, Par_ue, WL, WavelengthRound, CountsMean) %>%
  unique()

JazDataRound <- JazDataRound %>%
  filter(WavelengthRound >= 400 & WavelengthRound <= 700)
```

# Normalize Jaz emission to 439 nm

```{r Normalization Jaz emission at 439}
EmissionJazWW <- JazDataRound %>%
  group_by(WL, Strain, Par_ue) %>%
  filter(WavelengthRound == 439) %>%
  mutate(EmJaz439 = CountsMean) %>%
  select(Strain, EmJaz439, Par_ue) %>%
  ungroup()

JazDataRoundMeta <- JazDataRound %>%
  left_join(., EmissionJazWW) %>%
  mutate(EmNormJaz439 = CountsMean / EmJaz439)
```

# Merge Olis and Jaz data

```{r combine Olis spectra and Jaz data}
OLISSpectraMeta <- OLISSpectra %>%
  left_join(., JazDataRoundMeta, by = c("Par_ue" = "Par_ue", "nm" = "WavelengthRound", "Strain" = "Strain", "WL" = "WL")) %>%
  unique()
```

# Estimate PUR

```{r estimate PUR}
OLISSpectraMetaPUR <- OLISSpectraMeta %>%
  mutate(PURNorm = (AbsNorm440*EmNormJaz439)) %>% 
  group_by(WL, Strain, Par_ue, Photoperiod, E_days) %>%
  mutate(PURNormSum = sum(PURNorm)) %>% 
  mutate(SumEmNormJaz439 = sum(EmNormJaz439)) %>% 
  mutate(PUR = Par_ue*(PURNormSum/SumEmNormJaz439)) %>% 
  ungroup() 
```

# Calculate daily PUR photon dose

```{r calculate PUR Photon dose}
OLISSpectraMetaPUR1 <- OLISSpectraMetaPUR %>%
  filter(Photoperiod != "24") %>% 
  mutate(PURPhotonDose_day =(PUR/2)*Photoperiod*3600)

OLISSpectraMetaPUR2 <- OLISSpectraMetaPUR %>%
  filter(Photoperiod == "24") %>% 
  mutate(PURPhotonDose_day = PUR*Photoperiod*3600)

  OLISSpectraMetaPUR <-rbind(OLISSpectraMetaPUR1, OLISSpectraMetaPUR2)
  rm(OLISSpectraMetaPUR1,OLISSpectraMetaPUR2)
```

# Filter unrevelant data

```{r filter unrevelant data}
OLISSpectraMetaPUR<-OLISSpectraMetaPUR %>% 
  filter(PUR>=0) %>% 
  filter(PUR<=900)
```

# Preparing df for creating representative plot

```{r preparing selected data for creating representative plot}
OLISSpectraMetaPURShortBA127R<- OLISSpectraMetaPUR %>%
  filter(Photoperiod == "12") %>% 
  filter(Strain == "BA127R") %>% 
  filter(Par_ue == 300) %>% 
  filter(E_days == "4" | E_days == "8") %>% 
  mutate(E_days=case_when(E_days==4~3,
         E_days==8~7)) %>% 
  mutate(PURNormBA127R = PURNorm) %>% 
  select(-c(PURNorm, FilenameJaz, FilenameOlis))

OLISSpectraMetaPURShortBA77G<- OLISSpectraMetaPUR %>%
  filter(Photoperiod == "12") %>% 
  filter(Strain == "BA77G") %>% 
  filter(Par_ue == 300) %>% 
  filter(E_days == "3" | E_days == "7") %>% 
  mutate(PURNormBA77G = PURNorm) %>% 
  select(-c(PURNorm, FilenameJaz, FilenameOlis))

OLISSpectraMetaPURShort <- OLISSpectraMetaPURShortBA127R %>%
  full_join(., OLISSpectraMetaPURShortBA77G, by = c("SampleID" = "SampleID", "Run" = "Run", "Strain" = "Strain", "ExpDate" = "ExpDate", "Par_ue" = "Par_ue", "Photoperiod" = "Photoperiod", "MC"="MC", "Tube" = "Tube", "O2" = "O2", "WL" = "WL", "LightShape" = "LightShape", "ExpEndDate"="ExpEndDate", "nm" = "nm", "ObsDate" = "ObsDate", "Absorbance" = "Absorbance", "SumAb" = "SumAb", "Abs440" = "Abs440", "AbsNorm440" = "AbsNorm440", "SumAbNorm" = "SumAbNorm", "E_days" = "E_days", "CountsMean" = "CountsMean", "EmJaz439" = "EmJaz439", "EmNormJaz439" = "EmNormJaz439", "PURNormSum" = "PURNormSum", "SumEmNormJaz439" = "SumEmNormJaz439", "PUR" = "PUR", "PURPhotonDose_day" = "PURPhotonDose_day", "PARPhotonDose_day" = "PARPhotonDose_day")) 
```

```{r Changes strain names}
OLISSpectraMetaPURShort <- OLISSpectraMetaPURShort %>%
      mutate(Strain=case_when(Strain=="BA127R"~"PE-rich_127",
         Strain=="BA48R"~"PE-rich_048",
        Strain=="BA56G"~"PC-rich_056",
         Strain=="BA77G"~"PC-rich_077"))
```

# Create plot

```{r Create plot, fig.height = 10, fig.width = 8}
Par_ue.labs <- c("30 µE", "90 µE", "180 µE", "Phase of growth", "900 µE")
names(Par_ue.labs) <- c("30", "90", "180", "300", "900")

Photoperiod.labs <- c("8 h", "Photoperiod: 12 h", "16 h", "24 h")
names(Photoperiod.labs) <- c("8", "12", "16", "24")

E_days.labs <- c("Exponential", "Pre-stationary")
names(E_days.labs) <- c("3", "7")

Strain.labs <- c("PC-rich", "PE-rich")
names(Strain.labs) <- c("PC-rich_077", "PE-rich_127")

O2.labs <- c("Strain")
names(O2.labs) <- c(21)

data_textPUR0 <- data.frame(Par_ue = c(300, 300), E_days = c(3, 7), Strain  = c("PE-rich_127", "PE-rich_127"), Photoperiod = c(12), O2 = c(21), label = c('PAR = 300 µE', 'PAR = 300 µE'))

data_textPUR00 <- data.frame(Par_ue = c(300, 300), E_days = c(3, 7), Strain  = c("PC-rich_077", "PC-rich_077"), Photoperiod = c(12), O2 = c(21), label = c('PAR = 300 µE', 'PAR = 300 µE'))

data_textPUR1 <- data.frame(Par_ue = c(300, 300), E_days = c(3, 7), Strain  = c("PE-rich_127", "PE-rich_127"), Photoperiod = c(12), O2 = c(21), label = c('PUR = 176 µE', 'PUR = 129 µE'))

data_textPUR2 <- data.frame(Par_ue = c(300, 300), E_days = c(3, 7), Strain  = c("PC-rich_077", "PC-rich_077"), Photoperiod = c(12), O2 = c(21), label = c('PUR = 139 µE', 'PUR = 128 µE'))

data_textpigment1 <- data.frame(Par_ue = c(300), E_days = c(3), Strain  = c("PC-rich_077"), Photoperiod = c(12), O2 = c(21), label = c('Chl a 440 nm'))
data_textpigment3 <- data.frame(Par_ue = c(300), E_days = c(3), Strain  = c("PC-rich_077"), Photoperiod = c(12), O2 = c(21), label = c('Chl a 680 nm'))
data_textpigment4 <- data.frame(Par_ue = c(300), E_days = c(3), Strain  = c("PC-rich_077"), Photoperiod = c(12), O2 = c(21), label = c('PC 620 nm'))
data_textpigment5 <- data.frame(Par_ue = c(300), E_days = c(3), Strain  = c("PC-rich_077"), Photoperiod = c(12), O2 = c(21), label = c('Car 480 nm'))
data_textpigment6 <- data.frame(Par_ue = c(300), E_days = c(3), Strain  = c("PC-rich_077"), Photoperiod = c(12), O2 = c(21), label = c('PUB 500 nm'))

arrowdf1 <- tibble(Strain = "PC-rich_077", E_days = 3, Par_ue = 300, O2 = c(21))
arrowdf3 <- tibble(Strain = "PC-rich_077", E_days = 3, Par_ue = 300, O2 = c(21))
arrowdf4 <- tibble(Strain = "PC-rich_077", E_days = 3, Par_ue = 300, O2 = c(21))
arrowdf5 <- tibble(Strain = "PC-rich_077", E_days = 3, Par_ue = 300, O2 = c(21))
arrowdf6 <- tibble(Strain = "PC-rich_077", E_days = 3, Par_ue = 300, O2 = c(21))

data_textpigment7 <- data.frame(Par_ue = c(300), E_days = c(3), Strain  = c("PE-rich_127"), Photoperiod = c(12), O2 = c(21), label = c('Chl a 440 nm'))
data_textpigment8 <- data.frame(Par_ue = c(300), E_days = c(3), Strain  = c("PE-rich_127"), Photoperiod = c(12), O2 = c(21), label = c('Car 480 nm'))
data_textpigment9 <- data.frame(Par_ue = c(300), E_days = c(3), Strain  = c("PE-rich_127"), Photoperiod = c(12), O2 = c(21), label = c('PE 570 nm'))
data_textpigment10 <- data.frame(Par_ue = c(300), E_days = c(3), Strain  = c("PE-rich_127"), Photoperiod = c(12), O2 = c(21), label = c('Chl a 680 nm'))
data_textpigment11 <- data.frame(Par_ue = c(300), E_days = c(3), Strain  = c("PE-rich_127"), Photoperiod = c(12), O2 = c(21), label = c('PC 620 nm'))
data_textpigment12 <- data.frame(Par_ue = c(300), E_days = c(3), Strain  = c("PE-rich_127"), Photoperiod = c(12), O2 = c(21), label = c('PUB 500 nm'))

arrowdf7 <- tibble(Strain = "PE-rich_127", E_days = 3, Par_ue = 300, O2 = c(21))
arrowdf8 <- tibble(Strain = "PE-rich_127", E_days = 3, Par_ue = 300, O2 = c(21))
arrowdf9 <- tibble(Strain = "PE-rich_127", E_days = 3, Par_ue = 300, O2 = c(21))
arrowdf10 <- tibble(Strain = "PE-rich_127", E_days = 3, Par_ue = 300, O2 = c(21))
arrowdf11 <- tibble(Strain = "PE-rich_127", E_days = 3, Par_ue = 300, O2 = c(21))
arrowdf12 <- tibble(Strain = "PE-rich_127", E_days = 3, Par_ue = 300, O2 = c(21))

OLISSpectraMetaPURShort %>%
  filter(Photoperiod == "12") %>%
  filter(Strain == "PC-rich_077" | Strain == "PE-rich_127") %>%
  filter(Par_ue == 300) %>%
  filter(E_days == "3" | E_days == "7") %>%
  ggplot() +

  geom_area(aes(x = nm, y = EmNormJaz439, fill = "gray84"), alpha = 0.6, show.legend = F) +
  geom_area(aes(x = nm, y = PURNormBA127R, fill = "brown4"), alpha = 0.5, show.legend = F) +
  geom_area(aes(x = nm, y = PURNormBA77G, fill = "palegreen3"), alpha = 0.5, show.legend = F) +
  geom_line(aes(x = nm, y = AbsNorm440, colour = as.factor(Strain), linetype = as.factor(Strain)), show.legend = F, size = 0.4) +

  geom_text(data=data_textPUR0, aes(x=446, y=1.26, label=label), size=3.7) +
  geom_text(data=data_textPUR00, aes(x=446, y=1.26, label=label), size=3.7) +
  geom_text(data=data_textPUR1, aes(x=446, y=1.21, label=label), size=3.7) +
  geom_text(data=data_textPUR2, aes(x=446, y=1.21, label=label), size=3.7) +

  geom_text(data=data_textpigment1, aes(x=500, y=1.1, label=label), size=4, colour = "darkgreen") + #chla440
  geom_text(data=data_textpigment5, aes(x=500, y=0.8, label=label), size=4, colour = "orange") + #Car
  geom_text(data=data_textpigment3, aes(x=650, y=1.2, label=label), size=4, colour = "darkgreen") +#chla680
  geom_text(data=data_textpigment4, aes(x=590, y=0.95, label=label), size=4, colour = "deepskyblue4") + #PC
  geom_text(data=data_textpigment6, aes(x=485, y=0.25, label=label), size=4, colour = "sienna4") + #PUB
  geom_segment(data = arrowdf1,
               aes(x = 480, xend = 445, y = 1.05, yend = 1),
               colour = "darkgreen", size = 1, alpha=0.9, arrow = arrow(length=unit(0.2, 'cm'))) + #chla440
  geom_segment(data = arrowdf5,
               aes(x = 480, xend = 475, y = 0.75, yend = 0.54),
               colour = "orange", size = 1, alpha=0.9, arrow = arrow(length=unit(0.2, 'cm'))) + #Car
  geom_segment(data = arrowdf3,
               aes(x = 660, xend = 680, y = 1.14, yend = 0.95),
               colour = "darkgreen", size = 1, alpha=0.9, arrow = arrow(length=unit(0.2, 'cm'))) + #chla680
  geom_segment(data = arrowdf4,
               aes(x = 610, xend = 630, y = 0.90, yend = 0.82),
               colour = "deepskyblue4", size = 1, alpha=0.9, arrow = arrow(length=unit(0.2, 'cm'))) + #PC
  geom_segment(data = arrowdf4,
               aes(x = 480, xend = 494, y = 0.30, yend = 0.41),
               colour = "sienna4", size = 1, alpha=0.9, arrow = arrow(length=unit(0.2, 'cm'))) + #PUB
  geom_text(data=data_textpigment7, aes(x=500, y=1.1, label=label), size=4, colour = "darkgreen") + #chla440
  geom_text(data=data_textpigment8, aes(x=500, y=0.8, label=label), size=4, colour = "orange") + #Car
  geom_text(data=data_textpigment9, aes(x=580, y=1.0, label=label), size=4, colour = "darkmagenta") + #PE
  geom_text(data=data_textpigment10, aes(x=650, y=1.2, label=label), size=4, colour = "darkgreen") +#chla680
  geom_text(data=data_textpigment11, aes(x=600, y=0.17, label=label), size=4, colour = "deepskyblue4") + #PC
  geom_text(data=data_textpigment12, aes(x=495, y=0.30, label=label), size=4, colour = "sienna4") + #PUB
  geom_segment(data = arrowdf7,
               aes(x = 480, xend = 445, y = 1.05, yend = 1),
               colour = "darkgreen", size = 1, alpha=0.9, arrow = arrow(length=unit(0.2, 'cm'))) + #chla440
  geom_segment(data = arrowdf8,
               aes(x = 480, xend = 475, y = 0.75, yend = 0.55),
               colour = "orange", size = 1, alpha=0.9, arrow = arrow(length=unit(0.2, 'cm'))) + #Car
  geom_segment(data = arrowdf9,
               aes(x = 590, xend = 575, y = 0.95, yend = 0.85),
               colour = "darkmagenta", size = 1, alpha=0.9, arrow = arrow(length=unit(0.2, 'cm'))) + #PE
  geom_segment(data = arrowdf10,
               aes(x = 660, xend = 680, y = 1.14, yend = 0.85),
               colour = "darkgreen", size = 1, alpha=0.9, arrow = arrow(length=unit(0.2, 'cm'))) + #chla680
  geom_segment(data = arrowdf11,
               aes(x = 610, xend = 630, y = 0.2, yend = 0.3),
               colour = "deepskyblue4", size = 1, alpha=0.9, arrow = arrow(length=unit(0.2, 'cm'))) + #PC
  geom_segment(data = arrowdf12,
               aes(x = 490, xend = 500, y = 0.35, yend = 0.52),
               colour = "sienna4", size = 1, alpha=0.9, arrow = arrow(length=unit(0.2, 'cm'))) + #PUB
  scale_color_manual(values = c("PC-rich_077" ="palegreen3", "PE-rich_127" ="brown4")) +
  scale_linetype_manual(values = c("PC-rich_077" ="solid", "PE-rich_127" ="longdash")) +
  stat_wl_strip(aes(x = nm), ymin = -Inf, ymax = -0.025, alpha = 0.5) +
  scale_fill_identity() +
  labs(y = "Normalized absorbance", x = "Wavelength (nm)") +
  scale_y_continuous(breaks=seq(0, 1.2, by = 0.4)) +
  coord_cartesian(ylim = c (-0.01, 1.25)) +
  ggh4x::facet_nested(cols = vars(Par_ue, E_days), rows = vars(O2, Strain),labeller = labeller(Ex_WL = label_both, WL = label_both, E_days = E_days.labs, Photoperiod = Photoperiod.labs, Par_ue = Par_ue.labs, Strain = Strain.labs, O2 = O2.labs, )) +
  theme_bw() +
  theme(panel.grid.minor = element_blank(),
        panel.grid.major = element_blank(),
        panel.spacing.x = unit(0.4, 'cm'),
        axis.text = element_text(size=12),
        axis.text.x = element_text(size=12),
        axis.title = element_text(size=16),
        strip.background = element_rect(fill="white"),
        strip.text = element_text(size=12),
        axis.title.y = element_text(margin=margin(r=10)),
        axis.title.x = element_text(margin=margin(t=10)),
        legend.background = element_rect(fill="transparent"),
        legend.title = element_blank(),
        legend.direction="horizontal",
        legend.position = c(0.08,0.97),
        legend.spacing.x = unit(0.01, 'cm'),
        legend.text.align = 0,
        legend.key.width= unit(0.3, 'cm'),
        legend.text = element_text(size=10))
```

# Save plot

```{r save plot}
ggsave(file = file.path(PlotsPath, paste("OlisSpectra_Plot",".png",sep = "")), height=10, width= 8,  dpi = 300, limitsize = TRUE)
```

```{r Create plot, fig.height = 6, fig.width = 8}

Par_ue.labs <- c("30 µE", "90 µE", "180 µE", "Phase of growth", "900 µE")
names(Par_ue.labs) <- c("30", "90", "180", "300", "900")

Photoperiod.labs <- c("8 h", "Photoperiod: 12 h", "16 h", "24 h")
names(Photoperiod.labs) <- c("8", "12", "16", "24")

E_days.labs <- c("Exponential", "Pre-stationary")
names(E_days.labs) <- c("3", "7")

Strain.labs <- c("PE-rich_127", "PC-rich_077")
names(Strain.labs) <- c("PE-rich_127", "PC-rich_077")

data_textPUR0 <- data.frame(Par_ue = c(300, 300), E_days = c(3, 7), Strain  = c("PE-rich_127", "PE-rich_127"), Photoperiod = c(12), label = c('PAR = 300 µE', 'PAR = 300 µE'))

data_textPUR1 <- data.frame(Par_ue = c(300, 300), E_days = c(3, 7), Strain  = c("PE-rich_127", "PE-rich_127"), Photoperiod = c(12), label = c('PUR (PE-rich) = 176 µE', 'PUR (PE-rich) = 129 µE'))

data_textPUR2 <- data.frame(Par_ue = c(300, 300), E_days = c(3, 7), Strain  = c("PE-rich_127", "PE-rich_127"), Photoperiod = c(12), label = c('PUR (PC-rich) = 139 µE', 'PUR (PC-rich) = 128 µE'))

data_textpigment1 <- data.frame(Par_ue = c(300), E_days = c(3), Strain  = c("PE-rich_127"), Photoperiod = c(12), label = c('Chl a 440 nm'))
data_textpigment2 <- data.frame(Par_ue = c(300), E_days = c(3), Strain  = c("PE-rich_127"), Photoperiod = c(12), label = c('PE 570 nm'))
data_textpigment3 <- data.frame(Par_ue = c(300), E_days = c(3), Strain  = c("PE-rich_127"), Photoperiod = c(12), label = c('Chl a 680 nm'))
data_textpigment4 <- data.frame(Par_ue = c(300), E_days = c(3), Strain  = c("PE-rich_127"), Photoperiod = c(12), label = c('PC 620 nm'))
data_textpigment5 <- data.frame(Par_ue = c(300), E_days = c(3), Strain  = c("PE-rich_127"), Photoperiod = c(12), label = c('Car 480 nm'))
data_textpigment6 <- data.frame(Par_ue = c(300), E_days = c(3), Strain  = c("PE-rich_127"), Photoperiod = c(12), label = c('PUB 500 nm'))

arrowdf1 <- tibble(Strain = "PE-rich_127", E_days = 3, Par_ue = 300)
arrowdf2 <- tibble(Strain = "PE-rich_127", E_days = 3, Par_ue = 300)
arrowdf3 <- tibble(Strain = "PE-rich_127", E_days = 3, Par_ue = 300)
arrowdf4 <- tibble(Strain = "PE-rich_127", E_days = 3, Par_ue = 300)
arrowdf5 <- tibble(Strain = "PE-rich_127", E_days = 3, Par_ue = 300)
arrowdf6 <- tibble(Strain = "PE-rich_127", E_days = 3, Par_ue = 300)

OLISSpectraMetaPURShort %>%
  filter(Photoperiod == "12") %>%
  filter(Strain == "PE-rich_127" | Strain == "PC-rich_077") %>%
  filter(Par_ue == 300) %>%
  filter(E_days == "3" | E_days == "7") %>%
  ggplot() +
  geom_area(aes(x = nm, y = EmNormJaz439, fill = "gray84"), alpha = 0.6, show.legend = F) +
  geom_area(aes(x = nm, y = PURNormBA127R, fill = "brown4"), alpha = 0.5, show.legend = F) +
  geom_area(aes(x = nm, y = PURNormBA77G, fill = "palegreen3"), alpha = 0.5, show.legend = F) +
  geom_line(aes(x = nm, y = AbsNorm440, colour = as.factor(Strain), linetype = as.factor(Strain)), show.legend = F, size = 0.7) +

  geom_text(data=data_textPUR0, aes(x=435, y=1.26, label=label), size=3.5) +
  geom_text(data=data_textPUR1, aes(x=462, y=1.21, label=label), size=3.5) +
  geom_text(data=data_textPUR2, aes(x=462, y=1.16, label=label), size=3.5) +

  geom_text(data=data_textpigment1, aes(x=450, y=1.07, label=label), size=4, colour = "darkgreen") + #chla440
  geom_text(data=data_textpigment5, aes(x=430, y=0.36, label=label), size=4, colour = "coral3") + #Car
  geom_text(data=data_textpigment2, aes(x=520, y=0.92, label=label), size=4, colour = "darkmagenta") + #PE
  geom_text(data=data_textpigment3, aes(x=660, y=1.08, label=label), size=4, colour = "darkgreen") +#chla680
  geom_text(data=data_textpigment4, aes(x=605, y=1.0, label=label), size=4, colour = "deepskyblue4") + #PC
  geom_text(data=data_textpigment6, aes(x=485, y=0.08, label=label), size=4, colour = "sienna4") + #PUB

  geom_segment(data = arrowdf1,
               aes(x = 470, xend = 440, y = 1.04, yend = 1.02),
               colour = "darkgreen", size = 1, alpha=0.9, arrow = arrow(length=unit(0.2, 'cm'))) + #chla440
  geom_segment(data = arrowdf5,
               aes(x = 450, xend = 470, y = 0.4, yend = 0.48),
               colour = "coral3", size = 1, alpha=0.9, arrow = arrow(length=unit(0.2, 'cm'))) + #Car
  geom_segment(data = arrowdf2,
               aes(x = 538, xend = 568, y = 0.88, yend = 0.86),
               colour = "darkmagenta", size = 1, alpha=0.9, arrow = arrow(length=unit(0.2, 'cm'))) + #PE
  geom_segment(data = arrowdf3,
               aes(x = 697, xend = 680, y = 1.05, yend = 0.92),
               colour = "darkgreen", size = 1, alpha=0.9, arrow = arrow(length=unit(0.2, 'cm'))) + #chla680
  geom_segment(data = arrowdf4,
               aes(x = 632, xend = 630, y = 0.95, yend = 0.82),
               colour = "deepskyblue4", size = 1, alpha=0.9, arrow = arrow(length=unit(0.2, 'cm'))) + #PC
  geom_segment(data = arrowdf4,
               aes(x = 465, xend = 500, y = 0.12, yend = 0.52),
               colour = "sienna4", size = 1, alpha=0.9, arrow = arrow(length=unit(0.2, 'cm'))) + #PUB
  scale_color_manual(values = c("PE-rich_127" ="brown4", "PC-rich_077" ="palegreen3")) +
  scale_linetype_manual(values = c("PE-rich_127" ="longdash", "PC-rich_077" ="solid")) +
  stat_wl_strip(aes(x = nm), ymin = -Inf, ymax = -0.025, alpha = 0.5) + 
  scale_fill_identity() +
  labs(y = "Normalized absorbance", x = "Wavelength (nm)") +
  scale_y_continuous(breaks=seq(0, 1.2, by = 0.4)) +
  coord_cartesian(ylim = c (-0.01, 1.25)) +
  ggh4x::facet_nested(cols = vars(Par_ue, E_days), labeller = labeller(Ex_WL = label_both, WL = label_both, E_days = E_days.labs, Photoperiod = Photoperiod.labs, Par_ue = Par_ue.labs, Strain = Strain.labs)) +
  theme_bw() +
  theme(panel.grid.minor = element_blank(),
        panel.grid.major = element_blank(),
        panel.spacing.x = unit(0.4, 'cm'),
        axis.text = element_text(size=12),
        axis.text.x = element_text(size=12),
        axis.title = element_text(size=16),
        strip.background = element_rect(fill="white"),
        strip.text = element_text(size=12),
        axis.title.y = element_text(margin=margin(r=10)),
        axis.title.x = element_text(margin=margin(t=10)),
        legend.background = element_rect(fill="transparent"),
        legend.title = element_blank(),
        legend.direction="horizontal",
        legend.position = c(0.08,0.97),
        legend.spacing.x = unit(0.01, 'cm'),
        legend.text.align = 0,
        legend.key.width= unit(0.3, 'cm'),
        legend.text = element_text(size=10))
```

# Cleaning df before saving as rds and removed unnecessary files from the environment

```{r delete unused df from the environment}
rm(data_textpigment1,data_textpigment2,data_textpigment3,data_textpigment4,data_textpigment5,data_textpigment6, data_textpigment7, data_textpigment8, data_textpigment9, data_textpigment10, data_textpigment11, data_textpigment12, arrowdf1,arrowdf2,arrowdf3,arrowdf4,arrowdf5,arrowdf6, arrowdf7, arrowdf8, arrowdf9, arrowdf10, arrowdf11, arrowdf12, data_textPUR0, data_textPUR00, data_textPUR1,data_textPUR2, EmissionJazWW, JazData, JazDataRound, JazDataRoundMeta, OLISSpectraMetaPURShortBA127R, OLISSpectraMetaPURShortBA77G, OLISSpectraMeta, OLISSpectra)
```

# Create tidy df without all nm

```{r clean df, create smaller df - without absorbance and all nm}
OLISAllPURTidy<-OLISSpectraMetaPUR %>% 
  select(-c(Absorbance, nm, Abs440, AbsNorm440, EmJaz439, EmNormJaz439, PURNorm, CountsMean, SumAbNorm, PURNormSum, SumEmNormJaz439)) %>% 
  unique()

OLISAllPURTidy<-OLISAllPURTidy %>% 
  group_by(Strain, Par_ue, Photoperiod, E_days, MC, Run) %>%
  mutate(PURPARRatio = PUR/Par_ue) %>% 
ungroup()

OLISAllPURTidy <- OLISAllPURTidy %>% 
  mutate(E_hours = E_days*24) %>% 
  filter(PURPARRatio<=1) %>% 
  filter(PURPARRatio>0)
```

# Preparing df for creating plot

```{r create tidy PAR photon dose and facets for plot}
OLISAllPURTidy <- OLISAllPURTidy %>%
    mutate(PARPhotonDoseTidy=PARPhotonDose_day/100000) %>% 
    mutate(Strain=case_when(Strain=="BA127R"~"PE-rich_127",
         Strain=="BA48R"~"PE-rich_048",
        Strain=="BA56G"~"PC-rich_056",
         Strain=="BA77G"~"PC-rich_077"))

OLISAllPURTidy$facetsPar_ue = factor(OLISAllPURTidy$O2, labels = c("PAR~(µmol~photons~m^{-2}~s^{-1})"))
OLISAllPURTidy$facetsPhotoperiod = factor(OLISAllPURTidy$WL, labels = c("Photoperiod~(h)"))
OLISAllPURTidy$facetsStrain = factor(OLISAllPURTidy$WL, labels = c("Strain"))
OLISAllPURTidy$facetsPhase = factor(OLISAllPURTidy$WL, labels = c("Phase~of~growth"))
OLISAllPURTidy$facetsPARPhotonDose_day = factor(OLISAllPURTidy$WL, labels = c("PAR~photon~dose~(10^{5}~µmol~photons~m^{-2}~d^{-1})"))
```

# Create preliminary PUR/PAR ratio plot

```{r preliminary plot}
# lab1=c(expression("PC-rich_056"), expression("PC-rich_077"), expression("PE-rich_048"), expression("PE-rich_127"))
# lab2=c(expression("8 h"), expression("12 h"), expression("16 h"), expression("24 h"))
# lab3=c(expression("30 µE"), expression("90 µE"), expression("180 µE"), expression("300 µE"), expression("900 µE"))
# 
# OLISAllPURTidy %>%
#   ggplot() +
#   geom_point(aes(x = E_hours, y = PURPARRatio, colour = as.factor(Strain)), alpha = 0.9, size = 3, show.legend = T) +
#   scale_colour_discrete(type=c("seagreen4", "palegreen3", "brown1", "brown4"), name="", labels = lab1) +
#   scale_y_continuous(breaks=seq(0, 1.2, by = 0.4)) +
#   coord_cartesian(ylim = c(0, 1.7)) +
#   ggh4x::facet_nested(cols = vars(facetsPhotoperiod, Photoperiod), rows = vars(facetsPar_ue, Par_ue), labeller = label_parsed) +
#   theme_bw()
```

# Load tmaxAG catalog

```{r}
gs4_deauth()
tmaxAG<- read_sheet("https://docs.google.com/spreadsheets/d/1ksY7xlg9wOsICOBRmZkHPKdd9KOislNwPDzyuJ3UIUI/edit#gid=0")
as.data.frame(tmaxAG)
tmaxAG <- tmaxAG

tmaxAG<-tmaxAG %>% 
  mutate(Par_ue = as.numeric(Par_ue)) %>%
  mutate(Photoperiod = as.numeric(Photoperiod)) %>%
  mutate(tMaxAG = as.numeric(tMaxAG)) %>%
  mutate(dayRound_tmaxAG=tMaxAG/24) %>% 
  mutate(dayRound_tmaxAG = round(dayRound_tmaxAG, digits = 0)) %>% 
  mutate(Strain=case_when(Strain=="BA127R"~"PE-rich_127",
         Strain=="BA48R"~"PE-rich_048",
        Strain=="BA56G"~"PC-rich_056",
         Strain=="BA77G"~"PC-rich_077"))

tmaxAG$facetsPar_ue = factor(tmaxAG$O2, labels = c("PAR~(µmol~photons~m^{-2}~s^{-1})"))
tmaxAG$facetsPhotoperiod = factor(tmaxAG$WL, labels = c("Photoperiod~(h)"))
tmaxAG$facetsStrain = factor(tmaxAG$O2, labels = c("Strain"))
tmaxAG$facetsPARPhotonDose_day = factor(tmaxAG$WL, labels = c("PAR~photon~dose~(10^{5}~µmol~photons~m^{-2}~d^{-1})"))

tmaxAGno24<-tmaxAG %>% 
  filter(Photoperiod != "24") %>% 
  mutate(PARPhotonDose_day =(Par_ue/2)*Photoperiod*3600)
tmaxAG24<-tmaxAG %>% 
  filter(Photoperiod == "24") %>% 
  mutate(PARPhotonDose_day = Par_ue*Photoperiod*3600) 
tmaxAG <-rbind(tmaxAGno24, tmaxAG24)
  rm(tmaxAGno24, tmaxAG24)
```

# Select variable for choosen PARPhotonDose to create a plot

```{r select variable for choosen PARPhotonDose to create a plot}
tmaxAGSelect<-tmaxAG %>% 
  mutate(PARPhotonDoseTidy=PARPhotonDose_day/100000) %>% 
  filter(PARPhotonDoseTidy == 4.32000|
  PARPhotonDoseTidy == 6.48000|
  PARPhotonDoseTidy == 8.64000|
  PARPhotonDoseTidy == 43.20000|
  PARPhotonDoseTidy == 64.80000|
  PARPhotonDoseTidy == 86.40000) 
```

# Create 4 df for vertical lines (coloured as Strain) contained tMaxAG value

```{r select revelant columns}
tMaxAGLines_056Select<-tmaxAGSelect %>% 
  filter(Strain == "PC-rich_056") %>% 
  select(c(Par_ue, Photoperiod, facetsPARPhotonDose_day, PARPhotonDoseTidy, tMaxAG)) %>% 
  unique()
tMaxAGLines_077Select<-tmaxAGSelect %>% 
  filter(Strain == "PC-rich_077") %>% 
  select(c(Par_ue, Photoperiod, facetsPARPhotonDose_day, PARPhotonDoseTidy, tMaxAG)) %>% 
  unique()
tMaxAGLines_048Select<-tmaxAGSelect %>% 
  filter(Strain == "PE-rich_048") %>% 
  select(c(Par_ue, Photoperiod, facetsPARPhotonDose_day, PARPhotonDoseTidy, tMaxAG)) %>% 
  unique()
tMaxAGLines_127Select<-tmaxAGSelect %>% 
  filter(Strain == "PE-rich_127") %>% 
  select(c(Par_ue, Photoperiod, facetsPARPhotonDose_day, PARPhotonDoseTidy, tMaxAG)) %>% 
  unique()
```

# Create plot

```{r Create final PUR/PAR ratio plot, fig.height = 5, fig.width = 8}
lab1=c(expression("PC-rich_056"), expression("PC-rich_077"), expression("PE-rich_048"), expression("PE-rich_127"))
lab2=c(expression("8 h"), expression("12 h"), expression("16 h"), expression("24 h"))
lab3=c(expression("30 µE"), expression("90 µE"), expression("180 µE"), expression("300 µE"), expression("900 µE"))

OLISAllPURTidy %>%
  filter(PARPhotonDoseTidy == 4.32000|
  PARPhotonDoseTidy == 6.48000|
  PARPhotonDoseTidy == 8.64000|
  PARPhotonDoseTidy == 43.20000|
  PARPhotonDoseTidy == 64.80000|
  PARPhotonDoseTidy == 86.40000) %>%
  ggplot() +
  geom_point(aes(x = E_hours, y = PURPARRatio, colour = as.factor(Strain)), alpha = 0.9, size = 3, show.legend = T) +
  geom_vline(data = tMaxAGLines_056Select, aes(xintercept = tMaxAG), linetype="dotdash", colour = "seagreen4", size=0.4) +
  geom_vline(data = tMaxAGLines_077Select, aes(xintercept = tMaxAG), linetype="dotdash", colour = "palegreen3", size=0.4) +
  geom_vline(data = tMaxAGLines_048Select, aes(xintercept = tMaxAG), linetype="dotdash", colour = "brown4", size=0.4) +
  geom_vline(data = tMaxAGLines_127Select, aes(xintercept = tMaxAG), linetype="dotdash", colour = "brown1", size=0.4) +
  scale_colour_discrete(type=c("seagreen4", "palegreen3", "brown1", "brown4")) +
  labs(y = "PUR/PAR ratio", x = "Elapsed time (h)") +
  scale_y_continuous(breaks=seq(0, 1.2, by = 0.4)) +
  coord_cartesian(ylim = c(0, 1.5)) +
  ggh4x::facet_nested(cols = vars(facetsPARPhotonDose_day, PARPhotonDoseTidy), labeller = label_parsed) +
  theme_bw() +
  theme(panel.grid.minor = element_blank(),
        panel.grid.major = element_blank(),
        axis.text = element_text(size=12),
        axis.title = element_text(size=16),
        strip.background = element_rect(fill="white"),
        strip.text = element_text(size=12),
        axis.title.y = element_text(margin=margin(r=10)),
        axis.title.x = element_text(margin=margin(t=10)),
        legend.background = element_rect(fill="white"),
        legend.position = c(0.08,0.87),
        legend.key.height= unit(0.005, 'cm'),
        legend.spacing.y = unit(-0.005, "cm"),
        legend.spacing.x = unit(-0.005, "cm"),
        legend.title = element_blank(),
        legend.text = element_text(size=10))
```

# Cleaning df before saving as rds and removed unnecessary files from the environment

```{r}
rm(tMaxAGLines_056Select, tMaxAGLines_077Select, tMaxAGLines_048Select, tMaxAGLines_127Select, tmaxAG, tmaxAGSelect)
```

# Save rds for further analysis

```{r save rds}
saveRDS(OLISSpectraMetaPUR, file.path(DataOut, paste(Project, "Processed_OlisSpectraAll.Rds", sep = "_"), fsep = .Platform$file.sep), ascii = FALSE, version = NULL, compress = "xz", refhook = NULL)

saveRDS(OLISAllPURTidy, file.path(DataOut, paste(Project, "Processed_OlisSpectraTidy.Rds", sep = "_"), fsep = .Platform$file.sep), ascii = FALSE, version = NULL, compress = "xz", refhook = NULL)

saveRDS(OLISSpectraMetaPURShort, file.path(RDSPlotPath, paste(Project, "Plot_OlisSpectra.Rds", sep = "_"), fsep = .Platform$file.sep), ascii = FALSE, version = NULL, compress = "xz", refhook = NULL)
```

# Variable names used in Data Dictionary

```{r}
colnames(OLISSpectraMetaPUR)
```

# Variable names used in Data Dictionary

```{r}
colnames(OLISAllPURTidy)
```