### 2024-01-09 Code/Process_OlisJazEmData.Rmd
Removed PurParRatio from this rmd and added to the merge rmd

---
title: "Process_OlisJazEmData"
author:
- Mireille Savoie
- Sylwia Sliwinska-Wilczewska
- Douglas A. Campbell
date: "`r format(Sys.Date())`"
output:
bookdown::html_document2:
    code_folding: show
    keep_md: yes
    toc: TRUE
    toc_float: TRUE
    toc_depth: 6
    fig_caption: yes
bibliography: BalticPhotoperiod.bib
csl: plos-one.csl
editor_options: 
  markdown: 
    wrap: 72
---

# Set Chunk Options

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
knitr::opts_chunk$set(fig.path='Figs/')
```

# Introduction

Process_OlisSpectraJazEmData.Rmd processes and combines Baltic_Photoperiod_Imported_JazEmData.Rds from Data/ImportedData/ImportedJazEmData folder and Baltic_Photoperiod_Imported_OlisData.Rds from Data/ImportedData/ImportedOlisData folder. This .Rmd generates Baltic_Photoperiod_Processed_OlisSpectraAll.Rds and Baltic_Photoperiod_Processed_OlisSpectraTidy.Rds (both stored in Data/ProcessedData/ProcessedOlisJazData folder) and OlisSpectra_Plot.png (stored in Output/Plots folder).

# Load Libraries
```{r load libraries, warning = FALSE, echo=FALSE}

library(tidyverse)
library(lubridate)
library(broom)
library(knitr)
library(zoo)
library(photobiologyWavebands)

```

```{r set project variables}

Project <- "PICO"

DataInOlis <- file.path("..", "Data", "ImportedData", "ImportedOlisData", fsep = .Platform$file.sep)
DataInJaz <- file.path("..", "Data", "ImportedData", "ImportedJazEmData", fsep = .Platform$file.sep)
DataOut <- file.path("..", "Data", "ProcessedData", "ProcessedOlisJazData")
PlotsPath <- file.path("..", "Output", "Plots")

SectoHour <- 3600 #convert seconds to hours for PUR photon dose per day

```

# List imported Olis and Jaz files
```{r exported Rmd only first time in session}
list.files(path = DataInOlis, pattern = ".Rds", full.names = TRUE)
list.files(path = DataInJaz, pattern = ".Rds", full.names = TRUE)
```
# Read Olis whole cell absorbance and MultiCulti Jaz emission files
```{r read imported Olis and Jaz file}
OLISSpectraFile <- readRDS(file = file.path(DataInOlis,"PICO_Imported_OlisData.Rds"))

JazEmFile <- readRDS(file = file.path(DataInJaz,"PICO_Imported_JazEmData.Rds"))
```

# Filter out the white light emission data for mixed light color MC since not used. Calculate MC normalized emission to the emission maxima for each wavelength.
```{r calculate normalized emission for each WL}
JazEmFile <- JazEmFile %>% 
  filter(MC_SpectralID != "MCMIX004_WW",
         MCSetPar_uE == 30 |
         MCSetPar_uE == 15) %>% 
  select(-c( FilenameJaz, MC)) %>% 
  unnest(Spectra) %>%  
  group_by(WL) %>% 
  mutate(MC_NormEm = round(Emission / Emission[which.max(Emission)], digits = 4)) %>% 
  select(-c(ObsDate, MCSetPar_uE, MC_SpectralID, MCMeasuredPar_uE)) %>% 
ungroup()
```

# Unnest the Olis whole cell absorbance data and extract absorbance maxima for 675 nm (red) and 450 (blue) regions and extract nm (wavelength) at that absorbance maxima 
```{r}
OLISSpectraUnnest <- OLISSpectraFile %>% 
  unnest(AbsSpectra)

OlisSpectra_red <- OLISSpectraUnnest %>% 
  group_by(SampleID, E_days) %>%
  filter(OlisAbs_nm >= 625 & OlisAbs_nm <= 750) %>% 
   mutate(RedMaxima_nm = round(OlisAbs_nm[which.max(Absorbance)]),
          MaxAbs_Red = Absorbance[which.max(Absorbance)]) %>% 
  filter(OlisAbs_nm == 675) %>% # filter only 675 nm to get df that matches the nested df to join
  select(SampleID, Run, Strain, RedMaxima_nm, MaxAbs_Red, Par_ue, E_days) %>% 
     ungroup()

OlisSpectra_blue <- OLISSpectraUnnest %>% 
  group_by(SampleID, E_days) %>%
  filter(OlisAbs_nm >= 400 & OlisAbs_nm <= 460) %>% 
   mutate(BlueMaxima_nm = round(OlisAbs_nm[which.max(Absorbance)]),
          MaxAbs_Blue = Absorbance[which.max(Absorbance)])  %>% 
  filter(OlisAbs_nm == 440) %>% # filter only 440 nm to get df that matches the nested df to join
  select(SampleID, Run, Strain, BlueMaxima_nm, MaxAbs_Blue,  Par_ue, E_days) %>% 
  ungroup()

 OlisSpectraMeta <- left_join(OLISSpectraFile, OlisSpectra_red)
 OlisSpectraMeta <- left_join(OlisSpectraMeta, OlisSpectra_blue)

remove(OlisSpectra_red)
remove(OlisSpectra_blue)
remove(OLISSpectraUnnest)

```

# Left join Olis whole cell absorbance data with Jaz MC emission data. Emission measurements were performed on the mixed color MCMIX004 and the white light MC257
```{r join whole cell absorbance with MC emission data}

OlisSpectraMeta <- OlisSpectraMeta %>% 
  unnest(AbsSpectra)

OlisJazMetaData <- left_join(OlisSpectraMeta, JazEmFile, by = c("WL", "OlisAbs_nm" = "MCEm_nm")) %>% 
  mutate(MCEm_nm = OlisAbs_nm)

remove(OlisSpectraMeta)
remove(JazEmFile)
remove(OLISSpectraFile)
```

# Normalization. Used max peak of each spectra to normalize peaks to red Chla region (625 to 750 nm) and blue Chla region (400 to 460 nm).
```{r normalization }

OlisJazMetaData <- OlisJazMetaData %>% 
  group_by(SampleID, E_days) %>%
 mutate(AbsNorm440 = Absorbance / MaxAbs_Blue,
        AbsNorm675 = Absorbance / MaxAbs_Red) %>% 
  ungroup()

glimpse(OlisJazMetaData)
```

# Calculate Photosynthetic Usable Radiation (PUR) from MC normalized emission and normalized Olis whole cell absorbance spectrums.  
```{r PUR}
OlisJazMetaPUR <- OlisJazMetaData %>%
    filter(OlisAbs_nm >= 400 &
           OlisAbs_nm <= 700) %>%
  mutate(PURAbsWholeCell_Norm440 = MC_NormEm * AbsNorm440) %>% 
  group_by(SampleID, E_days) %>% 
  mutate(SumPURAbsWholeCell_Norm440 = sum(MC_NormEm * AbsNorm440),
         CalcPur_ue = round(Par_ue * (SumPURAbsWholeCell_Norm440 / sum(MC_NormEm)))) %>% 
ungroup() %>% 
  select(-SumPURAbsWholeCell_Norm440)

remove(OlisJazMetaData)
glimpse(OlisJazMetaPUR)

```

# Find Strain specific average PUR for each spectral color and each growth PAR
```{r Calc mean PUR}

OlisJazMetaPUR <- OlisJazMetaPUR %>%
  group_by(Strain, WL, Par_ue) %>% 
  mutate(Pur_ue = round(mean(CalcPur_ue)),
         PurParRatio = (Pur_ue / Par_ue))%>% 
ungroup()  %>% 
  select(-c(CalcPur_ue))

glimpse(OlisJazMetaPUR)
```

# Calculate daily PUR photon dose
```{r calculate PUR Photon dose}

OlisJazMetaPUR <- OlisJazMetaPUR %>%
 mutate(PURPhotonDose_day = case_when(LightShape == 'Sine' ~ Pur_ue/2 * SectoHour * Photoperiod,
                                  LightShape == 'Square' ~ Pur_ue * SectoHour * Photoperiod))

glimpse(OlisJazMetaPUR)
```

# Nest whole cell absorbance and emission data
```{r Nest spectra data}

OlisJazMetaPUR <- OlisJazMetaPUR %>% 
  group_by(SampleID, E_days) %>% 
  nest(PurData = c(OlisAbs_nm, Absorbance, AbsNorm440, AbsNorm675, PURAbsWholeCell_Norm440, BlueMaxima_nm, MaxAbs_Blue, RedMaxima_nm, MaxAbs_Red, MCEm_nm, Emission, MC_NormEm, EmMax_nm))

glimpse(OlisJazMetaPUR)
```

# Save nested rds for further analysis
```{r save rds}
saveRDS(OlisJazMetaPUR, file.path(DataOut, paste(Project, "Processed_OlisJazMetaPUR.Rds", sep = "_"), fsep = .Platform$file.sep), ascii = FALSE, version = NULL, compress = "xz", refhook = NULL)

```