### 2024-01-10 Code/Merge_GrowthPURData.Rmd
Removed redundant libraries and added ObsDate to merged DF

---
title: "Merge_PICO_Growth_PUR"
author: 
- Mireille Savoie
date: "`r format(Sys.Date())`"
output:
  bookdown::html_document2:
    code_folding: show
    keep_md: yes
    toc: TRUE
    toc_float: TRUE
    toc_depth: 6
    fig_caption: yes
    pandoc_args: 
      - '--lua-filter=scholarly-metadata.lua'
      - '--lua-filter=author-info-blocks.lua'
  bookdown::word_document2:
    code_folding: show
    keep_md: yes
    toc: TRUE
    toc_depth: 6
    fig_caption: yes
    pandoc_args: 
      - '--lua-filter=scholarly-metadata.lua'
      - '--lua-filter=author-info-blocks.lua'
bibliography: [Prochlorococcus_O2_NPQ.bib, RPackages.bib]
csl: plos-one.csl
editor_options: 
  markdown: 
    wrap: 72
---

## Set figure caption font size
```{css, echo=FALSE}
p.caption {
  font-size: 25px;
}
```

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = F)
knitr::opts_chunk$set(fig.path='Figs/')
```

```{r load libraries, include=FALSE}
# libraries; Note check actual dependencies
library(tidyverse)
library(lubridate)
library(broom)
library(knitr)

knitr::write_bib(c("tidyverse", "lubridate","knitr", "broom", "zoo", "RColorBrewer", "grid", "glue", "bookdown"), 'RPackages.bib')

```

```{r set project variables}
#"..", takes up a level in the directory path
Project <- "PICO"

DataInPUR <- file.path("..", "Data", "ProcessedData", "ProcessedOlisJazData", fsep = .Platform$file.sep)
DataInGrowth <- file.path("..", "Data", "CleanData", "CleanedMCData", fsep = .Platform$file.sep)
DataOut <- file.path("..", "Data", "CleanData")
PlotsPath <- file.path("..", "Output", "Plots")

SectoHour <- 3600 #seconds to hour
```

# Read in MC growth fits and PUR .Rds
```{r read ProcessFile}

NestedFitsData <- readRDS(file = file.path(DataInGrowth, "PICO_Processed_MCGrowthFits.Rds"))
NestedPURData <- readRDS(file = file.path(DataInPUR, "PICO_Processed_OlisJazMetaPUR.Rds"))

glimpse(NestedPURData)
```

# Join PUR data to wide dataframe containing growth estimates and PAR. Note: PUR data were calculated using whole cell spectras of cultures grown in high [O2] only. Assumed PUR results were same for all [O2], thus the calculated PUR results were joined to the appropriate matching PAR and WL for each strain.  
```{r Join PUR with PAR and growth fits data}
#add Purdata, e-days and parpurratios to PurParFits
PurData <- NestedPURData %>% 
  select(c("Strain", "SampleID", "WL", "Run",  "Par_ue",  "PurData", "ObsDate", "E_days", "PurParRatio", "PURPhotonDose_day")) %>% 
 group_by(Strain, WL, SampleID, Run, Par_ue) %>% 
  summarise(ObsDate, E_days, PurData, PurParRatio, PURPhotonDose_day) %>% 
ungroup

glimpse(PurData)

PURPARFits <- right_join(PurData, NestedFitsData,  by = c("Strain", "SampleID", "WL", "Run",  "Par_ue"), keep = NULL)

glimpse(PURPARFits)

PurDataSum <- NestedPURData %>% 
  select(c("Strain", "WL", "Par_ue", "Pur_ue")) %>% 
 group_by(Strain, WL, Par_ue, Pur_ue) %>% 
  summarise() %>% 
ungroup

glimpse(PurDataSum)

PURPARFits <- left_join(PURPARFits, PurDataSum, by = c("Strain", "WL", "Par_ue"))

glimpse(PURPARFits)

remove(PurDataSum)
remove(PurData)
```
# mutate PurParRatio to get ratios for all growth estimates as well as PURPhotonDose_day
```{r purparratio}
PURPARFits <- PURPARFits %>% 
  mutate(PurParRatio = Pur_ue / Par_ue,
         PURPhotonDose_day = case_when(LightShape == 'Sine' ~ Pur_ue/2 * SectoHour * Photoperiod,
                                  LightShape == 'Square' ~ Pur_ue * SectoHour * Photoperiod))

glimpse (PURPARFits)
```

```{r save Growth estimates with PUR and PAR combined }
saveRDS(object = PURPARFits, file = file.path(DataOut, paste(Project, "Merged_GrowthFitsPURPAR.Rds",  sep = "_")), ascii = FALSE, version = NULL, compress = "xz", refhook = NULL)
```