### 2024-01-17 Code/ManuscriptFigures.Rmd
Fixed order of PAR to increase from 30, 90, 180 uE when faceted by row

---
title: "ManuscriptFigures"
author: 
- Mireille Savoie
- Douglas A. Campbell
date: "`r format(Sys.Date())`"
output:
  bookdown::word_document2:
    code_folding: show
    keep_md: yes
    toc: TRUE
    toc_depth: 6
    fig_caption: yes
    pandoc_args: 
      - '--lua-filter=scholarly-metadata.lua'
      - '--lua-filter=author-info-blocks.lua'
  bookdown::html_document2:
    code_folding: show
    keep_md: yes
    toc: TRUE
    toc_float: TRUE
    toc_depth: 6
    fig_caption: yes
    pandoc_args: 
      - '--lua-filter=scholarly-metadata.lua'
      - '--lua-filter=author-info-blocks.lua'
bibliography: [Prochlorococcus_O2_NPQ.bib, RPackages.bib]
csl: plos-one.csl
editor_options: 
  markdown: 
    wrap: 72
---

## Set figure caption font size
```{css, echo=FALSE}
p.caption {
  font-size: 25px;
}
```
# Global chunk settings 
```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = F)
knitr::opts_chunk$set(fig.path='Figs/')
```

```{r load libraries, include=FALSE}

library(lubridate)
# library(stringr)
library(broom)
library(knitr)
# library(zoo)
library(RColorBrewer)
library(grid)
library(bookdown)
library(glue)
library(egg) # Adds Letters to grid plots
# library(ggspectra)
library(photobiologyWavebands)
# library(photobiology)
# library(ggpubr)
library(ggh4x)
library(cowplot)
library(tidyverse)

knitr::write_bib(c("tidyverse", "lubridate","knitr", "stringr", "broom", "zoo", "RColorBrewer", "grid", "glue", "bookdown","egg"), 'RPackages.bib')

```

```{r set project variables}
#"..", takes up a level in the directory path
Project <- "PICO"
DataIn <- file.path("..","Data", "CleanData")
PlotsPath <- file.path("..","Output", "Figures")
# DataOut <- file.path("..","GrowthAssessData")

```

# Read data
```{r read ProcessFile}

PURPARFits <- readRDS(file = file.path(DataIn,"PICO_Merged_GrowthFitsPURPAR.Rds"))

glimpse(PURPARFits)

```

# Make O2_uM and WL as factor levels from highest to lowest
```{r O2 levels}

PURPARFits <- PURPARFits %>% 
  mutate(O2_uM = factor(O2_uM, levels = c(250, 25, 2.5)),
         WL = factor(WL, levels = c("WW", 660, 530, 450)),
         deltaOD_Lmu_corr_day = deltaOD_Lmu_corr * 24,
         deltaOD_Lmu_se_day = deltaOD_Lmu_se * 24)
#  Par_ue = factor(Par_ue, levels =  c(180, 120, 100,90,60, 30, 15))
```

```{r set colours}

MyWavelengths = c("WW", 660, 530, 450)

MCMIXColours = c("black", w_length2rgb(660), w_length2rgb(530),  w_length2rgb(450))

names(MCMIXColours) <- MyWavelengths
MCMIXColours

ColoursStrain = c(MIT9313 = "#E69F00", SS120 = "#F0E442", MED4 = "#CC79A7" )

MyOxygen = c(2.5, 25, 250 )
# O2Colours = c("#99CCFF", "dodgerblue", "blue1")
O2Colours = c("#FFCC00", "#FF9900", "#663300")
names(O2Colours) <- MyOxygen
O2Colours

MyOxygen = c(2.5, 25, 250 )
O2LineType = c("dotted", "longdash", "solid")
names(O2LineType) <- MyOxygen
O2LineType

MyPhotoperiods = as.factor(c(4, 8, 12, 16))
ColoursPhotoperiod = c("gray88", "lightcyan3", "gray47", "gray8")
names(ColoursPhotoperiod) <- MyPhotoperiods
ColoursPhotoperiod

```

```{r set shapes}

MyPhotoperiods = as.factor(c(4, 8, 12, 16))
PhotoperiodShapes = c(6, 18, 16, 17)

names(PhotoperiodShapes) <- MyPhotoperiods
PhotoperiodShapes

O2Shapes <- c(13, 1, 19)

```

# Set Strip colors for WL x_strip is for WL faceted by columns and y_strip is for WL faceted by rows
```{r}
# x_strip <- strip_themed(background_x = elem_list_rect(fill = MCMIXColours, alpha = 0.8),
#                       background_y = elem_list_rect(fill = "grey27")) 
# 
# x_stripWW <- strip_themed(background_x = elem_list_rect(fill = "black"),
#                       background_y = elem_list_rect(fill = "#grey27"))
# 
# x_strip660 <- strip_themed(background_x = elem_list_rect(fill = "#730000"),
#                       background_y = elem_list_rect(fill = "#grey27"))
# 
# x_strip530 <- strip_themed(background_x = elem_list_rect(fill = "#00FF00"),
#                       background_y = elem_list_rect(fill = "#grey27"))
# 
# x_strip450 <- strip_themed(background_x = elem_list_rect(fill = "#0600FF"),
#                       background_y = elem_list_rect(fill = "#grey27"))
# 
# y_strip <- strip_themed(background_y = elem_list_rect(fill = MCMIXColours),
#                       background_x = elem_list_rect(fill = "#grey27")) 
# 
# y_stripWW <- strip_themed(background_y = elem_list_rect(fill = "black"),
#                       background_x = elem_list_rect(fill = "#grey27"))
# 
# y_strip660 <- strip_themed(background_y = elem_list_rect(fill = "#730000"),
#                       background_x = elem_list_rect(fill = "#grey27"))
# 
# y_strip530 <- strip_themed(background_y = elem_list_rect(fill = "#00FF00"),
#                       background_x = elem_list_rect(fill = "#grey27"))
# 
# y_strip450 <- strip_themed(background_y = elem_list_rect(fill = "#0600FF"),
#                       background_x = elem_list_rect(fill = "#grey27"))

```

# Remove redundant columns 
```{r}
PURPARFits <- PURPARFits %>% 
  select(!contains(c("trunc", "OD720", "OptodeMeasure"))) %>% 
  filter(Strain %in% c("MED4", "MIT9313", "SS120"),
         WL %in% c("WW", 450, 660))

glimpse(PURPARFits)
```

```{r PURPARRatioPlots, fig.height= 10, fig.width= 10}

PurParPlot <- PURPARFits %>%
  filter(Par_ue %in% c(30, 90, 180)) %>%
  mutate(Strain = factor(Strain, levels = c("MED4", "SS120", "MIT9313"))) %>% 
  ggplot() +
  geom_point(aes(x = Par_ue, y = Pur_ue, colour = as.factor(WL)), size = 3, alpha = 0.5) +
 geom_line(aes(x = Par_ue, y = Pur_ue,colour = as.factor(WL)), linewidth = 0.7) +
    geom_abline(slope = 1, intercept = 0, colour = "grey39", linetype = "dashed", linewidth = 1.2) +
  labs(y = "PUR (µE)" , x = "PAR (µE)") +
   scale_x_continuous(breaks=seq(0, 180, by = 30)) +
  # coord_cartesian(xlim = c (0, 180)) +
    scale_y_continuous(breaks=seq(0, 180, by = 30)) +
  coord_fixed(ylim = c (0, 180), xlim = c (0, 180), ratio = 1) +
   scale_colour_manual(values = MCMIXColours, labels = c("White LED", "660","530", "450" )) +
   geom_text(aes(x = 60, y = 180, label = paste(Strain), show.legend = FALSE, ),    # parse = TRUE,
    hjust = 0,
    vjust = 1,
    size = 8.0
  )  +
  facet_wrap(~Strain, ncol = 2) +
  theme_bw() +
  labs(colour = "Growth Waveband (nm)") +
  theme(strip.text = element_blank(),
        legend.text = element_text(size = 20),
        legend.title = element_text(size = 25),
        legend.position = c(0.80, 0.25),
        axis.title = element_text(hjust= 0.60, vjust = -0.2, size = 20),
        strip.text.x = element_blank(),
        axis.text = element_text(size = 15, color = "black"),)

PurParPlot <- tag_facet(PurParPlot, open = "", close = "", tag_pool = LETTERS, size = 8)

PurParPlot

```

```{r save ParPurPlots}

ggsave(file = file.path(PlotsPath, paste("PurParPlot",".png",sep = "")), plot = PurParPlot, device = NULL, scale = 1, height=20, width= 35, units = c("cm"),dpi = 300, limitsize = TRUE)

```

```{r PUR per day, fig.height= 10, fig.width= 10 }

scientific_10 <- function(x) {   parse(text=gsub("e\\+*", " %*% 10^", scales::scientific_format()(x))) }

# MED4NoPhotoPUR <- PURPARFits %>%
PURPARFits %>%
   filter(Strain == "MED4", 
         Par_ue %in% c(30,90,180)) %>%
   mutate(PhotoperiodLabel_title = "Photoperiod (h)",
         OxygenLabel = "Oxygen Concentration (µM)",
         Photoperiod = factor(Photoperiod, levels = c("16", "12", "8", "4" ))) %>%
  ggplot() +
  geom_point(aes(x = PURPhotonDose_day, y = ReplicateMean_deltaOD_Lmu_corr_day, shape = as.factor(Photoperiod), color = as.factor(WL), fill = as.factor(WL)), size = 6, alpha = 1.0) +
    geom_point(aes(x = PURPhotonDose_day, y = deltaOD_Lmu_corr_day, shape = as.factor(Photoperiod), color = as.factor(WL), fill = as.factor(WL)), size = 1.5, alpha = 0.8) +
 labs( y = "Chlorophyll Specific Growth Rate " ~  " ("  ~ d^-1~")", x = "PUR (µmol photons" ~m^-2 ~d^-1~")") +
scale_x_continuous(label=scientific_10) +
   scale_y_continuous(breaks = c(0,0.2, 0.4, 0.6, 0.8, 1.0), limits = c (0, 1.05)) +
  # coord_cartesian(ylim = c (0, 0.07)) +
  scale_colour_manual(values = MCMIXColours, labels = c("White LED", "660","530", "450" )) +
   scale_shape_manual(values = PhotoperiodShapes)  +
  ggh4x::facet_nested(rows = vars(OxygenLabel, O2_uM), 
                    labeller = labeller(Strain = label_both, WL = label_both, O2_uM = label_value, Photoperiod = label_value, PhotoperiodLabel_title = label_value, OxygenLabel = label_value)) +
 guides(linetype = "none",
        fill = "none",
        colour = guide_legend(ncol = 4)) +
 guides(shape = guide_legend(nrow = 1)) +
  theme_bw() +
   theme( strip.text = element_text(size = 30),
        axis.title = element_text(size = 30, color = "black"),
        axis.text.x = element_text(size = 20, color = "black", angle = 45, hjust = 1.0),
        axis.text.y = element_text(size = 20, color = "black", angle = 0, hjust = 1.0),
        strip.background=element_rect(color="grey87", fill = "grey82"),
        # strip.background=element_rect(fill = "white"),
        legend.key.size = unit(1, 'cm'), #change legend key size
        legend.key.height = unit(1, 'cm'), #change legend key height
        legend.key.width = unit(0.5, 'cm'), #change legend key width
        legend.title =  element_text(size = 20),  #change legend title font size
        legend.text = element_text(size = 20),   #change legend text font size
        legend.position = "top") +
  labs(color = "Growth Waveband (nm)",
       shape = "Photoperiod (h)")

# MED4NoPhotoPUR

# SS120NoPhotoPUR <- PURPARFits %>%
PURPARFits %>%
   filter(Strain == "SS120", 
           Par_ue %in% c(30,90,180)) %>%
   mutate(PhotoperiodLabel_title = "Photoperiod (h)",
         OxygenLabel = "Oxygen Concentration (µM)",
         Photoperiod = factor(Photoperiod, levels = c("16", "12", "8", "4" ))) %>%
  ggplot() +
  geom_point(aes(x = PURPhotonDose_day, y = ReplicateMean_deltaOD_Lmu_corr_day, shape = as.factor(Photoperiod), color = as.factor(WL), fill = as.factor(WL)), size = 6, alpha = 1.0) +
    geom_point(aes(x = PURPhotonDose_day, y = deltaOD_Lmu_corr_day, shape = as.factor(Photoperiod), color = as.factor(WL), fill = as.factor(WL)), size = 1.5, alpha = 0.8) +
 labs( y = "Chlorophyll Specific Growth Rate " ~  " ("  ~ d^-1~")", x = "PUR (µmol photons" ~m^-2 ~d^-1~")") +
scale_x_continuous(label=scientific_10) +
   scale_y_continuous(breaks = c(0,0.2, 0.4, 0.6, 0.8, 1.0), limits = c (0, 1.05)) +
  # coord_cartesian(ylim = c (0, 0.07)) +
  scale_colour_manual(values = MCMIXColours, labels = c("White LED", "660","530", "450" )) +
   scale_shape_manual(values = PhotoperiodShapes)  +
  ggh4x::facet_nested(rows = vars(OxygenLabel, O2_uM), 
                    labeller = labeller(Strain = label_both, WL = label_both, O2_uM = label_value, Photoperiod = label_value, PhotoperiodLabel_title = label_value, OxygenLabel = label_value)) +
 guides(linetype = "none",
        fill = "none",
        colour = guide_legend(ncol = 4)) +
 guides(shape = guide_legend(nrow = 1)) +
  theme_bw() +
   theme( strip.text = element_text(size = 30),
        axis.title = element_text(size = 30, color = "black"),
        axis.text.x = element_text(size = 20, color = "black", angle = 45, hjust = 1.0),
        axis.text.y = element_text(size = 20, color = "black", angle = 0, hjust = 1.0),
        strip.background=element_rect(color="grey87", fill = "grey82"),
        # strip.background=element_rect(fill = "white"),
        legend.key.size = unit(1, 'cm'), #change legend key size
        legend.key.height = unit(1, 'cm'), #change legend key height
        legend.key.width = unit(0.5, 'cm'), #change legend key width
        legend.title =  element_text(size = 20),  #change legend title font size
        legend.text = element_text(size = 20),   #change legend text font size
        legend.position = "top") +
  labs(color = "Growth Waveband (nm)",
       shape = "Photoperiod (h)")

# SS120NoPhotoPUR

# MIT9313NoPhotoPUR <- PURPARFits %>%
   PURPARFits %>%
     filter(Strain == "MIT9313", 
          Par_ue %in% c(30,90,180)) %>%
  mutate(PhotoperiodLabel_title = "Photoperiod (h)",
         OxygenLabel = "Oxygen Concentration (µM)",
         Photoperiod = factor(Photoperiod, levels = c("16", "12", "8", "4" ))) %>%
  ggplot() +
  geom_point(aes(x = PURPhotonDose_day, y = ReplicateMean_deltaOD_Lmu_corr_day, shape = as.factor(Photoperiod), color = as.factor(WL), fill = as.factor(WL)), size = 6, alpha = 1.0) +
    geom_point(aes(x = PURPhotonDose_day, y = deltaOD_Lmu_corr_day, shape = as.factor(Photoperiod), color = as.factor(WL), fill = as.factor(WL)), size = 1.5, alpha = 0.8) +
 # geom_line(aes(x = PURPhotonDose_day, y = ReplicateMean_deltaOD_Lmu_corr_day, shape = as.factor(Photoperiod), linetype = as.factor(WL)), size = 0.5) +
 labs( y = "Chlorophyll Specific Growth Rate " ~  " ("  ~ d^-1~")", x = "PUR (µmol photons" ~m^-2 ~d^-1~")") +
scale_x_continuous(label=scientific_10) +
   scale_y_continuous(breaks = c(0,0.2, 0.4, 0.6, 0.8, 1.0), limits = c (0, 1.05)) +
  # coord_cartesian(ylim = c (0, 0.07)) +
  scale_colour_manual(values = MCMIXColours, labels = c("White LED", "660","530", "450" )) +
   scale_shape_manual(values = PhotoperiodShapes)  +
  ggh4x::facet_nested(rows = vars(OxygenLabel, O2_uM), 
                    labeller = labeller(Strain = label_both, WL = label_both, O2_uM = label_value, Photoperiod = label_value, PhotoperiodLabel_title = label_value, OxygenLabel = label_value)) +
 guides(linetype = "none",
        fill = "none",
        colour = guide_legend(ncol = 4)) +
 guides(shape = guide_legend(nrow = 1)) +
  theme_bw() +
   theme( strip.text = element_text(size = 30),
        axis.title = element_text(size = 30, color = "black"),
        axis.text.x = element_text(size = 20, color = "black", angle = 45, hjust = 1.0),
        axis.text.y = element_text(size = 20, color = "black", angle = 0, hjust = 1.0),
        strip.background=element_rect(color="grey87", fill = "grey82"),
        # strip.background=element_rect(fill = "white"),
        legend.key.size = unit(1, 'cm'), #change legend key size
        legend.key.height = unit(1, 'cm'), #change legend key height
        legend.key.width = unit(0.5, 'cm'), #change legend key width
        legend.title =  element_text(size = 20),  #change legend title font size
        legend.text = element_text(size = 20),   #change legend text font size
        legend.position = "top") +
  labs(color = "Growth Waveband (nm)",
       shape = "Photoperiod (h)")

# MIT9313NoPhotoPUR

```

```{r, save O2ColorPURPlots}

# ggsave(file = file.path(PlotsPath, paste("MED4NoPhotoPUR",".png",sep = "")), plot = MED4NoPhotoPUR, device = NULL, scale = 1, height=25, width= 40, units = c("cm"),dpi = 300, limitsize = TRUE)
# 
# ggsave(file = file.path(PlotsPath, paste("MIT9313NoPhotoPUR",".png",sep = "")), plot = MIT9313NoPhotoPUR, device = NULL, scale = 1, height=25, width= 40, units = c("cm"),dpi = 300, limitsize = TRUE)
# 
# ggsave(file = file.path(PlotsPath, paste("SS120NoPhotoPUR",".png",sep = "")), plot = SS120NoPhotoPUR, device = NULL, scale = 1, height=25, width= 40, units = c("cm"),dpi = 300, limitsize = TRUE)

```

# PAR plots
```{r ParPlots, fig.height= 10, fig.width= 10}

MED4PAR <- PURPARFits %>%
   filter(Strain == "MED4",
           Par_ue %in% c(30,90,180),
          deltaOD_Lmu_corr == 0 |
          deltaOD_Lmu_corr != 0 & deltaOD_LseGrowthFlag == 1) %>%
  mutate(PhotoperiodLabel_title = "Photoperiod~(h)",
         OxygenLabel = "Oxygen~Concentration~(µM)",
         WL_label_title = "Growth~Wavelength",
         ParLabel = "PAR~(µmol~photons~m^{-2}~s^{-1})",
         Par_ue = factor(Par_ue, levels = c(30, 90, 180))) %>%
  ggplot() +
  geom_point(aes(x = Photoperiod, y = ReplicateMean_deltaOD_Lmu_corr_day, colour = as.factor(WL)), size = 5, alpha = 0.5) +
 geom_line(aes(x = Photoperiod, y = ReplicateMean_deltaOD_Lmu_corr_day, colour = as.factor(WL), linetype = as.factor(WL)), size = 0.5) +
    geom_point(aes(x = Photoperiod, y = (deltaOD_Lmu_corr_day), colour = as.factor(WL)), size = 2.5, alpha = 0.5) +
 labs( y = "Chlorophyll Specific Growth Rate " ~  " ("  ~ d^-1~")", x = "Photoperiod (h)") +
   geom_jitter(aes(x = Photoperiod, y = ReplicateMean_deltaOD_Lmu_corr_day, colour = as.factor(WL)), width = 0.1, size = 3) +
  scale_y_continuous(breaks = c(0,0.2, 0.4, 0.6, 0.8, 1.0), limits = c (0, 1.05)) +
  # coord_cartesian(ylim = c (0, 0.07)) +
  scale_colour_manual(values = MCMIXColours, labels = c("White LED", "660 nm","530 nm", "450 nm" )) +
   scale_x_continuous(breaks = c(4,8, 12, 16), limits = c (4, 16)) +
  ggh4x::facet_nested(cols = vars(ParLabel, Par_ue), rows = vars(OxygenLabel, O2_uM), 
                    labeller = label_parsed) + 
 guides(linetype = "none") +
   theme_bw() +
   theme( strip.text = element_text(size = 20),
       axis.title.x.bottom = element_text(size = 30, color = "black", vjust = 1.0),
      axis.title.y = element_text(size = 30, color = "black", vjust = 1.0),
       axis.title.x.top = element_text(size = 25, color = "black", vjust = 1.0),
       axis.text.x = element_text(size = 20, color = "black", angle = 0, hjust = 0.5, vjust = 1),
        axis.text.y = element_text(size = 20, color = "black", angle = 0, hjust = 1.0),
       axis.text.x.top = element_blank(),
       axis.ticks.x.top = element_blank(),
        # strip.background=element_rect(color="grey87", fill = "grey82"),
        # strip.background=element_rect(fill = "white"),
        legend.key.size = unit(1, 'cm'), #change legend key size
        legend.key.height = unit(1, 'cm'), #change legend key height
        legend.key.width = unit(0.5, 'cm'), #change legend key width
        # legend.title = element_text(size = 15),  #change legend title font size for right side legend placement
        # legend.text = element_text(size = 15),   #change legend text font size for right side legend placement
        legend.title = element_text(size = 25),  #change legend title font size
        legend.text = element_text(size = 25),   #change legend text font size
        legend.position = "top") +
  labs(color = "Growth Wavelength",
       fill = "Growth Wavelength")
MED4PAR

SS120PAR <- PURPARFits %>%
   filter(Strain == "SS120",
          Par_ue %in% c(30,90,180),
          deltaOD_Lmu_corr == 0 |
          deltaOD_Lmu_corr != 0 & deltaOD_LseGrowthFlag == 1) %>%
  mutate(PhotoperiodLabel_title = "Photoperiod~(h)",
         OxygenLabel = "Oxygen~Concentration~(µM)",
         WL_label_title = "Growth~Wavelength",
         ParLabel = "PAR~(µmol~photons~m^{-2}~s^{-1})",
         Par_ue = factor(Par_ue, levels = c(30, 90, 180))) %>%
  ggplot() +
  geom_point(aes(x = Photoperiod, y = ReplicateMean_deltaOD_Lmu_corr_day, colour = as.factor(WL)), size = 5, alpha = 0.5) +
 geom_line(aes(x = Photoperiod, y = ReplicateMean_deltaOD_Lmu_corr_day, colour = as.factor(WL), linetype = as.factor(WL)), size = 0.5) +
    geom_point(aes(x = Photoperiod, y = (deltaOD_Lmu_corr_day), colour = as.factor(WL)), size = 2.5, alpha = 0.5) +
 labs( y = "Chlorophyll Specific Growth Rate " ~  " ("  ~ d^-1~")", x = "Photoperiod (h)") +
   geom_jitter(aes(x = Photoperiod, y = ReplicateMean_deltaOD_Lmu_corr_day, colour = as.factor(WL)), width = 0.1, size = 3) +
  scale_y_continuous(breaks = c(0,0.2, 0.4, 0.6, 0.8, 1.0), limits = c (0, 1.05)) +
  # coord_cartesian(ylim = c (0, 0.07)) +
  scale_colour_manual(values = MCMIXColours, labels = c("White LED", "660 nm","530 nm", "450 nm" )) +
   scale_x_continuous(breaks = c(4,8, 12, 16), limits = c (4, 16)) +
  ggh4x::facet_nested(cols = vars(ParLabel, Par_ue), rows = vars(OxygenLabel, O2_uM), 
                    labeller = label_parsed) + 
 guides(linetype = "none") +
   theme_bw() +
   theme( strip.text = element_text(size = 20),
       axis.title.x.bottom = element_text(size = 30, color = "black", vjust = 1.0),
      axis.title.y = element_text(size = 30, color = "black", vjust = 1.0),
       axis.title.x.top = element_text(size = 25, color = "black", vjust = 1.0),
       axis.text.x = element_text(size = 20, color = "black", angle = 0, hjust = 0.5, vjust = 1),
        axis.text.y = element_text(size = 20, color = "black", angle = 0, hjust = 1.0),
       axis.text.x.top = element_blank(),
       axis.ticks.x.top = element_blank(),
        # strip.background=element_rect(color="grey87", fill = "grey82"),
        # strip.background=element_rect(fill = "white"),
        legend.key.size = unit(1, 'cm'), #change legend key size
        legend.key.height = unit(1, 'cm'), #change legend key height
        legend.key.width = unit(0.5, 'cm'), #change legend key width
        # legend.title = element_text(size = 15),  #change legend title font size for right side legend placement
        # legend.text = element_text(size = 15),   #change legend text font size for right side legend placement
        legend.title = element_text(size = 25),  #change legend title font size
        legend.text = element_text(size = 25),   #change legend text font size
        legend.position = "top") +
  labs(color = "Growth Wavelength",
       fill = "Growth Wavelength")

SS120PAR

MIT9313PAR <- PURPARFits %>%
   filter(Strain == "MIT9313",
            Par_ue %in% c(30,90,180),
          deltaOD_Lmu_corr == 0 |
          deltaOD_Lmu_corr != 0 & deltaOD_LseGrowthFlag == 1) %>%
   mutate(PhotoperiodLabel_title = "Photoperiod~(h)",
         OxygenLabel = "Oxygen~Concentration~(µM)",
         WL_label_title = "Growth~Wavelength",
         ParLabel = "PAR~(µmol~photons~m^{-2}~s^{-1})",
         Par_ue = factor(Par_ue, levels = c(30, 90, 180))) %>%
  ggplot() +
  geom_point(aes(x = Photoperiod, y = ReplicateMean_deltaOD_Lmu_corr_day, colour = as.factor(WL)), size = 5, alpha = 0.5) +
 geom_line(aes(x = Photoperiod, y = ReplicateMean_deltaOD_Lmu_corr_day, colour = as.factor(WL), linetype = as.factor(WL)), size = 0.5) +
    geom_point(aes(x = Photoperiod, y = (deltaOD_Lmu_corr_day), colour = as.factor(WL)), size = 2.5, alpha = 0.5) +
 labs( y = "Chlorophyll Specific Growth Rate " ~  " ("  ~ d^-1~")", x = "Photoperiod (h)") +
   geom_jitter(aes(x = Photoperiod, y = ReplicateMean_deltaOD_Lmu_corr_day, colour = as.factor(WL)), width = 0.1, size = 3) +
  scale_y_continuous(breaks = c(0,0.2, 0.4, 0.6, 0.8, 1.0), limits = c (0, 1.05)) +
  # coord_cartesian(ylim = c (0, 0.07)) +
  scale_colour_manual(values = MCMIXColours, labels = c("White LED", "660 nm","530 nm", "450 nm" )) +
   scale_x_continuous(breaks = c(4,8, 12, 16), limits = c (4, 16)) +
  ggh4x::facet_nested(cols = vars(ParLabel, Par_ue), rows = vars(OxygenLabel, O2_uM), 
                    labeller = label_parsed) + 
 guides(linetype = "none") +
   theme_bw() +
   theme( strip.text = element_text(size = 20),
       axis.title.x.bottom = element_text(size = 30, color = "black", vjust = 1.0),
      axis.title.y = element_text(size = 30, color = "black", vjust = 1.0),
       axis.title.x.top = element_text(size = 25, color = "black", vjust = 1.0),
       axis.text.x = element_text(size = 20, color = "black", angle = 0, hjust = 0.5, vjust = 1),
        axis.text.y = element_text(size = 20, color = "black", angle = 0, hjust = 1.0),
       axis.text.x.top = element_blank(),
       axis.ticks.x.top = element_blank(),
        # strip.background=element_rect(color="grey87", fill = "grey82"),
        # strip.background=element_rect(fill = "white"),
        legend.key.size = unit(1, 'cm'), #change legend key size
        legend.key.height = unit(1, 'cm'), #change legend key height
        legend.key.width = unit(0.5, 'cm'), #change legend key width
        # legend.title = element_text(size = 15),  #change legend title font size for right side legend placement
        # legend.text = element_text(size = 15),   #change legend text font size for right side legend placement
        legend.title = element_text(size = 25),  #change legend title font size
        legend.text = element_text(size = 25),   #change legend text font size
        legend.position = "top") +
  labs(color = "Growth Wavelength",
       fill = "Growth Wavelength")

MIT9313PAR
```

```{r, save PARPlots}

ggsave(file = file.path(PlotsPath, paste("MED4PAR",".png",sep = "")), plot = MED4PAR, device = NULL, scale = 1, height=25, width= 35, units = c("cm"),dpi = 300, limitsize = TRUE)

ggsave(file = file.path(PlotsPath, paste("SS120PAR",".png",sep = "")), plot = SS120PAR, device = NULL, scale = 1, height=25, width= 35, units = c("cm"),dpi = 300, limitsize = TRUE)

ggsave(file = file.path(PlotsPath, paste("MIT9313PAR",".png",sep = "")), plot = MIT9313PAR, device = NULL, scale = 1, height=25, width= 35, units = c("cm"),dpi = 300, limitsize = TRUE)

```

# Overlay whole cell absorbance with emission spectra
```{r MED4 All Spectral Emission,  fig.height = 10, fig.width = 10}

MyWavelengths = c("WW", 660, 530, 450)
OverlayColours = c("black", w_length2rgb(660), w_length2rgb(530),  w_length2rgb(450))

names(OverlayColours) <- MyWavelengths
OverlayColours

MED4OverlayPlots <- PURPARFits %>%
  filter(Strain == "MED4",
          Par_ue == 90,
          O2_Category == "High" & WL == 450 & Photoperiod == 12 |
          O2_Category == "High" & WL == 530 & Photoperiod == 16 |
          O2_Category == "Intermediate" & WL == "WW" & Photoperiod == 16 |
          O2_Category == "Intermediate" & WL == 660 & Photoperiod == 12) %>%
  mutate(WL_label = case_when(WL %in% c(450, 530, 660) ~ paste("Growth", "waveband", "=", WL, "nm", sep = " "),
        WL %in% c("WW") ~ paste("Growth", "waveband", "=", "White LED", sep = " "))) %>% 
   unnest(cols = c(PurData), names_sep = "_", keep_empty = TRUE) %>%
  pivot_longer(cols = c(PurData_PURAbsWholeCell_Norm440, PurData_AbsNorm440,  PurData_MC_NormEm), values_to = "Abs_Em_Value", names_to = "Abs_Em_Names") %>%
ggplot()+
  geom_area(aes(x = PurData_OlisAbs_nm, y = Abs_Em_Value, fill = as.factor(WL)), color = "black", 
            size = 0.5, alpha = 0.7, show.legend = FALSE, 
            data = . %>% filter(Abs_Em_Names == "PurData_PURAbsWholeCell_Norm440")) +
  geom_line(aes(x = PurData_OlisAbs_nm, y = Abs_Em_Value, color = Abs_Em_Names,  linetype = Abs_Em_Names), size = 0.85) +
  geom_line(aes(x = PurData_OlisAbs_nm, y = Abs_Em_Value, color = as.factor(WL),  linetype = Abs_Em_Names), size = 1.0, data = . %>% filter(Abs_Em_Names == "PurData_MC_NormEm")) +
  scale_color_manual(values = c(MCMIXColours, "PurData_PURAbsWholeCell_Norm440" = "black", "PurData_AbsNorm440" = "darkorchid3", "PurData_MC_NormEm" = "white")) +
  scale_linetype_manual(values = c("PurData_PURAbsWholeCell_Norm440" = "solid", "PurData_AbsNorm440" = "twodash",  "PurData_MC_NormEm" = "dotted"),
                        labels = c("LED emission", "Whole cell abs.", "PUR")) +
  scale_fill_manual(values = MCMIXColours) +
 # guides(colour = FALSE) +
  labs(y = "Normalized Absorbance and Emission", x = "Wavelength (nm)") +
  labs(linetype = "Spectra Type")+
  scale_y_continuous(breaks=seq(0, 1.5, by = 0.25)) +
 # coord_cartesian(ylim = c (0, 0.2)) +
   facet_wrap(vars(WL),
              labeller = labeller(strain = label_value, WL = label_value, Photoperiod = label_both, Par_ue = label_both, O2_Category = label_both, Pur_ue = label_both)) +
  geom_text(aes(x = 430, y = 1.1, label = WL_label, color = as.factor(WL)), show.legend = FALSE,
    # parse = TRUE,
    hjust = 0,
    vjust = 1,
    size = 5.0
  )  +
   geom_text(aes(x = 560, y = 0.95, label = paste("PAR", "=", Par_ue, sep = " "), show.legend = FALSE, ),
    # parse = TRUE,
    hjust = 0,
    vjust = 1,
    size = 5.0
  )  +
  geom_text(aes(x = 560, y = 0.85, label = paste("PUR", "=", Pur_ue,sep = " "), show.legend = FALSE, ),
    # parse = TRUE,
    hjust = 0,
    vjust = 1,
    size = 5.0
  )  +
  guides(color = "none") +
  theme_bw() +
 theme(strip.text = element_blank(),
       legend.background = element_rect(color = "transparent", fill = "transparent"),
        legend.text = element_text(size = 18),
        legend.title = element_text(size = 20),
        # legend.position = c(0.94,0.96),
        legend.position = "top",
        axis.title = element_text(vjust = 1.0, size = 20),
        # strip.text.x = element_text(size = 20),
        axis.text = element_text(size = 15, color = "black"))

MED4OverlayPlots <- tag_facet(MED4OverlayPlots, open = "", close = "", tag_pool = LETTERS, size = 8)

MED4OverlayPlots

```

```{r save MED4 PUR overlay plots}

# ggsave(file = file.path(PlotsPath, paste("MED4OverlayPlots",".png",sep = "")), plot = MED4OverlayPlots, device = NULL, scale = 1, height=20, width= 35, units = c("cm"),dpi = 300, limitsize = TRUE)
# 
# ggsave(file = file.path(PlotsPath, paste("MED4WWOverlayPlots",".png",sep = "")), plot = MED4WWOverlayPlots, device = NULL, scale = 1, height=25, width= 35, units = c("cm"),dpi = 300, limitsize = TRUE)

```

```{r MIT9313 All Spectral Emission,  fig.height = 10, fig.width = 10}

# MIT9313OverlayPlots <- PURPARFits %>%
PURPARFits %>%
   filter(Strain == "MIT9313",
          O2_Category == "High",
          Par_ue == 30,
          Photoperiod == 12 & WL == 450 |
          Photoperiod == 12 & WL == 530 |
          Photoperiod == 12 & WL == 660 |
          Photoperiod == 16 & WL == "WW") %>% 
  mutate(WL_label = case_when(WL %in% c(450, 530, 660) ~ paste("Growth", "wavelength", "=", WL, "nm", sep = " "),
        WL %in% c("WW") ~ paste("Growth", "wavelength", "=", "White LED", sep = " "))) %>% 
   unnest(cols = c(PurData), names_sep = "_", keep_empty = TRUE) %>%
  pivot_longer(cols = c(PurData_PURAbsWholeCell_Norm440, PurData_AbsNorm440,  PurData_MC_NormEm), values_to = "Abs_Em_Value", names_to = "Abs_Em_Names") %>%
ggplot()+
  geom_area(aes(x = PurData_OlisAbs_nm, y = Abs_Em_Value, fill = as.factor(WL)), color = "black", 
            size = 0.5, alpha = 0.7, show.legend = FALSE, 
            data = . %>% filter(Abs_Em_Names == "PurData_PURAbsWholeCell_Norm440")) +
  geom_line(aes(x = PurData_OlisAbs_nm, y = Abs_Em_Value, color = Abs_Em_Names,  linetype = Abs_Em_Names), size = 0.85) +
  geom_line(aes(x = PurData_OlisAbs_nm, y = Abs_Em_Value, color = as.factor(WL),  linetype = Abs_Em_Names), size = 1.0, data = . %>% filter(Abs_Em_Names == "PurData_MC_NormEm")) +
  scale_color_manual(values = c(MCMIXColours, "PurData_PURAbsWholeCell_Norm440" = "black", "PurData_AbsNorm440" = "darkorchid3", "PurData_MC_NormEm" = "white")) +
  scale_linetype_manual(values = c("PurData_PURAbsWholeCell_Norm440" = "solid", "PurData_AbsNorm440" = "twodash",  "PurData_MC_NormEm" = "dotted"),
                        labels = c("LED emission", "Whole cell abs.", "PUR")) +
  scale_fill_manual(values = MCMIXColours) +
 # guides(colour = FALSE) +
  labs(y = "Normalized Absorbance and Emission", x = "Wavelength (nm)") +
  labs(linetype = "Spectra Type")+
  scale_y_continuous(breaks=seq(0, 1.5, by = 0.25)) +
 # coord_cartesian(ylim = c (0, 0.2)) +
   facet_wrap(vars(WL),
              labeller = labeller(strain = label_value, WL = label_value, Photoperiod = label_both, Par_ue = label_both, O2_Category = label_both, Pur_ue = label_both)) +
  geom_text(aes(x = 430, y = 1.1, label = WL_label, color = as.factor(WL)), show.legend = FALSE,
    # parse = TRUE,
    hjust = 0,
    vjust = 1,
    size = 5.0
  )  +
   geom_text(aes(x = 560, y = 0.95, label = paste("PAR", "=", Par_ue, sep = " "), show.legend = FALSE, ),
    # parse = TRUE,
    hjust = 0,
    vjust = 1,
    size = 5.0
  )  +
  geom_text(aes(x = 560, y = 0.85, label = paste("PUR", "=", Pur_ue,sep = " "), show.legend = FALSE, ),
    # parse = TRUE,
    hjust = 0,
    vjust = 1,
    size = 5.0
  )  +
  guides(color = "none") +
  theme_bw() +
 theme(strip.text = element_blank(),
       legend.background = element_rect(color = "transparent", fill = "transparent"),
        legend.text = element_text(size = 18),
        legend.title = element_text(size = 20),
        # legend.position = c(0.94,0.96),
        legend.position = "top",
        axis.title = element_text(vjust = 1.0, size = 20),
        # strip.text.x = element_text(size = 20),
        axis.text = element_text(size = 15, color = "black"))

 # MIT9313OverlayPlots <- tag_facet(MIT9313OverlayPlots, open = "", close = "", tag_pool = LETTERS, size = 8)
 # 
 # MIT9313OverlayPlots
```

```{r save MIT9313 PUR overlay plots}

# ggsave(file = file.path(PlotsPath, paste("MIT9313OverlayPlots",".png",sep = "")), plot = MIT9313OverlayPlots, device = NULL, scale = 1, height=20, width= 35, units = c("cm"),dpi = 300, limitsize = TRUE)

```

```{r SS120 PUR overlay with legend,  fig.height = 10, fig.width = 10}
# SS120OverlayPlots <- PURPARFits %>%
PURPARFits %>%
   filter(Strain == "SS120",
          O2_Category == "High",
          Par_ue == 30,
          Photoperiod == 8 |
          Photoperiod == 12 & WL == 530,
          Run != 113 & Run != 112) %>%
  mutate(WL_label = case_when(WL %in% c(450, 530, 660) ~ paste("Growth", "wavelength", "=", WL, "nm", sep = " "),
        WL %in% c("WW") ~ paste("Growth", "wavelength", "=", "White LED", sep = " "))) %>% 
   unnest(cols = c(PurData), names_sep = "_", keep_empty = TRUE) %>%
  pivot_longer(cols = c(PurData_PURAbsWholeCell_Norm440, PurData_AbsNorm440,  PurData_MC_NormEm), values_to = "Abs_Em_Value", names_to = "Abs_Em_Names") %>%
ggplot()+
  geom_area(aes(x = PurData_OlisAbs_nm, y = Abs_Em_Value, fill = as.factor(WL)), color = "black", 
            size = 0.5, alpha = 0.7, show.legend = FALSE, 
            data = . %>% filter(Abs_Em_Names == "PurData_PURAbsWholeCell_Norm440")) +
  geom_line(aes(x = PurData_OlisAbs_nm, y = Abs_Em_Value, color = Abs_Em_Names,  linetype = Abs_Em_Names), size = 0.85) +
  geom_line(aes(x = PurData_OlisAbs_nm, y = Abs_Em_Value, color = as.factor(WL),  linetype = Abs_Em_Names), size = 1.0, data = . %>% filter(Abs_Em_Names == "PurData_MC_NormEm")) +
  scale_color_manual(values = c(MCMIXColours, "PurData_PURAbsWholeCell_Norm440" = "black", "PurData_AbsNorm440" = "darkorchid3", "PurData_MC_NormEm" = "white")) +
  scale_linetype_manual(values = c("PurData_PURAbsWholeCell_Norm440" = "solid", "PurData_AbsNorm440" = "twodash",  "PurData_MC_NormEm" = "dotted"),
                        labels = c("LED emission", "Whole cell abs.", "PUR")) +
  scale_fill_manual(values = MCMIXColours) +
 # guides(colour = FALSE) +
  labs(y = "Normalized Absorbance and Emission", x = "Wavelength (nm)") +
  labs(linetype = "Spectra Type")+
  scale_y_continuous(breaks=seq(0, 1.5, by = 0.25)) +
 # coord_cartesian(ylim = c (0, 0.2)) +
   facet_wrap(vars(WL),
              labeller = labeller(strain = label_value, WL = label_value, Photoperiod = label_both, Par_ue = label_both, O2_Category = label_both, Pur_ue = label_both))+
  geom_text(aes(x = 430, y = 1.1, label = WL_label, color = as.factor(WL)), show.legend = FALSE,
    # parse = TRUE,
    hjust = 0,
    vjust = 1,
    size = 5.0
  )  +
   geom_text(aes(x = 560, y = 0.95, label = paste("PAR", "=", Par_ue, sep = " "), show.legend = FALSE, ),
    # parse = TRUE,
    hjust = 0,
    vjust = 1,
    size = 5.0
  )  +
  geom_text(aes(x = 560, y = 0.85, label = paste("PUR", "=", Pur_ue,sep = " "), show.legend = FALSE, ),
    # parse = TRUE,
    hjust = 0,
    vjust = 1,
    size = 5.0
  )  +
  guides(color = "none") +
  theme_bw() +
 theme(strip.text = element_blank(),
       legend.background = element_rect(color = "transparent", fill = "transparent"),
        legend.text = element_text(size = 18),
        legend.title = element_text(size = 20),
        # legend.position = c(0.94,0.96),
        legend.position = "top",
        axis.title = element_text(vjust = 1.0, size = 20),
        # strip.text.x = element_text(size = 20),
        axis.text = element_text(size = 15, color = "black"))

 # SS120OverlayPlots <- tag_facet(SS120OverlayPlots, open = "", close = "", tag_pool = LETTERS, size = 8)
 # 
 # SS120OverlayPlots

```

```{r save SS120 PUR overlay plots}

# ggsave(file = file.path(PlotsPath, paste("SS120OverlayPlots",".png",sep = "")), plot = SS120OverlayPlots, device = NULL, scale = 1, height=20, width= 35, units = c("cm"),dpi = 300, limitsize = TRUE)

```