### 2024-02-13 Code/ManuscriptTargetProtein.Rmd
Edits to ManuscriptTargetProtein.Rmd and Manuscript_O2.Rmd figure names and captions

---
title: "ManuscriptTargetProtein"
author: "Mireille Savoie"
date: "`r format(Sys.Date())`"
output:
  bookdown::html_document2:
    code_folding: show
    keep_md: yes
    toc: TRUE
    toc_float: TRUE
    toc_depth: 6
    fig_caption: yes
    pandoc_args: 
      - '--lua-filter=scholarly-metadata.lua'
      - '--lua-filter=author-info-blocks.lua'
  bookdown::word_document2:
    code_folding: show
    keep_md: yes
    toc: TRUE
    toc_depth: 6
    fig_caption: yes
    pandoc_args: 
      - '--lua-filter=scholarly-metadata.lua'
      - '--lua-filter=author-info-blocks.lua'
bibliography: [Manuscript_O2.bib, RPackages.bib]
csl: plos-one.csl
editor_options: 
  markdown: 
    wrap: 72
---

# Introduction

*Some of the code used to create this R Notebook was refurbished from "PlateAbImport.Rmd" written by Maximilian Berthold, Douglas A. Campbell, Melissa L. Rioux, Sarah J Gore, and Alyson MacCormack.*

Importing and tidying data from a .csv containing already processed target proteins

```{r set project variables}

Project <- "PICO"
DataIn <- file.path("..","Data", "ProcessedData")
DataOut <- file.path("..", "Data", "CleanData")
PlotsOut <- file.path("..","Output", "Figures")

FileEncode <- "UTF-8" 
Delimiter <- ","

HeaderRows <- 0
#AbsWL = c(680, 720, 750) #select OD's of interest from plategrowth.Rds
```

```{r}
library(tidyverse)
library(lubridate)
library(stringr)
library(broom)
library(knitr)
# library(OneR)
# library(zoo)
# library(strucchange)

library(dplyr)
# library(magrittr)
library(googledrive)
library(googlesheets4)
library(readxl)
```

```{r read locally stored metadata from rds}

CultureCatalog <- readRDS(file = file.path("..", "Data","CultureCatalog.Rds"))

```

```{r read ProcessFile}

TargetProtein <- read.csv(file = file.path(DataIn,"2020_PICO_Multi_TargetProteins.csv"))

glimpse(TargetProtein)

TargetProtein <- TargetProtein %>%
  rename(SampleID = "ï..Sample")

glimpse(TargetProtein)

```

```{r set colours}
MyWavelengths = c(405, 450, 475, 530, 615, 660, 730, "WW")
MCMIXColours = c("violet", "darkblue", "dodgerblue", "darkgreen","orange","red","purple", "black")

names(MCMIXColours) <- MyWavelengths
MCMIXColours
```

```{r join protein data with culture meta data}

TargetProtein  <- TargetProtein %>% 
 left_join(., CultureCatalog, by = c("SampleID" = "SampleID"))

glimpse(TargetProtein)

```

```{r order protein by biological process}

TargetProtein  <- TargetProtein %>% 
 mutate(Target = factor(Target, levels = c("FtsH", "PsbA", "PetC", "PsaC", "RbcL")))

glimpse(TargetProtein)

```

```{r explore plots}
TargetProtein %>%
   filter(Strain == "MIT9313") %>%
  ggplot(aes(x=Target, y= Target_TotalProtein_fmole_ug)) +
  geom_col() +
  # ggplot() +
  # geom_point(aes(x = Par_ue, y = Target_TotalProtein_fmole_ug, colour = as.factor(WL)), size = 3, alpha = 0.4) +
  # scale_fill_viridis_d(fill = "f68060") %>% 
  # scale_colour_manual(values = MCMIXColours) +
 ggh4x::facet_nested(cols = vars(Strain, O2), rows = vars(WL,Par_ue), 
                    labeller = labeller(WL = label_both, O2 = label_both, Strain = label_both)) +
  theme_bw()

TargetProtein %>%
   filter(Strain == "MED4") %>%
  ggplot(aes(x=Target, y= Target_TotalProtein_fmole_ug)) +
  geom_col() +
  # ggplot() +
  # geom_point(aes(x = Par_ue, y = Target_TotalProtein_fmole_ug, colour = as.factor(WL)), size = 3, alpha = 0.4) +
  # scale_fill_viridis_d(fill = "f68060") %>% 
  # scale_colour_manual(values = MCMIXColours) +
 ggh4x::facet_nested(cols = vars(Strain, O2), rows = vars(WL,Par_ue), 
                    labeller = labeller(WL = label_both, O2 = label_both, Strain = label_both)) +
  theme_bw()

TargetProtein %>%
   filter(Strain == "MIT9313",
          O2 == 21) %>%
  ggplot(aes(x=Target, y= Target_TotalProtein_fmole_ug)) +
  geom_col() +
  # ggplot() +
  # geom_point(aes(x = Par_ue, y = Target_TotalProtein_fmole_ug, colour = as.factor(WL)), size = 3, alpha = 0.4) +
  # scale_fill_viridis_d(fill = "f68060") %>% 
  # scale_colour_manual(values = MCMIXColours) +
 ggh4x::facet_nested(cols = vars(Strain, O2), rows = vars(WL,Par_ue), 
                    labeller = labeller(WL = label_both, O2 = label_both, Strain = label_both)) +
  theme_bw()

TargetProtein %>%
   filter(Strain == "MED4",
          O2 == 21) %>%
  ggplot(aes(x=Target, y= Target_TotalProtein_fmole_ug)) +
  geom_col() +
  # ggplot() +
  # geom_point(aes(x = Par_ue, y = Target_TotalProtein_fmole_ug, colour = as.factor(WL)), size = 3, alpha = 0.4) +
  # scale_fill_viridis_d(fill = "f68060") %>% 
  # scale_colour_manual(values = MCMIXColours) +
 ggh4x::facet_nested(cols = vars(Strain, O2), rows = vars(WL,Par_ue), 
                    labeller = labeller(WL = label_both, O2 = label_both, Strain = label_both)) +
  theme_bw()

TargetProtein %>%
   filter(Strain == "MIT9313",
          O2 == 0) %>%
  ggplot(aes(x=Target, y= Target_TotalProtein_fmole_ug)) +
  geom_col() +
  # ggplot() +
  # geom_point(aes(x = Par_ue, y = Target_TotalProtein_fmole_ug, colour = as.factor(WL)), size = 3, alpha = 0.4) +
  # scale_fill_viridis_d(fill = "f68060") %>% 
  # scale_colour_manual(values = MCMIXColours) +
 ggh4x::facet_nested(cols = vars(Strain, O2), rows = vars(WL,Par_ue), 
                    labeller = labeller(WL = label_both, O2 = label_both, Strain = label_both)) +
  theme_bw()

TargetProtein %>%
   filter(Strain == "MED4",
          O2 == 0) %>%
  ggplot(aes(x=Target, y= Target_TotalProtein_fmole_ug)) +
  geom_col() +
  # ggplot() +
  # geom_point(aes(x = Par_ue, y = Target_TotalProtein_fmole_ug, colour = as.factor(WL)), size = 3, alpha = 0.4) +
  # scale_fill_viridis_d(fill = "f68060") %>% 
  # scale_colour_manual(values = MCMIXColours) +
 ggh4x::facet_nested(cols = vars(Strain, O2), rows = vars(WL,Par_ue), 
                    labeller = labeller(WL = label_both, O2 = label_both, Strain = label_both)) +
  theme_bw()

MIT9313ProteinAnotated <- TargetProtein %>%
  filter( Strain == "MIT9313") %>% 
  ggplot(aes(x=Target, y= Target_TotalProtein_fmole_ug)) +
  geom_col() +
   geom_text(aes(y = 0.05, label = paste(round(Target_TotalProtein_fmole_ug, 0)),  hjust = 0.5, vjust = -1.6, position = "stack", check_overlap = T), color = "Black") +
  # ggplot() +
  # geom_point(aes(x = Par_ue, y = Target_TotalProtein_fmole_ug, colour = as.factor(WL)), size = 3, alpha = 0.4) +
  # scale_fill_viridis_d(fill = "f68060") %>% 
  # scale_colour_manual(values = MCMIXColours) +
 ggh4x::facet_nested(cols = vars(Strain, O2), rows = vars(WL,Par_ue), 
                    labeller = labeller(Par_ue = label_both, WL = label_both, O2 = label_both, Strain = label_both)) +
  theme_bw()

MIT9313ProteinAnotated

MED4ProteinAnotated <- TargetProtein %>%
  filter( Strain == "MED4") %>% 
  ggplot(aes(x=Target, y= Target_TotalProtein_fmole_ug)) +
  geom_col() +
   geom_text(aes(y = 0.05, label = paste(round(Target_TotalProtein_fmole_ug, 0)),  hjust = 0.5, vjust = -1.5, position = "stack", check_overlap = T), color = "Black") +
  # ggplot() +
  # geom_point(aes(x = Par_ue, y = Target_TotalProtein_fmole_ug, colour = as.factor(WL)), size = 3, alpha = 0.4) +
  # scale_fill_viridis_d(fill = "f68060") %>% 
  # scale_colour_manual(values = MCMIXColours) +
 ggh4x::facet_nested(cols = vars(Strain, O2), rows = vars(WL,Par_ue), 
                    labeller = labeller(Par_ue = label_both, WL = label_both, O2 = label_both, Strain = label_both)) +
  theme_bw()

MED4ProteinAnotated
```

```{r }

 ggsave(file = file.path(PlotsOut, paste("MED4ProteinAnotated",".png",sep = "")), plot = MED4ProteinAttonated, device = NULL, scale = 1, height=15, width= 25, units = c("cm"),dpi = 300, limitsize = TRUE)

ggsave(file = file.path(PlotsOut, paste("MIT9313ProteinAnotated",".png",sep = "")), plot = MIT9313ProteinAttonated, device = NULL, scale = 1, height=15, width= 25, units = c("cm"),dpi = 300, limitsize = TRUE)

```

```{r, save rds}

saveRDS(TargetProtein, file.path(DataOut, paste(Project, "TargetProtein.Rds", sep = "_"), fsep = .Platform$file.sep))

# saveRDS(ChlTurnerCytometerAll, file.path(DataOut, paste(Project, "ChlTurnerMiSa.Rds", sep = "_"), fsep = .Platform$file.sep))
```