### 2024-02-16 Code/ManuscriptPURFits.Rmd
Fixed pur fit plot

---
title: "ManuscriptPURFits"
author: 
- Mireille Savoie
- Douglas A. Campbell
date: "`r format(Sys.Date())`"
output:
  bookdown::html_document2:
    code_folding: show
    keep_md: yes
    toc: TRUE
    toc_float: TRUE
    toc_depth: 6
    fig_caption: yes
    pandoc_args: 
      - '--lua-filter=scholarly-metadata.lua'
      - '--lua-filter=author-info-blocks.lua'
  bookdown::word_document2:
    code_folding: show
    keep_md: yes
    toc: TRUE
    toc_depth: 6
    fig_caption: yes
    pandoc_args: 
      - '--lua-filter=scholarly-metadata.lua'
      - '--lua-filter=author-info-blocks.lua'
bibliography: [Manuscript_O2.bib, RPackages.bib]
csl: plos-one.csl
editor_options: 
  markdown: 
    wrap: 72
---

To Do

## Set figure caption font size
```{css, echo=FALSE}
p.caption {
  font-size: 25px;
}
```

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = F)
knitr::opts_chunk$set(fig.path='Figs/')
```

```{r load libraries, include=FALSE}
# libraries; Note check actual dependencies
library(tidyverse)
library(lubridate)
#library(stringr)
library(broom)
library(knitr)
library(zoo)
#library(AICcmodavg)
# library(RColorBrewer)
library(grid)
library(gridExtra)
# library(bookdown)
# library(glue)
# library(egg)
# library(ggspectra)
# library(photobiologyWavebands)
# library(photobiology)
#library(ggpubr)
#library(ggh4x)
library(cowplot)
#library(jmv)
# library(pgirmess) # for the post hoc kruskalalmc test
#library(rcompanion) # for Scheirer-Ray-Hare non-parametric 2 way test, derived from Kruskal-Wallis
# library(FSA) # for Dunn test
#library(dunn.test)
# library(rstatix)
library(minpack.lm)

```

```{r set project variables}

#"..", takes up a level in the directory path
Project <- "PICO"
DataIn <- file.path("..","Data", "CleanData")
PlotsOut <- file.path("..","Output", "Figures")

```

Read .Rds
Multiple 'list' columns removed; this can sometimes happen if early steps generate 2 values for a single data frame column element; this converts the column to 'list' format to accomodate more than one value in a single data from column/row position.

```{r read ProcessFile}

PURPARFits <- readRDS(file = file.path(DataIn,"PICO_Merged_GrowthFitsPURPAR.Rds"))

glimpse(PURPARFits)

PURPARFits <- PURPARFits|>
  select(-c("tubedata")) |>
  select(-contains("OD720")) |>
  select(-c("ExpCul", "Optode", "OptodeCh", "ExpEndDate", "ExpEndHour")) |>
  select(-contains("trunc_")) |>
  select(-contains("predict")) |>
  select(-c("deltaOD_Lmax", "deltaOD_Lmu_raw", "deltaOD_Lintercept",  "deltaOD_Lmax_se",  "deltaOD_Lintercept_se")) |>
  select(-c("GrowthAmpdeltaOD_14days", "PurData")) |>
  select(-c("ReplicateMean_deltaOD_Lmu_corr", "ReplicateSD_deltaOD_Lmu_corr", "ReplicateSE_deltaOD_Lmu_corr" )) |>
 filter(Par_ue %in% c(30,90,180)) |>
  filter(!is.na(PURPhotonDose_day)) # PURPhotonDose_day NA mess up fits Strain = MED4 x 250 =  O2_uM

 # y = ReplicateMean_deltaOD_Lmu_corr_day

PURPARFits <- PURPARFits |>
  mutate(deltaOD_Lmu_corr_day = deltaOD_Lmu_corr * 24)

glimpse(PURPARFits)

```

Make O2_uM and WL as factor levels from highest to lowest and remove all green light (WL= 530) growth data
```{r O2 levels}

PURPARFits <- PURPARFits %>%
  filter(WL %in% c("WW", 660, 450)) %>% 
  mutate(O2_uM = factor(O2_uM, levels = c(250, 25, 2.5)),
         WL= factor(WL, levels = c("WW", 660, 450)),
        Photoperiod = factor(Photoperiod, levels = c("16", "12", "8", "4" )))

 # Par_ue = factor(Par_ue, levels =  c(180, 120, 100,90,60, 30, 15))
```

```{r set colours and shapes}

library(photobiology)

MyWavelengths = c("WW", 660, 530, 450)
MCMIXColours = c("black", w_length2rgb(660), w_length2rgb(530),  w_length2rgb(450))

names(MCMIXColours) <- MyWavelengths
MCMIXColours

MyPhotoperiods = as.factor(c(4, 8, 12, 16))
PhotoperiodShapes = c(6, 18, 16, 17)

names(PhotoperiodShapes) <- MyPhotoperiods
PhotoperiodShapes

```

```{r create dummy no growths for dark conditions}

#DummyDF for SS120; SS120 is a low light strain, based on literature it can only grow up to 300 uE WW; SS120 consistently died at various light colors and oxygen levels latest at ~PUR 400; it can safely be assumed it dies at par_ue = 0
Dummy_Strain <- c("SS120")
Dummy_Photoperiod <- c(4, 8, 12, 16)
Dummy_Par_ue<- c(0)
Dummy_O2_uM <- c(2.5, 25.0, 250.0)
Dummy_WL <- c(450, 660, "WW")
Dummy_deltaOD_Lmu_corr <- c(0)
Dummy_Pur_ue<- c(0)
Dummy_deltaOD_Lmu_corr_day <- c(0)
Dummy_PUR_day <- c(0)

DummyDF_SS120 <- expand_grid(
  Dummy_Strain, Dummy_Photoperiod, Dummy_Par_ue, Dummy_O2_uM, Dummy_WL, Dummy_deltaOD_Lmu_corr, Dummy_Pur_ue, Dummy_deltaOD_Lmu_corr_day,Dummy_PUR_day
  ) %>%
  mutate(SampleID = "DummySS120")

DummyDF_Expand_SS120 <- rbind(DummyDF_SS120, DummyDF_SS120, DummyDF_SS120) %>% 
        rename(Strain = "Dummy_Strain",
        Photoperiod = "Dummy_Photoperiod",
        Par_ue = "Dummy_Par_ue",
        O2_uM = "Dummy_O2_uM", 
        WL = "Dummy_WL", 
        deltaOD_Lmu_corr = "Dummy_deltaOD_Lmu_corr",
        Pur_ue = "Dummy_Pur_ue",
        deltaOD_Lmu_corr_day = "Dummy_deltaOD_Lmu_corr_day",
        PURPhotonDose_day = "Dummy_PUR_day")

Dummy_Strain <- c("MIT9313")

DummyDF_MIT9313 <- expand_grid(
  Dummy_Strain, Dummy_Photoperiod, Dummy_Par_ue, Dummy_O2_uM, Dummy_WL, Dummy_deltaOD_Lmu_corr, Dummy_Pur_ue, Dummy_deltaOD_Lmu_corr_day,Dummy_PUR_day
  ) %>%
  mutate(SampleID = "DummyMIT9313")

DummyDF_Expand_MIT9313 <- rbind(DummyDF_MIT9313, DummyDF_MIT9313, DummyDF_MIT9313) %>% 
  rename(Strain = "Dummy_Strain",
        Photoperiod = "Dummy_Photoperiod",
        Par_ue = "Dummy_Par_ue",
        O2_uM = "Dummy_O2_uM", 
        WL = "Dummy_WL", 
        deltaOD_Lmu_corr = "Dummy_deltaOD_Lmu_corr",
        Pur_ue = "Dummy_Pur_ue",
        deltaOD_Lmu_corr_day = "Dummy_deltaOD_Lmu_corr_day",
        PURPhotonDose_day = "Dummy_PUR_day")

Dummy_Strain <- c("MED4")

DummyDF_MED4 <- expand_grid(
  Dummy_Strain, Dummy_Photoperiod, Dummy_Par_ue, Dummy_O2_uM, Dummy_WL, Dummy_deltaOD_Lmu_corr, Dummy_Pur_ue, Dummy_deltaOD_Lmu_corr_day,Dummy_PUR_day
  ) %>%
  mutate(SampleID = "DummyMED4")

DummyDF_Expand_MED4 <- rbind(DummyDF_MED4, DummyDF_MED4, DummyDF_MED4) %>% 
  rename(Strain = "Dummy_Strain",
        Photoperiod = "Dummy_Photoperiod",
        Par_ue = "Dummy_Par_ue",
        O2_uM = "Dummy_O2_uM", 
        WL = "Dummy_WL", 
        deltaOD_Lmu_corr = "Dummy_deltaOD_Lmu_corr",
        Pur_ue = "Dummy_Pur_ue",
        deltaOD_Lmu_corr_day = "Dummy_deltaOD_Lmu_corr_day",
        PURPhotonDose_day = "Dummy_PUR_day")

DummyDF_Expand <- rbind(DummyDF_Expand_SS120, DummyDF_Expand_MIT9313, DummyDF_Expand_MED4)

remove(DummyDF_MED4)
remove(DummyDF_MIT9313)
remove(DummyDF_SS120)
remove(DummyDF_Expand_MED4)
remove(DummyDF_Expand_MIT9313)
remove(DummyDF_Expand_SS120)

```

# Join dummy df to growth df
```{r add dummy dark growth to DF}

library(plyr)
PURPARFits_dummy<- rbind.fill(PURPARFits, DummyDF_Expand) # Apply the rbind.fill plyr function

detach("package:plyr")

remove(DummyDF_Expand)
remove(PURPARFits)

```

```{r testplot}
 PURPARFits_dummy %>% 
  ggplot() +
  geom_point(aes(x = PURPhotonDose_day, y = deltaOD_Lmu_corr_day, colour = WL)) +
    scale_colour_manual(values = MCMIXColours, labels = c("White LED", "660", "450" )) +
  facet_grid(rows = vars(Par_ue), cols = vars(Strain)) +
  theme_bw()
```

Citations: Ma X, Liu Q, Zhang Z, Zhang Z, Zhou Z, Jiang Y, et al. (2021) Effects of photosynthetic models on the calculation results of photosynthetic response parameters in young Larix principis- rupprechtii Mayr. plantation. PLoS ONE 16(12): e0261683. https://doi.org/10.1371/journal.pone.0261683

https://www.frontiersin.org/articles/10.3389/fpls.2020.581851/full

I: incident light, umol photons m-2 s-1
a: alpha, initial slope of light response curve under limiting light; the 'affinity' of the culture for light.
Pmax: maximum rate of photosynthesis; umol O2 chl-1 s-1; the 'capacity' of the culture for photosynthesis.

R: dark respiration, umol O2 chl-1 s-1

b: beta 'photoinhibition' term for decline in oxygen evolution under increasing incident light.  This decrease can result from multiple mechanisms.

g: gamma fitting term related to curvature of response.

d: change in respiration under illumination.
```{r grc models}
# simple saturating function; rectangular hyperbola
grc <- function(I, a, Pmax){((a * I * Pmax)/((a*I) + Pmax))
}
  #(equivalent to mm <- function(I, KM, Pmax){(Pmax* I)/(KM + I)})

TestI = c(0:14)

plot(x = TestI, y = grc(I = TestI, a = 0.1, Pmax = 1))

# # function with beta photoinhibition;
#https://www.frontiersin.org/articles/10.3389/fpls.2020.581851/full
 grcbeta <- function(I, a, b, g){(a * ((1 - (b*I))/(1 + (g*I)))*I)
 }
# 
 plot(x = TestI, y = grcbeta(I = TestI, a = 0.1, b = 0.1, g = 0.01))

#Platt et al., 1980
#Harrison & Platt, 1986
grcplatt <- function(I, a, b, Pmax){Pmax * (1-exp(-a*I/Pmax)) * exp(-b*I/Pmax)}

plot(x = TestI, y = grcplatt(I = TestI, a = 0.1, b = 0.1, Pmax = 1))

# linear equation to deal with photoperiods and exclusively no growth
LinEqn <- function(x, M, Intercept){((x*M)+ Intercept)}

LinEqnStart <- list(M = 0, Intercept = 0)

```

Issue with MED4 data at 250 uM; lhs of model 42, right hand 40; for some reason lhs prediction missing for two of the rhs diel PUR values

Causes 'augment' to fail.
Get around other ways if needed.

Issue: OrgPur_day is 'NA' for 3 rows of MED4 x O2_um = 250
Substitute in PURPhotonDose_day which includes extrapolated estimates of PUR for strain x growth light that lack OLIS spectra

```{r growthPURfits}

# model with dummy dark, 0 PAR,  growth rates. Also removed data if no growth occurred for an O2 and photoperiod combination so that the model including all photoperiods would not be skewed towards zero.
PURPARStrainO2NestModels_filtered <- PURPARFits_dummy |>
  mutate(remove = case_when(Strain == "MED4" & Photoperiod == 4 ~ 1,
                            Strain == "MED4" & O2_uM == 25 & Photoperiod == 8 & WL == "WW" ~ 1,
                            Strain == "MIT9313" & O2_uM == 2.5 & Photoperiod == 16 ~ 1,
                            Strain == "MIT9313" & O2_uM == 2.5 & Photoperiod == 12 & WL == "WW" ~ 1,
                            Strain == "SS120"  & O2_uM == 2.5 & Photoperiod %in% c(16,8,4) ~ 1,
                            Strain == "SS120"  & O2_uM == 25 & Photoperiod %in% c(16) & WL == 450 ~ 1,
                            Strain == "SS120"  & O2_uM == 250 & Photoperiod %in% c(4) & WL == 450 ~ 1,
                            TRUE ~ 0)) |>
   filter(remove == 0) |>
  nest(data = -c("Strain","O2_uM")) |>
  mutate(GRCplatt_Model = map(data, possibly(~nlsLM(deltaOD_Lmu_corr_day ~ grcplatt(I = PURPhotonDose_day, a, b, Pmax),
                              data = .x,
                              start = list(a = 1e-7, b = 1e-8, Pmax = 1), 
                              lower = c(0, 0, 0),
                              upper = c(1,1,100),
                              control = list(maxiter = 500)
                         ), NULL)
                         )
         ) |>
  mutate(GRCplatt_Param = map(GRCplatt_Model, tidy),
         GRCplatt_Predict = map(GRCplatt_Model, possibly(augment)),
         GRCplatt_Glance = map(GRCplatt_Model, possibly(glance))
         ) |>
  mutate(GRClm_Model = map(data, possibly(~nlsLM(deltaOD_Lmu_corr_day ~ LinEqn(x = PURPhotonDose_day,  M, Intercept),
                              data = .x,
                              start = LinEqnStart
                         ), NULL)
                         )
         ) |>
  mutate(GRClm_Param = map(GRClm_Model, tidy),
         GRClm_Predict = map(GRClm_Model, possibly(augment)),
         GRClm_Glance = map(GRClm_Model, possibly(glance))
         )

# nest unfiltered data and join with filtered data to have all data to plot later 
AllData <- PURPARFits_dummy |>
    nest(Data_all = -c("Strain","O2_uM"))

PURPARStrainO2NestModels_filtered <- full_join(PURPARStrainO2NestModels_filtered, AllData)

# PURPARStrainO2NestModels <- PURPARFits_dummy |>
#   nest(data = -c("Strain","O2_uM")) |>
#   mutate(GRCplatt_Model = map(data, possibly(~nlsLM(deltaOD_Lmu_corr_day ~ grcplatt(I = PURPhotonDose_day, a, b, Pmax),
#                               data = .x,
#                               start = list(a = 1e-7, b = 1e-8, Pmax = 1), 
#                               lower = c(0, 0, 0),
#                               upper = c(1,1,100),
#                               control = list(maxiter = 500)
#                          ), NULL)
#                          )
#          ) |>
#   mutate(GRCplatt_Param = map(GRCplatt_Model, tidy),
#          GRCplatt_Predict = map(GRCplatt_Model, possibly(augment)),
#          GRCplatt_Glance = map(GRCplatt_Model, possibly(glance))
#          ) |>
#   mutate(GRClm_Model = map(data, possibly(~nlsLM(deltaOD_Lmu_corr_day ~ LinEqn(x = PURPhotonDose_day,  M, Intercept),
#                               data = .x,
#                               start = LinEqnStart
#                          ), NULL)
#                          )
#          ) |>
#   mutate(GRClm_Param = map(GRClm_Model, tidy),
#          GRClm_Predict = map(GRClm_Model, possibly(augment)),
#          GRClm_Glance = map(GRClm_Model, possibly(glance))
#          )

# # model with dummy dark, 0 PAR, 0 growth rates
# PURPARStrainO2PhotoNestModels <- PURPARFits_dummy  |>
#  nest(data = -c("Strain","O2_uM", "Photoperiod")) |>
#   mutate(GRCplatt_Model = map(data, possibly(~nlsLM(deltaOD_Lmu_corr_day ~ grcplatt(I = PURPhotonDose_day, a, b, Pmax),
#                               data = .x,
#                               start = list(a = 1e-8, b = 1e-9, Pmax = 1),
#                               lower = c(-0, 0, 0),
#                               upper = c(1,1,100),
#                               control = list(maxiter = 500)
#                          ), NULL)
#                          )
#          ) |>
#   mutate(GRCplatt_Param = map(GRCplatt_Model, tidy),
#          GRCplatt_Predict = map(GRCplatt_Model, possibly(augment)),
#          GRCplatt_Glance = map(GRCplatt_Model, possibly(glance))
#          ) |>
#   mutate(GRClm_Model = map(data, possibly(~nlsLM(deltaOD_Lmu_corr_day ~ LinEqn(x = PURPhotonDose_day,  M, Intercept),
#                               data = .x,
#                               start = LinEqnStart
#                          ), NULL)
#                          )
#          ) |>
#   mutate(GRClm_Param = map(GRClm_Model, tidy),
#          GRClm_Predict = map(GRClm_Model, possibly(augment)),
#          GRClm_Glance = map(GRClm_Model, possibly(glance))
#          )

PURPARStrainO2PhotoNestModels <- PURPARFits_dummy |>
 nest(data = -c("Strain","O2_uM", "Photoperiod")) |>
  mutate(GRCplatt_Model = map(data, possibly(~nlsLM(deltaOD_Lmu_corr_day ~ grcplatt(I = PURPhotonDose_day, a, b, Pmax),
                              data = .x,
                              start = list(a = 1e-8, b = 1e-9, Pmax = 1),
                              lower = c(-0, 0, 0),
                              upper = c(1,1,100),
                              control = list(maxiter = 500)
                         ), NULL)
                         )
         ) |>
  mutate(GRCplatt_Param = map(GRCplatt_Model, tidy),
         GRCplatt_Predict = map(GRCplatt_Model, possibly(augment)),
         GRCplatt_Glance = map(GRCplatt_Model, possibly(glance))
         ) |>
  mutate(GRClm_Model = map(data, possibly(~nlsLM(deltaOD_Lmu_corr_day ~ LinEqn(x = PURPhotonDose_day,  M, Intercept),
                              data = .x,
                              start = LinEqnStart
                         ), NULL)
                         )
         ) |>
  mutate(GRClm_Param = map(GRClm_Model, tidy),
         GRClm_Predict = map(GRClm_Model, possibly(augment)),
         GRClm_Glance = map(GRClm_Model, possibly(glance))
         )

# # model with dummy dark, 0 PAR,  growth rates 
# PURPARStrainO2WLNestModels <- PURPARFits_dummy |>
#   nest(data = -c("Strain","O2_uM", "WL")) |>
#   mutate(GRCplatt_Model = map(data, possibly(~nlsLM(deltaOD_Lmu_corr_day ~ grcplatt(I = PURPhotonDose_day, a, b, Pmax),
#                               data = .x,
#                               start = list(a = 1e-7, b = 1e-8, Pmax = 1),
#                               lower = c(0, 0, 0),
#                               upper = c(1,1,100),
#                               control = list(maxiter = 500)
#                          ), NULL)
#                          )
#          ) |>
#   mutate(GRCplatt_Param = map(GRCplatt_Model, tidy),
#          GRCplatt_Predict = map(GRCplatt_Model, possibly(augment)),
#          GRCplatt_Glance = map(GRCplatt_Model, possibly(glance))
#          )|>
#   mutate(GRClm_Model = map(data, possibly(~nlsLM(deltaOD_Lmu_corr_day ~ LinEqn(x = PURPhotonDose_day,  M, Intercept),
#                               data = .x,
#                               start = LinEqnStart
#                          ), NULL)
#                          )
#          ) |>
#   mutate(GRClm_Param = map(GRClm_Model, tidy),
#          GRClm_Predict = map(GRClm_Model, possibly(augment)),
#          GRClm_Glance = map(GRClm_Model, possibly(glance))
#          )
# 
# 
# # model with dummy dark, 0 PAR,  growth rates
# PURPARStrainO2PeriodWLNest <- PURPARFits_dummy |>
#   nest(data = -c("Strain","O2_uM", "Photoperiod", "WL")) |>
#   mutate(GRCplatt_Model = map(data, possibly(~nlsLM(deltaOD_Lmu_corr_day ~ grcplatt(I = PURPhotonDose_day, a, b, Pmax),
#                               data = .x,
#                               start = list(a = 1e-7, b = 1e-8, Pmax = 1),
#                               lower = c(0, 0, 0),
#                               upper = c(1,1,100),
#                               control = list(maxiter = 500)
#                          ), NULL)
#                          )
#          ) |>
#   mutate(GRCplatt_Param = map(GRCplatt_Model, tidy),
#          GRCplatt_Predict = map(GRCplatt_Model, possibly(augment)),
#          GRCplatt_Glance = map(GRCplatt_Model, possibly(glance))
#          )|>
#   mutate(GRClm_Model = map(data, possibly(~nlsLM(deltaOD_Lmu_corr_day ~ LinEqn(x = PURPhotonDose_day,  M, Intercept),
#                               data = .x,
#                               start = LinEqnStart
#                          ), NULL)
#                          )
#          ) |>
#   mutate(GRClm_Param = map(GRClm_Model, tidy),
#          GRClm_Predict = map(GRClm_Model, possibly(augment)),
#          GRClm_Glance = map(GRClm_Model, possibly(glance))
#          )
remove(AllData)

```

# Plot fits
```{r growthPURplots, fig.height= 10, fig.width= 10}

# PURPARStrainO2NestModels |>
#   filter(!map_lgl(GRC_Model, is.null)) |> #https://stackoverflow.com/questions/57336319/how-to-filter-out-null-elements-of-tibbles-list-column 
PURPARStrainO2NestModels_filtered|>
  filter(!map_lgl(GRCplatt_Model, is.null)) |>
  unnest(GRCplatt_Predict, data) |>
   mutate(O2_uM = factor(O2_uM, levels = c(250, 25, 2.5)),
         Strain = factor(Strain, levels = c("MED4", "SS120", "MIT9313")),
         Photoperiod = factor(Photoperiod, levels = c(16, 12, 8, 4))) |>
  ggplot() +
  geom_point(aes(x = PURPhotonDose_day, y = deltaOD_Lmu_corr_day, colour = Photoperiod)) + 
  geom_line(aes(x = PURPhotonDose_day, y = `.fitted`)) + 
  facet_grid(rows = vars(O2_uM), cols = vars(Strain)) + 
  labs(title = "PURPARStrainO2NestModels_filtered Platt") +
  theme_bw()

PURPARStrainO2PhotoNestModels |>
  filter(!map_lgl(GRCplatt_Model, is.null)) |>
  unnest(GRCplatt_Predict, data) |>
   mutate(O2_uM = factor(O2_uM, levels = c(250, 25, 2.5)),
         Strain = factor(Strain, levels = c("MED4", "SS120", "MIT9313")),
         Photoperiod = factor(Photoperiod, levels = c(16, 12, 8, 4))) |>
  ggplot() +
  geom_point(aes(x = PURPhotonDose_day, y = deltaOD_Lmu_corr_day, colour = WL)) +
  geom_line(aes(x = PURPhotonDose_day, y = `.fitted`)) +   
  scale_colour_manual(values = MCMIXColours) +
  facet_grid(rows = vars(O2_uM, Photoperiod), cols = vars(Strain)) +
    labs(title = "PURPARStrainO2PhotoNestModels Platt") +
  theme_bw()

PURPARStrainO2PhotoNestModels |>
  filter(!map_lgl(GRClm_Model, is.null)) |>
  unnest(GRClm_Predict, data) |>
   mutate(O2_uM = factor(O2_uM, levels = c(250, 25, 2.5)),
         Strain = factor(Strain, levels = c("MED4", "SS120", "MIT9313")),
         Photoperiod = factor(Photoperiod, levels = c(16, 12, 8, 4))) |>
  ggplot() +
  geom_point(aes(x = PURPhotonDose_day, y = deltaOD_Lmu_corr_day, colour = WL)) +
  geom_line(aes(x = PURPhotonDose_day, y = `.fitted`)) +
   scale_colour_manual(values = MCMIXColours) +
  facet_grid(rows = vars(O2_uM, Photoperiod), cols = vars(Strain)) +
    labs(title = "PURPARStrainO2PhotoNestModels Linear Model") +
  theme_bw()

PURPARFits_dummy |>
  # unnest(data) |>
   mutate(O2_uM = factor(O2_uM, levels = c(250, 25, 2.5)),
         Strain = factor(Strain, levels = c("MED4", "SS120", "MIT9313")),
         Photoperiod = factor(Photoperiod, levels = c(16, 12, 8, 4))) |>
  ggplot() +
  geom_point(aes(x = PURPhotonDose_day, y = deltaOD_Lmu_corr_day, colour = WL)) +
  # geom_line(aes(x = PURPhotonDose_day, y = `.fitted`)) +
     scale_colour_manual(values = MCMIXColours) +
    facet_grid(rows = vars(O2_uM, Photoperiod), cols = vars(Strain)) +
    labs(title = "PURPARStrainO2PhotoNestModels all data no filter") +
  theme_bw()

```

# Want platt model for all fits that did not fail and lm for platt models that returned a NULL
```{r}
# O2 and Photoperiod nested data frame containing the fits for each photoperiod for each strain
#filter to keep platt models
KeepPlattModel <- PURPARStrainO2PhotoNestModels|>
  select(-c(GRClm_Glance, GRClm_Model, GRClm_Param, GRClm_Predict)) |>
   filter(!map_lgl(GRCplatt_Model, is.null))

#filter to keep only the models that failed to fit a platt model, fits by lm remains. 
KeepLmModel <- PURPARStrainO2PhotoNestModels|>
  filter(map_lgl(GRCplatt_Model, is.null)) |>
 select(-c(GRCplatt_Glance, GRCplatt_Model, GRCplatt_Param, GRCplatt_Predict))

# Join  
StrainO2Photoperiod_AllModels <- full_join(KeepPlattModel, KeepLmModel)

 remove(KeepPlattModel)
 remove(KeepLmModel)

# O2 and Photoperiod nested data frame containing the fits for each photoperiod for each strain
 StrainO2Photoperiod_PlattLmModel <- StrainO2Photoperiod_AllModels |>
   pivot_longer(cols = -c(Strain, O2_uM, Photoperiod, data)) |>
  drop_na(value) |>
  filter(str_detect(name, pattern = "Model") | str_detect(name, pattern = "Glance") | str_detect(name, pattern = "Param") | str_detect(name, pattern = "Predict")) |>
  separate(col = name , into = c("ModelType", "Output"), sep = "_") |>
  pivot_wider(names_from = Output, values_from = value) |>
filter(!map_lgl(Model, is.null)) |>
  filter(!map_lgl(Glance, is.null)) |>
  mutate(AIC = map_dbl(Glance, ~ as.numeric(pluck(.x, "AIC")))) |>
  group_by(Strain, O2_uM, Photoperiod) |>
  filter(AIC == min(AIC)) |>
  ungroup() |>
  select(-c(Glance)) |>
  unnest(Param) |>
  select(-c(statistic)) |>
  pivot_wider(names_from = term, values_from = c(estimate, std.error, p.value))

# O2 nested data frame containing the single fit of pooled photoperiods 
 #filter to keep platt models
KeepPlattModel <- PURPARStrainO2NestModels_filtered|>
  select(-c(GRClm_Glance, GRClm_Model, GRClm_Param, GRClm_Predict)) |>
   filter(!map_lgl(GRCplatt_Model, is.null))

#filter to keep only the models that failed to fit a platt model, fits by lm remains. 
KeepLmModel <- PURPARStrainO2NestModels_filtered|>
  filter(map_lgl(GRCplatt_Model, is.null)) |>
 select(-c(GRCplatt_Glance, GRCplatt_Model, GRCplatt_Param, GRCplatt_Predict))

# Join  
StrainO2_AllModels <- full_join(KeepPlattModel, KeepLmModel)

# O2 nested data frame containing the single fit of pooled photoperiods 
 StrainO2filtered_PlattLmModel <- StrainO2_AllModels|>
  # select(-c(data)) |>
  pivot_longer(cols = -c(Strain, O2_uM, data, Data_all)) |>
  drop_na(value) |>
  filter(str_detect(name, pattern = "Model") | str_detect(name, pattern = "Glance") | str_detect(name, pattern = "Param") | str_detect(name, pattern = "Predict")) |>
  separate(col = name , into = c("ModelType", "Output"), sep = "_") |>
  pivot_wider(names_from = Output, values_from = value) |>
  filter(!map_lgl(Model, is.null)) |>
  filter(!map_lgl(Glance, is.null)) |>
  mutate(AIC = map_dbl(Glance, possibly(~ as.numeric(pluck(.x, "AIC"))))) |>
  group_by(Strain, O2_uM) |>
  filter(AIC == min(AIC)) |>
  ungroup() |>
  select(-c(Glance))|>
  unnest(Param) |>
  select(-c(statistic)) |>
  pivot_wider(names_from = term, values_from = c(estimate, std.error, p.value))

 remove(KeepPlattModel)
 remove(KeepLmModel)
 # remove(StrainO2Photoperiod_AllModels)
 # remove(StrainO2_AllModels)

```

Linear models do not have alpha, beta and pmax so inf and NA are generated. Change them to 0.
```{r lm platt parameters set to zero}

# O2 nested data frame containing the single fit of pooled photoperiods for each strain 
StrainO2filtered_PlattLmModel <- StrainO2filtered_PlattLmModel|>
  mutate(estimate_a = if_else(ModelType == "GRClm", 0, estimate_a),
         estimate_b = if_else(ModelType == "GRClm", 0, estimate_b),
         estimate_Pmax = if_else(ModelType == "GRClm", 0, estimate_Pmax),
         AIC = if_else(ModelType == "GRClm" & is.infinite(AIC), 0, AIC))

StrainO2filtered_PlattLmModel

# O2 and Photoperiod nested data frame containing the fits for each photoperiod for each strain
StrainO2Photoperiod_PlattLmModel <- StrainO2Photoperiod_PlattLmModel|>
  mutate(estimate_a = if_else(ModelType == "GRClm", 0, estimate_a),
         estimate_b = if_else(ModelType == "GRClm", 0, estimate_b),
         estimate_Pmax = if_else(ModelType == "GRClm", 0, estimate_Pmax),
         AIC = if_else(ModelType == "GRClm" & is.infinite(AIC), 0, AIC))

StrainO2Photoperiod_PlattLmModel

```

```{r plot models}

# O2 nested data frame containing the single fit of pooled photoperiods for each strain
StrainO2filtered_PlattLmModel |>
  unnest(Predict, data) |>
  mutate(O2_uM = factor(O2_uM, levels = c(250, 25, 2.5)),
         WL= factor(WL, levels = c("WW", 660, 450)),
         Strain = factor(Strain, levels = c("MED4", "SS120", "MIT9313")),
         Photoperiod = factor(Photoperiod, levels = c("16", "12", "8", "4"))) |>
  ggplot() +
  geom_point(aes(x = PURPhotonDose_day, y = deltaOD_Lmu_corr_day, shape = as.factor(Photoperiod), color = as.factor(WL), fill = as.factor(WL))) +
  geom_line(aes(x = PURPhotonDose_day, y = `.fitted`)) + 
   geom_text(aes(x = 500000, y = 0.75, label = paste(ModelType), show.legend = FALSE)
  ) +
     scale_colour_manual(values = MCMIXColours) +
   scale_shape_manual(values = PhotoperiodShapes)  +
  facet_grid(rows = vars(O2_uM), cols = vars(Strain)) +
    labs(title = "Single fit for each strain in each [O2]- all photoperiods included that had growth") +
  theme_bw()

# O2 and Photoperiod nested data frame containing the fits for each photoperiod for each strain
StrainO2Photoperiod_PlattLmModel |>
  unnest(Predict, data) |>
  mutate(O2_uM = factor(O2_uM, levels = c(250, 25, 2.5)),
         Strain = factor(Strain, levels = c("MED4", "SS120", "MIT9313")),
         Photoperiod = factor(Photoperiod, levels = c(16, 12, 8, 4))) |>
  ggplot() +
  geom_point(aes(x = PURPhotonDose_day, y = deltaOD_Lmu_corr_day, colour = WL)) +
  geom_line(aes(x = PURPhotonDose_day, y = `.fitted`)) + 
   geom_text(aes(x = 500000, y = 0.75, label = paste(ModelType), show.legend = FALSE)
  ) +
     scale_colour_manual(values = MCMIXColours) +
  facet_grid(rows = vars(O2_uM, Photoperiod), cols = vars(Strain)) +
    labs(title = "Fits to be used for ANOVA comparison to single fit across each strain and [O2]") +
  theme_bw()

```

# ANOVA comparison of single fit through pooled photoperiods for each strain and [O2] to the fit of each photoperiod in that pool. 
```{r compare 1 fit of all photoperiod to each photoperiods using platt or lm}

# make wider to join with DF that contains the 1 fit through all photoperiods
Wider_Photoperiods <- StrainO2Photoperiod_PlattLmModel %>% 
  select(Strain, Photoperiod, O2_uM, Model, Predict) %>% 
  pivot_wider(names_from = Photoperiod, values_from = c(Model, Predict))

# Wider_Photoperiods <- StrainO2Photoperiod_PlattLmModel %>% 
#   mutate(PhotoperiodLabel = Photoperiod) %>% 
#   select(Strain, Photoperiod, O2_uM, Model, Predict, data, PhotoperiodLabel) %>% 
#   pivot_wider(names_from = PhotoperiodLabel, values_from = c(Model, Predict)) %>% 
#     unnest(data) %>% 
#   group_by(Strain, O2_uM) %>% 
#     nest(Data_Photoperiod = -c(Strain, O2_uM, Model_16,Model_12, Model_8, Model_4, Predict_16, Predict_12, Predict_8, Predict_4)) %>% 
#   filter(!is.null(Model_16) & !is.null(Predict_16) |
#            !is.null(Model_12) & !is.null(Predict_12)|
#            !is.null(Model_8) & !is.null(Predict_8) |
#            !is.null(Model_4) & !is.null(Predict_4))

# Wider_Photoperiods <- StrainO2Photoperiod_PlattLmModel %>% 
#   select(Strain, Photoperiod, O2_uM, Model, Predict) %>% 
#   pivot_wider(names_from = Photoperiod, values_from = c(Model, Predict)) %>% 
#   unnest(Predict_16, keep_empty = TRUE) %>% 
#     mutate(Photoperiod = 16)%>% 
#     nest(Predict_16 = -c(Strain, O2_uM, Model_16, Model_12, Model_8, Model_4, Predict_12, Predict_8, Predict_4)) %>% 
#   unnest(Predict_12, keep_empty = TRUE) %>% 
#     mutate(Photoperiod = 12)%>% 
#     nest(Predict_12 = -c(Strain, O2_uM, Model_16, Model_12, Model_8, Model_4, Predict_16, Predict_8, Predict_4))%>% 
#   unnest(Predict_8, keep_empty = TRUE) %>% 
#     mutate(Photoperiod = 8)%>% 
#     nest(Predict_8 = -c(Strain, O2_uM, Model_16, Model_12, Model_8, Model_4, Predict_16, Predict_12, Predict_4))%>% 
#   unnest(Predict_4, keep_empty = TRUE) %>% 
#     mutate(Photoperiod = 4)%>% 
#     nest(Predict_4 = -c(Strain, O2_uM, Model_16, Model_12, Model_8, Model_4, Predict_16, Predict_12, Predict_8))

# this contains 1 fit through all photoperiods, the photoperiods with no growth were filtered out.
AllPhotoperiods <- StrainO2filtered_PlattLmModel %>% 
  select(Strain, O2_uM, Model, Predict, Data_all, data) %>% 
 rename(All = 'Model',
        Predict_pooled = 'Predict',
        Data_filtered = 'data')

AllFitsCombined <- left_join(Wider_Photoperiods, AllPhotoperiods)

ANOVA_AllPhotoperiods <- AllFitsCombined %>% 
  mutate(twelve = `Model_12`,
         four = `Model_4`,
         eight = `Model_8`,
         sixteen = `Model_16`) %>%
  mutate(fourvsAll = map(map2(four, All, possibly(anova, otherwise = NULL)), tidy),
         eightvsAll = map(map2(eight, All, possibly(anova, otherwise = NULL)), tidy),
         twelvevsAll = map(map2(twelve, All, possibly(anova, otherwise = NULL)), tidy),
         sixteenvsAll = map(map2(sixteen, All, possibly(anova, otherwise = NULL)), tidy)) %>% 
  select(Strain, O2_uM, fourvsAll, eightvsAll, twelvevsAll, sixteenvsAll) %>% 
  pivot_longer(cols = c(fourvsAll, eightvsAll, twelvevsAll, sixteenvsAll)) %>% 
  unnest(value, keep_empty = TRUE) %>% 
  filter(is.na(term) & is.na(p.value) |
           !is.na(term) & !is.na(p.value)) %>%
  select(Strain, O2_uM, name, p.value) %>% 
  separate(name, into = c("Photoperiod1", "AllPhotoperiod"), sep = "vs") %>% 
  mutate(pvalue = signif(`p.value`, digits = 2))

# %>% 
#   mutate(p.value = case_when(is.na(p.value) ~ 0,
#                              TRUE ~ p.value)) # Changed all failed fits to p.value = 0
  # mutate(pval = pluck(value, "p.value", 2))

ANOVA_AllPhotoperiods %>% 
  mutate(Photoperiod1 = as.numeric(case_when(Photoperiod1 == 'four' ~ 4,
                                  Photoperiod1 == 'eight' ~ 8,
                                  Photoperiod1 == 'twelve' ~ 12,
                                  Photoperiod1 == 'sixteen' ~ 16)),
         O2_uM = factor(O2_uM, levels = c(250, 25, 2.5)),
         Strain = factor(Strain, levels = c("MED4", "SS120", "MIT9313"))) %>%
    # mutate(Photoperiod1 = factor(Photoperiod1, levels = c('four','eight', 'twelve', 'sixteen'))) %>% 
  ggplot() +
  geom_text(aes(x = Photoperiod1, y = AllPhotoperiod, label = pvalue), data = . %>% filter(pvalue < 0.05), size = 2.5) +
  geom_text(aes(x = Photoperiod1, y = AllPhotoperiod), data = . %>% filter(pvalue > 0.05), label = "ns") +
  geom_text(aes(x = Photoperiod1, y = AllPhotoperiod), data = . %>% filter(is.na(pvalue)), label = "NA") +
  facet_grid(rows = vars(O2_uM), cols = vars(Strain)) +
  labs(title = "ANOVA compare all models",
       caption = "NA = atleast one model did not fit due to lack of data points; 
       ns = not significantly different from all photoperiod model (p value > 0.05); 
       numbers = significantly different pvalues from all photoperiod model (p value < 0.05)")

remove(Wider_Photoperiods)
remove(AllPhotoperiods)

```

# Fits are added that are not sig different individually,  next add fits that were sig different dynamically
```{r Plot fits, fig.height= 20, fig.width= 20 }

scientific_10 <- function(x) {   parse(text=gsub("e\\+*", " %*% 10^", scales::scientific_format()(x))) }

MED4PURFITPlot <- AllFitsCombined |>
    filter(Strain == "MED4") |>
  unnest(cols = c(Data_all), names_sep = "_", keep_empty = TRUE) |>
  mutate(PhotoperiodLabel_title = "Photoperiod (h)",
         OxygenLabel = "Oxygen Concentration (µM)",
         Photoperiod = factor(Data_all_Photoperiod, levels = c("16", "12", "8", "4" )),
         O2_uM = factor(O2_uM, levels = c(250, 25, 2.5)),
         WL = factor(Data_all_WL, levels = c("WW", 660, 450))) %>%
  ggplot() +
  geom_point(aes(x = Data_all_PURPhotonDose_day, y = Data_all_ReplicateMean_deltaOD_Lmu_corr_day, shape = as.factor(Photoperiod), color = as.factor(WL), fill = as.factor(WL)), size = 4, alpha = 1.0) +
    geom_point(aes(x = Data_all_PURPhotonDose_day, y = Data_all_deltaOD_Lmu_corr_day, shape = as.factor(Photoperiod), color = as.factor(WL), fill = as.factor(WL)), size = 2, alpha = 0.8) +
  geom_line(aes(x = PURPhotonDose_day, y = .fitted), size = 1, linetype = "solid", data = . %>% unnest(Predict_pooled) %>%  filter(Strain == "MED4")) + 
   geom_line(aes(x = PURPhotonDose_day, y = .fitted), size = 0.5, linetype = "dotted", data = . %>% unnest(Predict_8) %>%  filter(O2_uM != 2.5, Strain == "MED4")) +
   geom_line(aes(x = PURPhotonDose_day, y = .fitted), size = 0.5, linetype = "aa", data = . %>% unnest(Predict_4) %>%  filter(O2_uM != 2.5, Strain == "MED4")) +
 geom_text(aes(x = 1000, y = 0.75, label= "A     MED4 HLI"), data = . %>% filter(O2_uM == 250, Strain == "MED4"), size = 3, hjust = 0, vjust = 0) +
  scale_x_continuous(label=scientific_10) +
   scale_y_continuous(breaks = c(0,0.25, 0.5, 0.75, 1.0), limits = c (0, 1.05)) +
  scale_colour_manual(values = MCMIXColours, labels = c("White LED", 660, 450)) +
  scale_linetype_manual(values = c("twodash","dashed", "dotted", "aa"), labels = c("16 hr", "12 hr", "8 hr", "4 hr")) +
   scale_shape_manual(values = PhotoperiodShapes)  +
  ggh4x::facet_nested(rows = vars(OxygenLabel, O2_uM), 
                    labeller = labeller(Strain = label_both, O2_uM = label_value, OxygenLabel = label_value)) +
 guides(linetype = guide_legend(nrow = 1),
        fill = "none",
        colour = guide_legend(nrow = 1)) +
 guides(shape = guide_legend(nrow = 1)) +
  theme_bw() +
   theme( strip.text = element_text(size = 10),
        axis.title.y = element_blank(),
        axis.title.x.bottom = element_blank(),
        axis.text.x = element_blank(),
        axis.text.y = element_text(size = 7, color = "black", angle = 0, hjust = 1.0),
        strip.background=element_rect(color="grey87", fill = "grey82"),
        # strip.background=element_rect(fill = "white"),
        legend.key.size = unit(1, 'cm'), #change legend key size
        legend.key.height = unit(1, 'cm'), #change legend key height
        legend.key.width = unit(0.5, 'cm'), #change legend key width
        legend.title =  element_text(size = 15),  #change legend title font size
        legend.text = element_text(size = 15),   #change legend text font size
        legend.position = "top") +
  labs( shape = "Photoperiod (h)",
        color = "Growth Waveband (nm)",
       linetype = "Photoperiod (h) "
      )
MED4PURFITPlot

SS120PURFITPlot <- AllFitsCombined |>
    filter(Strain == "SS120") |>
  unnest(cols = c(Data_all), names_sep = "_", keep_empty = TRUE) |>
  mutate(PhotoperiodLabel_title = "Photoperiod (h)",
         OxygenLabel = "Oxygen Concentration (µM)",
         Photoperiod = factor(Data_all_Photoperiod, levels = c("16", "12", "8", "4" )),
         O2_uM = factor(O2_uM, levels = c(250, 25, 2.5)),
         WL = factor(Data_all_WL, levels = c("WW", 660, 450))) %>%
  ggplot() +
  geom_point(aes(x = Data_all_PURPhotonDose_day, y = Data_all_ReplicateMean_deltaOD_Lmu_corr_day, shape = as.factor(Photoperiod), color = as.factor(WL), fill = as.factor(WL)), size = 4, alpha = 1.0) +
    geom_point(aes(x = Data_all_PURPhotonDose_day, y = Data_all_deltaOD_Lmu_corr_day, shape = as.factor(Photoperiod), color = as.factor(WL), fill = as.factor(WL)), size = 2, alpha = 0.8) +
  geom_line(aes(x = PURPhotonDose_day, y = .fitted), size = 1, linetype = "solid", data = . %>% unnest(Predict_pooled) %>%  filter(Strain == "SS120")) + 
   geom_line(aes(x = PURPhotonDose_day, y = .fitted), size = 0.5, linetype = "dashed", data = . %>% unnest(Predict_12) %>%  filter(O2_uM == 250, Strain == "SS120")) +
   geom_line(aes(x = PURPhotonDose_day, y = .fitted), size = 0.5, linetype = "dotted", data = . %>% unnest(Predict_8) %>%  filter(Strain == "SS120")) +
   geom_line(aes(x = PURPhotonDose_day, y = .fitted), size = 0.5, linetype = "aa", data = . %>% unnest(Predict_4) %>%  filter(Strain == "SS120")) +
 geom_text(aes(x = 1000, y = 0.75, label= "B     SS120 LLII/III"), data = . %>% filter(O2_uM == 250, Strain == "SS120"), size = 3, hjust = 0, vjust = 0) +
  scale_x_continuous(label=scientific_10) +
   scale_y_continuous(breaks = c(0,0.25, 0.5, 0.75, 1.0), limits = c (0, 1.05)) +
  # coord_cartesian(ylim = c (0, 0.07)) +
  scale_colour_manual(values = MCMIXColours, labels = c("White LED", 660, 450)) +
  scale_linetype_manual(values = c("twodash","dashed", "dotted", "aa"), labels = c("16 hr", "12 hr", "8 hr", "4 hr")) +
   scale_shape_manual(values = PhotoperiodShapes)  +
  ggh4x::facet_nested(rows = vars(OxygenLabel, O2_uM), 
                    labeller = labeller(Strain = label_both, O2_uM = label_value, OxygenLabel = label_value)) +
 # guides(linetype = guide_legend(nrow = 4),
 #        fill = "none",
 #        colour = guide_legend(nrow = 4)) +
 # guides(shape = guide_legend(nrow = 4)) +
  theme_bw() +
   theme( strip.text = element_text(size = 10),
        axis.title.y = element_blank(),
        axis.title.x.bottom = element_blank(),
        axis.text.x = element_blank(),
        axis.text.y = element_text(size = 7, color = "black", angle = 0, hjust = 1.0),
        strip.background=element_rect(color="grey87", fill = "grey82"),
        # strip.background=element_rect(fill = "white"),
        legend.key.size = unit(1, 'cm'), #change legend key size
        legend.key.height = unit(1, 'cm'), #change legend key height
        legend.key.width = unit(0.5, 'cm'), #change legend key width
        # legend.title =  element_text(size = 20),  #change legend title font size
        # legend.text = element_text(size = 20),   #change legend text font size
        legend.position = "none") +
  labs( shape = "Photoperiod (h)",
        color = "Growth Waveband (nm)",
       linetype = "Photoperiod (h) "
      )
   SS120PURFITPlot

MIT9313PURFITPlot <- AllFitsCombined |>
   filter(Strain == "MIT9313") |>
  unnest(cols = c(Data_all), names_sep = "_", keep_empty = TRUE) |>
  mutate(PhotoperiodLabel_title = "Photoperiod (h)",
         OxygenLabel = "Oxygen Concentration (µM)",
         Photoperiod = factor(Data_all_Photoperiod, levels = c("16", "12", "8", "4" )),
         O2_uM = factor(O2_uM, levels = c(250, 25, 2.5)),
         WL = factor(Data_all_WL, levels = c("WW", 660, 450))) %>%
  ggplot() +
  geom_point(aes(x = Data_all_PURPhotonDose_day, y = Data_all_ReplicateMean_deltaOD_Lmu_corr_day, shape = as.factor(Photoperiod), color = as.factor(WL), fill = as.factor(WL)), size = 4, alpha = 1.0) +
    geom_point(aes(x = Data_all_PURPhotonDose_day, y = Data_all_deltaOD_Lmu_corr_day, shape = as.factor(Photoperiod), color = as.factor(WL), fill = as.factor(WL)), size = 2, alpha = 0.8) +
  geom_line(aes(x = PURPhotonDose_day, y = .fitted), size = 1, linetype = "solid", data = . %>% unnest(Predict_pooled) %>%  filter(Strain == "MIT9313")) + 
  geom_line(aes(x = PURPhotonDose_day, y = .fitted), size = 0.5, linetype = "twodash", data = . %>% unnest(Predict_16) %>%  filter(O2_uM == 2.5, Strain == "MIT9313")) +
   geom_line(aes(x = PURPhotonDose_day, y = .fitted), size = 0.5, linetype = "dashed", data = . %>% unnest(Predict_12) %>%  filter(O2_uM != 2.5, Strain == "MIT9313")) +
   geom_line(aes(x = PURPhotonDose_day, y = .fitted), size = 0.5, linetype = "dotted", data = . %>% unnest(Predict_8) %>%  filter(O2_uM == 25, Strain == "MIT9313")) +
   geom_line(aes(x = PURPhotonDose_day, y = .fitted), size = 0.5, linetype = "aa", data = . %>% unnest(Predict_4) %>%  filter(Strain == "MIT9313")) +
  labs(x = "PUR (µmol photons" ~m^-2 ~d^-1~")") +
 geom_text(aes(x = 1000, y = 0.75, label= "C     MIT9313 LLIV"), data = . %>% filter(O2_uM == 250, Strain == "MIT9313"), size = 3, hjust = 0, vjust = 0) +
  scale_x_continuous(breaks = c(0, 1000000,2000000, 3000000, 4000000),label=scientific_10, limits = c (-100, 5000000)) +
   scale_y_continuous(breaks = c(0, 0.25, 0.5, 0.75, 1.0), limits = c (0, 1.05)) +
  scale_colour_manual(values = MCMIXColours, labels = c("White LED", 660, 450)) +
  scale_linetype_manual(values = c("solid","twodash","dashed", "dotted", "aa"), labels = c("Pooled","16 hr", "12 hr", "8 hr", "4 hr")) +
   scale_shape_manual(values = PhotoperiodShapes)  +
  ggh4x::facet_nested(rows = vars(OxygenLabel, O2_uM), 
                    labeller = labeller(Strain = label_both, WL = label_both, O2_uM = label_value, OxygenLabel = label_value)) +
 guides(linetype = guide_legend(nrow = 4),
        fill = "none",
        colour = guide_legend(nrow = 4)) +
 guides(shape = guide_legend(nrow = 4)) +
  theme_bw() +
   theme( strip.text = element_text(size = 10),
          axis.title.y = element_blank(),
          axis.title.x = element_text(size = 15, color = "black"),
          axis.text.x = element_text(size = 10, color = "black", angle = 0, hjust = 1.0),
          axis.text.y = element_text(size = 7, color = "black", angle = 0, hjust = 1.0),
        strip.background=element_rect(color="grey87", fill = "grey82"),
        # strip.background=element_rect(fill = "white"),
        legend.key.size = unit(1, 'cm'), #change legend key size
        legend.key.height = unit(1, 'cm'), #change legend key height
        legend.key.width = unit(0.5, 'cm'), #change legend key width
        # legend.title =  element_text(size = 20),  #change legend title font size
        # legend.text = element_text(size = 20),   #change legend text font size
        legend.position = "none") +
  labs( shape = "Photoperiod (h)",
        color = "Growth Waveband (nm)",
       linetype = "Photoperiod (h) "
      )
MIT9313PURFITPlot

```

# Combine the 3 strains into 1 plot 
```{r combine plots}

AllStrainPurFitsPlots <- plot_grid(MED4PURFITPlot, SS120PURFITPlot, MIT9313PURFITPlot, ncol = 1)

#create common x and y labels
y.grob <- textGrob("Chlorophyll Specific Growth Rate " ~  " ("  ~ d^-1~")",
                   gp = gpar(col="black", fontsize = 15), rot = 90)

#add to plot
AllStrainPurFitsPlots <- grid.arrange(arrangeGrob(AllStrainPurFitsPlots, left = y.grob))

```

Would like to add Compact Letter Displays (cld) but do not know how. Some examples:
https://stackoverflow.com/questions/70854552/any-other-options-besides-the-traditional-cld-bar-graph/70863531#70863531
https://stackoverflow.com/questions/72756490/loop-tukey-post-hoc-letters-extraction/73274274#73274274

Need test, other than nesting, to determine if Photoperiod has a significant influence (and interaction) on growth rate?

GrowthRate function of PUR, 
  separate models for Strain & O2_uM,

GrowthRate function of PUR in interaction term with Photoperiod and/or light colour
  separate models for Strain & O2_uM,

```{r, save PURFitPlots}

ggsave(file = file.path(PlotsOut, paste("AllStrainPurFitsPlots",".png",sep = "")), plot = AllStrainPurFitsPlots, device = NULL, scale = 1, height=25, width= 40, units = c("cm"),dpi = 300, limitsize = TRUE)

ggsave(file = file.path(PlotsOut, paste("AllStrainPurFitsPlots",".tiff",sep = "")), plot = AllStrainPurFitsPlots, device = NULL, scale = 1, height=25, width= 40, units = c("cm"),dpi = 300, limitsize = TRUE)

# ggsave(file = file.path(PlotsOut, paste("MED4PURFITPlot",".png",sep = "")), plot = MED4PURFITPlot, device = NULL, scale = 1, height=25, width= 40, units = c("cm"),dpi = 300, limitsize = TRUE)
# 
# ggsave(file = file.path(PlotsOut, paste("SS120PURFITPlot",".png",sep = "")), plot = SS120PURFITPlot, device = NULL, scale = 1, height=25, width= 40, units = c("cm"),dpi = 300, limitsize = TRUE)
# 
# ggsave(file = file.path(PlotsOut, paste("MIT9313PURFITPlot",".png",sep = "")), plot = MIT9313PURFITPlot, device = NULL, scale = 1, height=25, width= 40, units = c("cm"),dpi = 300, limitsize = TRUE)
# 
# ggsave(file = file.path(PlotsOut, paste("MED4PURFITPlot",".tiff",sep = "")), plot = MED4PURFITPlot, device = NULL, scale = 1, height=25, width= 40, units = c("cm"),dpi = 300, limitsize = TRUE)
# 
# ggsave(file = file.path(PlotsOut, paste("SS120PURFITPlot",".tiff",sep = "")), plot = SS120PURFITPlot, device = NULL, scale = 1, height=25, width= 40, units = c("cm"),dpi = 300, limitsize = TRUE)
# 
# ggsave(file = file.path(PlotsOut, paste("MIT9313PURFITPlot",".tiff",sep = "")), plot = MIT9313PURFITPlot, device = NULL, scale = 1, height=25, width= 40, units = c("cm"),dpi = 300, limitsize = TRUE)

```