### 2024-02-26 Code/OceanProteinPlots.Rmd
added clade to captions for OceanProtein figures

---
title: "OceanProteinPlots"
---
```{r libraries}
library(readr)
library(tidyverse)
library(ggplot2)
library(grid)
library(gridExtra)
```

Imported Aurora .csv large files from DropBox, saved as .RDS for future use and smaller files to post to GitHub
```{r files paths}
#/Users/dcampbel/Dropbox/OceanProteinProchloro/KM1128/2020/Filtered Datasets
#poi <- read_csv(file.path("..", "..", "..","..", "Dropbox", "OceanProteinProchloro", "KM1128", "2020", "Filtered_Datasets", "poi.csv"))
#pro_proteins <- read_csv(file.path("..", "..", "..","..", "Dropbox", "OceanProteinProchloro", "KM1128", "2020", "Filtered_Datasets", "pro_proteins.csv"))

#saveRDS(poi, file = file.path("..", "Data", "ProcessedData", "poi.RDS"))
#saveRDS(pro_proteins, file = file.path("..", "Data", "ProcessedData", "pro_proteins.RDS"))

#Read in previously imported and saved data converted from large .csv to smaller .RDS
poi <- readRDS(file.path("..", "Data", "ProcessedData", "ProcessedOceanProtein", "poi.RDS"))
pro_proteins <- readRDS(file.path("..", "Data", "ProcessedData", "ProcessedOceanProtein", "pro_proteins.RDS"))
```

```{r strains}
HL<-c("AS9601", "MIT9312", "MED4", "MIT9515","MIT9215", "MIT9301", "MIT9202")
LL<-c("MIT9313", "MIT9303", "SS120", "MIT9211", "NATL1A", "NATL2A")

strains_list<-c("MED4","MIT9515","AS9601","MIT9202","MIT9215","MIT9301","MIT9312","NATL1A","NATL2A", "MIT9211","SS120","MIT9303","MIT9313")
```

```{r set depth light and O2}
#estimated from LightPhotoperiodColourImaging/WaterAttenuation.Rmd

DepthPAR <- c(130, 70, 40)
PAR <- c(30, 90, 180)
names(DepthPAR) <- PAR

DepthPAR
names(DepthPAR)

O2_uM <- c(2, 25, 250)

```

To analyze for presence/absence of additional proteins like AOX need to extract EC numbers from best_hit_annotation; difficult format delimited by ';' with variable numbers of fields.
EC 1.10.3.11, ubiquinol oxidase (non-electrogenic) 	cyanide-resistant bd-type ubiquinol oxidase, cyanide-resistant ubiquinol oxidase, cytochrome bo3 ubiquinol oxidase, ubiquinol oxidase (non-electrogenic)

EC 1.10.3.11 not detected by scan
KEGG EC 1.6.5.9  not detected; issue with search terms?
```{r extract hit info}
# #detect EC 1.10.3.11
# pro_proteins2 <- pro_proteins |>
#   mutate(AOX = str_detect(string = best_hit_annotation, pattern = "1.10.3.11"))
# 
# head(pro_proteins2)

# #detect KEGG EC 1.6.5.9  
# pro_proteins2 <- pro_proteins |>
#    mutate(AOX = str_detect(string = best_hit_annotation, pattern = "1.6.5.9"))
#  
#  head(pro_proteins2)
#  unique(pro_proteins2$AOX)
```

```{r pro_proteins plots}

pro_proteins |>
  ggplot()+
  geom_point(aes(x = O2, y= depth)) +
  scale_x_continuous(limits = c(0, 260)) +
  scale_y_reverse(limits = c(260, 0)) +
  geom_vline(xintercept = O2_uM,   linewidth = 2, alpha = 0.2) +
  geom_hline(yintercept = DepthPAR, linewidth = 2, alpha = 0.2) +
  labs(title = "All detections of Prochlorococcus proteins from OceanProteinPlots.Rmd, from pro_proteins",
       caption = "Horizontals approximate experimental growth light levels; Verticals indicate experimental growth oxygen levels") +
  theme_bw()

#ggsave(plot = last_plot(), filename = (file.path("..", "OutPut", "ProchloroAllProteinsDepthO2.png")))

pro_proteins |>
    filter(strain %in% c("MIT9313","MED4", "SS120")) |>
  ggplot()+
  geom_point(aes(x = O2, y= depth)) +
  scale_x_continuous(limits = c(0, 260)) +
  scale_y_reverse(limits = c(260, 0)) +
  geom_vline(xintercept = O2_uM,   linewidth = 2, alpha = 0.2) +
  geom_hline(yintercept = DepthPAR, linewidth = 2, alpha = 0.2) +
  labs(title = "All detections of MIT9313,MED4, SS120 proteins from OceanProteinPlots.Rmd, from pro_proteins",
       caption = "Horizontals approximate experimental growth light levels; Verticals indicate experimental growth oxygen levels") +
  theme_bw()

poi |>
  filter(strain %in% c("MIT9313","MED4", "SS120")) |>
  ggplot()+
  geom_point(aes(x = O2, y= depth)) +
  scale_x_continuous(limits = c(0, 260)) +
  scale_y_reverse(limits = c(260, 0)) +
  geom_vline(xintercept = O2_uM,   linewidth = 2, alpha = 0.2) +
  geom_hline(yintercept = DepthPAR, linewidth = 2, alpha = 0.2) +
  facet_grid(vars(strain)) +
  theme(panel.grid.minor = element_line(size = 0), panel.grid.major = element_line(size = 2)) +
  theme_bw()

```

```{r depth O2 plots}

#consider including ribosomal target for photoinhibition?
#consider assembling 'pcb' supercomplex?
#https://stackoverflow.com/questions/32555531/how-to-italicize-part-one-or-two-words-of-an-axis-title

PhotosynthLabels <- c("PSII", "Cytb6f", "PSI", "ATP Synthase", "RUBISCO")
ProteinMetabLabels <- c("FtsH1", "FtsH2", "FtsH3", "FtsH4", "Ribosome")
StrainLabels <- c("MED4", "SS120", "MIT9313")

ElectronTrans <- c("psii", "cytb6f", "psi", "rubisco", "atp")

ProteinMetab <- c("ftsh1", "ftsh2", "ftsh3", "ftsh4", "ribosome")

poi |>
  filter(complex %in% ElectronTrans) |>
  filter(strain %in% StrainLabels) |>
  ggplot() +
  geom_point(aes(x = O2, y= depth)) +
  scale_x_continuous(limits = c(0, 260)) +
  scale_y_reverse(limits = c(260, 0)) +
  geom_vline(xintercept = O2_uM,   linewidth = 2, alpha = 0.1) +
  geom_hline(yintercept = DepthPAR, linewidth = 2, alpha = 0.1) +
  facet_grid(factor(strain, levels = c("MED4", "SS120", "MIT9313")) ~ factor(complex, levels = c("psii", "cytb6f", "psi", "rubisco", "atp"), labels = PhotosynthLabels)) +
  labs(y = "Depth (m)",
       x = expression("O"[2] ~ "(µM)")) +
    # ) +
  theme_bw()

 ggsave(plot = last_plot(), filename = (file.path("..", "OutPut", "Figures", "ProchloroPhotosynthDepthO2.png")))
 ggsave(plot = last_plot(), filename = (file.path("..", "OutPut", "Figures", "ProchloroPhotosynthDepthO2.tiff")))

poi |>
  filter(complex %in% ProteinMetab) |>
  filter(strain %in% StrainLabels) |>
  ggplot() +
  geom_point(aes(x = O2, y= depth)) +
  scale_x_continuous(limits = c(0, 260)) +
  scale_y_reverse(limits = c(260, 0)) +
  geom_vline(xintercept = O2_uM,   linewidth = 2, alpha = 0.1) +
  geom_hline(yintercept = DepthPAR, linewidth = 2, alpha = 0.1) +
  facet_grid(factor(strain, levels = c("MED4", "SS120", "MIT9313")) ~ factor(complex, levels = c("ftsh1", "ftsh2", "ftsh3", "ftsh4", "ribosome"), labels = ProteinMetabLabels)) +
   labs(y = "Depth (m)",
        x = expression("O"[2] ~ "(µM)")) +
  theme_bw()

 ggsave(plot = last_plot(), filename = (file.path("..", "OutPut", "Figures", "ProchloroProteinMetabDepthO2.png")))
 ggsave(plot = last_plot(), filename = (file.path("..", "OutPut", "Figures", "ProchloroProteinMetabDepthO2.tiff")))

```