### 2024-03-01 Code/ManuscriptGenome.Rmd
Update ManuscriptGenome.Rmd

---
title: "R Notebook"
output: html_notebook
---

```{r, import libraries, include = FALSE}
library(tidyverse)
library(glue)

googlesheets4::gs4_deauth()

```

```{r, assign filepaths, include = FALSE}
DataIn <- file.path("..","Data", "CleanData")
Figures <- file.path("..", "Output", "Figures")
# BLAST <- file.path("..", "BLAST")
```

```{r define notin}
`%notin%` <- negate(`%in%`)
```

combinedHits can be found on https://doi.org/10.5061/dryad.kh1893284. Filtered to only include Prochlorococcus genomes.
```{r}
# DataIn <- read_csv(file.path(Data, "combinedHits.csv")) %>% 
#   group_by(EC, genome) %>%
#   mutate(Gene_count = n()) %>%
#   ungroup() %>%
#   group_by(genome) %>%
#   mutate(TotalDetectedEnzymes = n()) %>%
#   ungroup()  %>%
#   distinct(genome, EC, .keep_all = TRUE)
# 
MetaData <- googlesheets4::read_sheet("https://docs.google.com/spreadsheets/d/1kXm0GvMdTRbF12cG5Bps-2e_4MTjgUX7iUOjqiu-_c0/edit#gid=0") %>%
  dplyr::select(c(Genus, species, Strain, FileName, Ome, Taxa, Rad1_um, Rad2_um, Rad3_um, Flagella, GenomeSize_mbp, GeneModels_count, Latitude, Longitude, Marine, PennateCentric, ColonySpecies)) %>%
  mutate(Strain = as.character(Strain))
# 
# O2_Enzymes <- read_csv(file.path(Data, "OxygenEnzymes.csv"))
# 
# 
# KOROSEnzymes <- O2_Enzymes %>% 
#   # select(EC, KO, PriSec, ) %>%
#   separate_rows(KO, sep = ";")
# 
# O2_Enzymes_Km <- googlesheets4::read_sheet("https://docs.google.com/spreadsheets/d/1wH2PZpsIDyA_WGfMr53Bm_HwmR0z5W93ed-LgRdLPVw/edit#gid=0") %>%
#   group_by(EC) %>%
#   mutate(AvgKm_mM = mean(Km_mM)) %>%
#   ungroup()
# 
# MergedData <- full_join(x = MetaData, y = KOROSEnzymes, by = character()) %>%
#   left_join(., DataIn, by = c("FileName" = "genome", "EC" = "EC")) %>%
#   left_join(., O2_Enzymes_Km, by = c("EC")) %>%
#   filter(!is.na(FileName))# %>%
#   # filter(Taxa == "Prokaryote") #%>%
#   # complete(., nesting(FileName, EC), fill = list(Gene_count = 0))
# 
# # KOCombinedHits <- DataIn %>%
# #   separate_rows(KEGG_ko, sep = ",") %>%
# #   mutate_at("KEGG_ko", str_replace, "ko:", "") %>%
# #   filter(is.na(EC),
# #          !is.na(KEGG_ko),
# #          KEGG_ko != "-") %>%
# #   left_join(., 
# #             O2_Enzymes %>%
# #               select(EC, KO) %>%
# #               separate_rows(KO, sep = ";"), 
# #             by = c("KEGG_ko" = "KO")) %>%
# #   filter(!is.na(EC)) %>%
# #   select(-EC) %>%
# #   rename(EC = EC) %>%
# #   distinct(genome, query_name, EC, .keep_all = FALSE)
# 
# 
# 
# MergedData %<>%
#   mutate(Rad1_um = as.numeric(Rad1_um),
#          Rad2_um = as.numeric(Rad2_um),
#          Rad3_um = as.numeric(Rad3_um),
#          SA_um2 = 4*pi*(((((Rad1_um*Rad2_um)^1.6)+((Rad1_um*Rad3_um)^1.6)+((Rad2_um*Rad3_um)^1.6))/3)^0.625),
#          Volume_um3 = (4/3)*pi*Rad1_um*Rad2_um*Rad3_um,
#          SAVol_um = SA_um2/Volume_um3,
#          Gene_count = as.numeric(Gene_count),
#          #Diameter = as.numeric(Rad1_um*2),
#          # Radius_um = dplyr::rowwise(as.numeric(Rad1_um)) %>% min(.$Rad1_um, .$Rad2_um, .$Rad3_um),
#          Radius_um = as.numeric(Rad1_um),
#          log_Radius_um = log10(as.numeric(Rad1_um)),
#          log_GenomeSize_mbp = log10(as.numeric(GenomeSize_mbp)),
#          log_GeneModels_count = log10(as.numeric(GeneModels_count)),
#          log_Gene_count = log10(as.numeric(Gene_count)+1),
#          Latitude = as.numeric(as.character(Latitude)),
#          abs_Latitude = abs(Latitude),
#          log_Volume_um3 = log10(as.numeric(Volume_um3)),
#          log_SA_um2 = log10(as.numeric(SA_um2)),
#          GeneModels_count = as.numeric(GeneModels_count),
#          log_SAVol_um = log10(as.numeric(SAVol_um)), 
#          Name = interaction(Genus, species, Strain, sep = " ")) %>%
#   filter(Taxa != "Symbiodinium", 
#          Marine == "yes",
#          FileName != "Nitzschia_palea-dna-trans") 
# 
# ProchlorococcusGenomeData <- MergedData %>% 
#   filter(Genus == "Prochlorococcus")
# 
# saveRDS(object = ProchlorococcusGenomeData, file = file.path("..", "Data", "ProchlorococcusGenomeData.Rds"))
```

```{r}
MergedData <- readRDS(file.path(DataIn, "ProchlorococcusGenomeData.Rds"))
```

```{r, , fig.width = 7, fig.height = 5}
MergedData %>%
  filter(!is.na(Gene_count),
         Taxa == "Prokaryote") %>%
  group_by(EC) %>%
  mutate(Km_mM_min = min(Km_mM), 
         Km_mM_max = max(Km_mM)) %>% 
  ggplot() +
  geom_errorbar(aes(x = reorder(EnzymeName, Km_mM), ymin = log10(Km_mM_min), ymax = log10(Km_mM_max))) +
  geom_point(aes(x = reorder(EnzymeName, Km_mM), y = log10(Km_mM), color = log10(Km_mM))) +
  geom_hline(yintercept = log10(1e-3), linetype = "dotted") +
  geom_hline(yintercept = log10(5e-3), linetype = "dotted") +
  geom_hline(yintercept = log10(230e-3), linetype = "dotted", color = "red") +
  geom_hline(yintercept = log10(290e-3), linetype = "dotted", color = "red") +
  scale_color_gradient(low = "#0980F6", high = "#F67F09", na.value = "black") +
  theme_bw() +
  ggpubr::rotate_x_text(angle = 70) 
```

```{r add flv to MergedData}
# Values are according to paper https://doi.org/10.3390/life5010716
MergedData <- MergedData %>% 
  add_row(Strain = "CCMP1375", Ome = "Genome", Taxa = "Prokaryote", EnzymeName = "Flavodiiron (Flv1a/3a)", Gene_count = 1, Estimated = 0, Genus = "Prochlorococcus", Km_mM  = 200/1000) %>% 
  add_row(Strain = "CCMP1986", Ome = "Genome", Taxa = "Prokaryote", EnzymeName = "Flavodiiron (Flv1a/3a)", Gene_count = 1, Estimated = 0, Genus = "Prochlorococcus", Km_mM  = 200/1000) %>% 
  add_row(Strain = "MIT9313", Ome = "Genome", Taxa = "Prokaryote", EnzymeName = "Flavodiiron (Flv1a/3a)", Gene_count = 1, Estimated = 0, Genus = "Prochlorococcus", Km_mM  = 200/1000) %>% 
  add_row(Strain = "NATL2A", Ome = "Genome", Taxa = "Prokaryote", EnzymeName = "Flavodiiron (Flv1a/3a)", Gene_count = 1, Estimated = 0, Genus = "Prochlorococcus", Km_mM  = 200/1000) %>% 
  add_row(Strain = "CCMP1375", Ome = "Genome", Taxa = "Prokaryote", EnzymeName = "Flavodiiron (Flv1a/3a)", Gene_count = 1, Estimated = 0, Genus = "Prochlorococcus", Km_mM  = 2/1000) %>% 
  add_row(Strain = "CCMP1986", Ome = "Genome", Taxa = "Prokaryote", EnzymeName = "Flavodiiron (Flv1a/3a)", Gene_count = 1, Estimated = 0, Genus = "Prochlorococcus", Km_mM  = 2/1000) %>% 
  add_row(Strain = "MIT9313", Ome = "Genome", Taxa = "Prokaryote", EnzymeName = "Flavodiiron (Flv1a/3a)", Gene_count = 1, Estimated = 0, Genus = "Prochlorococcus", Km_mM  = 2/1000) %>% 
  add_row(Strain = "NATL2A", Ome = "Genome", Taxa = "Prokaryote", EnzymeName = "Flavodiiron (Flv1a/3a)", Gene_count = 1, Estimated = 0, Genus = "Prochlorococcus", Km_mM  = 2/1000) %>% 
  add_row(Strain = "CCMP1375", Ome = "Genome", Taxa = "Prokaryote", EnzymeName = "Flavodiiron (Flv1a/3a)", Gene_count = 1, Estimated = 0, Genus = "Prochlorococcus", Km_mM  = 7/1000) %>% 
  add_row(Strain = "CCMP1986", Ome = "Genome", Taxa = "Prokaryote", EnzymeName = "Flavodiiron (Flv1a/3a)", Gene_count = 1, Estimated = 0, Genus = "Prochlorococcus", Km_mM  = 7/1000) %>% 
  add_row(Strain = "MIT9313", Ome = "Genome", Taxa = "Prokaryote", EnzymeName = "Flavodiiron (Flv1a/3a)", Gene_count = 1, Estimated = 0, Genus = "Prochlorococcus", Km_mM  = 7/1000) %>% 
  add_row(Strain = "NATL2A", Ome = "Genome", Taxa = "Prokaryote", EnzymeName = "Flavodiiron (Flv1a/3a)", Gene_count = 1, Estimated = 0, Genus = "Prochlorococcus", Km_mM  = 7/1000)
```

```{r, ProchlorococcusEnzymeKms, fig.cap = ProchlorococcusEnzymeKms_cap, fig.width = 20, fig.height = 10}
ProchlorococcusEnzymeKms_cap <- glue("**Comparison of log~10~(K~m~ for O~2~ in mM) vs.genes encoding O~2~ utilizing enzymes in *Prochlorococcus***, faceted by their respective strains.",
                                     "Symbol color corresponds to number of genes encoding O~2~ utilizing enzymes.",
                                     "Symbol shape corresponds to whether the K~m~ was measured in the respective enzyme, or inferred from the K~m~ of a similar enzyme.",
                                     "Black vertical line corresponds to the range of K~m~ for O~2~.",
                                     "Vertical translucent bars correspond to concentrations of O~2~ in which *Prochlorococcus* strains were grown.",
                                     "K~m~ values were obtained from other organisms, excluding Ribulose bisphosphate carboxylase.",
                                     "Symbol absence means no sequences were detected in the genome."
                                     )

ProchlorococcusEnzymeKms <- MergedData %>%
  filter(!is.na(Gene_count),
         Genus == "Prochlorococcus",
         Strain %in% c("CCMP1375", "CCMP1986", "MIT9313", "NATL2A"),
         !is.na(Estimated),
         Taxa == "Prokaryote") %>% 
  mutate(StrainLabel = case_when(Strain == "CCMP1375" ~ "SS120",
                                 Strain == "CCMP1986" ~ "MED4",
                                 Strain == "MIT9313" ~ "MIT9313",
                                 Strain == "NATL2A" ~ "NATL2A")) %>%
  mutate(StrainLabel = factor(StrainLabel, levels = c("MED4", "NATL2A", "SS120", "MIT9313"))) %>%
  group_by(EC) %>%
  mutate(Km_mM_min = min(Km_mM, na.rm = TRUE), 
         Km_mM_max = max(Km_mM, na.rm = TRUE),
         Gene_count = as.character(Gene_count)) %>%
  mutate(Km_mM = case_when(Genus == "Prochlorococcus" & EC == "4.1.1.39" ~ 1.4,
                           TRUE ~ Km_mM),
         Km_mM_min = case_when(Genus == "Prochlorococcus" & EC == "4.1.1.39" ~ 1.4,
                           TRUE ~ Km_mM_min),
         Km_mM_max = case_when(Genus == "Prochlorococcus" & EC == "4.1.1.39" ~ 1.4,
                           TRUE ~ Km_mM_max)) %>% 
  ungroup() %>% 
  ggplot() +
  geom_errorbar(aes(x = reorder(EnzymeName, Km_mM * 1000), ymin = log10(Km_mM_min* 1000), ymax = log10(Km_mM_max* 1000))) +
  annotate(geom = "rect", ymin = log10(5e-4 * 1000), ymax =  log10(5e-3 * 1000), xmin = -Inf, xmax = Inf, alpha = 0.3, fill = "blue") +
  annotate(geom = "rect", ymax = log10(5e-2 * 1000), ymin =  log10(5e-3 * 1000), xmin = -Inf, xmax = Inf, alpha = 0.3, fill = "green") +
  annotate(geom = "rect", ymax = log10(280e-3 * 1000), ymin =  log10(230e-3 * 1000), xmin = -Inf, xmax = Inf, alpha = 0.3, fill = "#FF0000") +
  geom_point(aes(x = reorder(EnzymeName, Km_mM* 1000), y = log10(Km_mM* 1000), shape = as.logical(Estimated), size = Gene_count), stroke = 1) +
  geom_vline(aes(xintercept = reorder(EnzymeName, Km_mM* 1000)), size = 2,  data = . %>% filter(is.na(Km_mM)), linetype = "solid") +
  geom_text(aes(x = 18.5, y = 0.25, label = "[O2] ≈ 2.5 µM"), color = "blue", size = 6) +
  geom_text(aes(x = 18.5, y = 1.2, label = "[O2] ≈ 25 µM"), color = "darkgreen", size = 6) +
  geom_text(aes(x = 18.5, y = 2.5, label = "[O2] ≈ 250 µM"), color = "#FF0000", size = 6) +
  ggh4x::facet_nested(rows = vars(StrainLabel)) +
  theme_bw() +
  ggpubr::rotate_x_text(angle = 80) +
  scale_shape_manual(values = c(19,8), labels = c("FALSE" = "Measured", "TRUE" = "Inferred")) +
  scale_color_brewer(palette="Dark2") +
  labs(# title = ~K[m] ~O[2] ~  "for *Prochlorococcus* Enzymes using Oxygen", # "Estimates of Affinities of *Prochlorococcus* Enzymes for Oxygen" # "Affinities for Phytoplankton Enzymes for Oxygen", 
       x = "Enzyme Name",
       y = ~log[10] ~ "("~K[m] ~O[2] ~"(µM))",
       size = "Gene Count",
       shape = ""~K[m]) +
  guides(shape = guide_legend(override.aes = list(size=5))) +
   theme(axis.title = element_text(size = 45, color = "black"),
         axis.title.x = element_text(vjust=8),
         axis.title.y = element_blank(),
         title = element_text(size = 30),
         axis.text.x = element_text(size = 30, color = "black"),
        axis.text.y = element_text(size = 20, color = "black"),
        strip.text = element_text(size = 30),
        legend.key.size = unit(2, 'cm'), #change legend key size
        # legend.key.height = unit(2, 'cm'), #change legend key height
        # legend.key.width = unit(1, 'cm'), #change legend key width
        legend.title =  element_text(size = 30),  #change legend title font size
        legend.text = element_text(size = 30))   #change legend text font size

ProchlorococcusEnzymeKms
```

```{r, ProchlorococcusEnzymeKms ggsave, include = FALSE} 
ggsave(filename = file.path(Figures, "ProchlorococcusEnzymeKms.png"), plot = ProchlorococcusEnzymeKms, height = 23, width = 20)
```

```{r, ProchlorococcusEnzyme,  fig.width = 7, fig.height = 5}
# ProchlorococcusEnzyme <- MergedData %>%
#   filter(!is.na(Gene_count)) %>%
#   # filter(PriSec == "Primary") %>%
#   filter(FileName == "Prochlorococcus_marinus_MIT9313-aa-gen"| FileName == "Prochlorococcus_marinus_pastoris_CCMP1986-aa-gen" | FileName == "Prochlorococcus_marinus_marinus_CCMP1375-aa-gen") %>%
#   # group_by(EC) %>%
#   # mutate(ECNumberTotal = sum(Gene_count)) %>%
#   # ungroup() %>% View()
#   # filter(ECNumberTotal != 0) %>%
#   ggplot() +
#   geom_point(aes(x = reorder(EnzymeName, Km_mM), y = interaction(Genus, species, Strain, sep = " "), size = Gene_count, color = log10(AvgKm_mM))) +
#   facet_grid(rows = vars(Genus), scales = "free", space = "free") +
#   scale_color_gradient(low = "#0980F6", high = "#F67F09", na.value = "black") +
#   scale_x_discrete(position = "top") +
#   scale_radius(breaks = c(1,2,3,4,5), range = c(2.5,5)) +
#   theme_bw() +
#   ggpubr::rotate_x_text(hjust = 0, angle = 70) +
#   labs(x = "Enzyme Name",
#        y = "Organism",
#        size = "Gene Count", 
#        color = "log10(Km (mM))")
# 
# ProchlorococcusEnzyme
```

```{r, ProchlorococcusEnzyme ggsave, include = FALSE}
# ggsave(filename = file.path(Figures, "ProchlorococcusEnzyme.png"), plot = ProchlorococcusEnzyme, height = 10, width = 10)
```

```{r, DNARepair}
DNARepair <- MetaData %>% 
  left_join(., DataIn, by = c("FileName" = "genome")) %>% 
  filter(Genus == "Prochlorococcus") %>% 
  separate_rows(GOs, sep = ",") %>% 
  filter(Strain %in% c("CCMP1986", "CCMP1375", "MIT9313", "NATL2A")) %>% 
  filter(GOs %in% c("GO:0006281", "GO:0000716", "GO:0000725", "GO:0006284", "GO:0006289", "GO:0006298", "GO:0006302", "GO:0009432", "GO:0006974", "GO:0003904", "GO:0003914") | EC %in% c("4.1.99.3", "4.1.99.13", "1.13.11.33", "3.2.2.17", "5.6.2.3", "6.5.1.1", "6.5.1.2"))

DNARepairFig <- DNARepair %>% 
  mutate(EnzymeName = case_when(EC == "3.2.2.21" ~ "DNA-3-methyladenine glycosylase II",
                                EC == "3.1.22.4" ~ "crossover junction endodeoxyribonuclease",
                                EC == "6.5.1.2" ~ "DNA ligase (NAD+)",
                                EC == "6.5.1.1" ~ "DNA ligase (ATP)",
                                EC == "4.1.99.3" ~ "deoxyribodipyrimidine photo-lyase")) %>% 
  mutate(StrainLabel = case_when(Strain == "CCMP1375" ~ "SS120",
                                 Strain == "CCMP1986" ~ "MED4",
                                 Strain == "MIT9313" ~ "MIT9313",
                                 Strain == "NATL2A" ~ "NATL2A")) %>%
  mutate(StrainLabel = factor(StrainLabel, levels = c("MED4", "NATL2A", "SS120", "MIT9313"))) %>%
  ggplot() +
  geom_point(aes(x = EnzymeName, y = StrainLabel, size = (Gene_count))) + #y = interaction(Genus, species, StrainLabel, sep = " ")
  # facet_grid(rows = vars(Genus), scales = "free", space = "free") +
  # scale_color_gradient(low = "#0980F6", high = "#F67F09", na.value = "black") +
  scale_x_discrete(position = "bottom") +
  scale_radius(breaks = c(1,2,3,4,5), range = c(1,4)) +
  theme_bw() +
  ggpubr::rotate_x_text(hjust = 1, angle = 70)  +
  labs(x = "Enzyme Name",
       #y = "Organism",
       color = "Gene Count",
       size = "Gene Count") +
  facet_grid(rows = vars(StrainLabel), scale = "free") +
   theme(axis.text = element_text(size = 7, color = "black"),
         legend.position = "right",
         axis.text.y = element_blank(),
         axis.title.y = element_blank(),
         axis.title.x = element_text(vjust=10),
        axis.ticks.y = element_blank(),
        strip.text = element_text(size = 6),
        legend.box.background = element_rect(colour = "black")
  )

DNARepairFig
```

```{r, DNARepairFig ggsave, include = FALSE} 
ggsave(filename = file.path(Figures, "DNARepairFig.png"), plot = DNARepairFig, height =5 , width = 5)
```

```{r import LightEnzymes, include = FALSE}
LightEnzymes <- read_csv(file.path(Data, "LightEnzymes.csv")) %>% 
  select(-c(`...4`, `...5`)) %>% 
  filter(is.na(Remove))
```

```{r LightEnzymes, include=FALSE}
LightEnzymesList <- LightEnzymes$EC
LightEnzymesMerged <- MetaData %>% 
  filter(Genus == "Prochlorococcus" | Taxa == "Diatom") %>% 
  left_join(., DataIn, by = c("FileName" = "genome")) %>% 
  filter(EC %in% LightEnzymesList) %>% 
  left_join(., LightEnzymes, by = c("EC"), relationship = "many-to-many") %>% 
  filter(Description != "KM Value") %>% 
  separate_rows(GOs, sep = ",") %>% 
  mutate(EnzymeName = case_when(EC == "1.3.1.33" ~ "(light-dependent) protochlorophyllide reductase",
                                EC == "2.4.1.14" ~ "sucrose-phosphate synthase",
                                EC == "2.5.1.32" ~ "15-cis-phytoene synthase",
                                EC == "4.6.1.1" ~ "adenylate cyclase",
                                EC == "4.4.1.5" ~ "lactoylglutathione lyase",
                                EC == "4.2.3.5" ~ "chorismate synthase",
                                EC == "2.7.1.148" ~ "4-(cytidine 5'-diphospho)-2-C-methyl-D-erythritol kinase",
                                EC == "4.2.99.18" ~ "DNA-(apurinic or apyrimidinic site) lyase",
                                EC == "3.4.11.1" ~ "leucyl aminopeptidase",
                                EC == "2.4.1.255" ~ "protein O-GlcNAc transferase",
                                EC == "3.1.4.4" ~ "phospholipase D",
                                EC == "1.1.3.20" ~ "long-chain-alcohol oxidase",
                                EC == "1.7.1.1" ~ "nitrate reductase (NADH)",
                                EC == "2.3.1.87" ~ "aralkylamine N-acetyltransferase",
                                EC == "1.3.8.1" ~ "short-chain acyl-CoA dehydrogenase",
                                EC == "3.4.21.19" ~ "glutamyl endopeptidase",
                                EC == "4.6.1.2" ~ "guanylate cyclase",
                                EC == "1.7.7.2" ~ "ferredoxin-nitrate reductase",
                                EC == "4.2.1.84" ~ "nitrile hydratase",
                                EC == "1.10.3.3" ~ "L-ascorbate oxidase",
                                EC == "3.2.1.80" ~ "fructan beta-fructosidase",
                                EC == "2.3.1.37" ~ "5-aminolevulinate synthase",
                                EC == "3.1.4.35" ~ "3',5'-cyclic-GMP phosphodiesterase",
                                EC == "4.1.99.3" ~ "deoxyribodipyrimidine photo-lyase"))

```

```{r LightEnzymesFig, fig.cap = ProchlorococcusLightEnzymesFig_cap, echo = TRUE, fig.height = 5, fig.width = 8}
ProchlorococcusLightEnzymesFig_cap <- glue("**Summary of genes encoding enzymes activated or inhibited by light in *Prochlorococcus*.**",
                                           "Rows correspond to strains of *Prochlorococcus*", 
                                           "Symbol color and size correspond to counts of genes.",
                                           "Symbol absence means no sequences were detected in the genome."
                                           )

ProchlorococcusLightEnzymesFig <- LightEnzymesMerged %>% 
  filter(Strain %in% c("CCMP1986", "CCMP1375", "MIT9313", "NATL2A")) %>% 
  mutate(StrainLabel = case_when(Strain == "CCMP1375" ~ "SS120",
                                 Strain == "CCMP1986" ~ "MED4",
                                 Strain == "MIT9313" ~ "MIT9313",
                                 Strain == "NATL2A" ~ "NATL2A")) %>%
  mutate(StrainLabel = factor(StrainLabel, levels = c("MED4", "NATL2A", "SS120", "MIT9313"))) %>% 
  mutate(Compound = case_when(Compound == "blue-light regulator 2" ~ "light",
                              TRUE ~ Compound)) %>% 
  ggplot() +
  geom_point(aes(x = fct_reorder(EnzymeName, desc(Gene_count)), y = StrainLabel, size = as.character(Gene_count))) +
  ggh4x::facet_nested(cols = vars(Description, Compound),rows = vars(StrainLabel),  scale = "free", space = "free") + #

  labs(x = "Enzyme Name",
       # y = "Organism",
       size = "Gene Count",
       color = "Gene Count") +
  # scale_size_manual(values = c(1,2,3,4), breaks)
  theme_bw() +
  ggpubr::rotate_x_text(hjust = 1, angle = 70) +
  theme(axis.text = element_text(size = 7, color = "black"),
        axis.text.y = element_blank(),
        axis.title.y = element_blank(),
        axis.ticks.y = element_blank(),
  # axis.title = element_text(size = 40, color = "black"),
  #        title = element_text(size = 30),
  #     # axis.text.x = element_text(size = 30, color = "black"),
  #       axis.text.y = element_text(size = 30, color = "black"),
        strip.text = element_text(size = 6),
  legend.position = "bottom",
  legend.key.size = unit(1, 'cm'),
  legend.box.background = element_rect(colour = "black")
  #       # legend.key.size = unit(10, 'cm'), #change legend key size
  #       # legend.key.height = unit(2, 'cm'), #change legend key height
  #       # legend.key.width = unit(1, 'cm'), #change legend key width
  #       legend.title =  element_text(size = 30),  #change legend title font size
  #       legend.text = element_text(size = 30))   #change legend text font size
  )

 ProchlorococcusLightEnzymesFig
```

```{r, save ProchlorococcusLightEnzymesFig, ggsave, include = FALSE} 
ggsave(filename = file.path(Figures, "ProchlorococcusLightEnzymesFig.png"), plot = ProchlorococcusLightEnzymesFig, height = 6, width = 8)
```

```{r}
LightGOs <- MergedData %>% 
  separate_rows(GOs, sep = ",") %>% 
  filter(GOs %in% c("GO:0010018", "GO:0009785", "GO:0009882", "GO:0031516", "GO:0031517", "GO:0009883"))

LightKOs <- MergedData %>% 
  separate_rows(KEGG_ko, sep = ",") %>% 
  filter(KEGG_ko %in% c("ko:K02295", "ko:K12120", "ko:K12122", "ko:K12121", "ko:K12123"))
```

```{r, fig.width = 7, fig.height = 7}
MergedData %>%
  filter(!is.na(Gene_count)) %>%
  filter(PriSec == "Primary") %>%
  filter(Genus == "Prochlorococcus") %>%
  # filter(FileName == "Prochlorococcus_marinus_MIT9313-aa-gen"| FileName == "Prochlorococcus_marinus_pastoris_CCMP1986-aa-gen" | FileName == "Prochlorococcus_marinus_marinus_CCMP1375-aa-gen") %>%
  # group_by(EC) %>%
  # mutate(ECNumberTotal = sum(Gene_count)) %>%
  # ungroup() %>% View()
  # filter(ECNumberTotal != 0) %>%
  ggplot() +
  geom_point(aes(x = reorder(EnzymeName, Km_mM), y = FileName, size = Gene_count, color = log10(AvgKm_mM))) +
  facet_grid(rows = vars(Genus), scales = "free", space = "free") +
  scale_color_gradient(low = "#0980F6", high = "#F67F09", na.value = "black") +
  scale_x_discrete(position = "top") +
  scale_radius(breaks = c(1,2,3,4,5), range = c(1,4)) +
  theme_bw() +
  ggpubr::rotate_x_text(hjust = 0) 
```

```{r, fig.width = 7, fig.height = 7}
MergedData %>%
  filter(!is.na(Gene_count)) %>%
  filter(PriSec == "Primary") %>%
  # filter(FileName == "Prochlorococcus_marinus_MIT9313-aa-gen"| FileName == "Prochlorococcus_marinus_pastoris_CCMP1986-aa-gen" | FileName == "Prochlorococcus_marinus_marinus_CCMP1375-aa-gen") %>%
  # group_by(EC) %>%
  # mutate(ECNumberTotal = sum(Gene_count)) %>%
  # ungroup() %>% View()
  # filter(ECNumberTotal != 0) %>%
  ggplot() +
  geom_point(aes(x = reorder(EnzymeName, Km_mM), y = FileName, size = Gene_count, color = log10(AvgKm_mM))) +
  facet_grid(rows = vars(Genus), scales = "free", space = "free") +
  scale_color_gradient(low = "#0980F6", high = "#F67F09", na.value = "black") +
  scale_x_discrete(position = "top") +
  scale_radius(breaks = c(1,2,3,4,5), range = c(1,4)) +
  theme_bw() +
  ggpubr::rotate_x_text(hjust = 0) 
```