### 2024-03-15 Code/Plots_GAM.Rmd
changed how plots are generated and saved such that 'run all' is used instead of changing strain name in the chunk that generates and saves plots. Prone to errors.

---
title: "ManuscriptGAM"
author:
- Max Berthold
- Mireille Savoie
output:
bookdown::html_document2:
    code_folding: show
    keep_md: yes
    toc: TRUE
    toc_float: TRUE
    toc_depth: 6
    fig_caption: yes
bibliography: [Manuscript_O2, RPackages.bib]
csl: plos-one.csl
editor_options: 
  markdown: 
    wrap: 72
---

# Introduction
This .Rmd produces a Generalized Additive Model (GAM) to examine the relationship between the chlorophyll proxy (ΔOD) growth rate across the blue and red spectral wavebands, photoperiod and PAR levels for each P.marinus ecotype in this study. The gam function from the R package mgcv (Wood, 2022) was used to model the growth rate with smoothing terms to indicate the 90, 50 and 10% quantiles. "PICO_Merged_GrowthFitsPURPAR.Rds" is called in from \Data\CleanData\CleanedMCData folder and generated GAM plots are saved in \Output\Figures folder as .png.

# Set variables for project
```{r project data}
FP1 <- ".."
Project <- "PICO"
FileID <- "GAM"
Expert <- "MiSa" #MiSa SySl
GamFit <- "PICO_Merged_GrowthFitsPURPAR.Rds"

#"..", takes up a level in the directory path
DataIn <- file.path("..","Data", "CleanData")
DataOut <- file.path("..","Data", "CleanData")
PlotsOut <- file.path("..","Output", "Figures")
PlotsOutPLOS <- file.path("..", "OutPut", "PLOSFiguresRound1")

is_all_numeric <- function(x) { 
  !any(is.na(suppressWarnings(as.numeric(na.omit(x))))) & is.character(x) 
}
#https://stackoverflow.com/questions/22772279/converting-multiple-columns-from-character-to-numeric-format-in-r

GamPara <- "Lmu" # Stabds for growth estimates determined by logistic model

```

```{r load libraries, message=FALSE}
library(tidyverse)
library(lubridate)
library(broom)
library(knitr)
library(minpack.lm)
library(nlstools)
library(nlsMicrobio)
library(plotly)
library(mgcv)
library(tidymv)
library(CCA)
library(vegan)
library(cowplot)
library(glue)
library(gtable)
library(car)
library(MASS) # estimating linear models through generalized least squares (GLS), fitting negative binomial linear models, the robust fitting of linear models, and Kruskal's non-metric multidimensional scaling.
library(faraway) # LM package
library(grid)
library(gridExtra)
library(ggh4x)
library(ggspectra) 
library(photobiologyWavebands) 
# library(photobiology)
library(metR)
library(Rcpp) #'Rcpp' package provides R functions as well as C++ classes which offer a seamless integration of R and C++.

```

# Read data nad select relevant columns
```{r read data and select relevant columns}

GrowthFits <- readRDS(file = file.path(DataIn, GamFit))

GrowthFits <- GrowthFits %>% 
 dplyr::select(c(Filename, PrimaryOperator, Tube, SampleID, Run, Photoperiod, Strain, WL, O2_Category, O2_uM, LightShape, Pur_ue, Par_ue, PURPhotonDose_day, PARPhotonDose_day, OD720_Lmu_corr, OD720_Lmax, OD720_Lintercept, OD720_Lmu_se, OD720_Lmax_se, OD720_Lintercept_se, deltaOD_Lmu_corr, deltaOD_Lmax, deltaOD_Lintercept,  deltaOD_Lmu_se, deltaOD_Lmax_se, deltaOD_Lintercept_se))

ReplNo <- GrowthFits %>%
  distinct(Photoperiod, Strain, O2_uM, WL, Par_ue)

```

# make df longer for GAM
```{r set up df for GAM}
 Pivot_Longer_OD720_est <- GrowthFits %>% 
  dplyr::select(-c(OD720_Lmu_se, OD720_Lmax_se, OD720_Lintercept_se, deltaOD_Lmu_corr, deltaOD_Lmax, deltaOD_Lintercept,  deltaOD_Lmu_se, deltaOD_Lmax_se, deltaOD_Lintercept_se)) %>%
   pivot_longer(c(OD720_Lmu_corr, OD720_Lmax, OD720_Lintercept), names_to = c("Wavelength", "term"), values_to = c("estimate"), names_sep = '\\_')

# OD720 pivot longer Gmax, Gmu, Glag , Lmu, Lmax and Lintercept standard error
Pivot_Longer_OD720_se <- GrowthFits %>%
  dplyr::select(-c(OD720_Lmu_corr, OD720_Lmax, OD720_Lintercept, deltaOD_Lmu_corr, deltaOD_Lmax, deltaOD_Lintercept,  deltaOD_Lmu_se, deltaOD_Lmax_se, deltaOD_Lintercept_se)) %>%
   pivot_longer(c(OD720_Lmu_se, OD720_Lmax_se, OD720_Lintercept_se), names_to = c("Wavelength", "term"), values_to = c("se"), names_sep = '\\_')

# Join OD720 estimate and se DF 
Longer_OD720_Joined <- full_join(Pivot_Longer_OD720_est, Pivot_Longer_OD720_se)

remove(Pivot_Longer_OD720_est, Pivot_Longer_OD720_se)

# deltaOD pivot longer Gmax, Gmu, Glag, Lmu, Lmax and Lintercept estimates
Pivot_Longer_deltaOD_est <- GrowthFits %>%
  dplyr::select(-c(OD720_Lmu_corr, OD720_Lmax, OD720_Lintercept, OD720_Lmu_se, OD720_Lmax_se, OD720_Lintercept_se,  deltaOD_Lmu_se, deltaOD_Lmax_se, deltaOD_Lintercept_se)) %>%
   pivot_longer(c(deltaOD_Lmu_corr, deltaOD_Lmax, deltaOD_Lintercept), names_to = c("Wavelength", "term"), values_to = c("estimate"), names_sep = '\\_')

# deltaOD pivot longer Gmax, Gmu, Glag , Lmu, Lmax and Lintercept standard error
Pivot_Longer_deltaOD_se <- GrowthFits %>%
  dplyr::select(-c(OD720_Lmu_corr, OD720_Lmax, OD720_Lintercept, OD720_Lmu_se, OD720_Lmax_se, OD720_Lintercept_se, deltaOD_Lmu_corr, deltaOD_Lmax, deltaOD_Lintercept)) %>%
   pivot_longer(c(deltaOD_Lmu_se, deltaOD_Lmax_se, deltaOD_Lintercept_se), names_to = c("Wavelength", "term"), values_to = c("se"), names_sep = '\\_')

# Join deltaOD estimate and se DF 
Longer_deltaOD_Joined <- full_join(Pivot_Longer_deltaOD_est, Pivot_Longer_deltaOD_se)

remove(Pivot_Longer_deltaOD_est, Pivot_Longer_deltaOD_se)

#Join OD720 and deltaOD fits data
ComGrowthFits <- rbind(Longer_OD720_Joined, Longer_deltaOD_Joined) %>%
  mutate_if(is.numeric, round, digits = 3) %>% 
  filter(Strain %in% c("MED4", "SS120", "MIT9313"),
         WL %in% c(450, 660, "WW"))

remove(Longer_OD720_Joined, Longer_deltaOD_Joined)

glimpse(ComGrowthFits)

```

```{r DF preparation Prochloroccus}
GamWL = "deltaOD"
GamVar = "450"

#DummyDF for MED4; MED4 is a high light strain, based on literature it can grow up to 800 uE; we never exposed it to such high light levels and will not impose an upper end, but only a lower, i.e. Par_ue = 0, Lmu = 0, there are no reliable points for low O2 in shorter photoperiods
Strain <- c("MED4")
Photoperiod <- c(4, 8, 12, 16)
Par_ue<- c(0)
O2_uM <- c(2.5, 25.0, 250.0)
WL <- c(450, 530, 660, "WW")
estimate <- c(0)
term <- GamPara
Wavelength <- c("deltaOD", "OD720")
Tube <- c(9)

DummyDF_MED4 <- expand_grid(
  Strain, Photoperiod, Par_ue, O2_uM, WL, estimate, Wavelength, term, Tube
  ) %>%
  mutate(Lmu_seFlag = 1) 
DummyDF_Expand_MED4 <- rbind(DummyDF_MED4, DummyDF_MED4, DummyDF_MED4)

#DummyDF for MIT9313; MIT9313 is a low light strain, based on literature it can only grow up to 300 uE WW; MIT9313 consistently died at various light colors and oxygen levels latest at ~PAR 400; it can safely be assumed it dies at par_ue = 0
Strain <- c("MIT9313")
Photoperiod <- c(4, 8, 12, 16)
Par_ue<- c(0)
O2_uM <- c(2.5, 25.0, 250.0)
WL <- c(450, 530, 660, "WW")
estimate <- c(0)
term <- GamPara
Wavelength <- c("deltaOD", "OD720")
Tube <- c(9)

DummyDF_MIT9313 <- expand_grid(
  Strain, Photoperiod, Par_ue, O2_uM, WL, estimate, Wavelength, term, Tube
  ) %>%
  mutate(Lmu_seFlag = 1) 
DummyDF_Expand_MIT9313 <- rbind(DummyDF_MIT9313, DummyDF_MIT9313, DummyDF_MIT9313)

#DummyDF for SS120; SS120 is a low light strain, based on literature it can only grow up to 300 uE WW; SS120 consistently died at various light colors and oxygen levels latest at ~PAR 400; it can safely be assumed it dies at par_ue = 0
Strain <- c("SS120")
Photoperiod <- c(4, 8, 12, 16)
Par_ue<- c(0, 400)
O2_uM <- c(2.5, 25.0, 250.0)
WL <- c(450, 530, 660, "WW")
estimate <- c(0)
term <- GamPara
Wavelength <- c("deltaOD", "OD720")
Tube <- c(9)

DummyDF_SS120 <- expand_grid(
  Strain, Photoperiod, Par_ue, O2_uM, WL, estimate, Wavelength, term, Tube
  ) %>%
  mutate(Lmu_seFlag = 1) 
DummyDF_Expand_SS120 <- rbind(DummyDF_SS120, DummyDF_SS120, DummyDF_SS120)

DummyDF_Expand <- rbind(DummyDF_Expand_SS120, DummyDF_Expand_MIT9313, DummyDF_Expand_MED4)

PicoGrowthFits <-    ComGrowthFits # full_join(ComGrowthFits, DummyDF_Expand)

PicoLmuAb <- PicoGrowthFits %>%
  filter(term %in% GamPara) %>%
  filter(Wavelength %in% GamWL) %>%
  filter(WL %in% GamVar) %>%
  mutate(Strain = as.factor(Strain))

PicoLogGamAb <-  gam(estimate ~ s(Par_ue,  by = Strain, k = 3) + s(Photoperiod,  by = Strain, k = 3) + ti(Photoperiod, Par_ue, by = Strain, k = 3), data = PicoLmuAb, method = "REML")
summary(PicoLogGamAb)

PicoLogPar <- gam(estimate ~ s(Par_ue,  by = Strain, k = 3), data = PicoLmuAb, method = "REML")

PicoLogPhotPer <- gam(estimate ~ s(Photoperiod,  by = Strain, k = 3), data = PicoLmuAb, method = "REML")

AIC(PicoLogPar, PicoLogPhotPer, PicoLogGamAb)

ModelResid <- resid(PicoLogGamAb)
gam.check(PicoLogGamAb)
concurvity(PicoLogGamAb)
plot(PicoLogGamAb, residuals = TRUE, pch = 1, all.terms = TRUE)

#AIC of Gompertz-fitted model comparably low as to only-par and only-light model. 
#Next step, fitting Strain-specific models. 
#MED4 model to Par_ue-response falsely fits, due to badly fitted µ-rates -> cultures died, but model tried to fit anyway. Maybe filter everything below a certain R^2 in the fit?

```

# Generate GAM for red light conditions for all strains using chlorophyll proxy logistic growth rate. 
```{r GAM separate niches hypervolume for red light}
GamWL = c("deltaOD", "OD720") #deltaOD OD720
Proxy= "[~Chl]" #~Chl ~Cells 
GamParaMod = "Lmu"
GamVar = c("660")
ErrorTolGrowth = 30 # defined for the growth estimates
ErrorTol = 30 # defined for the GAM fit
# OxyLev = 25

#https://stackoverflow.com/questions/49570452/is-it-possible-to-specify-lower-bound-in-response-variable-during-smooth-with-ga 
possibGAM = possibly(.f = mgcv::gam, otherwise = NULL)
possibPRED = possibly(.f = mgcv::predict.gam, otherwise = NULL)

PAR_Photoperiod_Model_predi <- expand_grid(
  Par_ue = seq(from=min(10), 
              to=max(200), 
              length.out = 100),
  Photoperiod = seq(from=min(2), 
              to=max(18), 
              length.out = 100)
)

PicoGAM <- PicoGrowthFits %>%
  ungroup() %>%
  dplyr::mutate(Lmu_seFlag = case_when(se/estimate*100 < ErrorTolGrowth ~ 1, #filters faulty data where Standarderror larger +/- Errortolerance
                         FALSE ~ 0)) %>%
  dplyr::filter(
                estimate == 0 |
                estimate != 0 & Lmu_seFlag == 1,
                term %in% GamParaMod, 
                Wavelength %in% GamWL
                ) %>%
  mutate(estimate = estimate*24, #to convert growth from per hour to per day
         se = se*24) %>%
  group_by(Strain, Wavelength, WL, O2_uM) %>%
  nest() %>%
  dplyr::arrange(Strain, Wavelength, WL, O2_uM) %>%
  mutate(GamModel = map(data, ~possibGAM(estimate ~ s(Par_ue, k = 3) + s(Photoperiod, k = 3) + ti(Photoperiod, Par_ue, k = 3),  data = .x,
                                    family = "gaussian"(link = "log"), 
                                    method = "REML")),
         predicted = map(GamModel, ~possibPRED(.x, newdata = PAR_Photoperiod_Model_predi, type = 'response', se.fit=TRUE)),
         augmented = map(GamModel, ~augment(.x))
         )

PicoGAM_Pred <- PicoGAM %>%
  dplyr::select(Strain, Wavelength, WL, O2_uM, predicted) %>%
  dplyr::filter(predicted != "NULL") %>%
  rowwise() %>%
  mutate(db = list(cbind(predicted, PAR_Photoperiod_Model_predi))) %>%
  dplyr::select(-predicted) %>%
  unnest(col = db) %>%
  group_by(Strain, Wavelength, WL, O2_uM) %>%
  dplyr::filter(fit < 1.2) %>% #with two exceptions, growth rates were never higher than 1.2; GAM model creates some unrealistic high growth rates due to low data availability
  mutate(
    IsFitGood = case_when(se.fit/fit*100 < ErrorTol ~ 1, #flags faulty data where Standarderror larger +/- Errortolerance
                         se.fit/fit*100 > ErrorTol ~ 0
                    )
    #se_erro_per = se.fit/fit*100
  # ,fit = case_when(fit >= quantile(fit, 0.75, na.rm = TRUE) ~ fit*1, #leaves only top-10%, that means fastest growth rates in the dataframe
  #                 FALSE ~ fit)
  ) %>% 
  dplyr::filter(IsFitGood == 1)

```

# Generate GAM for blue light conditions for all strains using chlorophyll proxy logistic growth rate. 
```{r GAM separate niches hypervolume for blue light}
GamWL = c("deltaOD", "OD720") #deltaOD OD720
Proxy= "[~Chl]" #~Chl ~Cells 
GamParaMod = "Lmu"
GamVar = c("450")
ErrorTolGrowth = 30 # defined for the growth estimates
ErrorTol = 30 # defined for the GAM fit
# OxyLev = 25

#https://stackoverflow.com/questions/49570452/is-it-possible-to-specify-lower-bound-in-response-variable-during-smooth-with-ga 
possibGAM = possibly(.f = mgcv::gam, otherwise = NULL)
possibPRED = possibly(.f = mgcv::predict.gam, otherwise = NULL)

PAR_Photoperiod_Model_predi <- expand_grid(
  Par_ue = seq(from=min(10), 
              to=max(200), 
              length.out = 100),
  Photoperiod = seq(from=min(2), 
              to=max(18), 
              length.out = 100)
)

PicoGAM <- PicoGrowthFits %>%
  ungroup() %>%
  dplyr::mutate(Lmu_seFlag = case_when(se/estimate*100 < ErrorTolGrowth ~ 1, #filters faulty data where Standarderror larger +/- Errortolerance
                         FALSE ~ 0)) %>%
  dplyr::filter(
                estimate == 0 |
                estimate != 0 & Lmu_seFlag == 1,
                term %in% GamParaMod, 
                Wavelength %in% GamWL
                ) %>%
  mutate(estimate = estimate*24, #to convert growth from per hour to per day
         se = se*24) %>%
  group_by(Strain, Wavelength, WL, O2_uM) %>%
  nest() %>%
  dplyr::arrange(Strain, Wavelength, WL, O2_uM) %>%
  mutate(GamModel = map(data, ~possibGAM(estimate ~ s(Par_ue, k = 3) + s(Photoperiod, k = 3) + ti(Photoperiod, Par_ue, k = 3),  data = .x,
                                    family = "gaussian"(link = "log"), 
                                    method = "REML")),
         predicted = map(GamModel, ~possibPRED(.x, newdata = PAR_Photoperiod_Model_predi, type = 'response', se.fit=TRUE)),
         augmented = map(GamModel, ~augment(.x))
         )

PicoGAM_Pred <- PicoGAM %>%
  dplyr::select(Strain, Wavelength, WL, O2_uM, predicted) %>%
  dplyr::filter(predicted != "NULL") %>%
  rowwise() %>%
  mutate(db = list(cbind(predicted, PAR_Photoperiod_Model_predi))) %>%
  dplyr::select(-predicted) %>%
  unnest(col = db) %>%
  group_by(Strain, Wavelength, WL, O2_uM) %>%
  dplyr::filter(fit < 1.2) %>% #with two exceptions, growth rates were never higher than 1.2; GAM model creates some unrealistic high growth rates due to low data availability
  mutate(
    IsFitGood = case_when(se.fit/fit*100 < ErrorTol ~ 1, #flags faulty data where Standarderror larger +/- Errortolerance
                         se.fit/fit*100 > ErrorTol ~ 0
                    )
    #se_erro_per = se.fit/fit*100
  # ,fit = case_when(fit >= quantile(fit, 0.75, na.rm = TRUE) ~ fit*1, #leaves only top-10%, that means fastest growth rates in the dataframe
  #                 FALSE ~ fit)
  ) %>% 
  dplyr::filter(IsFitGood == 1)

```

# Plot red light GAM for MED4
``` {r MED4GAM Plot Red WL}
GamStrain = "MED4" 
GamColor = "Red" 
WL_nm = 660 
PlotWL <- "deltaOD"

maxMu <- PicoGAM_Pred %>%
  ungroup() %>%
  filter(Wavelength %in% PlotWL) %>%
  slice(which.max(fit)) #base::max does not work with package photobiology loaded simultaneously 
maxMu <- maxMu$fit

minMu <- PicoGAM_Pred %>%
  ungroup() %>%
  filter(Wavelength %in% PlotWL) %>%
  slice(which.min(fit)) #base::max does not work with package photobiology loaded simultaneously 
  #        fit == min(fit, na.rm = TRUE)) %>%
  # distinct(fit)
minMu <- minMu$fit

PicoGAM_Plot_Prep <- PicoGAM_Pred %>%
  filter(Wavelength %in% PlotWL,
         Strain %in% GamStrain,
         WL %in% WL_nm) %>%
  mutate(
         OxygenLabel = "Oxygen concentration (uM)",
         O2_level = case_when(O2_uM == 2.5 ~ "low",
                              O2_uM == 25 ~ "intermediate",
                              O2_uM == 250 ~ "high"),
         Color_name = case_when(WL == "450" ~ 'Blue',
                                WL == "530" ~ 'Green',
                                WL == "660" ~ 'Red',
                                WL == "WW" ~ 'White')) %>%
  unite(WL_O2, Color_name, O2_level, sep = "_", remove = FALSE)

WL_O2_Colors = c(Blue = w_length2rgb(450), green = w_length2rgb(530), red = w_length2rgb(660), white= "#FFCC00")

ColourDF <- data.frame(
  ColorName = c('Blue', 'Green', 'Red', 'White'),
  WL_O2_Colors = c( w_length2rgb(450), w_length2rgb(530), w_length2rgb(660), "#FFCC00")
) %>%
  filter(ColorName %in% GamColor) %>%
  dplyr::select(WL_O2_Colors)

ContourBreaks <- PicoGAM_Plot_Prep %>%
  filter(
         Strain %in% GamStrain,
         Wavelength %in% PlotWL
  ) %>%
  group_by(O2_uM) %>%
  mutate(upper10 = quantile(fit, 0.9, na.rm = TRUE),
         median50 = quantile(fit, 0.5, na.rm = TRUE),
         lower10 = quantile(fit, 0.1, na.rm = TRUE)) %>%
  ungroup() %>%
  dplyr::select(O2_uM, upper10, median50, lower10) %>%
  distinct() %>%
  pivot_longer(cols = c(upper10, median50, lower10), names_to = 'quantile', values_to = 'quantileValue') %>%
  mutate(quantileValue = round(quantileValue, digits = 2))

##############
###plot
#https://stackoverflow.com/questions/28775036/ggplot-line-graph-with-na-values
#https://stackoverflow.com/questions/6906661/ggplot2-make-missing-value-in-geom-tile-not-blank
#https://stackoverflow.com/questions/24822621/multiple-legends-in-a-plot-with-geom-tile
#https://stackoverflow.com/questions/43395219/plotting-multiple-layers-with-geom-raster-or-geom-tile-or-geom-rect
#https://stackoverflow.com/questions/59949635/ggplot-align-grid-lines-with-end-of-geom-tile
#https://stackoverflow.com/questions/22945651/remove-space-between-plotted-data-and-the-axes
#https://stackoverflow.com/questions/5293715/how-to-use-greek-symbols-in-ggplot2

HyperChlPlotLow <- PicoGAM_Plot_Prep %>%
  filter(
         Strain %in% GamStrain,
         Wavelength %in% PlotWL,
         O2_uM == "2.5"
  )
ContourLines <- ContourBreaks %>%
  filter(O2_uM == "2.5") 
ContourLines <- ContourLines$quantileValue

(ContourLowO2 <- ggplot(na.omit(HyperChlPlotLow), aes(Photoperiod, Par_ue, z = fit, fill = fit)) + geom_tile() +
  scale_fill_gradientn(name = bquote("µ ("~ d^-1~")"), colours=c("white", ColourDF), limits  = range(minMu, maxMu), na.value = "white") +
    metR::geom_contour2(aes(label = stat(level)),
                        breaks = ContourLines,
                        skip = 0) +
    scale_x_continuous(limits = c(4, 16), breaks = c(4, 8, 12, 16)) +
    scale_y_continuous(limits = c(10, 200)) +
     xlab(NULL)+
      ylab(NULL)+
    theme_classic() +
    theme(axis.text=element_text(size=15),
          axis.title.x = element_text(size=20),
           axis.title.y = element_text(size=35)))

HyperChlPlotInt <- PicoGAM_Plot_Prep %>%
  filter(
         Strain %in% GamStrain,
         Wavelength %in% PlotWL,
         O2_uM == "25"
  )
ContourLines <- ContourBreaks %>%
  filter(O2_uM == "25") 
ContourLines <- ContourLines$quantileValue

(ContourIntO2 <- ggplot(na.omit(HyperChlPlotInt), aes(Photoperiod, Par_ue, z = fit, fill = fit)) + geom_tile() +
  scale_fill_gradientn(name = bquote("µ ("~ d^-1~")"), colours=c("white", ColourDF), limits  = range(minMu, maxMu), na.value = "white") +
    metR::geom_contour2(aes(label = stat(level)),
                        breaks = ContourLines,
                        skip = 0) +
    scale_x_continuous(limits = c(4, 16), breaks = c(4, 8, 12, 16)) +
    scale_y_continuous(limits = c(10, 200)) +
    xlab(NULL)+
      ylab(NULL)+
   theme_classic() +
  theme(axis.text=element_text(size=15),
          axis.title.x = element_text(size=20),
           axis.title.y = element_text(size=23)))

HyperChlPlotHigh <- PicoGAM_Plot_Prep %>%
  filter(
         Strain %in% GamStrain,
         Wavelength %in% PlotWL,
         O2_uM == "250"
  )
ContourLines <- ContourBreaks %>%
  filter(O2_uM == "250") 
ContourLines <- ContourLines$quantileValue

(ContourHighO2 <- ggplot(na.omit(HyperChlPlotHigh), aes(Photoperiod, Par_ue, z = fit, fill = fit)) + geom_tile() +
    scale_fill_gradientn(name = bquote("µ ("~ d^-1~")"), colours=c("white", ColourDF), limits  = range(minMu, maxMu), na.value = "white") +
    metR::geom_contour2(aes(label = stat(level)),
                        breaks = ContourLines,
                        skip = 0) +
    scale_x_continuous(limits = c(4, 16), breaks = c(4, 8, 12, 16)) +
    scale_y_continuous(limits = c(10, 200)) +
    xlab(NULL)+
      ylab(NULL)+
    #https://stackoverflow.com/questions/15074127/use-expression-with-a-variable-r
    theme_classic() +
   theme(axis.text=element_text(size=15),
          axis.title.x = element_text(size=20),
           axis.title.y = element_text(size=35)))

g_legend<-function(a.gplot){
  tmp <- ggplot_gtable(ggplot_build(a.gplot))
  leg <- which(sapply(tmp$grobs, function(x) x$name) == "guide-box")
  legend <- tmp$grobs[[leg]]
  legend
}

(g_dummy1 <- ggplot(subset(HyperChlPlotLow, Color_name == w_length2rgb(WL_nm)), aes(Photoperiod, Par_ue, fill=fit)) + geom_tile()
+ scale_fill_gradientn(name = bquote("µ ("~ d^-1~")"), colours=c("white", w_length2rgb(WL_nm)),
  limits  = range(minMu, 1.0), na.value = "white") + theme_classic())
legend1 <- g_legend(g_dummy1)
(g_dummy2 <- ggplot(subset(HyperChlPlotInt, Color_name == w_length2rgb(WL_nm)), aes(Photoperiod, Par_ue, fill=fit)) + geom_tile()
+ scale_fill_gradientn(name = bquote("µ ("~ d^-1~")"), colours=c("white", w_length2rgb(WL_nm)),
  limits  = range(minMu, 1.0), na.value = "white") + theme_classic())
legend2 <- g_legend(g_dummy2)
(g_dummy3 <- ggplot(subset(HyperChlPlotHigh, Color_name == w_length2rgb(WL_nm)), aes(Photoperiod, Par_ue, fill=fit)) + geom_tile()
+ scale_fill_gradientn(name = bquote("µ ("~ d^-1~")"), colours=c("white", w_length2rgb(WL_nm)),
  limits  = range(minMu, 1.0), na.value = "white") + theme_classic())
legend3 <- g_legend(g_dummy3)

# comment rows that are not included in grid arrange 
LayoutNiche <- rbind(
                      # c(3, 3, 3, 3, 6), # For 2.5 uM O2     # remove comment if want 2.5 uM O2 plotted
                     # c(3, 3, 3, 3, 6), # For 2.5 uM O2    # remove comment if want 2.5 uM O2 plotted
                     c(2, 2, 2, 2, 5),   # For 25 uM O2
                     c(2, 2, 2, 2, 5),   # For 25 uM O2
                     c(1, 1, 1, 1, 4),   # For 250 uM O2
                     c(1, 1, 1, 1, 4)    # For 250 uM O2
                     )

#LayoutNiche <- rbind(c(1, 1, 1, 2, 2, 2, 3, 3, 3, 4))

p3_Red <- grid.arrange(
                       # ContourLowO2 +   labs(title = bquote("C.   " ~ .(GamStrain) ~ '  2.5 µM' ~ "[" ~ O[2]~ "]"), size = 25) + theme(legend.position = 'none'),   # remove comment if want 2.5 uM O2 plotted
                       ContourIntO2 + labs(title = bquote("B.   " ~ .(GamStrain) ~ '  25 µM' ~ "[" ~ O[2]~ "]"), size = 25) +
                         theme(legend.position = 'none'), 
                       ContourHighO2 + labs(title = bquote("A.   " ~ .(GamStrain) ~ '  250 µM' ~ "[" ~ O[2]~ "]")) + 
                         theme(legend.position = 'none'),
                       # legend1,   # remove comment if want 2.5 uM O2 plotted
                        legend2, 
                        legend3, 
                       layout_matrix = LayoutNiche)

```

# Plot blue light GAM for MED4
``` {r MED4GAM Plot Blue WL}
GamStrain = "MED4" 
GamColor = "Blue"
WL_nm = 450 
PlotWL <- "deltaOD"

maxMu <- PicoGAM_Pred %>%
  ungroup() %>%
  filter(Wavelength %in% PlotWL) %>%
  slice(which.max(fit)) #base::max does not work with package photobiology loaded simultaneously 
maxMu <- maxMu$fit

minMu <- PicoGAM_Pred %>%
  ungroup() %>%
  filter(Wavelength %in% PlotWL) %>%
  slice(which.min(fit)) #base::max does not work with package photobiology loaded simultaneously 
  #        fit == min(fit, na.rm = TRUE)) %>%
  # distinct(fit)
minMu <- minMu$fit

PicoGAM_Plot_Prep <- PicoGAM_Pred %>%
  filter(Wavelength %in% PlotWL,
         Strain %in% GamStrain,
         WL %in% WL_nm) %>%
  mutate(
         OxygenLabel = "Oxygen concentration (uM)",
         O2_level = case_when(O2_uM == 2.5 ~ "low",
                              O2_uM == 25 ~ "intermediate",
                              O2_uM == 250 ~ "high"),
         Color_name = case_when(WL == "450" ~ 'Blue',
                                WL == "530" ~ 'Green',
                                WL == "660" ~ 'Red',
                                WL == "WW" ~ 'White')) %>%
  unite(WL_O2, Color_name, O2_level, sep = "_", remove = FALSE)

WL_O2_Colors = c(Blue = w_length2rgb(450), green = w_length2rgb(530), red = w_length2rgb(660), white= "#FFCC00")

ColourDF <- data.frame(
  ColorName = c('Blue', 'Green', 'Red', 'White'),
  WL_O2_Colors = c( w_length2rgb(450), w_length2rgb(530), w_length2rgb(660), "#FFCC00")
) %>%
  filter(ColorName %in% GamColor) %>%
  dplyr::select(WL_O2_Colors)

ContourBreaks <- PicoGAM_Plot_Prep %>%
  filter(
         Strain %in% GamStrain,
         Wavelength %in% PlotWL
  ) %>%
  group_by(O2_uM) %>%
  mutate(upper10 = quantile(fit, 0.9, na.rm = TRUE),
         median50 = quantile(fit, 0.5, na.rm = TRUE),
         lower10 = quantile(fit, 0.1, na.rm = TRUE)) %>%
  ungroup() %>%
  dplyr::select(O2_uM, upper10, median50, lower10) %>%
  distinct() %>%
  pivot_longer(cols = c(upper10, median50, lower10), names_to = 'quantile', values_to = 'quantileValue') %>%
  mutate(quantileValue = round(quantileValue, digits = 2))

##############
###plot
#https://stackoverflow.com/questions/28775036/ggplot-line-graph-with-na-values
#https://stackoverflow.com/questions/6906661/ggplot2-make-missing-value-in-geom-tile-not-blank
#https://stackoverflow.com/questions/24822621/multiple-legends-in-a-plot-with-geom-tile
#https://stackoverflow.com/questions/43395219/plotting-multiple-layers-with-geom-raster-or-geom-tile-or-geom-rect
#https://stackoverflow.com/questions/59949635/ggplot-align-grid-lines-with-end-of-geom-tile
#https://stackoverflow.com/questions/22945651/remove-space-between-plotted-data-and-the-axes
#https://stackoverflow.com/questions/5293715/how-to-use-greek-symbols-in-ggplot2

HyperChlPlotLow <- PicoGAM_Plot_Prep %>%
  filter(
         Strain %in% GamStrain,
         Wavelength %in% PlotWL,
         O2_uM == "2.5"
  )
ContourLines <- ContourBreaks %>%
  filter(O2_uM == "2.5") 
ContourLines <- ContourLines$quantileValue

(ContourLowO2 <- ggplot(na.omit(HyperChlPlotLow), aes(Photoperiod, Par_ue, z = fit, fill = fit)) + geom_tile() +
    scale_fill_gradientn(name = bquote("µ ("~ d^-1~")"), colours=c(low = "white", intermediate = "steelblue1", high = "royalblue4"), limits  = range(minMu, maxMu), na.value = "white") +
    metR::geom_contour2(aes(label = stat(level)),
                        breaks = ContourLines,
                        skip = 0) +
    scale_x_continuous(limits = c(4, 16), breaks = c(4, 8, 12, 16)) +
    scale_y_continuous(limits = c(10, 200)) +
    xlab(NULL)+
      ylab(NULL)+
   theme_classic() +
   theme(axis.text=element_text(size=15),
          axis.title.x = element_text(size=20),
           axis.title.y = element_text(size=35)))

HyperChlPlotInt <- PicoGAM_Plot_Prep %>%
  filter(
         Strain %in% GamStrain,
         Wavelength %in% PlotWL,
         O2_uM == "25"
  )
ContourLines <- ContourBreaks %>%
  filter(O2_uM == "25") 
ContourLines <- ContourLines$quantileValue

(ContourIntO2 <- ggplot(na.omit(HyperChlPlotInt), aes(Photoperiod, Par_ue, z = fit, fill = fit)) + geom_tile() +
    scale_fill_gradientn(name = bquote("µ ("~ d^-1~")"), colours=c(low = "white", intermediate = "steelblue1", high = "royalblue4"), limits  = range(minMu, maxMu), na.value = "white") +
    metR::geom_contour2(aes(label = stat(level)),
                        breaks = ContourLines,
                        skip = 0) +
    scale_x_continuous(limits = c(4, 16), breaks = c(4, 8, 12, 16)) +
    scale_y_continuous(limits = c(10, 200)) +
    xlab(NULL)+
      ylab(NULL)+ 
  theme_classic() +
   theme(axis.text=element_text(size=15),
          axis.title.x = element_text(size=20),
           axis.title.y = element_text(size=35)))

HyperChlPlotHigh <- PicoGAM_Plot_Prep %>%
  filter(
         Strain %in% GamStrain,
         Wavelength %in% PlotWL,
         O2_uM == "250"
  )
ContourLines <- ContourBreaks %>%
  filter(O2_uM == "250") 
ContourLines <- ContourLines$quantileValue

(ContourHighO2 <- ggplot(na.omit(HyperChlPlotHigh), aes(Photoperiod, Par_ue, z = fit, fill = fit)) + geom_tile() +
    scale_fill_gradientn(name = bquote("µ ("~ d^-1~")"), colours=c(low = "white", intermediate = "steelblue1", high = "royalblue4"), limits  = range(minMu, maxMu), na.value = "white") +
    metR::geom_contour2(aes(label = stat(level)),
                        breaks = ContourLines,
                        skip = 0) +
    scale_x_continuous(limits = c(4, 16), breaks = c(4, 8, 12, 16)) +
    scale_y_continuous(limits = c(10, 200)) +
    xlab(NULL)+
      ylab(NULL)+ #https://stackoverflow.com/questions/15074127/use-expression-with-a-variable-r
theme_classic() +
   theme(axis.text=element_text(size=15),
          axis.title.x = element_text(size=20),
           axis.title.y = element_text(size=35)))

g_legend<-function(a.gplot){
  tmp <- ggplot_gtable(ggplot_build(a.gplot))
  leg <- which(sapply(tmp$grobs, function(x) x$name) == "guide-box")
  legend <- tmp$grobs[[leg]]
  legend
}

(g_dummy1 <- ggplot(subset(HyperChlPlotLow, Color_name == w_length2rgb(WL_nm)), aes(Photoperiod, Par_ue, fill=fit)) + geom_tile()
 + scale_fill_gradientn(name = bquote("µ ("~ d^-1~")"), colours=c(low = "white", intermediate = "steelblue1", high = "royalblue4"),
 limits  = range(minMu, 1.0), na.value = "white") + theme_classic())
legend1 <- g_legend(g_dummy1)
(g_dummy2 <- ggplot(subset(HyperChlPlotInt, Color_name == w_length2rgb(WL_nm)), aes(Photoperiod, Par_ue, fill=fit)) + geom_tile()
 + scale_fill_gradientn(name = bquote("µ ("~ d^-1~")"), colours=c(low = "white", intermediate = "steelblue1", high = "royalblue4"),
  limits  = range(minMu, 1.0), na.value = "white") + theme_classic())
legend2 <- g_legend(g_dummy2)
(g_dummy3 <- ggplot(subset(HyperChlPlotHigh, Color_name == w_length2rgb(WL_nm)), aes(Photoperiod, Par_ue, fill=fit)) + geom_tile()
+ scale_fill_gradientn(name = bquote("µ ("~ d^-1~")"), colours=c(low = "white", intermediate = "steelblue1", high = "royalblue4"),
  limits  = range(minMu, 1.0), na.value = "white") + theme_classic())
legend3 <- g_legend(g_dummy3)

# comment rows that are not included in grid arrange 
LayoutNiche <- rbind(
                     # c(3, 3, 3, 3, 6), # For 2.5 uM O2   # remove comment if want 2.5 uM O2
                     # c(3, 3, 3, 3, 6), # For 2.5 uM O2   # remove comment if want 2.5 uM O2
                     c(2, 2, 2, 2, 5),   # For 25 uM O2
                     c(2, 2, 2, 2, 5),   # For 25 uM O2
                     c(1, 1, 1, 1, 4),   # For 250 uM O2
                     c(1, 1, 1, 1, 4)    # For 250 uM O2
                     )

#LayoutNiche <- rbind(c(1, 1, 1, 2, 2, 2, 3, 3, 3, 4))

p3_Blue <- grid.arrange(
                         # ContourLowO2 +   labs(title = bquote("F.   " ~ .(GamStrain) ~ '  2.5 µM' ~ "[" ~ O[2]~ "]"), size = 25) + theme(legend.position = 'none', axis.title.y = element_blank(), axis.text.y = element_blank()),  
                        ContourIntO2 +  labs(title = bquote("D.   " ~ .(GamStrain) ~ '  25 µM' ~ "[" ~ O[2]~ "]"), size = 25) +
                          theme(legend.position = 'none', axis.title.y = element_blank(), axis.text.y = element_blank()), 
                        ContourHighO2 + labs(title = bquote("C.   " ~ .(GamStrain) ~ '  250 µM' ~ "[" ~ O[2]~ "]")) + 
                          theme(legend.position = 'none', axis.title.y = element_blank(), axis.text.y = element_blank()),
                        # legend1,    # remove comment if want 2.5 uM O2
                        legend2, 
                        legend3, 
                        layout_matrix = LayoutNiche)

```

```{r combine red and blue light MED4GAM plots}

#combine red and blue GAM plots
CombinedGAM <- plot_grid(p3_Red, p3_Blue, nrow = 1, align = "v")

# common x-axis title
x.grob <- textGrob("Photoperiod (h)", 
                   gp = gpar(fontsize = 20))
# common y-axis title
y.grob <- textGrob("PAR (µmol photons"~ m^-2~ ~s^-1~ ")", 
                   gp = gpar(fontface="bold", fontsize = 20), rot = 90)

# add titles to plot
LabelledCombinedGam <- grid.arrange(arrangeGrob(CombinedGAM, left = y.grob, bottom = x.grob))

```

```{r save MED4GAM plots}
# MED4GAM,png for manuscript
 ggsave(file = file.path(paste(PlotsOut, sep = "/"), paste(GamStrain, "GAM", ".png", sep = ""), fsep = .Platform$file.sep), plot = LabelledCombinedGam,  height = 10, width = 10)

# MED4GAM is Fig4.tiff for PLOSONE submission
 ggsave(file = file.path(PlotsOutPLOS, "Fig4.tiff", fsep = .Platform$file.sep), plot = LabelledCombinedGam, device = NULL, scale = 1,  height = 10, width = 10, compression = "lzw")

```

# Plot red light GAM for SS120
``` {r SS120GAM Plot Red WL}
GamStrain = "SS120" 
GamColor = "Red" 
WL_nm = 660 
PlotWL <- "deltaOD"

maxMu <- PicoGAM_Pred %>%
  ungroup() %>%
  filter(Wavelength %in% PlotWL) %>%
  slice(which.max(fit)) #base::max does not work with package photobiology loaded simultaneously 
maxMu <- maxMu$fit

minMu <- PicoGAM_Pred %>%
  ungroup() %>%
  filter(Wavelength %in% PlotWL) %>%
  slice(which.min(fit)) #base::max does not work with package photobiology loaded simultaneously 
  #        fit == min(fit, na.rm = TRUE)) %>%
  # distinct(fit)
minMu <- minMu$fit

PicoGAM_Plot_Prep <- PicoGAM_Pred %>%
  filter(Wavelength %in% PlotWL,
         Strain %in% GamStrain,
         WL %in% WL_nm) %>%
  mutate(
         OxygenLabel = "Oxygen concentration (uM)",
         O2_level = case_when(O2_uM == 2.5 ~ "low",
                              O2_uM == 25 ~ "intermediate",
                              O2_uM == 250 ~ "high"),
         Color_name = case_when(WL == "450" ~ 'Blue',
                                WL == "530" ~ 'Green',
                                WL == "660" ~ 'Red',
                                WL == "WW" ~ 'White')) %>%
  unite(WL_O2, Color_name, O2_level, sep = "_", remove = FALSE)

WL_O2_Colors = c(Blue = w_length2rgb(450), green = w_length2rgb(530), red = w_length2rgb(660), white= "#FFCC00")

ColourDF <- data.frame(
  ColorName = c('Blue', 'Green', 'Red', 'White'),
  WL_O2_Colors = c( w_length2rgb(450), w_length2rgb(530), w_length2rgb(660), "#FFCC00")
) %>%
  filter(ColorName %in% GamColor) %>%
  dplyr::select(WL_O2_Colors)

ContourBreaks <- PicoGAM_Plot_Prep %>%
  filter(
         Strain %in% GamStrain,
         Wavelength %in% PlotWL
  ) %>%
  group_by(O2_uM) %>%
  mutate(upper10 = quantile(fit, 0.9, na.rm = TRUE),
         median50 = quantile(fit, 0.5, na.rm = TRUE),
         lower10 = quantile(fit, 0.1, na.rm = TRUE)) %>%
  ungroup() %>%
  dplyr::select(O2_uM, upper10, median50, lower10) %>%
  distinct() %>%
  pivot_longer(cols = c(upper10, median50, lower10), names_to = 'quantile', values_to = 'quantileValue') %>%
  mutate(quantileValue = round(quantileValue, digits = 2))

##############
###plot
#https://stackoverflow.com/questions/28775036/ggplot-line-graph-with-na-values
#https://stackoverflow.com/questions/6906661/ggplot2-make-missing-value-in-geom-tile-not-blank
#https://stackoverflow.com/questions/24822621/multiple-legends-in-a-plot-with-geom-tile
#https://stackoverflow.com/questions/43395219/plotting-multiple-layers-with-geom-raster-or-geom-tile-or-geom-rect
#https://stackoverflow.com/questions/59949635/ggplot-align-grid-lines-with-end-of-geom-tile
#https://stackoverflow.com/questions/22945651/remove-space-between-plotted-data-and-the-axes
#https://stackoverflow.com/questions/5293715/how-to-use-greek-symbols-in-ggplot2

HyperChlPlotLow <- PicoGAM_Plot_Prep %>%
  filter(
         Strain %in% GamStrain,
         Wavelength %in% PlotWL,
         O2_uM == "2.5"
  )
ContourLines <- ContourBreaks %>%
  filter(O2_uM == "2.5") 
ContourLines <- ContourLines$quantileValue

(ContourLowO2 <- ggplot(na.omit(HyperChlPlotLow), aes(Photoperiod, Par_ue, z = fit, fill = fit)) + geom_tile() +
  scale_fill_gradientn(name = bquote("µ ("~ d^-1~")"), colours=c("white", ColourDF), limits  = range(minMu, maxMu), na.value = "white") +
    metR::geom_contour2(aes(label = stat(level)),
                        breaks = ContourLines,
                        skip = 0) +
    scale_x_continuous(limits = c(4, 16), breaks = c(4, 8, 12, 16)) +
    scale_y_continuous(limits = c(10, 200)) +
     xlab(NULL)+
    theme_classic() +
    theme(axis.text=element_text(size=15),
          axis.title.x = element_text(size=20),
           axis.title.y = element_text(size=35)))

HyperChlPlotInt <- PicoGAM_Plot_Prep %>%
  filter(
         Strain %in% GamStrain,
         Wavelength %in% PlotWL,
         O2_uM == "25"
  )
ContourLines <- ContourBreaks %>%
  filter(O2_uM == "25") 
ContourLines <- ContourLines$quantileValue

(ContourIntO2 <- ggplot(na.omit(HyperChlPlotInt), aes(Photoperiod, Par_ue, z = fit, fill = fit)) + geom_tile() +
  scale_fill_gradientn(name = bquote("µ ("~ d^-1~")"), colours=c("white", ColourDF), limits  = range(minMu, maxMu), na.value = "white") +
    metR::geom_contour2(aes(label = stat(level)),
                        breaks = ContourLines,
                        skip = 0) +
    scale_x_continuous(limits = c(4, 16), breaks = c(4, 8, 12, 16)) +
    scale_y_continuous(limits = c(10, 200)) +
    xlab(NULL)+
      ylab(NULL)+
   theme_classic() +
  theme(axis.text=element_text(size=15),
          axis.title.x = element_text(size=20),
           axis.title.y = element_text(size=23)))

HyperChlPlotHigh <- PicoGAM_Plot_Prep %>%
  filter(
         Strain %in% GamStrain,
         Wavelength %in% PlotWL,
         O2_uM == "250"
  )
ContourLines <- ContourBreaks %>%
  filter(O2_uM == "250") 
ContourLines <- ContourLines$quantileValue

(ContourHighO2 <- ggplot(na.omit(HyperChlPlotHigh), aes(Photoperiod, Par_ue, z = fit, fill = fit)) + geom_tile() +
    scale_fill_gradientn(name = bquote("µ ("~ d^-1~")"), colours=c("white", ColourDF), limits  = range(minMu, maxMu), na.value = "white") +
    metR::geom_contour2(aes(label = stat(level)),
                        breaks = ContourLines,
                        skip = 0) +
    scale_x_continuous(limits = c(4, 16), breaks = c(4, 8, 12, 16)) +
    scale_y_continuous(limits = c(10, 200)) +
    xlab(NULL)+
      ylab(NULL)+ #https://stackoverflow.com/questions/15074127/use-expression-with-a-variable-r
    theme_classic() +
   theme(axis.text=element_text(size=15),
          axis.title.x = element_text(size=20),
           axis.title.y = element_text(size=35)))

g_legend<-function(a.gplot){
  tmp <- ggplot_gtable(ggplot_build(a.gplot))
  leg <- which(sapply(tmp$grobs, function(x) x$name) == "guide-box")
  legend <- tmp$grobs[[leg]]
  legend
}

(g_dummy1 <- ggplot(subset(HyperChlPlotLow, Color_name == w_length2rgb(WL_nm)), aes(Photoperiod, Par_ue, fill=fit)) + geom_tile()
+ scale_fill_gradientn(name = bquote("µ ("~ d^-1~")"), colours=c("white", w_length2rgb(WL_nm)),
  limits  = range(minMu, 1.0), na.value = "white") + theme_classic())
legend1 <- g_legend(g_dummy1)
(g_dummy2 <- ggplot(subset(HyperChlPlotInt, Color_name == w_length2rgb(WL_nm)), aes(Photoperiod, Par_ue, fill=fit)) + geom_tile()
+ scale_fill_gradientn(name = bquote("µ ("~ d^-1~")"), colours=c("white", w_length2rgb(WL_nm)),
  limits  = range(minMu, 1.0), na.value = "white") + theme_classic())
legend2 <- g_legend(g_dummy2)
(g_dummy3 <- ggplot(subset(HyperChlPlotHigh, Color_name == w_length2rgb(WL_nm)), aes(Photoperiod, Par_ue, fill=fit)) + geom_tile()
+ scale_fill_gradientn(name = bquote("µ ("~ d^-1~")"), colours=c("white", w_length2rgb(WL_nm)),
  limits  = range(minMu, 1.0), na.value = "white") + theme_classic())
legend3 <- g_legend(g_dummy3)

# comment rows that are not included in grid arrange 
LayoutNiche <- rbind(
                      # c(3, 3, 3, 3, 6), # For 2.5 uM O2     # remove comment if want 2.5 uM O2
                     # c(3, 3, 3, 3, 6), # For 2.5 uM O2    # remove comment if want 2.5 uM O2
                     c(2, 2, 2, 2, 5),   # For 25 uM O2
                     c(2, 2, 2, 2, 5),   # For 25 uM O2
                     c(1, 1, 1, 1, 4),   # For 250 uM O2
                     c(1, 1, 1, 1, 4)    # For 250 uM O2
                     )

#LayoutNiche <- rbind(c(1, 1, 1, 2, 2, 2, 3, 3, 3, 4))

p3_Red <- grid.arrange(
                       # ContourLowO2 +   labs(title = bquote("C.   " ~ .(GamStrain) ~ '  2.5 µM' ~ "[" ~ O[2]~ "]"), size = 25) + theme(legend.position = 'none'),   # remove comment for MIT9313
                       ContourIntO2 + labs(title = bquote("B.   " ~ .(GamStrain) ~ '  25 µM' ~ "[" ~ O[2]~ "]"), size = 25) +
                         theme(legend.position = 'none'), 
                       ContourHighO2 + labs(title = bquote("A.   " ~ .(GamStrain) ~ '  250 µM' ~ "[" ~ O[2]~ "]")) + 
                         theme(legend.position = 'none'),
                       # legend1,   # remove comment if want 2.5 uM O2
                        legend2, 
                        legend3, 
                       layout_matrix = LayoutNiche)

```

# Plot blue light SS120GAM
``` {r GAM Plot Blue WL}
GamStrain = "SS120"
GamColor = "Blue"
WL_nm = 450 
PlotWL <- "deltaOD"

maxMu <- PicoGAM_Pred %>%
  ungroup() %>%
  filter(Wavelength %in% PlotWL) %>%
  slice(which.max(fit)) #base::max does not work with package photobiology loaded simultaneously 
maxMu <- maxMu$fit

minMu <- PicoGAM_Pred %>%
  ungroup() %>%
  filter(Wavelength %in% PlotWL) %>%
  slice(which.min(fit)) #base::max does not work with package photobiology loaded simultaneously 
  #        fit == min(fit, na.rm = TRUE)) %>%
  # distinct(fit)
minMu <- minMu$fit

PicoGAM_Plot_Prep <- PicoGAM_Pred %>%
  filter(Wavelength %in% PlotWL,
         Strain %in% GamStrain,
         WL %in% WL_nm) %>%
  mutate(
         OxygenLabel = "Oxygen concentration (uM)",
         O2_level = case_when(O2_uM == 2.5 ~ "low",
                              O2_uM == 25 ~ "intermediate",
                              O2_uM == 250 ~ "high"),
         Color_name = case_when(WL == "450" ~ 'Blue',
                                WL == "530" ~ 'Green',
                                WL == "660" ~ 'Red',
                                WL == "WW" ~ 'White')) %>%
  unite(WL_O2, Color_name, O2_level, sep = "_", remove = FALSE)

WL_O2_Colors = c(Blue = w_length2rgb(450), green = w_length2rgb(530), red = w_length2rgb(660), white= "#FFCC00")

ColourDF <- data.frame(
  ColorName = c('Blue', 'Green', 'Red', 'White'),
  WL_O2_Colors = c( w_length2rgb(450), w_length2rgb(530), w_length2rgb(660), "#FFCC00")
) %>%
  filter(ColorName %in% GamColor) %>%
  dplyr::select(WL_O2_Colors)

ContourBreaks <- PicoGAM_Plot_Prep %>%
  filter(
         Strain %in% GamStrain,
         Wavelength %in% PlotWL
  ) %>%
  group_by(O2_uM) %>%
  mutate(upper10 = quantile(fit, 0.9, na.rm = TRUE),
         median50 = quantile(fit, 0.5, na.rm = TRUE),
         lower10 = quantile(fit, 0.1, na.rm = TRUE)) %>%
  ungroup() %>%
  dplyr::select(O2_uM, upper10, median50, lower10) %>%
  distinct() %>%
  pivot_longer(cols = c(upper10, median50, lower10), names_to = 'quantile', values_to = 'quantileValue') %>%
  mutate(quantileValue = round(quantileValue, digits = 2))

##############
###plot
#https://stackoverflow.com/questions/28775036/ggplot-line-graph-with-na-values
#https://stackoverflow.com/questions/6906661/ggplot2-make-missing-value-in-geom-tile-not-blank
#https://stackoverflow.com/questions/24822621/multiple-legends-in-a-plot-with-geom-tile
#https://stackoverflow.com/questions/43395219/plotting-multiple-layers-with-geom-raster-or-geom-tile-or-geom-rect
#https://stackoverflow.com/questions/59949635/ggplot-align-grid-lines-with-end-of-geom-tile
#https://stackoverflow.com/questions/22945651/remove-space-between-plotted-data-and-the-axes
#https://stackoverflow.com/questions/5293715/how-to-use-greek-symbols-in-ggplot2

HyperChlPlotLow <- PicoGAM_Plot_Prep %>%
  filter(
         Strain %in% GamStrain,
         Wavelength %in% PlotWL,
         O2_uM == "2.5"
  )
ContourLines <- ContourBreaks %>%
  filter(O2_uM == "2.5") 
ContourLines <- ContourLines$quantileValue

(ContourLowO2 <- ggplot(na.omit(HyperChlPlotLow), aes(Photoperiod, Par_ue, z = fit, fill = fit)) + geom_tile() +
    scale_fill_gradientn(name = bquote("µ ("~ d^-1~")"), colours=c(low = "white", intermediate = "steelblue1", high = "royalblue4"), limits  = range(minMu, maxMu), na.value = "white") +
    metR::geom_contour2(aes(label = stat(level)),
                        breaks = ContourLines,
                        skip = 0) +
    scale_x_continuous(limits = c(4, 16), breaks = c(4, 8, 12, 16)) +
    scale_y_continuous(limits = c(10, 200)) +
    xlab(NULL)+
      ylab(NULL)+
   theme_classic() +
   theme(axis.text=element_text(size=15),
          axis.title.x = element_text(size=20),
           axis.title.y = element_text(size=35)))

HyperChlPlotInt <- PicoGAM_Plot_Prep %>%
  filter(
         Strain %in% GamStrain,
         Wavelength %in% PlotWL,
         O2_uM == "25"
  )
ContourLines <- ContourBreaks %>%
  filter(O2_uM == "25") 
ContourLines <- ContourLines$quantileValue

(ContourIntO2 <- ggplot(na.omit(HyperChlPlotInt), aes(Photoperiod, Par_ue, z = fit, fill = fit)) + geom_tile() +
    scale_fill_gradientn(name = bquote("µ ("~ d^-1~")"), colours=c(low = "white", intermediate = "steelblue1", high = "royalblue4"), limits  = range(minMu, maxMu), na.value = "white") +
    metR::geom_contour2(aes(label = stat(level)),
                        breaks = ContourLines,
                        skip = 0) +
    scale_x_continuous(limits = c(4, 16), breaks = c(4, 8, 12, 16)) +
    scale_y_continuous(limits = c(10, 200)) +
    xlab(NULL)+
      ylab(NULL)+
  theme_classic() +
   theme(axis.text=element_text(size=15),
          axis.title.x = element_text(size=20),
           axis.title.y = element_text(size=35)))

HyperChlPlotHigh <- PicoGAM_Plot_Prep %>%
  filter(
         Strain %in% GamStrain,
         #DM == "Chl",
         Wavelength %in% PlotWL,
         O2_uM == "250"
  )
ContourLines <- ContourBreaks %>%
  filter(O2_uM == "250") 
ContourLines <- ContourLines$quantileValue

(ContourHighO2 <- ggplot(na.omit(HyperChlPlotHigh), aes(Photoperiod, Par_ue, z = fit, fill = fit)) + geom_tile() +
    scale_fill_gradientn(name = bquote("µ ("~ d^-1~")"), colours=c(low = "white", intermediate = "steelblue1", high = "royalblue4"), limits  = range(minMu, maxMu), na.value = "white") +
    metR::geom_contour2(aes(label = stat(level)),
                        breaks = ContourLines,
                        skip = 0) +
    scale_x_continuous(limits = c(4, 16), breaks = c(4, 8, 12, 16)) +
    scale_y_continuous(limits = c(10, 200)) +
    xlab(NULL)+
      ylab(NULL)+ #https://stackoverflow.com/questions/15074127/use-expression-with-a-variable-r
theme_classic() +
   theme(axis.text=element_text(size=15),
          axis.title.x = element_text(size=20),
           axis.title.y = element_text(size=35)))

g_legend<-function(a.gplot){
  tmp <- ggplot_gtable(ggplot_build(a.gplot))
  leg <- which(sapply(tmp$grobs, function(x) x$name) == "guide-box")
  legend <- tmp$grobs[[leg]]
  legend
}

(g_dummy1 <- ggplot(subset(HyperChlPlotLow, Color_name == w_length2rgb(WL_nm)), aes(Photoperiod, Par_ue, fill=fit)) + geom_tile()
 + scale_fill_gradientn(name = bquote("µ ("~ d^-1~")"), colours=c(low = "white", intermediate = "steelblue1", high = "royalblue4"),
 limits  = range(minMu, 1.0), na.value = "white") + theme_classic())
legend1 <- g_legend(g_dummy1)
(g_dummy2 <- ggplot(subset(HyperChlPlotInt, Color_name == w_length2rgb(WL_nm)), aes(Photoperiod, Par_ue, fill=fit)) + geom_tile()
 + scale_fill_gradientn(name = bquote("µ ("~ d^-1~")"), colours=c(low = "white", intermediate = "steelblue1", high = "royalblue4"),
  limits  = range(minMu, 1.0), na.value = "white") + theme_classic())
legend2 <- g_legend(g_dummy2)
(g_dummy3 <- ggplot(subset(HyperChlPlotHigh, Color_name == w_length2rgb(WL_nm)), aes(Photoperiod, Par_ue, fill=fit)) + geom_tile()
+ scale_fill_gradientn(name = bquote("µ ("~ d^-1~")"), colours=c(low = "white", intermediate = "steelblue1", high = "royalblue4"),
  limits  = range(minMu, 1.0), na.value = "white") + theme_classic())
legend3 <- g_legend(g_dummy3)

# comment rows that are not included in grid arrange 
LayoutNiche <- rbind(
                     # c(3, 3, 3, 3, 6), # For 2.5 uM O2   # remove comment if want 2.5 uM O2
                     # c(3, 3, 3, 3, 6), # For 2.5 uM O2   # remove comment if want 2.5 uM O2
                     c(2, 2, 2, 2, 5),   # For 25 uM O2
                     c(2, 2, 2, 2, 5),   # For 25 uM O2
                     c(1, 1, 1, 1, 4),   # For 250 uM O2
                     c(1, 1, 1, 1, 4)    # For 250 uM O2
                     )

#LayoutNiche <- rbind(c(1, 1, 1, 2, 2, 2, 3, 3, 3, 4))

p3_Blue <- grid.arrange(
                         # ContourLowO2 + labs(title = bquote("F.   " ~ .(GamStrain) ~ '  2.5 µM' ~ "[" ~ O[2]~ "]"), size = 25) + theme(legend.position = 'none', axis.title.y = element_blank(), axis.text.y = element_blank()),  # remove comment if want 2.5 uM O2
                        ContourIntO2 +  labs(title = bquote("D.   " ~ .(GamStrain) ~ '  25 µM' ~ "[" ~ O[2]~ "]"), size = 25) +
                          theme(legend.position = 'none', axis.title.y = element_blank(), axis.text.y = element_blank()), 
                        ContourHighO2 + labs(title = bquote("C.   " ~ .(GamStrain) ~ '  250 µM' ~ "[" ~ O[2]~ "]")) + 
                          theme(legend.position = 'none', axis.title.y = element_blank(), axis.text.y = element_blank()),
                        # legend1,    # remove comment if want 2.5 uM O2
                        legend2, 
                        legend3, 
                        layout_matrix = LayoutNiche)

```

```{r combine red and blue light SS120GAM plots}

#combine red and blue GAM plots
CombinedGAM <- plot_grid(p3_Red, p3_Blue, nrow = 1, align = "v")

# common x-axis title
x.grob <- textGrob("Photoperiod (h)", 
                   gp = gpar(fontsize = 20))
# common y-axis title
y.grob <- textGrob("PAR (µmol photons"~ m^-2~ ~s^-1~ ")", 
                   gp = gpar(fontface="bold", fontsize = 20), rot = 90)

# add titles to plot
LabelledCombinedGam <- grid.arrange(arrangeGrob(CombinedGAM, left = y.grob, bottom = x.grob))

```

```{r save SS120GAM plots}

# SS120GAM.png for manuscript
 ggsave(file = file.path(paste(PlotsOut, sep = "/"), paste(GamStrain, "GAM", ".png", sep = ""), fsep = .Platform$file.sep), plot = LabelledCombinedGam,  height = 10, width = 10)

# SS120GAM is Fig6.tiff for PLOSONE submission
 ggsave(file = file.path(PlotsOutPLOS, "Fig6.tiff", fsep = .Platform$file.sep), plot = LabelledCombinedGam, device = NULL, scale = 1,  height = 10, width = 10, compression = "lzw")

```

# Plot red light MIT9313 GAM
``` {r MIT9313GAM Plot Red WL}
GamStrain = "MIT9313"
GamColor = "Red" 
WL_nm = 660 
PlotWL <- "deltaOD"

maxMu <- PicoGAM_Pred %>%
  ungroup() %>%
  filter(Wavelength %in% PlotWL) %>%
  slice(which.max(fit)) #base::max does not work with package photobiology loaded simultaneously 
maxMu <- maxMu$fit

minMu <- PicoGAM_Pred %>%
  ungroup() %>%
  filter(Wavelength %in% PlotWL) %>%
  slice(which.min(fit)) #base::max does not work with package photobiology loaded simultaneously 
  #        fit == min(fit, na.rm = TRUE)) %>%
  # distinct(fit)
minMu <- minMu$fit

PicoGAM_Plot_Prep <- PicoGAM_Pred %>%
  filter(Wavelength %in% PlotWL,
         Strain %in% GamStrain,
         WL %in% WL_nm) %>%
  mutate(
         OxygenLabel = "Oxygen concentration (uM)",
         O2_level = case_when(O2_uM == 2.5 ~ "low",
                              O2_uM == 25 ~ "intermediate",
                              O2_uM == 250 ~ "high"),
         Color_name = case_when(WL == "450" ~ 'Blue',
                                WL == "530" ~ 'Green',
                                WL == "660" ~ 'Red',
                                WL == "WW" ~ 'White')) %>%
  unite(WL_O2, Color_name, O2_level, sep = "_", remove = FALSE)

WL_O2_Colors = c(Blue = w_length2rgb(450), green = w_length2rgb(530), red = w_length2rgb(660), white= "#FFCC00")

ColourDF <- data.frame(
  ColorName = c('Blue', 'Green', 'Red', 'White'),
  WL_O2_Colors = c( w_length2rgb(450), w_length2rgb(530), w_length2rgb(660), "#FFCC00")
) %>%
  filter(ColorName %in% GamColor) %>%
  dplyr::select(WL_O2_Colors)

ContourBreaks <- PicoGAM_Plot_Prep %>%
  filter(
         Strain %in% GamStrain,
         Wavelength %in% PlotWL
  ) %>%
  group_by(O2_uM) %>%
  mutate(upper10 = quantile(fit, 0.9, na.rm = TRUE),
         median50 = quantile(fit, 0.5, na.rm = TRUE),
         lower10 = quantile(fit, 0.1, na.rm = TRUE)) %>%
  ungroup() %>%
  dplyr::select(O2_uM, upper10, median50, lower10) %>%
  distinct() %>%
  pivot_longer(cols = c(upper10, median50, lower10), names_to = 'quantile', values_to = 'quantileValue') %>%
  mutate(quantileValue = round(quantileValue, digits = 2))

##############
###plot
#https://stackoverflow.com/questions/28775036/ggplot-line-graph-with-na-values
#https://stackoverflow.com/questions/6906661/ggplot2-make-missing-value-in-geom-tile-not-blank
#https://stackoverflow.com/questions/24822621/multiple-legends-in-a-plot-with-geom-tile
#https://stackoverflow.com/questions/43395219/plotting-multiple-layers-with-geom-raster-or-geom-tile-or-geom-rect
#https://stackoverflow.com/questions/59949635/ggplot-align-grid-lines-with-end-of-geom-tile
#https://stackoverflow.com/questions/22945651/remove-space-between-plotted-data-and-the-axes
#https://stackoverflow.com/questions/5293715/how-to-use-greek-symbols-in-ggplot2

HyperChlPlotLow <- PicoGAM_Plot_Prep %>%
  filter(
         Strain %in% GamStrain,
         Wavelength %in% PlotWL,
         O2_uM == "2.5"
  )
ContourLines <- ContourBreaks %>%
  filter(O2_uM == "2.5") 
ContourLines <- ContourLines$quantileValue

(ContourLowO2 <- ggplot(na.omit(HyperChlPlotLow), aes(Photoperiod, Par_ue, z = fit, fill = fit)) + geom_tile() +
  scale_fill_gradientn(name = bquote("µ ("~ d^-1~")"), colours=c("white", ColourDF), limits  = range(minMu, maxMu), na.value = "white") +
    metR::geom_contour2(aes(label = stat(level)),
                        breaks = ContourLines,
                        skip = 0) +
    scale_x_continuous(limits = c(4, 16), breaks = c(4, 8, 12, 16)) +
    scale_y_continuous(limits = c(10, 200)) +
     xlab(NULL)+
      ylab(NULL)+
    theme_classic() +
    theme(axis.text=element_text(size=15),
          axis.title.x = element_text(size=20),
           axis.title.y = element_text(size=35)))

HyperChlPlotInt <- PicoGAM_Plot_Prep %>%
  filter(
         Strain %in% GamStrain,
         Wavelength %in% PlotWL,
         O2_uM == "25"
  )
ContourLines <- ContourBreaks %>%
  filter(O2_uM == "25") 
ContourLines <- ContourLines$quantileValue

(ContourIntO2 <- ggplot(na.omit(HyperChlPlotInt), aes(Photoperiod, Par_ue, z = fit, fill = fit)) + geom_tile() +
  scale_fill_gradientn(name = bquote("µ ("~ d^-1~")"), colours=c("white", ColourDF), limits  = range(minMu, maxMu), na.value = "white") +
    metR::geom_contour2(aes(label = stat(level)),
                        breaks = ContourLines,
                        skip = 0) +
    scale_x_continuous(limits = c(4, 16), breaks = c(4, 8, 12, 16)) +
    scale_y_continuous(limits = c(10, 200)) +
    xlab(NULL)+
      ylab(NULL)+
   theme_classic() +
  theme(axis.text=element_text(size=15),
          axis.title.x = element_text(size=20),
           axis.title.y = element_text(size=23)))

HyperChlPlotHigh <- PicoGAM_Plot_Prep %>%
  filter(
         Strain %in% GamStrain,
         Wavelength %in% PlotWL,
         O2_uM == "250"
  )
ContourLines <- ContourBreaks %>%
  filter(O2_uM == "250") 
ContourLines <- ContourLines$quantileValue

(ContourHighO2 <- ggplot(na.omit(HyperChlPlotHigh), aes(Photoperiod, Par_ue, z = fit, fill = fit)) + geom_tile() +
    scale_fill_gradientn(name = bquote("µ ("~ d^-1~")"), colours=c("white", ColourDF), limits  = range(minMu, maxMu), na.value = "white") +
    metR::geom_contour2(aes(label = stat(level)),
                        breaks = ContourLines,
                        skip = 0) +
    scale_x_continuous(limits = c(4, 16), breaks = c(4, 8, 12, 16)) +
    scale_y_continuous(limits = c(10, 200)) +
    xlab(NULL)+
      ylab(NULL)+
    #https://stackoverflow.com/questions/15074127/use-expression-with-a-variable-r
    theme_classic() +
   theme(axis.text=element_text(size=15),
          axis.title.x = element_text(size=20),
           axis.title.y = element_text(size=35)))

g_legend<-function(a.gplot){
  tmp <- ggplot_gtable(ggplot_build(a.gplot))
  leg <- which(sapply(tmp$grobs, function(x) x$name) == "guide-box")
  legend <- tmp$grobs[[leg]]
  legend
}

(g_dummy1 <- ggplot(subset(HyperChlPlotLow, Color_name == w_length2rgb(WL_nm)), aes(Photoperiod, Par_ue, fill=fit)) + geom_tile()
+ scale_fill_gradientn(name = bquote("µ ("~ d^-1~")"), colours=c("white", w_length2rgb(WL_nm)),
  limits  = range(minMu, 1.0), na.value = "white") + theme_classic())
legend1 <- g_legend(g_dummy1)
(g_dummy2 <- ggplot(subset(HyperChlPlotInt, Color_name == w_length2rgb(WL_nm)), aes(Photoperiod, Par_ue, fill=fit)) + geom_tile()
+ scale_fill_gradientn(name = bquote("µ ("~ d^-1~")"), colours=c("white", w_length2rgb(WL_nm)),
  limits  = range(minMu, 1.0), na.value = "white") + theme_classic())
legend2 <- g_legend(g_dummy2)
(g_dummy3 <- ggplot(subset(HyperChlPlotHigh, Color_name == w_length2rgb(WL_nm)), aes(Photoperiod, Par_ue, fill=fit)) + geom_tile()
+ scale_fill_gradientn(name = bquote("µ ("~ d^-1~")"), colours=c("white", w_length2rgb(WL_nm)),
  limits  = range(minMu, 1.0), na.value = "white") + theme_classic())
legend3 <- g_legend(g_dummy3)

# comment rows that are not included in grid arrange 
LayoutNiche <- rbind(
                     c(3, 3, 3, 3, 6), # For 2.5 uM O2    
                     c(3, 3, 3, 3, 6), # For 2.5 uM O2  
                     c(2, 2, 2, 2, 5),   # For 25 uM O2
                     c(2, 2, 2, 2, 5),   # For 25 uM O2
                     c(1, 1, 1, 1, 4),   # For 250 uM O2
                     c(1, 1, 1, 1, 4)    # For 250 uM O2
                     )

#LayoutNiche <- rbind(c(1, 1, 1, 2, 2, 2, 3, 3, 3, 4))

p3_Red <- grid.arrange(
                       ContourLowO2 + labs(title = bquote("C.   " ~ .(GamStrain) ~ '  2.5 µM' ~ "[" ~ O[2]~ "]"), size = 25) +
                         theme(legend.position = 'none'),  
                       ContourIntO2 + labs(title = bquote("B.   " ~ .(GamStrain) ~ '  25 µM' ~ "[" ~ O[2]~ "]"), size = 25) +
                         theme(legend.position = 'none'), 
                       ContourHighO2 + labs(title = bquote("A.   " ~ .(GamStrain) ~ '  250 µM' ~ "[" ~ O[2]~ "]")) + 
                         theme(legend.position = 'none'), 
                        legend1, 
                        legend2, 
                        legend3, 
                       layout_matrix = LayoutNiche)

```

# Plot blue light MIT9313 GAM
``` {r MIT9313GAM Plot Blue WL}
GamStrain = "MIT9313"
GamColor = "Blue"
WL_nm = 450 
PlotWL <- "deltaOD"

maxMu <- PicoGAM_Pred %>%
  ungroup() %>%
  filter(Wavelength %in% PlotWL) %>%
  slice(which.max(fit)) #base::max does not work with package photobiology loaded simultaneously 
maxMu <- maxMu$fit

minMu <- PicoGAM_Pred %>%
  ungroup() %>%
  filter(Wavelength %in% PlotWL) %>%
  slice(which.min(fit)) #base::max does not work with package photobiology loaded simultaneously 
  #        fit == min(fit, na.rm = TRUE)) %>%
  # distinct(fit)
minMu <- minMu$fit

PicoGAM_Plot_Prep <- PicoGAM_Pred %>%
  filter(Wavelength %in% PlotWL,
         Strain %in% GamStrain,
         WL %in% WL_nm) %>%
  mutate(
         OxygenLabel = "Oxygen concentration (uM)",
         O2_level = case_when(O2_uM == 2.5 ~ "low",
                              O2_uM == 25 ~ "intermediate",
                              O2_uM == 250 ~ "high"),
         Color_name = case_when(WL == "450" ~ 'Blue',
                                WL == "530" ~ 'Green',
                                WL == "660" ~ 'Red',
                                WL == "WW" ~ 'White')) %>%
  unite(WL_O2, Color_name, O2_level, sep = "_", remove = FALSE)

WL_O2_Colors = c(Blue = w_length2rgb(450), green = w_length2rgb(530), red = w_length2rgb(660), white= "#FFCC00")

ColourDF <- data.frame(
  ColorName = c('Blue', 'Green', 'Red', 'White'),
  WL_O2_Colors = c( w_length2rgb(450), w_length2rgb(530), w_length2rgb(660), "#FFCC00")
) %>%
  filter(ColorName %in% GamColor) %>%
  dplyr::select(WL_O2_Colors)

ContourBreaks <- PicoGAM_Plot_Prep %>%
  filter(
         Strain %in% GamStrain,
         Wavelength %in% PlotWL
  ) %>%
  group_by(O2_uM) %>%
  mutate(upper10 = quantile(fit, 0.9, na.rm = TRUE),
         median50 = quantile(fit, 0.5, na.rm = TRUE),
         lower10 = quantile(fit, 0.1, na.rm = TRUE)) %>%
  ungroup() %>%
  dplyr::select(O2_uM, upper10, median50, lower10) %>%
  distinct() %>%
  pivot_longer(cols = c(upper10, median50, lower10), names_to = 'quantile', values_to = 'quantileValue') %>%
  mutate(quantileValue = round(quantileValue, digits = 2))

##############
###plot
#https://stackoverflow.com/questions/28775036/ggplot-line-graph-with-na-values
#https://stackoverflow.com/questions/6906661/ggplot2-make-missing-value-in-geom-tile-not-blank
#https://stackoverflow.com/questions/24822621/multiple-legends-in-a-plot-with-geom-tile
#https://stackoverflow.com/questions/43395219/plotting-multiple-layers-with-geom-raster-or-geom-tile-or-geom-rect
#https://stackoverflow.com/questions/59949635/ggplot-align-grid-lines-with-end-of-geom-tile
#https://stackoverflow.com/questions/22945651/remove-space-between-plotted-data-and-the-axes
#https://stackoverflow.com/questions/5293715/how-to-use-greek-symbols-in-ggplot2

HyperChlPlotLow <- PicoGAM_Plot_Prep %>%
  filter(
         Strain %in% GamStrain,
         Wavelength %in% PlotWL,
         O2_uM == "2.5"
  )
ContourLines <- ContourBreaks %>%
  filter(O2_uM == "2.5") 
ContourLines <- ContourLines$quantileValue

(ContourLowO2 <- ggplot(na.omit(HyperChlPlotLow), aes(Photoperiod, Par_ue, z = fit, fill = fit)) + geom_tile() +
    scale_fill_gradientn(name = bquote("µ ("~ d^-1~")"), colours=c(low = "white", intermediate = "steelblue1", high = "royalblue4"), limits  = range(minMu, maxMu), na.value = "white") +
    metR::geom_contour2(aes(label = stat(level)),
                        breaks = ContourLines,
                        skip = 0) +
    scale_x_continuous(limits = c(4, 16), breaks = c(4, 8, 12, 16)) +
    scale_y_continuous(limits = c(10, 200)) +
    xlab(NULL)+
      ylab(NULL)+
   theme_classic() +
   theme(axis.text=element_text(size=15),
          axis.title.x = element_text(size=20),
           axis.title.y = element_text(size=35)))

HyperChlPlotInt <- PicoGAM_Plot_Prep %>%
  filter(
         Strain %in% GamStrain,
         Wavelength %in% PlotWL,
         O2_uM == "25"
  )
ContourLines <- ContourBreaks %>%
  filter(O2_uM == "25") 
ContourLines <- ContourLines$quantileValue

(ContourIntO2 <- ggplot(na.omit(HyperChlPlotInt), aes(Photoperiod, Par_ue, z = fit, fill = fit)) + geom_tile() +
    scale_fill_gradientn(name = bquote("µ ("~ d^-1~")"), colours=c(low = "white", intermediate = "steelblue1", high = "royalblue4"), limits  = range(minMu, maxMu), na.value = "white") +
    metR::geom_contour2(aes(label = stat(level)),
                        breaks = ContourLines,
                        skip = 0) +
    scale_x_continuous(limits = c(4, 16), breaks = c(4, 8, 12, 16)) +
    scale_y_continuous(limits = c(10, 200)) +
    xlab(NULL)+
      ylab(NULL)+
  theme_classic() +
   theme(axis.text=element_text(size=15),
          axis.title.x = element_text(size=20),
           axis.title.y = element_text(size=35)))

HyperChlPlotHigh <- PicoGAM_Plot_Prep %>%
  filter(
         Strain %in% GamStrain,
         Wavelength %in% PlotWL,
         O2_uM == "250"
  )
ContourLines <- ContourBreaks %>%
  filter(O2_uM == "250") 
ContourLines <- ContourLines$quantileValue

(ContourHighO2 <- ggplot(na.omit(HyperChlPlotHigh), aes(Photoperiod, Par_ue, z = fit, fill = fit)) + geom_tile() +
    scale_fill_gradientn(name = bquote("µ ("~ d^-1~")"), colours=c(low = "white", intermediate = "steelblue1", high = "royalblue4"), limits  = range(minMu, maxMu), na.value = "white") +
    metR::geom_contour2(aes(label = stat(level)),
                        breaks = ContourLines,
                        skip = 0) +
    scale_x_continuous(limits = c(4, 16), breaks = c(4, 8, 12, 16)) +
    scale_y_continuous(limits = c(10, 200)) +
    xlab(NULL)+
      ylab(NULL)+
    #https://stackoverflow.com/questions/15074127/use-expression-with-a-variable-r
theme_classic() +
   theme(axis.text=element_text(size=15),
          axis.title.x = element_text(size=20),
           axis.title.y = element_text(size=35)))

g_legend<-function(a.gplot){
  tmp <- ggplot_gtable(ggplot_build(a.gplot))
  leg <- which(sapply(tmp$grobs, function(x) x$name) == "guide-box")
  legend <- tmp$grobs[[leg]]
  legend
}

(g_dummy1 <- ggplot(subset(HyperChlPlotLow, Color_name == w_length2rgb(WL_nm)), aes(Photoperiod, Par_ue, fill=fit)) + geom_tile()
 + scale_fill_gradientn(name = bquote("µ ("~ d^-1~")"), colours=c(low = "white", intermediate = "steelblue1", high = "royalblue4"),
 limits  = range(minMu, 1.0), na.value = "white") + theme_classic())
legend1 <- g_legend(g_dummy1)
(g_dummy2 <- ggplot(subset(HyperChlPlotInt, Color_name == w_length2rgb(WL_nm)), aes(Photoperiod, Par_ue, fill=fit)) + geom_tile()
 + scale_fill_gradientn(name = bquote("µ ("~ d^-1~")"), colours=c(low = "white", intermediate = "steelblue1", high = "royalblue4"),
  limits  = range(minMu, 1.0), na.value = "white") + theme_classic())
legend2 <- g_legend(g_dummy2)
(g_dummy3 <- ggplot(subset(HyperChlPlotHigh, Color_name == w_length2rgb(WL_nm)), aes(Photoperiod, Par_ue, fill=fit)) + geom_tile()
+ scale_fill_gradientn(name = bquote("µ ("~ d^-1~")"), colours=c(low = "white", intermediate = "steelblue1", high = "royalblue4"),
  limits  = range(minMu, 1.0), na.value = "white") + theme_classic())
legend3 <- g_legend(g_dummy3)

# comment rows that are not included in grid arrange 
LayoutNiche <- rbind(
                     c(3, 3, 3, 3, 6), # For 2.5 uM O2   
                     c(3, 3, 3, 3, 6), # For 2.5 uM O2  
                     c(2, 2, 2, 2, 5),   # For 25 uM O2
                     c(2, 2, 2, 2, 5),   # For 25 uM O2
                     c(1, 1, 1, 1, 4),   # For 250 uM O2
                     c(1, 1, 1, 1, 4)    # For 250 uM O2
                     )

#LayoutNiche <- rbind(c(1, 1, 1, 2, 2, 2, 3, 3, 3, 4))

p3_Blue <- grid.arrange(
                         ContourLowO2 + labs(title = bquote("F.   " ~ .(GamStrain) ~ '  2.5 µM' ~ "[" ~ O[2]~ "]"), size = 25) +
                           theme(legend.position = 'none', axis.title.y = element_blank(), axis.text.y = element_blank()),
                        ContourIntO2 +  labs(title = bquote("E.   " ~ .(GamStrain) ~ '  25 µM' ~ "[" ~ O[2]~ "]"), size = 25) + 
                          theme(legend.position = 'none', axis.title.y = element_blank(), axis.text.y = element_blank()), 
                        ContourHighO2 + labs(title = bquote("D.   " ~ .(GamStrain) ~ '  250 µM' ~ "[" ~ O[2]~ "]")) + 
                          theme(legend.position = 'none', axis.title.y = element_blank(), axis.text.y = element_blank()),                          
                        legend1,    
                        legend2, 
                        legend3, 
                        layout_matrix = LayoutNiche)  
```

```{r combine red and blue light MIT9313GAM plots}

#combine red and blue GAM plots
CombinedGAM <- plot_grid(p3_Red, p3_Blue, nrow = 1, align = "v")

# common x-axis title
x.grob <- textGrob("Photoperiod (h)", 
                   gp = gpar(fontsize = 20))

# common y-axis title
y.grob <- textGrob("PAR (µmol photons"~ m^-2~ ~s^-1~ ")", 
                   gp = gpar(fontface="bold", fontsize = 20), rot = 90)

# add titles to plot
LabelledCombinedGam <- grid.arrange(arrangeGrob(CombinedGAM, left = y.grob, bottom = x.grob))

```

```{r save MIT9313GAM plots}

# MIT9313GAM.png for manuscript
 ggsave(file = file.path(paste(PlotsOut, sep = "/"), paste(GamStrain, "GAM", ".png", sep = ""), fsep = .Platform$file.sep), plot = LabelledCombinedGam,  height = 10, width = 10)

# MIT9313 GAM is Fig8.tiff for PLOSONE submission
 ggsave(file = file.path(PlotsOutPLOS, "Fig8.tiff", fsep = .Platform$file.sep), plot = LabelledCombinedGam, device = NULL, scale = 1,  height = 10, width = 10,  compression = "lzw")

```