### 2024-05-27 Output/PLOSRound2/ResponsetoReviewers.txt
Doug edits done

PONE-D-24-11318
Prochlorococcus marinus responses to light and oxygen
PLOS ONE

Jos? A. Fern?ndez Robledo, Ph.D. 
Academic Editor 
PLOS ONE

Dear Dr. Fern?ndez Robledo,

We thank you and the reviewers for constructive feedback, which we have addressed through revisions, as outlined below.

?	separate file labeled 'Revised Manuscript with Track Changes'.
?	An unmarked version of your revised paper without tracked changes. You should upload this as a separate file labeled 'Manuscript'.

1. Please ensure that your manuscript meets PLOS ONE's style requirements, including those for file 
naming. The PLOS ONE style templates can be found at  
https://journals.plos.org/plosone/s/file?id=wjVg/PLOSOne_formatting_sample_main_body.pdf and  
https://journals.plos.org/plosone/s/file?id=ba62/PLOSOne_formatting_sample_title_authors_affiliation
s.pdf

2. Please note that PLOS ONE has specific guidelines on code sharing for submissions in which author-
generated code underpins the findings in the manuscript. In these cases, all author-generated code 
must be made available without restrictions upon publication of the work. Please review our guidelines 
at https://journals.plos.org/plosone/s/materials-and-software-sharing#loc-sharing-code and ensure that 
your code is shared in a way that follows best practice and facilitates reproducibility and reuse.

3. Thank you for uploading your study's underlying data set. Unfortunately, the repository you have 
noted in your Data Availability statement does not qualify as an acceptable data repository according to 
PLOS's standards.

At this time, please upload the minimal data set necessary to replicate your study's findings to a stable, 
public repository (such as figshare or Dryad) and provide us with the relevant URLs, DOIs, or accession 
numbers that may be used to access these data. For a list of recommended repositories and additional 
information on PLOS standards for data deposition, please see 
https://journals.plos.org/plosone/s/recommended-repositories.

4. We note that you have referenced (Campbell, unpub) on page 17, which has currently not yet been 
accepted for publication. Please remove this from your References and amend this to state in the body 
of your manuscript: (ie ?Bewick et al. [Unpublished]?) as detailed online in our guide for authors 
http://journals.plos.org/plosone/s/submission-guidelines#loc-reference-style

We removed this citation.

5. We notice that your supplementary [Supplemental Figures 1-5 and S1 Table] are included in the 
manuscript file. Please remove them and upload them with the file type 'Supporting Information'. Please 
ensure that each Supporting Information file has a legend listed in the manuscript after the references 
list.

6. Please review your reference list to ensure that it is complete and correct. If you have cited papers 
that have been retracted, please include the rationale for doing so in the manuscript text, or remove 
these references and replace them with relevant current references. Any changes to the reference list 
should be mentioned in the rebuttal letter that accompanies your revised manuscript. If you need to cite 
a retracted article, indicate the article?s retracted status in the References list and also include a citation 
and full reference for the retraction notice.

XXX Check XXX

Reviewers' comments:

Reviewer's Responses to Questions
Comments to the Author

Reviewer #1: This manuscript presents a helpful expansion of our knowledge of the ecological tolerance 
ranges of several Prochlorococcus ecotypes. It is especially interesting in its implications for range 
expansion of this genus under potential future conditions, although this angle could be expanded on in 
my opinion, as I explain below.

Specific Comments:

The abstract seems to omit any discussion of the conclusions/bigger picture of the study.

Line 55: LLII/III and LLIV are not the ?oldest? clades, but the most basal or early branching.

We have edited the text in the revised manuscript: "Clades LLII and LLIII, including cultured strain SS120, are grouped together as early branching phylogenetic lineage diversifying in the P. marinus radiation, with a preference for low light."
Line: XXX

Line 369: The authors invoke ?limitations in the annotation process? to explain the lack of evidence for 
MED4 ATP synthase. What does this mean? The MED4 genome surely contains ATP synthase genes.

We agree that the MED4 genome includes genes encoding ATP Synthase, but in the Ocean Protein Portal database no ATP Synthase peptides are annotated as coming from MED4. This likely reflects an annotation assignment issue because the ATP Synthase proteins are highly conserved, so assignments to strains or clades of Prochlorococcus may be ambiguous.
We have edited the text in the revised manuscript: "Though present in the genome of MED4, the absence of proteins annotated as ATP Synthase complex for Clade HLI, compared to annotated detections of ATP Synthase across the other three clades suggests limitations in the OPP annotation process for highly conserved protein sequences."
Line: XXX

Paragraph starting L363: Why is there no discussion of the Rubisco data? Or especially the absence 
thereof at low O2 for ecotypes HL1 and LL1? Since Rubisco is one of the most abundant proteins in a 
photosynthesizing cell, does this indicate that these ecotypes are present but not actively fixing carbon 
under these conditions? Also, is it worth noting that LLIV is the only variety to be detected at all in the 
deepest low O2 site -- which would appear to be consistent with subsequent findings in this study?

 XXX Doug address RUBISCO XXX

We revised the text from lines 668 to XXX to incorporate these helpful observations.

Figures 3 through 8: These figures are obviously the crux of the results in this paper, but as they are 
currently presented it is relatively difficult to think clearly about them. I feel like they could be combined 
into 2 or 3 full page figures which might make it easier to make the relevant comparisons; either by 
combining the two types of figure (growth rates and model fits) for each ecotype, or else all the growth 
rate curves together in one figure and all the model plots together in another.

Thank you for this suggestion. We have combined the growth rate figure with the GAM model for each strain to faciliate comparisons.

I'm also not sure I understand why the red light data is relevant -- perhaps by moving it to supplemental, it would be easier to compile these data into compound figures?

We briefly motivated the inclusion of the red light growth trials, as mechanistically informative, at line 706:
" As mentioned, growth trials under red light, although not representative of *Prochlorococcus* niches, are mechanistically informative [@murphyPhotoinactivationPhotosystemII2017] regarding photoinactivation of PSII."

Through the Discussion, notably under the discussion of the GAM models and responses to growth rate to increasing PUR, we return to the responses to red light to show that some 'high light' limitations on growth are interpretable as photoinhibition of PSII, which is more severe under blue, compared to red, light, notably at lines 1023-1032 (xXXX).

I also wonder if the data from the model fits could be used to predict which of the three ecotypes would 
prevail (have the highest growth rate) across the parameter space depicted, resulting in another figure 
showing something like "zones of exclusion" for each ecotype, which would be interesting in terms of 
predictive power. One could easily imagine applying those "zones of exclusion" to a map of the actual 
ocean at various depths, or even comparing current with projected future conditions of spreading 
OMZs. This is merely a suggestion -- I don't expect the authors to necessarily do this, but I think if they 
chose to it would make the manuscript of broader interest.

Thank you for the comment.
We added a table YY of potential strain growth in temperate seasonal and depth niches, in future temperate warmed oceans.
The GAM model panels in figures XYZ show 'zones of exclusion' from light level, photoperiods, and oxygen regimes, as unshaded white areas where no growth is predicted.
We indeed contemplating attempting global predicted distribution maps, but the interacting complexities of season, depth, changing extents of OMZ and extrapolation from culture growth, render it beyond the scope of this study.

XXX

L482, 536, 548: some experiments were not performed ?due to time constraints?. I'm not sure I 
understand this, although I am interpreting it along the lines of "the person doing the experiments 
graduated"? In any case, it seems like these particularly treatments were unlikely to grow, so it's not 
important that they were omitted -- but maybe there is a better way to phrase this that doesn?t sound 
so fishy?

Thank you for this suggestion. 
We edited the text:
Line: XXX

Figure S4 ? It is very hard to see the fit lines other than the pooled fit. Consider using different colors 
and solid lines since this is online.

We edited the figure to clarify the fit lines:
Line: XXX

Table 1: The trees described in the legend should be included in the supplemental material.

Thank you for bringing this to our attention. We clarified the wording on line XXXX of the revised manuscripted. "Protein homologies for FtsH isoforms between *Prochlorococcus* and *Synechocystis* were extracted from @bonisteelStrainSpecificDifferences2018."

Fig 11: Despite the fact that it is explained in the legend, the figure should probably indicate that O2 
values are log Km, or else put actual values on the log-scale Y axis. Also, the axis should indicate the 
units for Km. Finally, why is the red band so narrow compared to the green and blue?

We apologize for the omissions. We have revised the axis labels to make the log scaling more noticeable. The narrowness of the red bar (spanning 230 to 280 ?M) results from the log scaling of the y axis.

L757: The authors should specify that this is a DNA repair ligase, not the DNA ligase that seals Okazaki 
fragments during DNA replication (assuming this is the case).

Thank you for your comment, the revised text now reads "DNA repair ligase".  
Line: XXX

L 758: Photolyase should not have been relevant in the lab experiments because there is no UV light 
emitted (AFAIK) by the bulbs used in the Multicultivator. Also, it is unlikely that the absence of these two 
genes is sufficient to explain why these strains can?t grow at the ocean?s surface.

Thank you. The revised manuscript now reads: "The absence of genes encoding deoxyribodipyrimidine photolyase and DNA ligase (ATP) in P. marinus MIT9313 and P. marinus SS120 may, in part, explain why these two strains cannot tolerate growth under full [O2] and high light, found at the ocean surface."
Line: XXX
XXXXX Mireille to address the wording pertaining to photolyase. We did not imply that photolyase was responsible for the results we found in lab but I will confirm this. XXXXXXXXXX

Fig 12: Phosphoglycolate phosphatase is a major enzyme in the photorespiration pathway ? what does 
it have to do with DNA repair? There is evidence that 2-phosphoglycolate causes oxidative stress, but I 
don?t think this enzyme is properly a DNA repair enzyme. 
XXX Naaman to removeXXXXX

Summary and Conclusion: I feel like the authors really need to expand their discussion of how these 
findings apply to the future ocean. Where are OMZs likely to expand? Are there specific examples of 
regions where LLIV might expand to the surface? What would it mean for LLIV to become competitive 
against HLI and HLII? I think this is likely to be similar to scenarios where Synechococcus may 
outcompete Prochlorococcus, since LLIV is quite similar to Syn in terms of size and genome repertoire. 
I'm not sure where exactly such a discussion should end up, but I think it is important that the authors 
try to "seal the deal" a bit better in this concluding section and relate the findings back to the issues laid 
out in the introduction.

We added the following to the Summary and Conclusion section:
"In warming oceans *P. marinus* clades will differentially expand into new regions depending upon their different, interacting capacities to tolerate and exploit photoperiods, light levels, and [O~2~]. Competition among clades will be driven not simply by light levels. Clade HLI (including MED4) is excluded from short photoperiod regimes, typical of temperate winter, or light attenuated depths. In contrast, clade LLIV (including MIT9313) might actually exploit higher light niches under OMZ conditions, where low O~2~ relieves the stresses of oxidation stress and PSII photoinhibition."
Line:XXX

Reviewer #2: Savoie et al describe a beautiful and extensive set of detailed Prochlorococcus growth rate 
measurements for three strains, representing distinct ecotypes, under varied conditions of light and 
oxygen concentration. The work adds a lot of interesting and useful information concerning the impact 
of varied photoperiods, wavelengths/PUR, and oxygen concentrations to the literature. These data are 
also connected to hypotheses about the potential role of DNA repair and PSII maintenance to explain 
the results. The methods are appropriately detailed, and the experiments appear to have been done 
with great care. The data are available in a usable form.

My primary concern with the manuscript comes from the way the results are framed throughout. The 
authors describe results as informing about the growth of Prochlorococcus across a matrix of different 
conditions. However, my interpretation of the methods is that the authors took from a maintenance 
culture kept under standard growth conditions and then immediately transferred them into the 
Multicultivator to look at responses. This is of course completely valid, but extensive experience by 
many in the field (e.g. described in Moore et al 2007 Limn. Oceanog. Methods and elsewhere) is that 
these cells typically require multiple generations to acclimate to new growth conditions. I read these 
data as more reflecting the immediate response of Prochlorococcus to environmental changes. It is 
quite conceivable that some of these strains may exhibit a different growth rate than what you report if 
allowed to gradually acclimate to their new condition rather than directly move them. I would suggest 
that the framing and interpretation of the paper should reflect this. I also have no idea whether this 
might have influenced the surprising observations of MIT9313 growth at relatively high PAR. Global 
change scenarios would presumably give cells many generations to physiologically acclimate to 
differences (as well as to evolutionarily adapt).

Thank you for highlighting this issue. 
We added the following in the materials and methods section of the revised manuscript:
"Each tube containing 70 mL of Pro99 media
was inoculated around mid day of the 12 h maintenance photoperiod with 10 mL of growing maintenance preculture, to reach a starting OD680 of approximately 0.020. The tubes containing the cultures were then placed in the Multicultivator water bath set at 22?C, sparged with the experimental [O~2~], and kept at low light until late afternoon. Cultures were then in the dark for 12 to 16 hr until the photoregime of a sinusoidal photoperiod commenced the following morning, reaching peak PAR at noon each day. Cultures thus took approximately 24 h to move gradually from maintenance photoregime to the peak PAR of the experimental photoregime.  Cultures were grown for  xx to yy days, until they reached stationary phase at OD680 of approximately 0.4 to 0.8 after approximately 5 generations of growth."
Line: XXX

We added the following in the results and discussion section of the revised manuscript:
"We implemented measures to minimize shock to cultures from exposure to experimental growth conditions by inoculating them the day before the experiment began, employing a sinusoidal photoperiod with a gradual increase in PAR exposure, and extracting exponential growth rates from logistic curves fit over approximately 5 generations of growth, to accommodate multiple generations to acclimate to the imposed growth conditions (Moore et al. 2007)."
Line: XXX

This reviewer found the manuscript to be quite difficult to read as I kept wanting to understand how 
different strains were responding to similar types of variables, which made me have to constantly flip 
through and reread parts of the manuscript to find the information. Establishing a clearer overall 
summary narrative early on, as opposed to presenting a laundry list of data, might help. For instance, 
more clearly motivating the questions driving the analyses in each section would help; many paragraphs 
simply start with ?Figure X shows??. But, the organization of the paper is of course up to the authors.

XXXX In response to reviewer 1 and reviewer 2 we have combined the growth rate figure with the GAM model for each strain to faciliate comparisons. We have edited the abstract, results and discussion to better convey the motivations and findings of the study. In particular lines  XXX to YYY and XXXXX

Minor points:

General note: the manuscript text at some points refers broadly to ?growth?, and other places is more 
specific about drawing conclusions about ?growth rate?. I would suggest being clearer in some way that 
you are focused on growth rate and not growth yield.

We edited the text to use 'growth rate' throughout.

40: Despite _its_ small cell size (?)

We have edited the text:
Line: XXX

56: ?Affinity? for low light? Consider rephrasing

We have edited the text in the revised manuscript: " Clades LLII and LLIII, including cultured strain SS120, are
grouped together as early branching phylogenetic lineage in
the *P. marinus* radiation, with a preference for low light."
Line: XXX

84: Follett et al 2022 PNAS showed that temp is not the only factor influencing poleward issues - 
ecological interactions may further impact poleward shift during warming

Thank you for  bringing this to our attention. We have added the following text to the revised manuscript: "Follett et al 2022, however, model interactions of heterotrophic bacteria  that may influence latitudinal expansions of *P. marinus*."
Line XXX:

199: Could you please state more about why you are confident in this assignment? What validation or 
metric leads you to this?

Thank you for noting that this needs further clarification. 
We have edited the paragraph for clarity in the revised manuscript adding: "While the accuracy of strain specific protein annotations are limited due to the high conservation of the target protein complexes, the two step approach of peptide-to-spectrum matching using deep paired metagenomics, does assign proteins at the level of clades."
Lines: XXX

209: It seems overly general to blame the culturing difficulties solely on having reduced genomes ? in 
this context, it is the lack of just one gene (catalase) that matters, and reduced genomes reflect an 
outcome of the ultimate suite of selective pressures acting on these systems.

Thank you for your comment. We have edited the text in the revised manuscript: "*Prochlorococcus* remain challenging to culture at high densities or under fluctuating environments, partially due to their dependence upon mutualistic heterotrophic bacteria to detoxify
reactive oxygen species [48,49]." 
Line XXX:

363: Could the authors comment somewhere on why you took this proteomic approach to examine the 
baseline distribution of Pro across O2 and depth, as opposed to (or at least in concert with) the 
extensive metagenomic, amplicon, qPCR, etc datasets of ecotype abundances available?

Thank you for your comment. Co-authors, Aurora MattisonXXX and Dr. Amanda Cockshutt, used the metaproteomics database as an extension of their earlier studies on the functional implications of differential protein allocations across different Prochlorococcus strains [Zorz, Bonisteel]. Those studies indeed motivated the design of our growth matrix. Extending the study to include metagonomic analyses is valid but is beyond our scope.
We have edited the text in the revised manuscript: "Our analyses utilized a proteomics dataset; however, alternate approaches, such as metagenomics or metatranscriptomics, could have been employed to analyze ecotype abundances using the TARA Oceans [@pesantOpenScienceResources2015] dataset."
Line: XXX

407: growth -> grew?

We have edited the text: "The growth rate for Prochlorococcus marinus MED4, clade HLI, under 250 ?M O2, increased with higher imposed PAR and longer photoperiods (Fig 3A), across all spectral wavebands."
Line XXX:

428: How many replicates were carried out for most experiments? Perhaps I just missed it, but please be 
sure this information is clear in the methods

Thank you for this comment. In this work we used 1 to 3 replicates depending on the experimental conditions. Specific number of replicates (shown as small dots in Figures (3 - 5)).
We have edited the text in the revised manuscript:" Consistent absence of growth of some strains under some levels of photoperiod, PAR, or [O~2~] meant we completed 268 growth factor treatment combinations with 1 to 3 replicates."
Line: XXX

431, and generally throughout the paper: Would the authors care to comment on any connection 
between the differences in blue vs red light growth and the underlying amount of energy input from 
these different wavelengths?

XXX DOUGXXX 
XXX Energy- blue have shorter WL so high energy, red longer WL so lower energy??? XXXX
XXX Can add something on line 827 to expand XXXXX

Fig 4 and related figures: It might be useful to plot the specific measured points within the countour plot

Thank you for your suggestion. We have combined the growth rate figure with the GAM model for each strain. We believe this will aid the reader with interpreting both plots.

685: The lack of detected FtsH could simply be a factor of low overall spectral counts and detection 
limits in the dataset; please provide more context to help the reader understand whether this could be 
just a small number problem with respect to MIT9313 in low O2 samples; Fig. 10 seems to just show a 
binary detected/not detected value.