### 2024-06-10 Code/Plots_ProNicheMap.Rmd
Larger fonts for figures

---
title: "Import_Map"
author:
- Sylwia Sliwinska-Wilczewska
- Douglas A. Campbell
date: "`r format(Sys.Date())`"
output:
bookdown::html_document2:
    code_folding: show
    keep_md: yes
    toc: TRUE
    toc_float: TRUE
    toc_depth: 6
    fig_caption: yes
bibliography: Manuscript_O2.bib
csl: plos-one.csl
editor_options: 
  markdown: 
    wrap: 72
---

# Set Chunk Options

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
knitr::opts_chunk$set(fig.path='Figs/')
```

# Introduction

# Load Libraries and set Project Variables

```{r load libraries} 
library(tidyverse)
library(broom)
library(OneR)
library(zoo)
library(ggpubr)
library(data.table)
library(ggspectra)
library(photobiologyWavebands)
library(photobiology)
library(ggforce) #ellipse
library("cowplot")
library("patchwork") # merging plots
library(grid) # for making one strip bigger - grid.draw

```

```{r set project variables}
Project <- "prochlorococcus_O2"

PlotsOut <- file.path("..","Output", "Figures")
PlotsOutPLOS <- file.path("..", "OutPut", "PLOSRound2")
```

```{r read photoperiod depth data}

PhotoperiodDepth <- readRDS(file = file.path("..", "Data", "ProcessedData", "DepthPhotoperiodTableFilter.Rds"))

```

```{r create df for full oxygen and OMZ potential niches}

# add column for each clade with 1 for potential and 0 for no potential niche based on PAR at corresponding depths (50 and 100 m)   for temporal and equatorial zones.
PhotoperiodDepthOMZ <- PhotoperiodDepth |>
  select(-c(ToD, DepthPARabs, DepthPARSine, DepthPARabsNA, Photicperiod:PhoticperiodSineSatPAR_h)) |>
  filter(DepthNames %in% c("50 m", "100 m")) |>
  mutate(DepthNames = factor(DepthNames, levels = c("100 m", "50 m")),
         lat = factor(lat, levels = c("45", "0"))) |>
  relocate(DepthNames, .after = lat) |>
  mutate(CladeHLI = case_when(DayofYear == "Winter Solstice" & lat == 45 ~ 0,
                              DayofYear == "Equinox" & DepthNames == "100 m" & lat == 45 ~ 0,
                              DayofYear == "Summer Solstice" & DepthNames == "100 m" & lat == 0  ~ 0,
                              TRUE ~ 1)) |>
  mutate(CladeLLII_LLIII = case_when(DayofYear == "Winter Solstice" & DepthNames == "100 m" & lat == 45 ~ 0,
                                     DayofYear == "Winter Solstice" & DepthNames == "50 m" & lat == 0 ~ 0,
                                     DayofYear == "Equinox" & DepthNames == "50 m" ~ 0,
                                     DayofYear == "Summer Solstice" & lat == 45  ~ 0,
                                     DayofYear == "Summer Solstice" & DepthNames == "50 m" & lat == 0  ~ 0,
                              TRUE ~ 1 ))  |>
  mutate(CladeLLIV = case_when(DayofYear == "Winter Solstice" & DepthNames == "100 m" & lat == 45 ~ 0,
                              TRUE ~ 1))

PhotoperiodDepthOMZ

PhotoperiodDepthFullO2 <- PhotoperiodDepth |>
  select(-c(ToD, DepthPARabs, DepthPARSine, DepthPARabsNA, Photicperiod:PhoticperiodSineSatPAR_h)) |>
  filter(DepthNames %in% c("50 m", "100 m")) |>
  mutate(DepthNames = factor(DepthNames, levels = c("100 m", "50 m")),
         lat = factor(lat, levels = c("45", "0"))) |>
  relocate(DepthNames, .after = lat) |>
  mutate(CladeHLI = case_when(DayofYear == "Winter Solstice" & DepthNames == "100 m" & lat == 45 ~ 0,
                              TRUE ~ 1)) |>
  mutate(CladeLLII_LLIII = case_when(DayofYear == "Winter Solstice" ~ 0,
                                     DayofYear == "Equinox" & DepthNames == "50 m" ~ 0,
                                     DayofYear == "Summer Solstice" & lat == 45  ~ 0,
                                     DayofYear == "Summer Solstice" & DepthNames == "50 m" & lat == 0  ~ 0,
                              TRUE ~ 1 ))  |>
  mutate(CladeLLIV = case_when(DayofYear == "Winter Solstice" & DepthNames == "100 m" & lat == 45 ~ 0,
                               DayofYear == "Winter Solstice" & DepthNames == "50 m" & lat == 0 ~ 0,
                               DayofYear == "Equinox" & DepthNames == "50 m"  ~ 0,
                               DayofYear == "Summer Solstice" & lat == 45  ~ 0,
                               DayofYear == "Summer Solstice" & DepthNames == "50 m" & lat == 0  ~ 0,
                              TRUE ~ 1))

PhotoperiodDepthFullO2

```

```{r read spectral depth backdrop}

# library("png")
# DepthBackdrop <- readPNG("DepthBackdrop.png")
# # Get dimensions of the image
# width <- dim(DepthBackdrop)[2]
# height <- dim(DepthBackdrop)[1]
# 
# # Plot the image
# plot(1:width, type = "n", xlim = c(0, width), ylim = c(0, height),
#      xlab = "", ylab = "", axes =FALSE)
# rasterImage(DepthBackdrop, 0, 0, width, height) 
# points(PhotoperiodDepthFullO2$PhoticperiodPeakSatPAR_h, PhotoperiodDepthFullO2$DepthNames, col = "red", pch = 16)

```

#fig.height = 10, fig.width = 8, warning = FALSE
```{r, Niches figures}
# 
# 
# NichesFullO2North <- PhotoperiodDepthFullO2 %>%
#   filter(lat == 45) %>% 
# mutate(TextLabel = (case_when(CladeHLI == 1 & CladeLLII_LLIII == 0 & CladeLLIV == 0 ~ "Clade HLI",
#                                 CladeHLI == 1 & CladeLLII_LLIII == 1 & CladeLLIV == 0 ~ "Clades HLI, LLII/III",
#                                 CladeHLI == 1 & CladeLLII_LLIII == 1 & CladeLLIV == 1 ~ "Clades HLI, LLII/III, LLIV",
#                                 CladeLLII_LLIII == 1 & CladeHLI == 0 & CladeLLIV == 0 ~ "Clade LLII/III",
#                                 CladeLLII_LLIII == 1 & CladeHLI == 0 & CladeLLIV == 1 ~ "Clades LLII/III, LLIV",
#                                 CladeLLIV == 1 & CladeHLI == 0 & CladeLLII_LLIII == 0 ~ "Clade LLIV",
#                                 CladeLLIV == 1 & CladeHLI == 1 & CladeLLII_LLIII == 0 ~ "Clades HLI, LLIV",
#                                 CladeLLIV == 0 & CladeHLI == 0 & CladeLLII_LLIII == 0 ~ "")),
#         TextLocation = 5) %>% 
# mutate(DepthNames = factor(DepthNames, levels = c("50 m", "100 m")),
#          lat = factor(lat, levels = c("45", "0"))) %>%
# ggplot()+
#   geom_point(aes(x = PhoticperiodPeakSatPAR_h, y = DepthNames),
#             size = 0.5, alpha = 0, show.legend = TRUE)+
#   labs(title = "Full Oxygen") +
#   scale_x_continuous(breaks=seq(0, 25, by = 5)) +
#     facet_grid(rows = vars(lat),
#               cols = vars(DayofYear),
#               labeller = labeller(lat = label_both, DayofYear = label_both)) +
#    geom_text(aes(x = TextLocation, y = DepthNames, label = paste(TextLabel), show.legend = FALSE ),
#    # parse = TRUE,
#     hjust = 0,
#     vjust = 1,
#     size = 3
#   )  +
#   guides(color = "none") +
#   theme_bw()+
#  theme(strip.text = element_text(size = 10),
#         axis.title = element_blank(),
#         # strip.text.x = element_text(size = 20),
#         axis.text.y = element_text(size = 10),
#         axis.text.x = element_blank(),
#         axis.ticks.x.bottom = element_blank(),
#         axis.ticks.y.left = element_blank(),
#         panel.grid.minor = element_blank(),
#         panel.grid.major = element_blank(),
#         plot.margin = unit(c(0,30,0,0), "pt"))
# 
# NichesFullO2North
# 
# NichesFullO2Equat <- PhotoperiodDepthFullO2 %>%
#   filter(lat == 0) %>% 
# mutate(TextLabel = (case_when(CladeHLI == 1 & CladeLLII_LLIII == 0 & CladeLLIV == 0 ~ "Clade HLI",
#                                 CladeHLI == 1 & CladeLLII_LLIII == 1 & CladeLLIV == 0 ~ "Clades HLI, LLII/III",
#                                 CladeHLI == 1 & CladeLLII_LLIII == 1 & CladeLLIV == 1 ~ "Clades HLI, LLII/III, LLIV",
#                                 CladeLLII_LLIII == 1 & CladeHLI == 0 & CladeLLIV == 0 ~ "Clade LLII/III",
#                                 CladeLLII_LLIII == 1 & CladeHLI == 0 & CladeLLIV == 1 ~ "Clades LLII/III, LLIV",
#                                 CladeLLIV == 1 & CladeHLI == 0 & CladeLLII_LLIII == 0 ~ "Clade LLIV",
#                                 CladeLLIV == 1 & CladeHLI == 1 & CladeLLII_LLIII == 0 ~ "Clades HLI, LLIV",
#                                 CladeLLIV == 0 & CladeHLI == 0 & CladeLLII_LLIII == 0 ~ "")),
#         TextLocation = 5) %>% 
# mutate(DepthNames = factor(DepthNames, levels = c("50 m", "100 m")),
#          lat = factor(lat, levels = c("45", "0"))) %>%
# ggplot()+
#   geom_point(aes(x = PhoticperiodPeakSatPAR_h, y = DepthNames),
#             size = 0.5, alpha = 0, show.legend = TRUE)+
#   # labs(title = "Full Oxygen") +
#   scale_x_continuous(breaks=seq(0, 25, by = 5)) +
#     facet_grid(rows = vars(lat),
#               cols = vars(DayofYear),
#               labeller = labeller(lat = label_both)) +
#    geom_text(aes(x = TextLocation, y = DepthNames, label = paste(TextLabel), show.legend = FALSE ),
#    # parse = TRUE,
#     hjust = 0,
#     vjust = 1,
#     size = 3
#   )  +
#   guides(color = "none") +
#   theme_bw() +
#  theme(strip.text.x = element_blank(),
#        strip.text.y = element_text(size = 10),
#         axis.title = element_blank(),
#         # strip.text.x = element_text(size = 20),
#         axis.text.y = element_text(size = 10),
#         # axis.text.y = element_blank(),
#         # axis.text.y = element_text(size = 5),
#          axis.text.x = element_blank(),
#         axis.ticks.x.bottom = element_blank(),
#         axis.ticks.y.left = element_blank(),
#         panel.grid.minor = element_blank(),
#         panel.grid.major = element_blank())
# 
# NichesFullO2Equat
# 
# 
# NichesOMZNorth <- PhotoperiodDepthOMZ %>%
#   filter(lat == 45) %>% 
# mutate(TextLabel = (case_when(CladeHLI == 1 & CladeLLII_LLIII == 0 & CladeLLIV == 0 ~ "Clade HLI",
#                                 CladeHLI == 1 & CladeLLII_LLIII == 1 & CladeLLIV == 0 ~ "Clades HLI, LLII/III",
#                                 CladeHLI == 1 & CladeLLII_LLIII == 1 & CladeLLIV == 1 ~ "Clades HLI, LLII/III, LLIV",
#                                 CladeLLII_LLIII == 1 & CladeHLI == 0 & CladeLLIV == 0 ~ "Clade LLII/III",
#                                 CladeLLII_LLIII == 1 & CladeHLI == 0 & CladeLLIV == 1 ~ "Clades LLII/III, LLIV",
#                                 CladeLLIV == 1 & CladeHLI == 0 & CladeLLII_LLIII == 0 ~ "Clade LLIV",
#                                 CladeLLIV == 1 & CladeHLI == 1 & CladeLLII_LLIII == 0 ~ "Clades HLI, LLIV",
#                                 CladeLLIV == 0 & CladeHLI == 0 & CladeLLII_LLIII == 0 ~ "")),
#         TextLocation = 5) %>% 
# mutate(DepthNames = factor(DepthNames, levels = c("50 m", "100 m")),
#          lat = factor(lat, levels = c("45", "0"))) %>%
# ggplot()+
#   geom_point(aes(x = PhoticperiodPeakSatPAR_h, y = DepthNames),
#             size = 0.5, alpha = 0, show.legend = TRUE)+
#    labs(title = "OMZ") +
#   scale_x_continuous(breaks=seq(0, 25, by = 5)) +
#     facet_grid(rows = vars(lat),
#               cols = vars(DayofYear),
#               labeller = labeller(lat = label_both)) +
#    geom_text(aes(x = TextLocation, y = DepthNames, label = paste(TextLabel), show.legend = FALSE ),
#    # parse = TRUE,
#     hjust = 0,
#     vjust = 1,
#     size = 3
#   )  +
#   guides(color = "none") +
#   theme_bw() +
#  theme(strip.text.x = element_text(size = 10),
#        strip.text.y = element_text(size = 10),
#         axis.title = element_blank(),
#         # strip.text.x = element_text(size = 20),
#         axis.text.y = element_text(size = 10),
#         # axis.text.y = element_blank(),
#         axis.text.x = element_blank(),
#         axis.ticks.x.bottom = element_blank(),
#         axis.ticks.y.left = element_blank(),
#         panel.grid.minor = element_blank(),
#         panel.grid.major = element_blank())
# 
# NichesOMZNorth
# 
# 
# 
# 
# NichesOMZEquat <- PhotoperiodDepthOMZ %>%
#   filter(lat == 0) %>% 
# mutate(TextLabel = (case_when(CladeHLI == 1 & CladeLLII_LLIII == 0 & CladeLLIV == 0 ~ "Clade HLI",
#                                 CladeHLI == 1 & CladeLLII_LLIII == 1 & CladeLLIV == 0 ~ "Clades HLI, LLII/III",
#                                 CladeHLI == 1 & CladeLLII_LLIII == 1 & CladeLLIV == 1 ~ "Clades HLI, LLII/III, LLIV",
#                                 CladeLLII_LLIII == 1 & CladeHLI == 0 & CladeLLIV == 0 ~ "Clade LLII/III",
#                                 CladeLLII_LLIII == 1 & CladeHLI == 0 & CladeLLIV == 1 ~ "Clades LLII/III, LLIV",
#                                 CladeLLIV == 1 & CladeHLI == 0 & CladeLLII_LLIII == 0 ~ "Clade LLIV",
#                                 CladeLLIV == 1 & CladeHLI == 1 & CladeLLII_LLIII == 0 ~ "Clades HLI, LLIV",
#                                 CladeLLIV == 0 & CladeHLI == 0 & CladeLLII_LLIII == 0 ~ "")),
#         TextLocation = 5) %>% 
# mutate(DepthNames = factor(DepthNames, levels = c("50 m", "100 m")),
#          lat = factor(lat, levels = c("45", "0"))) %>%
# ggplot()+
#   geom_point(aes(x = PhoticperiodPeakSatPAR_h, y = DepthNames),
#             size = 0.5, alpha = 0, show.legend = TRUE)+
#   # labs(title = "Full Oxygen") +
#   scale_x_continuous(breaks=seq(0, 25, by = 5)) +
#     facet_grid(rows = vars(lat),
#               cols = vars(DayofYear),
#               labeller = labeller(lat = label_both)) +
#    geom_text(aes(x = TextLocation, y = DepthNames, label = paste(TextLabel), show.legend = FALSE ),
#    # parse = TRUE,
#     hjust = 0,
#     vjust = 1,
#     size = 3
#   )  +
#   guides(color = "none") +
#   theme_bw() +
#  theme(strip.text.x = element_blank(),
#        strip.text.y = element_text(size = 10),
#         axis.title = element_blank(),
#         # strip.text.x = element_text(size = 20),
#         axis.text.y = element_text(size = 10),
#         # axis.text.y = element_blank(),
#          axis.text.x = element_blank(),
#         axis.ticks.x.bottom = element_blank(),
#         axis.ticks.y.left = element_blank(),
#         panel.grid.minor = element_blank(),
#         panel.grid.major = element_blank())
# 
# NichesOMZEquat

```

# Create map
https://www.geeksforgeeks.org/how-to-make-world-map-with-ggplot2-in-r/

```{r world map}
world_coordinates <- map_data("world") 
# world_coordinates$facetsFuture <- 'Future~scenario'

data_textA <- data.frame(label = c('A'))
data_textB <- data.frame(label = c('B'))
# data_textC <- data.frame(label = c('C'))
```

```{r create maps, fig.height = 10, fig.width = 6, warning = FALSE}

d=data.frame(x1=c(-Inf), x2=c(Inf), y1=c(35), y2=c(55))
d2=data.frame(x1=c(-Inf), x2=c(Inf), y1=c(-35), y2=c(-55))
d3=data.frame(x1=c(-Inf), x2=c(Inf), y1=c(-Inf), y2=c(Inf))
d4=data.frame(x1=c(-Inf), x2=c(Inf), y1=c(-15), y2=c(15))
d5=data.frame(x1=c(-Inf), x2=c(Inf), y1=c(-40), y2=c(40))

FullO2map<-ggplot() + 
  geom_rect(data=d3, mapping=aes(xmin=x1, xmax=x2, ymin=y1, ymax=y2), fill = "lightskyblue1", color="lightskyblue3", alpha=0.5) +
  geom_rect(data=d, mapping=aes(xmin=x1, xmax=x2, ymin=y1, ymax=y2), fill = "khaki1", color="khaki2", alpha=0.8) +
  geom_rect(data=d2, mapping=aes(xmin=x1, xmax=x2, ymin=y1, ymax=y2), fill = "khaki1", color="khaki2", alpha=0.8) +
    geom_rect(data=d4, mapping=aes(xmin=x1, xmax=x2, ymin=y1, ymax=y2), fill = "khaki1", color="khaki2", alpha=0.8) +
  geom_hline(yintercept = 40, fill = "transparent", color="olivedrab3", linetype = "dashed", alpha=1, size = 1) +
  geom_hline(yintercept = -40, fill = "transparent", color="olivedrab3", linetype = "dashed", alpha=1, size = 1) +
  geom_map(data = world_coordinates, map = world_coordinates, aes(long, lat, map_id = region), color = "white", fill = "gray50", size = 0.2) +
    geom_text(data=data_textA, aes(x=180, y=120, label=label), size=7, parse=TRUE) +
  scale_y_continuous(breaks=seq(-90, 90, by = 45)) +
  coord_cartesian(ylim = c(-90, 120)) +
  labs(y = "Latitude", x = "Longitude") +
  # ggh4x::facet_nested(rows = vars(facetsFuture), labeller = label_parsed, switch = "y") +
  theme_bw() +
  theme(panel.grid.minor = element_blank(),
        panel.grid.major = element_blank(),
        strip.placement = "outside",
        strip.switch.pad.grid = unit(1, "cm"),
        axis.text = element_text(size=14),
        axis.title = element_text(size=14),
        strip.background = element_rect(fill="white"),
        strip.text = element_text(size=18),
        axis.title.y = element_text(vjust = 1),
        #axis.title.x = element_text(margin=margin(t=10)),
        axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        legend.background = element_rect(fill="transparent"),
        legend.position = c(0.12,0.76),
        legend.key.height= unit(0.005, 'cm'),
        legend.spacing.x = unit(0.01, 'cm'),
        legend.text = element_text(size=14))
FullO2map

OMZmap<-ggplot() + 
  geom_rect(data=d3, mapping=aes(xmin=x1, xmax=x2, ymin=y1, ymax=y2), fill = "lightskyblue1", color="lightskyblue3", alpha=0.5) +
  geom_rect(data=d, mapping=aes(xmin=x1, xmax=x2, ymin=y1, ymax=y2), fill = "khaki1", color="khaki2", alpha=0.8) +
  geom_rect(data=d2, mapping=aes(xmin=x1, xmax=x2, ymin=y1, ymax=y2), fill = "khaki1", color="khaki2", alpha=0.8) +
  geom_rect(data=d4, mapping=aes(xmin=x1, xmax=x2, ymin=y1, ymax=y2), fill = "khaki1", color="khaki2", alpha=0.8) +
  geom_hline(yintercept = 40, fill = "transparent", color="olivedrab3", linetype = "dashed", alpha=1, size = 1) +
  geom_hline(yintercept = -40, fill = "transparent", color="olivedrab3", linetype = "dashed", alpha=1, size = 1) +
  geom_map(data = world_coordinates, map = world_coordinates, aes(long, lat, map_id = region), color = "white", fill = "gray50", size = 0.2) +
     geom_text(data=data_textB, aes(x=180, y=120, label=label), size=7, parse=TRUE) +
  scale_y_continuous(breaks=seq(-90, 90, by = 45)) +
  coord_cartesian(ylim = c(-90, 120)) +
  labs(y = "Latitude", x = "Longitude") +
  # ggh4x::facet_nested(rows = vars(facetsFuture), labeller = label_parsed, switch = "y") +
  theme_bw() +
  theme(panel.grid.minor = element_blank(),
        panel.grid.major = element_blank(),
        strip.placement = "outside",
        strip.switch.pad.grid = unit(1, "cm"),
        axis.text = element_text(size=14),
        axis.title = element_text(size=14),
        strip.background = element_rect(fill="white"),
        strip.text = element_text(size=18),
        axis.title.y = element_text(vjust = 1),
        #axis.title.x = element_text(margin=margin(t=10)),
        axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        legend.background = element_rect(fill="transparent"),
        legend.position = c(0.12,0.76),
        legend.key.height= unit(0.005, 'cm'),
        legend.spacing.x = unit(0.01, 'cm'),
        legend.text = element_text(size=14))
OMZmap

```

-------------------------------------------------------------------------------------------------------------------
---------------------------------------------- # SySl changes -----------------------------------------------------------
----------------------------------------------------------------------------------------------------------------

fig.height = 10, fig.width = 6, warning = FALSE
```{r create map, fig.height = 16, fig.width = 10, warning = FALSE}

d=data.frame(x1=c(-Inf), x2=c(Inf), y1=c(35), y2=c(55))
d2=data.frame(x1=c(-Inf), x2=c(Inf), y1=c(-35), y2=c(-55))
d3=data.frame(x1=c(-Inf), x2=c(Inf), y1=c(-Inf), y2=c(Inf))
d4=data.frame(x1=c(-Inf), x2=c(Inf), y1=c(-15), y2=c(15))
d5=data.frame(x1=c(-Inf), x2=c(Inf), y1=c(-40), y2=c(40))

map<-ggplot() + 
  geom_rect(data=d3, mapping=aes(xmin=x1, xmax=x2, ymin=y1, ymax=y2), fill = "lightskyblue1", color="lightskyblue3", alpha=0.5) +
  geom_rect(data=d, mapping=aes(xmin=x1, xmax=x2, ymin=y1, ymax=y2), fill = "khaki1", color="khaki2", alpha=0.8) +
  geom_rect(data=d2, mapping=aes(xmin=x1, xmax=x2, ymin=y1, ymax=y2), fill = "khaki1", color="khaki2", alpha=0.8) +
    geom_rect(data=d4, mapping=aes(xmin=x1, xmax=x2, ymin=y1, ymax=y2), fill = "khaki1", color="khaki2", alpha=0.8) +
  geom_hline(yintercept = 40, fill = "transparent", color="chartreuse4", linetype = "dashed", alpha=1, size = 1) +
  geom_hline(yintercept = -40, fill = "transparent", color="chartreuse4", linetype = "dashed", alpha=1, size = 1) +
  geom_map(data = world_coordinates, map = world_coordinates, aes(long, lat, map_id = region), color = "gray50", fill = "gray50", size = 0.2) +
   # geom_text(data=data_textA, aes(x=180, y=120, label=label), size=7, parse=TRUE) +
  scale_y_continuous(breaks=seq(-90, 90, by = 45)) +
  coord_cartesian(ylim = c(-90, 120)) +
  coord_fixed(ratio = 2, xlim = NULL, ylim = NULL, expand = TRUE, clip = "on") +
  labs(y = "Latitude", x = "Longitude") +
  # ggh4x::facet_nested(rows = vars(facetsFuture), labeller = label_parsed, switch = "y") +
  theme_bw() +
  theme(panel.grid.minor = element_blank(),
        panel.grid.major = element_blank(),
        axis.text = element_text(size=25),
        axis.title = element_text(size=28),
        strip.background = element_rect(fill="white"),
        strip.text = element_text(size=20),
        #axis.title.x = element_text(margin=margin(t=10)),
        axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        legend.background = element_rect(fill="transparent"),
        legend.position = c(0.12,0.76),
        legend.key.height= unit(0.005, 'cm'),
        legend.spacing.x = unit(0.01, 'cm'),
        legend.text = element_text(size=14)) 
map

```

```{r create globe shape map, fig.height = 8, fig.width = 14, warning = FALSE}
world_coordinates2 <- map_data("world") %>% 
  filter(! long > 180)

mapCircle<-ggplot() + 
  geom_rect(data=d3, mapping=aes(xmin=x1, xmax=x2, ymin=y1, ymax=y2), fill = "lightskyblue1", color="lightskyblue3", alpha=0.5) +
  geom_rect(data=d, mapping=aes(xmin=x1, xmax=x2, ymin=y1, ymax=y2), fill = "khaki1", color="khaki2", alpha=0.8) +
  geom_rect(data=d2, mapping=aes(xmin=x1, xmax=x2, ymin=y1, ymax=y2), fill = "khaki1", color="khaki2", alpha=0.8) +
    geom_rect(data=d4, mapping=aes(xmin=x1, xmax=x2, ymin=y1, ymax=y2), fill = "khaki1", color="khaki2", alpha=0.8) +
  geom_hline(yintercept = 40, fill = "transparent", color="chartreuse4", linetype = "dashed", alpha=1, size = 1) +
  geom_hline(yintercept = -40, fill = "transparent", color="chartreuse4", linetype = "dashed", alpha=1, size = 1) +
  geom_map(data = world_coordinates2, map = world_coordinates2, aes(long, lat, map_id = region), color = "gray50", fill = "gray50", size = 0.2) +
   # geom_text(data=data_textA, aes(x=180, y=120, label=label), size=7, parse=TRUE) +
  scale_y_continuous(breaks=seq(-90, 90, by = 45)) +
  coord_cartesian(ylim = c(-90, 90)) +
  #coord_fixed(ratio = 2, xlim = NULL, ylim = NULL, expand = TRUE, clip = "on") +
  labs(y = "Latitude", x = "Longitude") +
  # ggh4x::facet_nested(rows = vars(facetsFuture), labeller = label_parsed, switch = "y") +
  theme_bw() +
  theme(panel.grid.minor = element_blank(),
        panel.grid.major = element_blank(),
        axis.text = element_text(size=20),
        axis.title = element_text(size=28),
        strip.background = element_rect(fill="white"),
        strip.text = element_text(size=20),
        #axis.title.x = element_text(margin=margin(t=10)),
        axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        legend.background = element_rect(fill="transparent"),
        legend.position = c(0.12,0.76),
        legend.key.height= unit(0.005, 'cm'),
        legend.spacing.x = unit(0.01, 'cm'),
        legend.text = element_text(size=14)) +
    coord_map("moll") 
mapCircle

```

```{r, Niches figures, fig.height = 4, fig.width = 8, warning = FALSE}

d=data.frame(x=c(-Inf,700,700), y=c(-Inf,-Inf,3), t=c('a', 'a', 'a'), r=c(1,2,3))

PhotoperiodDepthFullO2<-PhotoperiodDepthFullO2
PhotoperiodDepthFullO2$nm <- c(400, 400, 400, 400, 400, 400, 700, 700, 700, 700, 700, 700)

PhotoperiodDepthOMZ<-PhotoperiodDepthOMZ
PhotoperiodDepthOMZ$nm <- c(400, 400, 400, 400, 400, 400, 700, 700, 700, 700, 700, 700)

NichesFullO2North <- PhotoperiodDepthFullO2 %>%
  filter(lat == 45) %>% 
mutate(TextLabel = (case_when(CladeHLI == 1 & CladeLLII_LLIII == 0 & CladeLLIV == 0 ~ "HLI",
                                CladeHLI == 1 & CladeLLII_LLIII == 1 & CladeLLIV == 0 ~ "HLI, LLII/III",
                                CladeHLI == 1 & CladeLLII_LLIII == 1 & CladeLLIV == 1 ~ "HLI, LLII/III, LLIV",
                                CladeLLII_LLIII == 1 & CladeHLI == 0 & CladeLLIV == 0 ~ "LLII/III",
                                CladeLLII_LLIII == 1 & CladeHLI == 0 & CladeLLIV == 1 ~ "LLII/III, LLIV",
                                CladeLLIV == 1 & CladeHLI == 0 & CladeLLII_LLIII == 0 ~ "LLIV",
                                CladeLLIV == 1 & CladeHLI == 1 & CladeLLII_LLIII == 0 ~ "HLI, LLIV",
                                CladeLLIV == 0 & CladeHLI == 0 & CladeLLII_LLIII == 0 ~ "")),
        TextLocation = 400) %>% 
ggplot()+
  geom_point(aes(x = nm, y = DepthNames), size = 0.5, alpha = 0, show.legend = TRUE)+
  # stat_wl_strip(aes(x = nm), ymin = -Inf, ymax = Inf, alpha = 0.3, PhotoperiodDepthFullO2) +
  # scale_fill_identity() +
  # scale_x_continuous(breaks=seq(0, 25, by = 5)) +
  # coord_cartesian(xlim = c(0, 20)) +
  coord_fixed(ratio = 55, xlim = NULL, ylim = NULL, expand = TRUE, clip = "on") +
  #geom_ellipse(aes(x0 = 555, y0 = 1.43, a = 10, b = 2, angle = pi / 1), fill = "transparent", colour = "transparent") +
  #geom_polygon(data=d, mapping=aes(x=x, y=y, group=t), fill = "white", show.legend = F) +
  geom_text(aes(x = TextLocation, y = DepthNames, label = paste(TextLabel), show.legend = FALSE), hjust = 0, size = 8)  +
  ggh4x::facet_nested(cols = vars(DayofYear), labeller = labeller(lat = label_both)) +
  guides(color = "none") +
  labs(title = "Full Oxygen Zones") +
  theme_bw() +
  theme(plot.title = element_text(size = 30),
        panel.grid.minor = element_blank(),
        panel.grid.major = element_blank(),
        axis.text = element_text(size=28),
        #axis.title = element_text(size=14),
        strip.background = element_rect(fill="white"),
        strip.text = element_text(size=28),
        #axis.title.y = element_text(vjust = 1),
        #axis.title.x = element_text(margin=margin(t=10)),
        axis.title = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        legend.background = element_rect(fill="transparent"),
        legend.position = c(0.12,0.76),
        legend.key.height= unit(0.005, 'cm'),
        legend.spacing.x = unit(0.01, 'cm'),
        legend.text = element_text(size=14)) 
NichesFullO2North

NichesFullO2Equat <- PhotoperiodDepthFullO2 %>%
  filter(lat == 0) %>% 
mutate(TextLabel = (case_when(CladeHLI == 1 & CladeLLII_LLIII == 0 & CladeLLIV == 0 ~ "HLI",
                                CladeHLI == 1 & CladeLLII_LLIII == 1 & CladeLLIV == 0 ~ "HLI, LLII/III",
                                CladeHLI == 1 & CladeLLII_LLIII == 1 & CladeLLIV == 1 ~ "HLI, LLII/III, LLIV",
                                CladeLLII_LLIII == 1 & CladeHLI == 0 & CladeLLIV == 0 ~ "LLII/III",
                                CladeLLII_LLIII == 1 & CladeHLI == 0 & CladeLLIV == 1 ~ "LLII/III, LLIV",
                                CladeLLIV == 1 & CladeHLI == 0 & CladeLLII_LLIII == 0 ~ "LLIV",
                                CladeLLIV == 1 & CladeHLI == 1 & CladeLLII_LLIII == 0 ~ "HLI, LLIV",
                                CladeLLIV == 0 & CladeHLI == 0 & CladeLLII_LLIII == 0 ~ "")),
        TextLocation = 400) %>% 
ggplot()+
  geom_point(aes(x = nm, y = DepthNames), size = 0.5, alpha = 0, show.legend = TRUE)+
  # stat_wl_strip(aes(x = nm), ymin = -Inf, ymax = Inf, alpha = 0.3, PhotoperiodDepthFullO2) +
  # scale_fill_identity() +
  # scale_x_continuous(breaks=seq(0, 25, by = 5)) +
  # coord_cartesian(xlim = c(0, 20)) +
  coord_fixed(ratio = 55, xlim = NULL, ylim = NULL, expand = TRUE, clip = "on") +
  #geom_ellipse(aes(x0 = 555, y0 = 1.43, a = 10, b = 2, angle = pi / 1), fill = "transparent", colour = "transparent") +
  #geom_polygon(data=d, mapping=aes(x=x, y=y, group=t), fill = "white", show.legend = F) +
  geom_text(aes(x = TextLocation, y = DepthNames, label = paste(TextLabel), show.legend = FALSE), hjust = 0, size = 8)  +
  ggh4x::facet_nested(cols = vars(DayofYear), labeller = labeller(lat = label_both)) +
  guides(color = "none") +
  theme_bw() +
  theme(plot.title = element_text(vjust = 10),
        panel.grid.minor = element_blank(),
        panel.grid.major = element_blank(),
        axis.text = element_text(size=28),
        #axis.title = element_text(size=14),
        strip.background = element_rect(fill="white"),
        strip.text = element_text(size=28),
        #axis.title.y = element_text(vjust = 1),
        #axis.title.x = element_text(margin=margin(t=10)),
        axis.title = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        legend.background = element_rect(fill="transparent"),
        legend.position = c(0.12,0.76),
        legend.key.height= unit(0.005, 'cm'),
        legend.spacing.x = unit(0.01, 'cm'),
        legend.text = element_text(size=14))

NichesFullO2Equat

NichesOMZNorth <- PhotoperiodDepthOMZ %>%
  filter(lat == 45) %>% 
mutate(TextLabel = (case_when(CladeHLI == 1 & CladeLLII_LLIII == 0 & CladeLLIV == 0 ~ "HLI",
                                CladeHLI == 1 & CladeLLII_LLIII == 1 & CladeLLIV == 0 ~ "HLI, LLII/III",
                                CladeHLI == 1 & CladeLLII_LLIII == 1 & CladeLLIV == 1 ~ "HLI, LLII/III, LLIV",
                                CladeLLII_LLIII == 1 & CladeHLI == 0 & CladeLLIV == 0 ~ "LLII/III",
                                CladeLLII_LLIII == 1 & CladeHLI == 0 & CladeLLIV == 1 ~ "LLII/III, LLIV",
                                CladeLLIV == 1 & CladeHLI == 0 & CladeLLII_LLIII == 0 ~ "LLIV",
                                CladeLLIV == 1 & CladeHLI == 1 & CladeLLII_LLIII == 0 ~ "HLI, LLIV",
                                CladeLLIV == 0 & CladeHLI == 0 & CladeLLII_LLIII == 0 ~ "")),
        TextLocation = 400) %>% 
ggplot()+
  geom_point(aes(x = nm, y = DepthNames), size = 0.5, alpha = 0, show.legend = TRUE)+
  # stat_wl_strip(aes(x = nm), ymin = -Inf, ymax = Inf, alpha = 0.3, PhotoperiodDepthOMZ) +
  # scale_fill_identity() +
  # scale_x_continuous(breaks=seq(0, 25, by = 5)) +
  # coord_cartesian(xlim = c(0, 20)) +
  coord_fixed(ratio = 55, xlim = NULL, ylim = NULL, expand = TRUE, clip = "on") +
  #geom_ellipse(aes(x0 = 555, y0 = 1.43, a = 10, b = 2, angle = pi / 1), fill = "transparent", colour = "transparent") +
  #geom_polygon(data=d, mapping=aes(x=x, y=y, group=t), fill = "white", show.legend = F) +
  geom_text(aes(x = TextLocation, y = DepthNames, label = paste(TextLabel), show.legend = FALSE), hjust = 0, size = 8)  +
  ggh4x::facet_nested(cols = vars(DayofYear), labeller = labeller(lat = label_both)) +
  guides(color = "none") +
  labs(title = "Oxygen Minimum Zones") +
  theme_bw() +
  theme(plot.title = element_text(size = 30),
        panel.grid.minor = element_blank(),
        panel.grid.major = element_blank(),
        axis.text = element_text(size=28),
        #axis.title = element_text(size=14),
        strip.background = element_rect(fill="white"),
        strip.text = element_text(size=28),
        #axis.title.y = element_text(vjust = 1),
        #axis.title.x = element_text(margin=margin(t=10)),
        axis.title = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        legend.background = element_rect(fill="transparent"),
        legend.position = c(0.12,0.76),
        legend.key.height= unit(0.005, 'cm'),
        legend.spacing.x = unit(0.01, 'cm'),
        legend.text = element_text(size=14))

NichesOMZNorth

NichesOMZEquat <- PhotoperiodDepthOMZ %>%
  filter(lat == 0) %>% 
mutate(TextLabel = (case_when(CladeHLI == 1 & CladeLLII_LLIII == 0 & CladeLLIV == 0 ~ "HLI",
                                CladeHLI == 1 & CladeLLII_LLIII == 1 & CladeLLIV == 0 ~ "HLI, LLII/III",
                                CladeHLI == 1 & CladeLLII_LLIII == 1 & CladeLLIV == 1 ~ "HLI, LLII/III, LLIV",
                                CladeLLII_LLIII == 1 & CladeHLI == 0 & CladeLLIV == 0 ~ "LLII/III",
                                CladeLLII_LLIII == 1 & CladeHLI == 0 & CladeLLIV == 1 ~ "LLII/III, LLIV",
                                CladeLLIV == 1 & CladeHLI == 0 & CladeLLII_LLIII == 0 ~ "LLIV",
                                CladeLLIV == 1 & CladeHLI == 1 & CladeLLII_LLIII == 0 ~ "HLI, LLIV",
                                CladeLLIV == 0 & CladeHLI == 0 & CladeLLII_LLIII == 0 ~ "")),
        TextLocation = 400) %>%

ggplot()+
  geom_point(aes(x = nm, y = DepthNames), size = 0.5, alpha = 0, show.legend = TRUE)+
  # stat_wl_strip(aes(x = nm), ymin = -Inf, ymax = Inf, alpha = 0.3, PhotoperiodDepthOMZ) +
  # scale_fill_identity() +
  # scale_x_continuous(breaks=seq(0, 25, by = 5)) +
  # coord_cartesian(xlim = c(0, 20)) +
  coord_fixed(ratio = 55, xlim = NULL, ylim = NULL, expand = TRUE, clip = "on") +
  #geom_ellipse(aes(x0 = 555, y0 = 1.43, a = 10, b = 2, angle = pi / 1), fill = "transparent", colour = "transparent") +
  #geom_polygon(data=d, mapping=aes(x=x, y=y, group=t), fill = "white", show.legend = F) +
  geom_text(aes(x = TextLocation, y = DepthNames, label = paste(TextLabel), show.legend = FALSE), hjust = 0, size = 8)  +
  ggh4x::facet_nested(cols = vars(DayofYear), labeller = labeller(lat = label_both)) +
  guides(color = "none") +
  theme_bw() +
  theme(plot.title = element_text(vjust = 10),
        panel.grid.minor = element_blank(),
        panel.grid.major = element_blank(),
        axis.text = element_text(size=28),
        #axis.title = element_text(size=14),
        strip.background = element_rect(fill="white"),
        strip.text = element_text(size=28),
        #axis.title.y = element_text(vjust = 1),
        #axis.title.x = element_text(margin=margin(t=10)),
        axis.title = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        legend.background = element_rect(fill="transparent"),
        legend.position = c(0.12,0.76),
        legend.key.height= unit(0.005, 'cm'),
        legend.spacing.x = unit(0.01, 'cm'),
        legend.text = element_text(size=14))

NichesOMZEquat

BlankPanel <- PhotoperiodDepthOMZ %>%
  filter(lat == 0) %>% 
mutate(TextLabel = (case_when(CladeHLI == 1 & CladeLLII_LLIII == 0 & CladeLLIV == 0 ~ "HLI",
                                CladeHLI == 1 & CladeLLII_LLIII == 1 & CladeLLIV == 0 ~ "HLI, LLII/III",
                                CladeHLI == 1 & CladeLLII_LLIII == 1 & CladeLLIV == 1 ~ "HLI, LLII/III, LLIV",
                                CladeLLII_LLIII == 1 & CladeHLI == 0 & CladeLLIV == 0 ~ "LLII/III",
                                CladeLLII_LLIII == 1 & CladeHLI == 0 & CladeLLIV == 1 ~ "LLII/III, LLIV",
                                CladeLLIV == 1 & CladeHLI == 0 & CladeLLII_LLIII == 0 ~ "LLIV",
                                CladeLLIV == 1 & CladeHLI == 1 & CladeLLII_LLIII == 0 ~ "HLI, LLIV",
                                CladeLLIV == 0 & CladeHLI == 0 & CladeLLII_LLIII == 0 ~ "")),
        TextLocation = 400) %>%

ggplot()+
  geom_point(aes(x = nm, y = DepthNames), size = 0.5, alpha = 0, show.legend = TRUE)+
  # stat_wl_strip(aes(x = nm), ymin = -Inf, ymax = Inf, alpha = 0.3, PhotoperiodDepthOMZ) +
  # scale_fill_identity() +
  # scale_x_continuous(breaks=seq(0, 25, by = 5)) +
  # coord_cartesian(xlim = c(0, 20)) +
  coord_fixed(ratio = 55, xlim = NULL, ylim = NULL, expand = TRUE, clip = "on") +
  #geom_ellipse(aes(x0 = 555, y0 = 1.43, a = 10, b = 2, angle = pi / 1), fill = "transparent", colour = "transparent") +
  #geom_polygon(data=d, mapping=aes(x=x, y=y, group=t), fill = "white", show.legend = F) +
  # geom_text(aes(x = TextLocation, y = DepthNames, label = paste(TextLabel), show.legend = FALSE), hjust = 0, size = 8)  +
  ggh4x::facet_nested(cols = vars(DayofYear), labeller = labeller(lat = label_both)) +
  guides(color = "none") +
  theme_bw() +
  theme(plot.title = element_blank(),
        panel.grid.minor = element_blank(),
        panel.grid.major = element_blank(),
        axis.text = element_blank(),
        #axis.title = element_text(size=14),
        strip.background = element_blank(),
        strip.text = element_blank(),
        #axis.title.y = element_text(vjust = 1),
        #axis.title.x = element_text(margin=margin(t=10)),
        axis.title = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks = element_blank(),
        panel.border =  element_blank(),
        legend.background = element_rect(fill="transparent"),
        legend.position = c(0.12,0.76),
        legend.key.height= unit(0.005, 'cm'),
        legend.spacing.x = unit(0.01, 'cm'),
        legend.text = element_blank())

BlankPanel

```

#Cool web about merging plots
http://www.sthda.com/english/articles/24-ggpubr-publication-ready-plots/81-ggplot2-easy-way-to-mix-multiple-graphs-on-the-same-page/

```{r join figures, fig.height = 4, fig.width = 8, warning = FALSE}

NichesMap_1 <- ggdraw() +
  draw_plot(NichesFullO2North, x = 0, y = 0.65, width = .3, height = .3) +
  draw_plot(NichesFullO2Equat, x = 0, y = 0.455, width = .3, height = .3) +
  draw_plot(map, x = .1, y = .2, width = 0.8, height = 0.8) +
  draw_plot(NichesOMZNorth, x = 0.7, y = 0.65, width = .3, height = 0.3) +
  draw_plot(NichesOMZEquat, x = 0.7, y = 0.455, width = .3, height = 0.3)

```

```{r save plot}

# ggsave(file = file.path(PlotsOut, paste("NichesMap",".png",sep = "")), height=12, width= 28,  dpi = 300, limitsize = TRUE)
# 
# ggsave(file = file.path(PlotsOutPLOS, "Fig10.tiff"), plot = NichesMap, device = NULL, scale = 1, height=20, width= 30, units = c("cm"),dpi = 300, compression = "lzw", limitsize = TRUE)

```

```{r}
ggdraw() +
  draw_plot(NichesFullO2North, x = 0, y = 0.68, width = .26, height = .26) +
  draw_plot(NichesFullO2Equat, x = 0, y = 0.51, width = .26, height = .26) +
  draw_plot(mapCircle, x = .26, y = .29, width = 0.46, height = 0.7) +
  draw_plot(NichesOMZNorth, x = 0.72, y = 0.68, width = .26, height = 0.26) +
  draw_plot(NichesOMZEquat, x = 0.72, y = 0.51, width = .26, height = 0.26)

# ggsave(file = file.path(PlotsOut, paste("NichesMap_2",".png",sep = "")), height=12, width= 28,  dpi = 300, limitsize = TRUE)
```

```{r, fig.height = 8, fig.width = 16, warning = FALSE}

# P1<-ggdraw() +
#   draw_plot(NichesFullO2North, x = 0.001, y = 0.695, width = .36, height = .22) +
#   draw_plot(NichesFullO2Equat, x = 0.001, y = 0.485, width = .36, height = .22) +
#   draw_plot(map, x = .179, y = .2, width = 0.8, height = 0.8) 
# P1
# 
# P2<-ggdraw() +
#   draw_plot(NichesOMZNorth, x = 0.001, y = 0.695, width = .36, height = .22) +
#   draw_plot(NichesOMZEquat, x = 0.001, y = 0.485, width = .36, height = .22) +
#   draw_plot(map, x = .179, y = .2, width = 0.8, height = 0.8) 
# P2
# 
# 
# 
# ggarrange(P1, P2, ncol = 1, nrow = 2,
#           labels = c("A", "B"),  
#           font.label = list(size = 25, color = "black"),
#           align = "hv")

# ggsave(file = file.path(PlotsOut, paste("NichesMap_3",".png",sep = "")), height=18, width= 20,  dpi = 300, limitsize = TRUE, bg='white')
# 
# ggsave(file = file.path(PlotsOutPLOS, "Fig10.tiff"), device = NULL, scale = 1, height=55, width= 55, units = c("cm"),dpi = 300, compression = "lzw", limitsize = TRUE, bg='white')
```

```{r NichesMap, fig.height = 8, fig.width = 16, warning = FALSE}

RightSide <-ggarrange(BlankPanel,NichesFullO2North, NichesFullO2Equat, BlankPanel,  ncol = 1, nrow = 4, heights = c(1, 2,2,3))
FullO2 <- ggarrange(map, RightSide, ncol = 2, nrow = 1)
FullO2

RightSide <-ggarrange(BlankPanel, NichesOMZNorth, NichesOMZEquat, BlankPanel,  ncol = 1, nrow = 4 , heights = c(1, 2,2,3))
FullOMZ <- ggarrange(map, RightSide,  ncol = 2, nrow = 1)
FullOMZ

NichesMap<- ggarrange(FullO2, FullOMZ, ncol = 1, nrow = 2,
          labels = c("A", "B"),  
          font.label = list(size = 30, color = "black"),
          align = "hv")
NichesMap

ggsave(file = file.path(PlotsOut, paste("NichesMap",".png",sep = "")), plot = NichesMap, height=18, width= 20,  dpi = 300, limitsize = TRUE,  bg='white')

ggsave(file = file.path(PlotsOutPLOS, "Fig10.tiff"), plot = NichesMap, device = NULL, scale = 1, height=55, width= 55, units = c("cm"),dpi = 300, compression = "lzw", limitsize = TRUE, bg='white')
```

```{r, fig.height = 8, fig.width = 16, warning = FALSE}

P1<-ggdraw() +
  draw_plot(NichesFullO2North, x = 0.001, y = 0.750, width = .350, height = .22) +
  draw_plot(NichesFullO2Equat, x = 0.001, y = 0.510, width = .350, height = .22) +
  draw_plot(mapCircle, x = .350, y = .28, width = 0.7, height = 0.7) 
P1

P2<-ggdraw() +
  draw_plot(NichesOMZNorth, x = 0.001, y = 0.750, width = .350, height = .22) +
  draw_plot(NichesOMZEquat, x = 0.001, y = 0.510, width = .350, height = .22) +
  draw_plot(mapCircle, x = .350, y = .28, width = 0.7, height = 0.7) 
P2

ggarrange(P1, P2, ncol = 1, nrow = 2)

ggsave(file = file.path(PlotsOut, paste("NichesMap_4",".png",sep = "")), height=18, width= 21,  dpi = 300, limitsize = TRUE)

```

-----------------------------------------------------------------------------------------------------------------
------------------------------------------- SySl End ----------------------------------------------------------
---------------------------------------------------------------------------------------------------------------